### 2022-02-02 figures_01302022.Rmd
connect local folder to existing repo test

---
title: "Figures"
author: "Christopher Cook"
date: "1/21/2022"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(results = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

```{r, include = FALSE}
library(tidyverse)
library(data.table)
devtools::load_all(".")
```

```{r}
#############
### histogram showing data spread for invivoPKfit cp values
#############

data.params.1 <- read.csv("inst/ext/PK.fit.table.1comp.12212021.csv")
data.params.2 <- read.csv("inst/ext/PK.fit.table.2comp.12212021.csv")

data.set.1 <- read.csv("inst/ext/processed.1comp.data.12212021.csv")
data.set.2 <- read.csv("inst/ext/processed.2comp.data.12212021.csv")

data.params.1 <- data.params.1 %>% filter(param.value.type == "Fitted geometric mean")
data.params.1$X <- NULL
data.params.1 <- data.params.1 %>% filter(Data.Analyzed != "Joint Analysis")

data.params.2 <- data.params.2 %>% filter(param.value.type == "Fitted geometric mean")
data.params.2$X <- NULL
data.params.2 <- data.params.2 %>% filter(Data.Analyzed != "Joint Analysis")

data.set.1$Reference <- as.character(data.set.1$Reference)
data.set.1$kelim <- NULL
data.set.1$Vdist <- NULL
data.set.1$Fgutabs <- NULL
data.set.1$kgutabs <- NULL

data.set.2$Reference <- as.character(data.set.2$Reference)
data.set.2$kelim <- NULL
data.set.2$Vdist <- NULL
data.set.2$Fgutabs <- NULL
data.set.2$kgutabs <- NULL

data.set.2$V1 <- NULL
data.set.2$Ralphatokelim <- NULL
data.set.2$Fbetaofalpha <- NULL

full.data.1 <- left_join(data.set.1, data.params.1, by = c("Compound", "CAS", "Reference"))
full.data.1 <- full.data.1 %>% filter(AIC != Inf)

full.data.2 <- left_join(data.set.2, data.params.2, by = c("Compound", "CAS", "Reference"))
full.data.2 <- full.data.2 %>% filter(AIC != Inf)

### function that find cp
test.1 <- function(full.data.1) {plyr::adply(full.data.1, 1, mutate, cp = cp_1comp(time = Time,
                                                                                   params = full.data.1[1,],
                                                                                   dose = Dose,
                                                                                   iv.dose = iv))}

test.2 <- function(full.data.2) {plyr::adply(full.data.2, 1, mutate, cp = cp_2comp(time = Time,
                                                                                   params = full.data.2[1,],
                                                                                   dose = Dose,
                                                                                   iv.dose = iv))}

### have to split data frame into ones with unique parameter values because cp_1comp takes constants
split_df.1 <- split(full.data.1, list(full.data.1$Compound,
                                      full.data.1$Reference,
                                      full.data.1$CAS),
                    drop = TRUE)

test_df.1 <- lapply(split_df.1, test.1)

maybe.1 <- do.call(rbind, test_df.1)

### nomralize cp by Value
maybe.1$cp_norm <- maybe.1$cp / maybe.1$Value
maybe.1$type <- "1-Compartment"

split_df.2 <- split(full.data.2, list(full.data.2$Compound,
                                      full.data.2$Reference,
                                      full.data.2$CAS),
                    drop = TRUE)

test_df.2 <- lapply(split_df.2, test.2)

maybe.2 <- do.call(rbind, test_df.2)

maybe.2$cp_norm <- maybe.2$cp / maybe.2$Value
maybe.2$type <- "2-Compartment"

# maybe.2 <- maybe.2 %>%
#   filter(cp != 1e-20)

maybe.1 <- maybe.1[, intersect(colnames(maybe.2), colnames(maybe.1))]
maybe.2 <- maybe.2[, intersect(colnames(maybe.1), colnames(maybe.2))]

maybe <- rbind(maybe.1, maybe.2)

p <- ggplot(data = maybe,
            mapping = aes(x = cp_norm)) +
  geom_histogram() +
  geom_vline(xintercept = c(0.5, 2),
             color = "Red",
             linetype = "dashed") +
  # xlim(0, 1e42) +
  # ylim(0, 400) +
  # scale_y_log10()+
  theme(title = element_text(size = 16),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14)) +
  scale_x_log10(limits = c(0.01, 1000))+
  theme_bw() +
  labs(x = "Predicted/Observed",
       y = "Number of Observations") +
  facet_wrap(. ~ type)

p
```

```{r}
### histogram of data variability

data <- as.data.table(read.csv("inst/ext/processed.1comp.data.12212021.csv"))
# data <- as.data.table(read.csv("inst/ext/processed.2comp.data.12212021.csv"))

# data <- data %>% filter(Route == "iv")

### function  that finds mean conc of each replicate timepoint
find_mean_conc <- function(data) {

  ### only use data with replicate timepoints
  data <- data[ ave(1:nrow(data), data$Time, FUN=length) > 1 , ]

  if(nrow(data) == 0) {
    return(NULL)
  } else {
    ### something that deletes replicate timepoints where all Values are NA
    time_split <- split(data, data$Time)

    # time_split_list <- keep(time_split, ~all(!is.na(.$Value)))

    time_split_list <- purrr::discard(time_split, ~all(is.na(.$Value)))
    # ~nrow(.) > 0)

    if(length(time_split_list) == 0) {
      return(NULL)
    } else {

      data <- do.call(rbind, time_split_list)
      # data$conc <- as.numeric(data$conc)
      data$Value[is.na(data$Value)] <- data$LOQ[is.na(data$Value)]

      ### take the average Value for each set of timepoints
      data <- data %>%
        group_by(Time) %>%
        do(mutate(., mean_conc = mean(.$Value, na.rm = TRUE)))

      data <- data %>%
        group_by(Time) %>%
        do(mutate(., range_conc = max(.$Value) - min(.$Value)))

      ### create normalized conc values by dividing conc by mean_conc
      data$conc_norm <- data$Value/data$mean_conc

      data <- data[order(data$Time),]

      return(data)
    }
  }
}

data.list <- list()
for(this.compound in unique(data[, Compound])) {
  this.compound.data <- data[Compound == this.compound]
  # print(this.compound)
  ### list of columns by which to split data.set
  col_list <- list(this.compound.data$Reference,
                   this.compound.data$Dose,
                   this.compound.data$Route,
                   this.compound.data$Media)

  ### split data.set into list of data.tables
  split_list <- split(this.compound.data, col_list)

  ### remove data.tables with 0 rows
  split_list <- keep(split_list, ~nrow(.) > 0)

  mean_conc <- lapply(split_list, find_mean_conc)
  test <- do.call(rbind, mean_conc)

  data.list[[this.compound]] <- test
}

### combine data.tables back into one large data.set
data.set <- do.call(rbind, data.list)

plot_fxn <- function(i){
  test_plot <- ggplot(data = i,
                      mapping = aes(x = conc_norm)) +
    # geom_histogram(aes(y = ..density..), binwidth = 0.25) +    geom_histogram(aes(y = ..density..), binwidth = 0.25) +
    # geom_density() +
    geom_histogram() +
    geom_vline(xintercept = c(0.5, 2),
               color = "Red",
               linetype = "dashed") +
    xlim(c(0, 4)) +
    theme_bw() +
    # theme(title = element_text(size = 16),
    #       axis.text = element_text(size = 12),
    #       axis.title = element_text(size = 14)) +
    labs(x = "Normalized concentration",
         y = "Number of observations")
}

my_hist <- plot_fxn(data.set)
my_hist

nrow(data.set[data.set$conc_norm > 2, ])
nrow(data.set[data.set$conc_norm < 0.5, ])

nrow(data.set[data.set$conc_norm >= 0.5 & data.set$conc_norm <= 2, ]) / nrow(data.set)

nrow(data.set)

```

```{r}

### Model Error vs Data Variability Density Plot

rownames(maybe.1) <- c()
rownames(maybe.2) <- c()

data.set$Reference <- as.character(data.set$Reference)

data.set$X <- NULL

data.set$kelim <- NULL
data.set$Vdist <- NULL
data.set$Fgutabs <- NULL
data.set$kgutabs <- NULL

data.set$type <- "1-Compartment"
test.1 <- left_join(data.set, maybe.1)

### trying to find number of points in each plot panel for 1-compartment
test.1.center <- test.1 %>% filter(conc_norm >= 0.5 & conc_norm <= 2)
test.1.center <- test.1.center %>% filter(cp_norm >= 0.5 & cp_norm <= 2)
ratio.1.center <- nrow(test.1.center)/nrow(test.1 %>% filter(!is.na(cp_norm)))

test.1.green <- test.1 %>% filter(conc_norm < 0.5 | conc_norm > 2)
test.1.green <- test.1.green %>% filter(cp_norm < 0.5 | cp_norm > 2)
ratio.1.green <- nrow(test.1.green)/nrow(test.1 %>% filter(!is.na(cp_norm)))

test.1.gray <- test.1 %>% filter(conc_norm < 0.5 | conc_norm > 2)
test.1.gray <- test.1.gray %>% filter(cp_norm >= 0.5 & cp_norm <= 2)
ratio.1.gray <- nrow(test.1.gray)/nrow(test.1 %>% filter(!is.na(cp_norm)))

test.1.tan <- test.1 %>% filter(conc_norm >= 0.5 & conc_norm <= 2)
test.1.tan <- test.1.tan %>% filter(cp_norm < 0.5 | cp_norm > 2)
ratio.1.tan <- nrow(test.1.tan)/nrow(test.1 %>% filter(!is.na(cp_norm)))

###########
###########

data.set$type <- "2-Compartment"
test.2 <- left_join(data.set, maybe.2)

### trying to find number of points in each plot panel for 2-compartment
test.2.center <- test.2 %>% filter(conc_norm >= 0.5 & conc_norm <= 2)
test.2.center <- test.2.center %>% filter(cp_norm >= 0.5 & cp_norm <= 2)
ratio.2.center <- nrow(test.2.center)/nrow(test.2 %>% filter(!is.na(cp_norm)))

test.2.green <- test.2 %>% filter(conc_norm < 0.5 | conc_norm > 2)
test.2.green <- test.2.green %>% filter(cp_norm < 0.5 | cp_norm > 2)
ratio.2.green <- nrow(test.2.green)/nrow(test.2 %>% filter(!is.na(cp_norm)))

test.2.gray <- test.2 %>% filter(conc_norm < 0.5 | conc_norm > 2)
test.2.gray <- test.2.gray %>% filter(cp_norm >= 0.5 & cp_norm <= 2)
ratio.2.gray <- nrow(test.2.gray)/nrow(test.2 %>% filter(!is.na(cp_norm)))

test.2.tan <- test.2 %>% filter(conc_norm >= 0.5 & conc_norm <= 2)
test.2.tan <- test.2.tan %>% filter(cp_norm < 0.5 | cp_norm > 2)
ratio.2.tan <- nrow(test.2.tan)/nrow(test.2 %>% filter(!is.na(cp_norm)))

# test.2 <- test.2 %>% filter(cp_norm > 1e-12)

test <- rbind(test.1, test.2)

p <- ggplot(data = test,
            mapping = aes(x = conc_norm, y = cp_norm)) +
  annotate(geom = "rect", xmin = 0, xmax = 0.5, ymin = 2, ymax = Inf, alpha = 0.3, fill = "#86A397") +
  annotate(geom = "rect", xmin = 0, xmax = 0.5, ymin = 0, ymax = 0.5, alpha = 0.3, fill = "#86A397") +
  annotate(geom = "rect", xmin = 2, xmax = Inf, ymin = 0, ymax = 0.5, alpha = 0.3, fill = "#86A397") +
  annotate(geom = "rect", xmin = 2, xmax = Inf, ymin = 2, ymax = Inf, alpha = 0.3, fill = "#86A397") +
  annotate(geom = "rect", xmin = 0, xmax = 0.5, ymin = 0.5, ymax = 2, alpha = 0.3, fill = "#361D2E") +
  annotate(geom = "rect", xmin = 2, xmax = Inf, ymin = 0.5, ymax = 2, alpha = 0.3, fill = "#361D2E") +
  annotate(geom = "rect", xmin = 0.5, xmax = 2, ymin = 2, ymax = Inf, alpha = 0.3, fill = "#E1B07E") +
  annotate(geom = "rect", xmin = 0.5, xmax = 2, ymin = 0, ymax = 0.5, alpha = 0.3, fill = "#E1B07E") +
  theme_bw() +
  theme(panel.spacing = unit(2, "lines")) +
  # stat_density_2d() +
  geom_bin2d(bins = 100) +
  scale_fill_distiller(palette = "Blues") +
  geom_vline(xintercept = c(0.5, 2),
             linetype = "dashed") +
  geom_hline(yintercept = c(0.5, 2),
             linetype = "dashed") +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = "Data Variability",
       y = "Model Error") +
  facet_wrap(. ~type)

p
```

```{r}

### sigma distibution histogram

data1 <- read.csv("inst/ext/PK.fit.table.1comp.12212021.csv") %>%
  filter(param.value.type == "Fitted geometric mean" & AIC != Inf & Data.Analyzed != "Joint Analysis")

data2 <- read.csv("inst/ext/PK.fit.table.2comp.12212021.csv") %>%
  filter(param.value.type == "Fitted geometric mean" & AIC != Inf & Data.Analyzed != "Joint Analysis")

data1 <- subset(data1, select = sigma_value)
data1$model <- 1

data2 <- subset(data2, select = sigma_value)
data2$model <- 2

data <- rbind(data1, data2)

test_plot <- ggplot(data = data,
                    mapping = aes(x = sigma_value)) +
  # geom_histogram(binwidth = 0.25) +
  geom_histogram() +
  # geom_density() +
  # scale_y_log10()+
  scale_x_log10() +
  theme_bw() +
  labs(x = "Sigma",
       y = "Number of observations") +
  facet_wrap(.~ model)
  # stat_function(fun = dlnorm, args = list(mean = mean(log(data$sigma_value)), sd = sd(log(data$sigma_value))))

test_plot
```

```{r}

### Orders of magnitude histogram

data <- read.csv("inst/ext/processed.1comp.data.12212021.csv")

data$X <- NULL

split_df <- split(data, list(data$Compound,
                             data$Reference,
                             data$CAS,
                             data$Dose,
                             data$Species,
                             data$Route,
                             data$Media),
                  drop = TRUE)

find_mag <- function(data) {

  data$Value <- abs(data$Value)

  diff <- log10(max(data$Value, na.rm = TRUE)) - log10(min(data$Value, na.rm = TRUE))

  diff <- round(diff)

  return(diff)

}

thing <- lapply(split_df, find_mag)

thing <- pivot_longer(as.data.frame(thing), cols = everything())

p <- ggplot(data = thing, mapping = aes(x = value)) +
  geom_histogram(bins = 5, binwidth = 0.5) +
  xlab("Orders of magnitude") + theme_bw()
p
```

```{r}

### number of observations vs sigma value

one_comp_data <- read.csv("inst/ext/processed.1comp.data.12212021.csv") %>%
  filter(if_all(Value, ~!is.na(.x)))

onecomp <- read.csv("inst/ext/PK.fit.table.1comp.12212021.csv") %>%
  filter(AIC != Inf)

two_comp_data <- read.csv("inst/ext/processed.2comp.data.12212021.csv") %>%
  filter(if_all(Value, ~!is.na(.x)))

twocomp <- read.csv("inst/ext/PK.fit.table.2comp.12212021.csv") %>%
  filter(AIC != Inf)

one_comp_count <- one_comp_data %>% count(CAS, Compound, Reference, name = "Observations")

two_comp_count <- two_comp_data %>% count(CAS, Compound, Reference, name = "Observations")

#######
#######

onecomp_no_joint <- onecomp %>%
  filter(param.value.type == "Fitted geometric mean" & Data.Analyzed != "Joint Analysis" & AIC != Inf)

one_comp_count$Reference <- as.character(one_comp_count$Reference)

together_one <- left_join(one_comp_count, onecomp_no_joint, by = c("CAS", "Compound", "Reference"))

together_one <- subset(together_one, select = c(CAS, Compound, Reference, Observations, sigma_id, sigma_value))

together_one$model <- 1

twocomp_no_joint <- twocomp %>%
  filter(param.value.type == "Fitted geometric mean" & Data.Analyzed != "Joint Analysis" & AIC != Inf)

two_comp_count$Reference <- as.character(two_comp_count$Reference)

together_two <- left_join(two_comp_count, twocomp_no_joint, by = c("CAS", "Compound", "Reference"))

together_two <- subset(together_two, select = c(CAS, Compound, Reference, Observations, sigma_id, sigma_value))

together_two$model <- 2

together <- rbind(together_one, together_two)

p <- ggplot(data = together, mapping = aes(x = Observations, y = sigma_value)) +
  geom_point() +
  scale_y_log10() +
  labs(x = "Number of observations",
       y = "Sigma") +
  theme_bw() +
  facet_wrap(.~ model)

p
```

```{r}

### model error vs sigma

p <- ggplot(data = maybe, mapping = aes(x = sigma_value, y = cp_norm)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  labs( x = "Sigma",
        y = "Predicted/Obseved") +
  theme_bw() +
  facet_wrap(.~ model)

p

```

```{r}

### sigma estimates, joint vs individual

onecomp <- as.data.table(read.csv("inst/ext/PK.fit.table.1comp.12212021.csv") %>%
  filter(param.value.type == "Fitted geometric mean"))

twocomp <- as.data.table(read.csv("inst/ext/PK.fit.table.2comp.12212021.csv") %>%
  filter(param.value.type == "Fitted geometric mean"))

data_1 <- onecomp
data_2 <- twocomp

find_rep_sigma <- function(data) {

  ### only use data with replicate timepoints
  data <- data[ ave(1:nrow(data), data$sigma_id, FUN=length) > 1 , ]

  if(nrow(data) == 0) {
    return(NULL)
  } else {

    return(data)
  }
}

split_list_1 <- split(data_1, data_1$Compound)
rep_sigma_1 <- lapply(split_list_1, find_rep_sigma)
data_1 <- do.call(rbind, rep_sigma_1)
data_1 <- subset(data_1, select = c(param.value.type, sigma_id, sigma_value, Data.Analyzed, Compound))
data_1$Data.Analyzed[data_1$Data.Analyzed != "Joint Analysis"] <- "Ind Analysis"
data_1 <- data_1 %>% pivot_wider(names_from = Data.Analyzed, values_from = sigma_value)
data_1 <- data_1 %>% filter(sigma_id != "sigma2.14")
data_1$`Joint Analysis` <- as.numeric(data_1$`Joint Analysis`)
data_1$`Ind Analysis` <- as.numeric(data_1$`Ind Analysis`)
data_1$model <- "1-Compartment"

split_list_2 <- split(data_2, data_2$Compound)
rep_sigma_2 <- lapply(split_list_2, find_rep_sigma)
data_2 <- do.call(rbind, rep_sigma_2)
data_2 <- subset(data_2, select = c(param.value.type, sigma_id, sigma_value, Data.Analyzed, Compound))
data_2$Data.Analyzed[data_2$Data.Analyzed != "Joint Analysis"] <- "Ind Analysis"
data_2 <- data_2 %>% pivot_wider(names_from = Data.Analyzed, values_from = sigma_value)
data_2 <- data_2 %>% filter(sigma_id != "sigma2.14")
data_2$`Joint Analysis` <- as.numeric(data_2$`Joint Analysis`)
data_2$`Ind Analysis` <- as.numeric(data_2$`Ind Analysis`)
data_2$model <- "2-Compartment"

data <- rbind(data_1, data_2)

p <- ggplot(data, aes(x = `Ind Analysis`, y = `Joint Analysis`)) +
  geom_point() +
  scale_y_log10() +
  scale_x_log10() +
  geom_abline(slope = 1) +
  facet_wrap(. ~ model)
p
```

```{r}
data1 <- read.csv("inst/ext/processed.1comp.data.12212021.csv")

data1$X <- NULL

# data1 <- data1 %>% filter(Compound == "tamoxifen")

split_df_1 <- split(data1, list(data1$Compound,
                             data1$Reference,
                             data1$CAS,
                             data1$Dose,
                             data1$Species,
                             data1$Route,
                             data1$Media),
                  drop = TRUE)

find_normalized_max <- function(data) {

  data$normalized_max <- max(data$Value, na.rm = TRUE)/data$LOQ

  return(data)

}

thing <- lapply(split_df_1, find_normalized_max)

thing <- do.call(rbind, thing)

row.names(thing) <- c()

thing <- thing %>%
  distinct(Compound,
           Reference,
           CAS,
           Dose,
           Species,
           Route,
           Media,
           normalized_max)

# thing <- pivot_longer(as.data.frame(thing), cols = everything())

p1 <- ggplot(data = thing, mapping = aes(x = normalized_max)) +
  geom_histogram(bins = 15) +
  scale_x_log10() +
  xlab("Max Concentration / LOQ") + theme_bw() +
  geom_vline(xintercept = 24.8482,
             color = "red",
             linetype = "dashed")
p1

###################################
###################################

find_normalized_mean <- function(data) {

  data$normalized_mean <- mean(data$Value, na.rm = TRUE)/data$LOQ

  return(data)

}

thing <- lapply(split_df_1, find_normalized_mean)

thing <- do.call(rbind, thing)

p2 <- ggplot(data = thing, mapping = aes(x = normalized_mean)) +
  geom_histogram(bins = 15) +
  scale_x_log10() +
  xlab("Mean Concentration / LOQ") + 
  theme_bw() +
  geom_vline(xintercept = 7.395552,
             color = "red",
             linetype = "dashed")
p2

###################################
###################################

data.params.1 <- read.csv("inst/ext/PK.fit.table.1comp.12212021.csv")
# data.params.2 <- read.csv("inst/ext/PK.fit.table.2comp.12212021.csv")

data.set.1 <- read.csv("inst/ext/processed.1comp.data.12212021.csv")
# data.set.2 <- read.csv("inst/ext/processed.2comp.data.12212021.csv")

data.params.1 <- data.params.1 %>% filter(param.value.type == "Fitted geometric mean")
data.params.1$X <- NULL
data.params.1 <- data.params.1 %>% filter(Data.Analyzed != "Joint Analysis")

# data.params.2 <- data.params.2 %>% filter(param.value.type == "Fitted geometric mean")
# data.params.2$X <- NULL
# data.params.2 <- data.params.2 %>% filter(Data.Analyzed != "Joint Analysis")

data.set.1$Reference <- as.character(data.set.1$Reference)
data.set.1$kelim <- NULL
data.set.1$Vdist <- NULL
data.set.1$Fgutabs <- NULL
data.set.1$kgutabs <- NULL

# data.set.2$Reference <- as.character(data.set.2$Reference)
# data.set.2$kelim <- NULL
# data.set.2$Vdist <- NULL
# data.set.2$Fgutabs <- NULL
# data.set.2$kgutabs <- NULL
# 
# data.set.2$V1 <- NULL
# data.set.2$Ralphatokelim <- NULL
# data.set.2$Fbetaofalpha <- NULL

full.data.1 <- left_join(data.set.1, data.params.1, by = c("Compound", "CAS", "Reference"))
full.data.1 <- full.data.1 %>% filter(AIC != Inf)

# full.data.2 <- left_join(data.set.2, data.params.2, by = c("Compound", "CAS", "Reference"))
# full.data.2 <- full.data.2 %>% filter(AIC != Inf)

split.full.data.1 <- split(full.data.1, list(full.data.1$Compound,
                             full.data.1$Reference,
                             full.data.1$CAS,
                             full.data.1$Dose,
                             full.data.1$Species.x,
                             full.data.1$Route,
                             full.data.1$Media),
                  drop = TRUE)

find_halflife_count <- function(data) {

  data_subset <- data %>% filter(!is.na(Value))

  max_time <- max(unique(data_subset$Time))

  data$halflife_count <- max_time/data$halflife

  return(data)

}

thing <- lapply(split.full.data.1, find_halflife_count)

thing <- do.call(rbind, thing)

thing <- thing %>%
  distinct(Compound,
           Reference,
           CAS,
           Dose,
           Species.x,
           Route,
           Media,
           halflife_count)

p3 <- ggplot(data = thing, mapping = aes(x = halflife_count)) +
  geom_histogram() +
  # scale_x_log10() +
  xlab("halflife count") + 
  theme_bw() +
  geom_vline(xintercept = 330.7598,
             color = "red",
             linetype = "dashed")
p3

```

```{r}

### bunch of polymeric stuff that might be useless

# data <- read.csv("inst/ext/processed.1comp.data.12212021.csv")
# 
# data$X <- NULL
# 
# split_df <- split(data, list(data$Compound,
#                              data$Reference,
#                              data$CAS,
#                              data$Dose,
#                              data$Species,
#                              data$Route,
#                              data$Media,
#                              data$LOQ),
#                   drop = TRUE)
# 
# # split_df <- purrr::discard(split_df, ~all(is.na(.$Value)))
# 
# # split_df <- purrr::discard(split_df, ~sum(is.na(.$Value)) < 3)
# 
# fit_poly <- function(data) {
# 
#   tryCatch({
# 
#     line <- lm(data = data, formula = Value ~ poly(Time))
# 
#     RSS <- c(crossprod(line$residuals))
#     MSE <- RSS / length(line$residuals)
#     RMSE <- sqrt(MSE)
# 
#     data$RMSE <- RMSE
# 
#     return(data)
#   },
#   error = function(cond) {
#     message("poly fit failed.")
#     return(NULL)
#   })
# 
# }
# 
# work <- lapply(split_df, fit_poly)
# 
# yay <- do.call(rbind, work)
# 
# yay$Reference <- as.character(yay$Reference)
# full <- left_join(yay, maybe, by = c(c("Species" = "Species.x"),
#                                      "Compound",
#                                      "CAS",
#                                      "Reference",
#                                      "Dose",
#                                      "Route",
#                                      "Media",
#                                      "Source"))
# 
# full <- full %>%
#   filter(!is.na(type))
# just_rmse <- distinct(full, Compound, RMSE)
# #
# p1 <- ggplot(data = just_rmse, mapping = aes(x = RMSE)) +
#   geom_histogram(bins = 50) +
#   scale_x_log10() +
#   theme_bw() +
#   labs(title = "RMSE of polynomial models fit through series data")
# 
# p1
# 
# p2 <- ggplot(full, aes(x = RMSE, y = sigma_value)) +
#   geom_point() +
#   geom_abline() +
#   scale_y_log10() +
#   scale_x_log10() +
#   theme_bw() +
#   facet_wrap(.~ type)
# 
# 
# p2
# 
# 
# p3 <- ggplot(data = full, mapping = aes(x = RMSE, y = cp_norm)) +
#   # geom_point() +
#   geom_bin2d() +
#   scale_x_log10() +
#   scale_y_log10() +
#   labs(x = "RMSE",
#        y = "Predicted/Obseved") +
#   theme_bw() +
#   facet_wrap(.~ model)
# 
# p3

```

## Discussion