### 2022-02-02 figures_01312020.Rmd

---
title: "<i>invivoPKfit</i> and <i>CvTdb</i> figures"
author: "Christopher Cook"
date: "1/31/2022"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(results = TRUE)
knitr::opts_chunk$set(message = FALSE)
```

```{r}
rm(list = ls())
```

```{r, include = FALSE}
library(tidyverse)
library(data.table)
devtools::load_all(".")
```

```{r, run_invivoPKfit, eval = FALSE}

#################
#################

data <- read.csv("inst/ext/cvt_query_for_john_27may2021_v2.csv")

#################
#################

### fix compound names
data$analyte_name_original[data$analyte_name_original == "Methyleugonal "] <- "Methyleugonal"
data$analyte_name_original[data$analyte_name_original == "benzo[a]pyrene"] <- "benzo(a)pyrene"
data$analyte_name_original[data$analyte_name_original == "2,4-D"] <- "2,4-Dichlorophenoxyacetic acid"
data$analyte_name_original[data$analyte_name_original == "Diclofenac"] <- "Hexobarbital"
data$analyte_name_original[data$analyte_name_original == "Pentadecafluorooctanoic acid"] <- "Perfluorooctanoic acid"

### make fk_reference_document equal fk_extraction document if fk_reference_document is NULL
data$fk_reference_document_id[data$fk_reference_document_id == "NULL"] <- data$fk_extraction_document_id[data$fk_reference_document_id == "NULL"]

### do 1-Compartment fit
system.time(PK.fit.table.1comp <- fit_all(

  data.set = data,
  model = "1compartment",
  modelfun = "analytic",
  compound.col = "analyte_name_original",
  cas.col = "dsstox_casrn",
  reference.col = "fk_reference_document_id",
  species.col = "species",
  species.weight.col = "weight_kg",
  species.weight.units.default = "kg",
  dose.col = "dose_level_normalized",
  time.col = "time_hr",
  time.units.default = "hr",
  media.col = "conc_medium_normalized",
  media.units.default = "normalized",
  value.col = "conc",
  units.default = NA,
  route.col = "administration_route_normalized",
  source.col = "fk_extraction_document_id",
  loq.default = NA,
  subject.default = NA,
  info.default = NA,
  ratio.data.to.dose = 0.001
))

### do 2-Compartment fit
system.time(PK.fit.table.2comp <- fit_all(

  data.set = data,
  model = "1compartment",
  modelfun = "analytic",
  compound.col = "analyte_name_original",
  cas.col = "dsstox_casrn",
  reference.col = "fk_reference_document_id",
  species.col = "species",
  species.weight.col = "weight_kg",
  species.weight.units.default = "kg",
  dose.col = "dose_level_normalized",
  time.col = "time_hr",
  time.units.default = "hr",
  media.col = "conc_medium_normalized",
  media.units.default = "normalized",
  value.col = "conc",
  units.default = NA,
  route.col = "administration_route_normalized",
  source.col = "fk_extraction_document_id",
  loq.default = NA,
  subject.default = NA,
  info.default = NA,
  ratio.data.to.dose = 0.001
))

write.csv(PK.fit.table.1comp[[1]] <- "inst/ext/PK.fit.table.1comp.DATE.csv")
write.csv(PK.fit.table.1comp[[2]] <- "inst/ext/processed.1comp.data.DATE.csv")

write.csv(PK.fit.table.2comp[[1]] <- "inst/ext/PK.fit.table.1comp.DATE.csv")
write.csv(PK.fit.table.2comp[[2]] <- "inst/ext/processed.1comp.data.DATE.csv")
```

```{r, make_rdata, eval = FALSE}

### save data
save(pk_output_1comp_12212021, file = "figures_rdata/pk_output_1comp_12212021.Rdata")
save(pk_output_2comp_12212021, file = "figures_rdata/pk_output_2comp_12212021.Rdata")
save(raw_1comp_12212021, file = "figures_rdata/raw_1comp_12212021.Rdata")
save(raw_2comp_12212021, file = "figures_rdata/raw_2comp_12212021.Rdata")

### save cp data
save(cp_df, file = "figures_rdata/cp_df.Rdata")
```

```{r, read_data}

### read in PK output files
pk_output_1comp_12212021 <- read.csv("inst/ext/PK.fit.table.1comp.12212021.csv")
pk_output_2comp_12212021 <- read.csv("inst/ext/PK.fit.table.2comp.12212021.csv")

### read in processed raw data files
raw_1comp_12212021 <- read.csv("inst/ext/processed.1comp.data.12212021.csv")
raw_2comp_12212021 <- read.csv("inst/ext/processed.2comp.data.12212021.csv")

### load all of this wonderful data
load("figures_rdata/pk_output_1comp_12212021.Rdata")
load("figures_rdata/pk_output_2comp_12212021.Rdata")
load("figures_rdata/raw_1comp_12212021.Rdata")
load("figures_rdata/raw_2comp_12212021.Rdata")
load("figures_rdata/cp_df_12212021.Rdata")
load("figures_rdata/cvt_variability_df_12212021.Rdata")
```

```{r, make_invivoPKfit_plots, eval = FALSE}

### 1 compartment
junk <- plot_conctime(PK.fit.table = as.data.table(pk_output_1comp_12212021),
                      data.set = as.data.table(raw_1comp_12212021),
                      model = "1compartment")

### 2 compartment
junk <- plot_conctime(PK.fit.table = as.data.table(pk_output_2comp_12212021),
                      data.set = as.data.table(raw_2comp_12212021),
                      model = "2compartment")

```

```{r make_cp_rdata, eval = FALSE}
################################
### histogram showing data spread of invivoPKfit cp values
################################

###########
### this is 1-compartment data
###########

###################
###################

### subset to only actual model outputs and discard "Joint Analysis" rows
pk_data1 <- pk_output_1comp_12212021 %>%
  filter(param.value.type == "Fitted geometric mean" & Data.Analyzed != "Joint Analysis")

raw_data1 <- dplyr::select(raw_1comp_12212021, -c("kelim", 
                                                  "Vdist", 
                                                  "Fgutabs", 
                                                  "kgutabs"))

###################
###################

raw_data11$Reference <- as.character(raw_data1$Reference)

full_data_1 <- left_join(raw_data1, pk_data1, by = c("Compound", "CAS", "Reference"))
full_data_1 <- full_data_1 %>% filter(AIC != Inf)

### function that finds cp
find_cp_1 <- function(full_data_1) {plyr::adply(full_data_1, 1, mutate, cp = cp_1comp(time = Time,
                                                                                      params = full_data_1[1,],
                                                                                      dose = Dose,
                                                                                      iv.dose = iv))}

### have to split data frame into ones with unique parameter values because cp_1comp takes constants
split_df_1 <- split(full_data_1, list(full_data_1$Compound,
                                      full_data_1$Reference,
                                      full_data_1$CAS),
                    drop = TRUE)

cp_df_1 <- lapply(split_df_1, find_cp_1)

cp_df_1 <- do.call(rbind, cp_df_1)

### nomralize cp by Value
cp_df_1$cp_norm <- cp_df_1$cp / cp_df_1$Value
cp_df_1$type <- "1-Compartment"

###########
### this is 2-compartment data
###########

#######################
#######################

### subset to only actual model outputs and discard "Joint Analysis" rows
pk_data2 <- pk_output_2comp_12212021 %>%
  filter(param.value.type == "Fitted geometric mean" & Data.Analyzed != "Joint Analysis")

raw_data2 <- dplyr::select(raw_2comp_12212021, -c("kelim", 
                                                  "Vdist", 
                                                  "Fgutabs", 
                                                  "kgutabs",
                                                  "V1",
                                                  "Ralphatokelim",
                                                  "Fbetaofalpha"))

########################
########################

raw_data2$Reference <- as.character(raw_data2$Reference)

full_data_2 <- left_join(raw_data2, pk_data2, by = c("Compound", "CAS", "Reference"))
full_data_2 <- full_data_2 %>% filter(AIC != Inf)
split_df_2 <- split(full_data_2, list(full_data_2$Compound,
                                      full_data_2$Reference,
                                      full_data_2$CAS),
                    drop = TRUE)

### function that finds cp
find_cp_2 <- function(full_data_2) {plyr::adply(full_data_2, 1, mutate, cp = cp_2comp(time = Time,
                                                                                      params = full_data_2[1,],
                                                                                      dose = Dose,
                                                                                      iv.dose = iv))}
cp_df_2 <- lapply(split_df_2, find_cp_2)

cp_df_2 <- do.call(rbind, cp_df_2)

cp_df_2$cp_norm <- cp_df_2$cp / cp_df_2$Value
cp_df_2$type <- "2-Compartment"

# cp_df_2 <- cp_df_2 %>%
#   filter(cp != 1e-20)

### make both data.frames have the same column names so they are rbind-able
cp_df_1 <- cp_df_1[, intersect(colnames(cp_df_2), colnames(cp_df_1))]
cp_df_2 <- cp_df_2[, intersect(colnames(cp_df_1), colnames(cp_df_2))]

cp_df <- rbind(cp_df_1, cp_df_2)

cp_df_12212021 <- cp_df

save(cp_df_12212021, file = "figures_rdata/cp_df_12212021.Rdata")
```

```{r, make_cp_histogram}

##################
##################

data <- cp_df_12212021

##################
##################

p <- ggplot(data = data,
            mapping = aes(x = cp_norm)) +
  geom_histogram(bins = 30) +
  geom_vline(xintercept = c(0.5, 2),
             color = "Red",
             linetype = "dashed") +
  # xlim(0, 1e42) +
  # ylim(0, 400) +
  # scale_y_log10()+
  theme(title = element_text(size = 16),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14)) +
  scale_x_log10(limits = c(0.01, 1000)) +
  theme_bw() +
  labs(x = "Predicted/Observed",
       y = "Number of Observations") +
  facet_wrap(. ~ type)

p
```

```{r, make_cvt_variability_rdata, eval = FALSE}

##################
##################

data <- as.data.table(raw_1comp_12212021)

##################
##################

### function that finds mean conc of each replicate timepoint
find_mean_conc <- function(data) {

  ### only use data with replicate timepoints
  data <- data[ ave(1:nrow(data), data$Time, fun = length) > 1 , ]

  if(nrow(data) == 0) {
    return(NULL)
  } else {

    time_split <- split(data, data$Time)

    ### discard replicate timepoints where all concentrations are NA
    time_split_list <- purrr::discard(time_split, ~all(is.na(.$Value)))

    if(length(time_split_list) == 0) {
      return(NULL)
    } else {

      data <- do.call(rbind, time_split_list)

      ### where some replicate concentration data are NA, set them equal to the LOQ
      data$Value[is.na(data$Value)] <- data$LOQ[is.na(data$Value)]

      ### take the average Value for each set of timepoints
      data <- data %>%
        group_by(Time) %>%
        do(mutate(., mean_conc = mean(.$Value, na.rm = TRUE)))

      data <- data %>%
        group_by(Time) %>%
        do(mutate(., range_conc = max(.$Value) - min(.$Value)))

      ### create normalized conc values by dividing conc by mean_conc
      data$conc_norm <- data$Value/data$mean_conc

      data <- data[order(data$Time), ]

      return(data)
    }
  }
}

data.list <- list()
for(this.compound in unique(data[, Compound])) {
  this.compound.data <- data[Compound == this.compound]

  ### list of columns by which to split data.set
  col_list <- list(this.compound.data$Reference,
                   this.compound.data$Dose,
                   this.compound.data$Route,
                   this.compound.data$Media)

  ### split data.set into list of data.tables
  split_list <- split(this.compound.data, col_list)

  ### remove data.tables with 0 rows
  split_list <- keep(split_list, ~nrow(.) > 0)

  ### use find_mean_conc fn
  mean_conc <- lapply(split_list, find_mean_conc)

  test <- do.call(rbind, mean_conc)

  data.list[[this.compound]] <- test
}

### combine data.tables back into one large data.set
cvt_variability_df_12212021 <- do.call(rbind, data.list)

save(cvt_variability_df_12212021, file = "figures_rdata/cvt_variability_df_12212021.Rdata")
```

```{r, make_cvt_variability_histogram}

#################
#################

data <- cvt_variability_df_12212021

#################
#################

p <- ggplot(data, aes(x = conc_norm)) +
  geom_histogram() +
  xlim(c(0, 4)) +
  geom_vline(xintercept = c(0.5, 2),
             color = "Red",
             linetype = "dashed") +
  theme_bw() +
  labs(x = "Normalized concentration",
       y = "Number of observations")
p
```

```{r, make_density_plot}

##################
##################

pk_data <- cp_df_12212021

cvt_data <- cvt_variability_df_12212021

##################
##################

row.names(pk_data) <- c()

cvt_data <- dplyr::select(cvt_data, -c("X",
                                       "kelim", 
                                       "Vdist", 
                                       "Fgutabs", 
                                       "kgutabs"))

### split pk_data by model so you can properly join with raw data
model_split <- split(pk_data, pk_data$model)

cvt_data$Reference <- as.character(cvt_data$Reference)

cvt_data$type <- "1-Compartment"
full_data_1 <- left_join(cvt_data, model_split$`1compartment`)

cvt_data$type <- "2-Compartment"
full_data_2 <- left_join(cvt_data, model_split$`2compartment`)

### put data back together
full_data <- rbind(full_data_1, full_data_2)

###########
###########
###########

### Find number of points in each plot panel for 1-compartment
full_data_1_center <- full_data_1 %>% filter(conc_norm >= 0.5 & conc_norm <= 2)
full_data_1_center <- full_data_1_center %>% filter(cp_norm >= 0.5 & cp_norm <= 2)
ratio_1_center <- nrow(full_data_1_center) / nrow(full_data_1 %>% filter(!is.na(cp_norm)))

full_data_1_green <- full_data_1 %>% filter(conc_norm < 0.5 | conc_norm > 2)
full_data_1_green <- full_data_1_green %>% filter(cp_norm < 0.5 | cp_norm > 2)
ratio_1_green <- nrow(full_data_1_green) / nrow(full_data_1_green %>% filter(!is.na(cp_norm)))

full_data_1_gray <- full_data_1 %>% filter(conc_norm < 0.5 | conc_norm > 2)
full_data_1_gray <- full_data_1_gray %>% filter(cp_norm >= 0.5 & cp_norm <= 2)
ratio_1_gray <- nrow(full_data_1_gray) / nrow(full_data_1 %>% filter(!is.na(cp_norm)))

full_data_1_tan <- full_data_1 %>% filter(conc_norm >= 0.5 & conc_norm <= 2)
full_data_1_tan <- full_data_1_tan %>% filter(cp_norm < 0.5 | cp_norm > 2)
ratio_1_tan <- nrow(full_data_1_tan) / nrow(full_data_1 %>% filter(!is.na(cp_norm)))

###########
###########

### Find number of points in each plot panel for 2-compartment
full_data_2_center <- full_data_2 %>% filter(conc_norm >= 0.5 & conc_norm <= 2)
full_data_2_center <- full_data_2_center %>% filter(cp_norm >= 0.5 & cp_norm <= 2)
ratio_2_center <- nrow(full_data_2_center) / nrow(full_data_2 %>% filter(!is.na(cp_norm)))

full_data_2_green <- full_data_2 %>% filter(conc_norm < 0.5 | conc_norm > 2)
full_data_2_green <- full_data_2 %>% filter(cp_norm < 0.5 | cp_norm > 2)
ratio_2_green <- nrow(full_data_2_green) / nrow(full_data_2_green %>% filter(!is.na(cp_norm)))

full_data_2_gray <- full_data_2 %>% filter(conc_norm < 0.5 | conc_norm > 2)
full_data_2_gray <- full_data_2_gray %>% filter(cp_norm >= 0.5 & cp_norm <= 2)
ratio_2_gray <- nrow(full_data_2_gray) / nrow(full_data_2 %>% filter(!is.na(cp_norm)))

full_data_2_tan <- full_data_2 %>% filter(conc_norm >= 0.5 & conc_norm <= 2)
full_data_2_tan <- full_data_2_tan %>% filter(cp_norm < 0.5 | cp_norm > 2)
ratio_2_tan <- nrow(full_data_2_tan) / nrow(full_data_2 %>% filter(!is.na(cp_norm)))

###########
###########
###########

p <- ggplot(data = full_data,
            mapping = aes(x = conc_norm, y = cp_norm)) +
  annotate(geom = "rect", xmin = 0, xmax = 0.5, ymin = 2, ymax = Inf, alpha = 0.3, fill = "#86A397") +
  annotate(geom = "rect", xmin = 0, xmax = 0.5, ymin = 0, ymax = 0.5, alpha = 0.3, fill = "#86A397") +
  annotate(geom = "rect", xmin = 2, xmax = Inf, ymin = 0, ymax = 0.5, alpha = 0.3, fill = "#86A397") +
  annotate(geom = "rect", xmin = 2, xmax = Inf, ymin = 2, ymax = Inf, alpha = 0.3, fill = "#86A397") +
  annotate(geom = "rect", xmin = 0, xmax = 0.5, ymin = 0.5, ymax = 2, alpha = 0.3, fill = "#361D2E") +
  annotate(geom = "rect", xmin = 2, xmax = Inf, ymin = 0.5, ymax = 2, alpha = 0.3, fill = "#361D2E") +
  annotate(geom = "rect", xmin = 0.5, xmax = 2, ymin = 2, ymax = Inf, alpha = 0.3, fill = "#E1B07E") +
  annotate(geom = "rect", xmin = 0.5, xmax = 2, ymin = 0, ymax = 0.5, alpha = 0.3, fill = "#E1B07E") +
  theme_bw() +
  theme(panel.spacing = unit(2, "lines")) +
  # stat_density_2d() +
  geom_bin2d(bins = 100) +
  scale_fill_distiller(palette = "Blues") +
  geom_vline(xintercept = c(0.5, 2),
             linetype = "dashed") +
  geom_hline(yintercept = c(0.5, 2),
             linetype = "dashed") +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = "Data Variability",
       y = "Model Error") +
  facet_wrap(. ~ type)

p
```

```{r, make_sigma_distribution_histogram}

##################
##################

data1 <- pk_output_1comp_12212021 %>%
  filter(param.value.type == "Fitted geometric mean" & AIC != Inf & Data.Analyzed != "Joint Analysis")

data2 <- pk_output_2comp_12212021 %>%
  filter(param.value.type == "Fitted geometric mean" & AIC != Inf & Data.Analyzed != "Joint Analysis")

##################
##################

### assign new column, with either "1-Compartment" or "2-Compartment"
data1 <- subset(data1, select = sigma_value)
data1$model <- "1-Compartment"

data2 <- subset(data2, select = sigma_value)
data2$model <- "2-Compartment"

data <- rbind(data1, data2)

p <- ggplot(data = data,
            mapping = aes(x = sigma_value)) +
  geom_histogram(bins = 15) +
  # geom_density() +
  # scale_y_log10()+
  scale_x_log10() +
  theme_bw() +
  labs(x = "Sigma",
       y = "Number of observations") +
  facet_wrap(.~ model)
# stat_function(fun = dlnorm, args = list(mean = mean(log(data$sigma_value)), sd = sd(log(data$sigma_value))))

p
```

```{r make_magnitude_histogram}

##################
##################

data <- raw_1comp_12212021

##################
##################

data$X <- NULL

split_df <- split(data, list(data$Compound,
                             data$Reference,
                             data$CAS,
                             data$Dose,
                             data$Species,
                             data$Route,
                             data$Media),
                  drop = TRUE)

### find number of orders of magnitude
find_mag <- function(data) {

  data$Value <- abs(data$Value)

  diff <- log10(max(data$Value, na.rm = TRUE)) - log10(min(data$Value, na.rm = TRUE))

  diff <- round(diff)

  return(diff)

}

data <- lapply(split_df, find_mag)

data <- pivot_longer(as.data.frame(data), cols = everything())

p <- ggplot(data = data, mapping = aes(x = value)) +
  geom_histogram(bins = 5, binwidth = 0.5) +
  xlab("Orders of magnitude") + theme_bw()

p
```

```{r make_joint_sigma_plot}

##################
##################

data1 <- as.data.table(pk_output_1comp_12212021 %>%
                         filter(param.value.type == "Fitted geometric mean"))

data2 <- as.data.table(pk_output_2comp_12212021 %>%
                         filter(param.value.type == "Fitted geometric mean"))

##################
##################

### find data with repeated sigma_id values
find_rep_sigma <- function(data) {

  data <- data[ ave(1:nrow(data), data$sigma_id, FUN=length) > 1 , ]

  if(nrow(data) == 0) {
    return(NULL)
  } else {

    return(data)
  }
}

split_list_1 <- split(data1, data1$Compound)
rep_sigma_1 <- lapply(split_list_1, find_rep_sigma)
data1 <- do.call(rbind, rep_sigma_1)
data1 <- subset(data1, select = c(param.value.type, sigma_id, sigma_value, Data.Analyzed, Compound))
data1$Data.Analyzed[data1$Data.Analyzed != "Joint Analysis"] <- "Ind Analysis"
data1 <- data1 %>% pivot_wider(names_from = Data.Analyzed, values_from = sigma_value)
data1 <- data1 %>% filter(sigma_id != "sigma2.14")
data1$`Joint Analysis` <- as.numeric(data1$`Joint Analysis`)
data1$`Ind Analysis` <- as.numeric(data1$`Ind Analysis`)
data1$model <- "1-Compartment"

split_list_2 <- split(data2, data2$Compound)
rep_sigma_2 <- lapply(split_list_2, find_rep_sigma)
data2 <- do.call(rbind, rep_sigma_2)
data2 <- subset(data2, select = c(param.value.type, sigma_id, sigma_value, Data.Analyzed, Compound))
data2$Data.Analyzed[data2$Data.Analyzed != "Joint Analysis"] <- "Ind Analysis"
data2 <- data2 %>% pivot_wider(names_from = Data.Analyzed, values_from = sigma_value)
data2 <- data2 %>% filter(sigma_id != "sigma2.14")
data2$`Joint Analysis` <- as.numeric(data2$`Joint Analysis`)
data2$`Ind Analysis` <- as.numeric(data2$`Ind Analysis`)
data2$model <- "2-Compartment"

data <- rbind(data1, data2)

p <- ggplot(data, aes(x = `Ind Analysis`, y = `Joint Analysis`)) +
  geom_point() +
  scale_y_log10() +
  scale_x_log10() +
  geom_abline(slope = 1) +
  facet_wrap(. ~ model)
p
```

```{r}

###########
### one compartment
###########

##################
##################

raw_data1 <- raw_1comp_12212021

##################
##################

pk_data1 <- pk_output_1comp_12212021 %>% 
  filter(param.value.type == "Fitted geometric mean" & Data.Analyzed != "Joint Analysis")

pk_data1$X <- NULL

raw_data1$Reference <- as.character(raw_data1$Reference)

raw_data1 <- dplyr::select(raw_data1, -c("X",
                                         "kelim",
                                         "Vdist",
                                         "Fgutabs",
                                         "kgutabs"))

# pk_data2 <- pk_output_2comp_12212021

full_data_1 <- left_join(raw_data1, pk_data1, by = c("Compound", "CAS", "Reference"))

full_data_1 <- full_data_1 %>% filter(AIC != Inf)

find_halflife_count <- function(data) {

  data_subset <- data %>% filter(!is.na(Value))

  max_time <- max(unique(data_subset$Time))

  data$halflife_count <- max_time/data$halflife

  return(data)

}

split_full_data_1 <- split(full_data_1, list(full_data_1$Compound,
                                             full_data_1$Reference,
                                             full_data_1$CAS,
                                             full_data_1$Dose,
                                             full_data_1$Species.x,
                                             full_data_1$Route,
                                             full_data_1$Media),
                           drop = TRUE)

full_data_1_hl <- lapply(split_full_data_1, find_halflife_count)

full_data_1_hl <- do.call(rbind, full_data_1_hl)

full_data_1_hl <- full_data_1_hl %>%
  distinct(Compound,
           Reference,
           CAS,
           Dose,
           Species.x,
           Route,
           Media,
           halflife_count)

full_data_1_hl$model <- "1-Compartment"

###########
### two compartment
###########

##################
##################

raw_data2 <- raw_2comp_12212021

pk_data2 <- pk_output_2comp_12212021 %>% 
  filter(param.value.type == "Fitted geometric mean" & Data.Analyzed != "Joint Analysis")

##################
##################

pk_data2$X <- NULL

raw_data2$Reference <- as.character(raw_data2$Reference)

raw_data2 <- dplyr::select(raw_data2, -c("kelim", 
                                         "Vdist", 
                                         "Fgutabs", 
                                         "kgutabs",
                                         "V1",
                                         "Ralphatokelim",
                                         "Fbetaofalpha"))

# pk_data2 <- pk_output_2comp_12212021

full_data_2 <- left_join(raw_data2, pk_data2, by = c("Compound", "CAS", "Reference"))

full_data_2 <- full_data_2 %>% filter(AIC != Inf)

find_halflife_count <- function(data) {

  data_subset <- data %>% filter(!is.na(Value))

  max_time <- max(unique(data_subset$Time))

  data$halflife_count <- max_time/data$halflife

  return(data)

}

split_full_data_2 <- split(full_data_2, list(full_data_2$Compound,
                                             full_data_2$Reference,
                                             full_data_2$CAS,
                                             full_data_2$Dose,
                                             full_data_2$Species.x,
                                             full_data_2$Route,
                                             full_data_2$Media),
                           drop = TRUE)

full_data_2_hl <- lapply(split_full_data_2, find_halflife_count)

full_data_2_hl <- do.call(rbind, full_data_2_hl)

full_data_2_hl <- full_data_2_hl %>%
  distinct(Compound,
           Reference,
           CAS,
           Dose,
           Species.x,
           Route,
           Media,
           halflife_count)

full_data_2_hl$model <- "2-Compartment"

###########
###########
###########

full_data_hl <- rbind(full_data_1_hl, full_data_2_hl)

###########

p <- ggplot(data = full_data_hl, mapping = aes(x = halflife_count)) +
  geom_histogram() +
  scale_x_log10() +
  xlab("halflife count") + 
  theme_bw() +
  # geom_vline(xintercept = 330.7598,
  #            color = "red",
  #            linetype = "dashed") + 
  facet_wrap(. ~ model)
p
```