### 2022-02-23 cook_sot_march.Rmd
copied poster files from main branch

---
title: <b>EPA's Concentration <i>versus</i> Time Database&colon; <br>
    A Resource for Extrapolating Toxicokinetic Trends Across Chemicals</b>
author:
  - name: Christopher Cook
    affil: 1,2
  - name: Lucas Albrecht
    affil: 1
  - name: Grace Cary
    affil: 1,2
  - name: Brenda Edwards
    affil: 1
  - name: Derik Haggard
    affil: 1
  - name: Nancy Hanley
    affil: 1
  - name: Michael F. Hughes
    affil: 1
  - name: Anna Jarnagin
    affil: 1,2
  - name: Tirumula D. Kodavanti
    affil: 1,2
  - name: Evgenia Koral-Bexell
    affil: 1,2
  - name: Anna Kreutz
    affil: 1,2
  - name: Mayla Ngo
    affil: 1,2
  - name: Caitlyn Patullo
    affil: 1,2
  - name: Caroline Ring
    affil: 1
  - name: Risa Sayre
    affil: 1,3
  - name: Bhaskar Sharma
    affil: 1
  - name: Jonathan Wall
    affil: 1
  - name: Hiroshi Yamazaki
    affil: 4
  - name: John Wambaugh
    affil: 1
affiliation:
  - num:
    address: "<br> ^1^Center for Computational Toxicology and Exposure, Office of Research and Development, United States Environmental Protection Agency, Durham, North Carolina"
  - num:
    address: "^2^Oak Ridge Institute for Science and Education, Oak Ridge, Tennessee"
  - num:
    address: "^3^Department of Environmental Sciences and Engineering, University of North Carolina at Chapel Hill, Chapel Hill, North Carolina"
  - num:
    address: "^4^Showa Pharmaceutical University, Tokyo, Japan"
column_numbers: 5
logoleft_name: "poster/epa_logo_white.png"
output: 
  posterdown::posterdown_html:
    self_contained: FALSE
titlebox_height: "18%"
body_textsize: "45px"
primary_colour: "#005780"
columnline_col: "#001d2b"
title_textsize: "100pt"
author_textsize: "50pt"
affilition_textsize: "30pt"
titlebox_bordercol: "##005780"
titlebox_borderwidth: "0.1in"
sectitle_bordercol: "#001d2b"
sectitle_borderwidth: "1mm"
font_family: Arimo
titletext_fontfamily: Arimo
bibliography: "poster/references.bib"
nocite: "@*"
poster_height: "43in"
poster_width: "89in"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

```{r, include = FALSE}
library(tidyverse)
library(data.table)
devtools::load_all(".")
```

```{r, load_data}

### load all of this wonderful data
load("figures_rdata/pk_output_1comp_12212021.Rdata")
load("figures_rdata/pk_output_2comp_12212021.Rdata")
load("figures_rdata/raw_1comp_12212021.Rdata")
load("figures_rdata/raw_2comp_12212021.Rdata")
load("figures_rdata/cp_df_12212021.Rdata")
load("figures_rdata/cvt_variability_df_12212021.Rdata")
load("figures_rdata/pk_lombardo_02072022.Rdata")
```

# 1 Expansion of the Concentration *versus* Time Database

## **1a) What is the CvTdb?**

* The Concentration *versus* Time Database  (CvTdb) is a database of standardized CvT data (time-course data of chemical compound concentrations measured *in vivo*) from hundreds of diverse publications, and serves as a platform for assessing PK/TK models on a large set of data. (Sayre, Grulke, and Wambaugh, 2020).

```{r, out.width="60%", fig.align="center", fig.cap="Typical plot from which CvT data is extracted. From Sowa and Steibert (1985)."}
knitr::include_graphics("poster/tk_curve.png")
```

* The CvTdb contains 181 unique analytes, including 54 pharmaceuticals (~30%). 96 analytes are present in EPA's ToxCast dataset (~53%).

## **1b) Integration of Data Shared by Other Institutions**

* Showa Pharmaceutical University shared CvT data consisting of over 200 unique compounds (Kamiya *et al.*, 2020; Kamiya *et al.*, 2021).

* If YOU have CvT data, please consider collaborating with us! Visit [https://github.com/USEPA/CompTox-PK-CvTdb](https://github.com/USEPA/CompTox-PK-CvTdb) or email Wambaugh.John@epa.gov

## **1c) In-House Curation Workflow**
1) Identify publications with CvT data using PubMed queries

2) Manually extract CvT data using [WebPlotDigitizer](https://automeris.io/WebPlotDigitizer/)

3) Curate CvT data and metadata using EPA's *.xlsx* template

4) Screen template using *CvTdbQA*, an R Shiny application that checks template for format and reconstructs plots based on reported CvT data

5) Manually check template for format and data correctness

6) Normalize metadata units using an R script workflow, while also retaining original values

7) Push template to CvTdb

# 2 Toxicokinetic Trends Across Chemicals

* We can explore trends across chemicals and references using the CvTdb.

* "In PBPK modeling, predictions that are, on average, within a **factor of 2** of the experimental
data have frequently been considered adequate" (WHO, 2010).

* In Figure 2, 84.9% of replicate timepoint data in the CvTdb is within a factor of two of their respective mean. The "factor of two" rule may not be appropriate if the data is spread beyond a factor of two.

```{r, make_cvt_variability_histogram}

#################
#################

data <- cvt_variability_df_12212021

#################
#################

p <- ggplot(data, aes(x = conc_norm)) +
  geom_histogram() +
  xlim(c(0, 4)) +
  geom_vline(xintercept = c(0.5, 2),
             color = "Red",
             linetype = "dashed") +
  theme_bw() +
  labs(x = "Normalized concentration",
       y = "Number of observations")
p
```

# 3 Introducing *invivoPKfit*

## **3a What is *invivoPKfit*?**

* R package that fits 1- and 2- compartment models to CvT data and extrapolates TK parameter values (e.g., half-life, elimination rate, and volume of distribution)

```{r, out.width="30%", fig.align="center", fig.cap="Typical plot from which CvT data is extracted. From Sowa and Steibert (1985)."}
knitr::include_graphics("poster/pk_workflow.png")
```

## **3b TK Model Descriptions**
|    **<span style="text-decoration:underline">1-Compartment Model</span>**

```{r, fig.cap = "1-Compartment Model. Adapted from Wambaugh et al. (2018).", fig.align = "center",  out.width = "60%"}
knitr::include_graphics("poster/1_comp.png")
```

$$
C_{iv}(dose,\ t)\ =\ \frac{dose}{V_d} e ^ {(-k_{elim}\ *\ t)}
$$

$$
C_{oral}(dose,\ t)\ =\ \frac{F_{bio}\ *\ dose}{V_{d}}\ *\ \frac{k_{abs}\ *\ (e ^ {-k_{elim}\ *\ t}\ -\ e ^ {-k_{abs}\ *\ t})}{k_{abs}\ -\ k_{elim}}   
$$

* Parameters include $V_d$ (volume of distribution), $k_{elim}$ (elimination rate), $k_{abs}$ (absorption rate; oral only), and $F_{bio}$ (fraction bioavailable; oral only).

|    **<span style="text-decoration:underline">2-Compartment Model</span>**

```{r, fig.cap = "2-Compartment Model. Adapted from Wambaugh et al. (2018).", fig.align = "center", out.width = "60%"}
knitr::include_graphics("poster/2_comp.png")
```

$$
C_{iv}(dose,\ t)\ =\ A\ *\ e^{-\alpha\ *\ t}\ +\ B\ *\ e^{-\beta\ *\ t}
$$  
$$
C_{oral}(dose,\ t)\ =\ A\ *\ e^{-\alpha\ *\ t}\ +\ B\ *\ e^{-\beta\ *\ t}\ +\ C\ *\ e^{-ka\ *\ t}
$$
* Parameters include $V1$ (volume of primary compartment), $k_{elim}$, $k_{12}$ (distribution rate to deep compartment), $k_{21}$ distribution rate from deep compartment), $k_{abs}$ (oral only), $F_{bio}$ (oral only).

## **3c Statistical Model Description**

* *invivoPKfit* calculates the probability of $C_{obs}$ given a PK model $M(p,\ t)$ and a statistical model $S(\sigma)$, where $P\ =\ L(C_{obs},\ M,\ S)$, and assumes that $C_{obs}$ are log-normally distributed around the concentrations predicted by $M(p,\ t)$ (Cox and Hinkley, 1979).

$$
ln\ L(x,\ \mu,\ \sigma)\ =\ \Sigma_{i=1,n}ln\frac{1}{x\sigma \sqrt{2\pi}}e^{-\frac{1}{2}\left( \frac{ln\ x\ -\ \mu)}{\sigma} \right)^2}
$$

$$
ln\ L(x,\ \mu,\ \sigma)\ =\ \Sigma_{i=1,n}ln\frac{1}{C_{obs,i}\sigma \sqrt{2\pi}}e^{-\frac{1}{2}\left( \frac{ln\ C_{obs,i}\ -\ M(p,t_i)}{\sigma} \right)^2}
$$

```{r, warning = FALSE, fig.align = "center", out.width = "60%", fig.cap = "Log-Normal Distribution"}
load(file = "poster/lognormal.Rdata")
p9
```

* The R package, [*optimx*](https://cran.r-project.org/web/packages/optimx/index.html) (Nash and Varadhan, 2011; Nash, 2014) optimizes the log-likelihood equation and estimates $p$ and $\sigma$

* We used the Hessian of the likelihood function to estimate standard deviations of the log parameter (Dovi, 1991)

# 4 ** *invivoPKfit* Analyses**

## **4a Predicted Concentrations**

* 58.4% of 12134 concentrations predicted by the 1-compartment model fit within a factor of two of the observed concentrations.

* 64.8% of 9684 concentrations predicted by the 2-compartment model fit within a factor of two of the observed concentrations.

```{r, make_cp_histogram}

##################
##################

data <- cp_df_12212021

##################
##################

p <- ggplot(data = data,
            mapping = aes(x = cp_norm)) +
  geom_histogram(bins = 30) +
  geom_vline(xintercept = c(0.5, 2),
             color = "Red",
             linetype = "dashed") +
  # xlim(0, 1e42) +
  # ylim(0, 400) +
  # scale_y_log10()+
  theme(title = element_text(size = 16),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14)) +
  scale_x_log10(limits = c(0.01, 1000)) +
  theme_bw() +
  labs(x = "Predicted/Observed",
       y = "Number of Observations") +
  facet_wrap(. ~ type)

p
```

## **4b *invivoPKfit* Performance *versus* Data Variability**

* Can start to gauge how data variability influences the performance of a model

```{r, fig.cap = "2-Compartment Model. Adapted from Wambaugh et al. (2018).", fig.align = "center", out.width = "100%"}
knitr::include_graphics("poster/density_plot.png")
```

## **4c Diagnosing Poor Fits**

# What's Next?

# References

# Acknowledgements/Disclaimers