### 2022-10-26 MainAnalysis.Rmd
Most of MainAnalysis.Rmd working now

---
title: "<i>invivoPKfit</i> and <i>CvTdb</i> figures"
author: "Christopher Cook"
date: "1/31/2022"
output:
  html_document: default
  word_document: default
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(results = TRUE)
knitr::opts_chunk$set(message = FALSE)
```

```{r, include = FALSE}
library(tidyverse)
library(data.table)
library(readxl)
devtools::load_all("invivopkfit")
```

```{r, load_cvt, eval = TRUE}

### this block runs the model fits
# and does some pre-processing to clean up the data

#################
#################

# data <- read.csv("inst/ext/cvt_query_for_john_27may2021_v2.csv")
# data <- read.csv("inst/ext/cvtdb_invivopkfit_02242022.csv")
# data <- read.csv("inst/ext/cvtdb_invivopkfit_04142022.csv")
cvt.data <- read.csv("inst/ext/cvtdb_invivopkfit_04262022.csv")

### these compounds had no CAS in the cvtdb
# double check and they should not have been added
# so I removed them

cvt.data <- cvt.data %>%
  filter(analyte_name_original != "Methemoglobin") %>%
  filter(analyte_name_original != "N-desmethyltamoxifen")

#################
#################

### fix compound names
cvt.data$analyte_name_original[cvt.data$analyte_name_original == "Methyleugonal "] <- "Methyleugonal"
cvt.data$analyte_name_original[cvt.data$analyte_name_original == "benzo[a]pyrene"] <- "benzo(a)pyrene"
cvt.data$analyte_name_original[cvt.data$analyte_name_original == "2,4-D"] <- "2,4-Dichlorophenoxyacetic acid"
cvt.data$analyte_name_original[cvt.data$analyte_name_original == "Diclofenac"] <- "Hexobarbital"
cvt.data$analyte_name_original[cvt.data$analyte_name_original == "Pentadecafluorooctanoic acid"] <- "Perfluorooctanoic acid"
cvt.data$analyte_name_original[cvt.data$analyte_name_original == "3−hydroxybenzo[a]pyrene"] <- "3−hydroxybenzo(a)pyrene"

### make fk_reference_document equal fk_extraction document if fk_reference_document is NULL
cvt.data$fk_reference_document_id[cvt.data$fk_reference_document_id == "NULL"] <- 
  cvt.data$fk_extraction_document_id[cvt.data$fk_reference_document_id == "NULL"]

### this is basically my debug line,
# where I can focus on one problematic compound 
# and see what's going wrong
# data <- data %>% filter(dsstox_casrn == "68291-97-4" | dsstox_casrn == "82626-48-0")
# data <- data %>% filter(dsstox_casrn == "123-91-1")
```

Calculate a limit of quantification (90% of lowest observation)
```{r, calc_loq, eval = TRUE}
source.col <- "fk_study_id"
cas.col <- "dsstox_casrn"
media.col <- "conc_medium_normalized"
value.col <- "conc"
for (this.study in unique(cvt.data[,source.col]))
{
  this.subset1 <- subset(cvt.data,cvt.data[,source.col]==this.study)
  for (this.chem in unique(this.subset1[,cas.col]))
  {
    this.subset2 <- subset(this.subset1,this.subset1[,cas.col]==this.chem)
    for (this.media in unique(this.subset2[,media.col]))
    {
      this.subset3 <- subset(this.subset2,this.subset2[media.col]==this.media & 
        !is.na(this.subset2[,value.col]))
      this.subset3 <- subset(this.subset3,this.subset3[,value.col]>0)
      this.loq <- suppressWarnings(min(this.subset3[,value.col], na.rm=TRUE))
      this.loq <- this.loq*0.9
      cvt.data[
        cvt.data[,source.col]==this.study & 
        cvt.data[,cas.col]==this.chem & 
        cvt.data[,media.col]==this.media,"calc_loq"] <- this.loq
    }
  }
}
cvt.data <- subset(cvt.data,!is.infinite(calc_loq))

```

```{r, run_invivoPKfit_1comp, eval = TRUE}
THIS.DATE <- "102422"
### do 1-Compartment fit
system.time(PK.fit.table.1comp <- fit_all(
  data.set = cvt.data,
  model = "1compartment",
  modelfun = "analytic",
  compound.col = "analyte_name_original",
  cas.col = "dsstox_casrn",
  reference.col = "fk_reference_document_id",
  species.col = "species",
  species.weight.col = "weight_kg",
  species.weight.units.default = "kg",
  dose.col = "dose_level_normalized",
  time.col = "time_hr",
  time.units.default = "hr",
  media.col = "conc_medium_normalized",
  media.units.default = "normalized",
  value.col = "conc",
  units.default = NA,
  route.col = "administration_route_normalized",
  source.col = "document_id",
  loq.default = NA,
  subject.default = NA,
  info.default = NA
))
```

```{r, plot_1comp, eval=FALSE}
plot_conctime(PK.fit.table.1comp[[1]],cvt.data,"1compartment")
head(subset(PK.fit.table.1comp[[1]],Vdist>10000&!is.infinite(AIC)&param.value.type=="Fitted geometric mean"))
```

```{r, run_invivoPKfit_2comp, eval = TRUE}
### do 2-Compartment fit
system.time(PK.fit.table.2comp <- fit_all(
  data.set = cvt.data,
  model = "2compartment",
  modelfun = "analytic",
  compound.col = "analyte_name_original",
  cas.col = "dsstox_casrn",
  reference.col = "fk_reference_document_id",
  species.col = "species",
  species.weight.col = "weight_kg",
  species.weight.units.default = "kg",
  dose.col = "dose_level_normalized",
  time.col = "time_hr",
  time.units.default = "hr",
  media.col = "conc_medium_normalized",
  media.units.default = "normalized",
  value.col = "conc",
  units.default = NA,
  route.col = "administration_route_normalized",
  source.col = "document_id",
  loq.default = NA,
  subject.default = NA,
  info.default = NA
))
```
```{r, plot_2comp, eval=FALSE}
plot_conctime(PK.fit.table.2comp[[1]],cvt.data,"2compartment")
head(subset(PK.fit.table.2comp[[1]],Vdist>10000&!is.infinite(AIC)&param.value.type=="Fitted geometric mean"))
```

```{r, run_invivoPKfit_flat, eval = TRUE}
### do flat fit
system.time(PK.fit.table.flat <- fit_all(
  data.set = cvt.data,
  model = "flat",
  modelfun = "analytic",
  compound.col = "analyte_name_original",
  cas.col = "dsstox_casrn",
  reference.col = "fk_reference_document_id",
  species.col = "species",
  species.weight.col = "weight_kg",
  species.weight.units.default = "kg",
  dose.col = "dose_level_normalized",
  time.col = "time_hr",
  time.units.default = "hr",
  media.col = "conc_medium_normalized",
  media.units.default = "normalized",
  value.col = "conc",
  units.default = NA,
  route.col = "administration_route_normalized",
  source.col = "document_id",
  loq.default = NA,
  subject.default = NA,
  info.default = NA
))
```

```{r, write_invivoPKfits_to_file, eval = TRUE}
write.csv(PK.fit.table.1comp[[1]], paste("inst/ext/PK.fit.table.1comp.",THIS.DATE,".csv",sep=""), row.names = FALSE)
write.csv(PK.fit.table.1comp[[2]], paste("inst/ext/processed.1comp.data.",THIS.DATE,".csv",sep=""), row.names = FALSE)

write.csv(PK.fit.table.2comp[[1]], paste("inst/ext/PK.fit.table.2comp.",THIS.DATE,".csv",sep=""), row.names = FALSE)
write.csv(PK.fit.table.2comp[[2]], paste("inst/ext/processed.2comp.data.",THIS.DATE,".csv",sep=""), row.names = FALSE)

write.csv(PK.fit.table.flat[[1]], paste("inst/ext/PK.fit.table.flat.",THIS.DATE,".csv",sep=""), row.names = FALSE)
write.csv(PK.fit.table.flat[[2]], paste("inst/ext/processed.flat.data.",THIS.DATE,".csv",sep=""), row.names = FALSE)
```

```{r, read_data}

### this block reads the exported model fit csv files from run_invivoPKfit

pk_output_2comp <- read.csv(paste("inst/ext/PK.fit.table.2comp.",THIS.DATE,".csv",sep=""))
pk_output_flat <- read.csv(paste("inst/ext/PK.fit.table.flat.",THIS.DATE,".csv",sep=""))
pk_output_1comp <- read.csv(paste("inst/ext/PK.fit.table.1comp.",THIS.DATE,".csv",sep=""))

raw_1comp <- read.csv(paste("inst/ext/processed.1comp.data.",THIS.DATE,".csv",sep=""))
raw_2comp <- read.csv(paste("inst/ext/processed.2comp.data.",THIS.DATE,".csv",sep=""))
raw_flat <- read.csv(paste("inst/ext/processed.flat.data.",THIS.DATE,".csv",sep=""))

pk_lombardo <- read_xlsx("inst/ext/Supplemental_82966_revised_corrected.xlsx")
```

```{r, make_rdata, eval = FALSE}

### this block takes the csv files in read_data
# and saves them as Rdata files

### save data
paste("inst/ext/PK.fit.table.1comp.",THIS.DATE,".RData",sep="")
save(pk_output_1comp, file = paste("figures_rdata/pk_output_1comp_",THIS.DATE,".RData",sep=""))
save(pk_output_2comp, file = paste("figures_rdata/pk_output_2comp_",THIS.DATE,".RData",sep=""))
save(pk_output_flat, file = paste("figures_rdata/pk_output_flat_",THIS.DATE,".RData",sep=""))
save(raw_1comp, file = paste("figures_rdata/raw_1comp_",THIS.DATE,".RData",sep=""))
save(raw_2comp, file = paste("figures_rdata/raw_2comp_",THIS.DATE,".RData",sep=""))
save(raw_flat, file = paste("figures_rdata/raw_flat_",THIS.DATE,".RData",sep=""))

save(pk_lombardo, file = paste("figures_rdata/pk_lombardo",THIS.DATE,".RData",sep=""))
```

```{r, make_invivoPKfit_plots, eval = TRUE}

### this block produces the pdf files of the model fits
# plotted on top of the original data

### 1 compartment

# pk_output_1comp_03012022 <- pk_output_1comp_03012022 %>%
#   filter(Compound == "tamoxifen") %>%
#   filter(Data.Analyzed != "Joint Analysis")

# raw_1comp_03012022 <- raw_1comp_03012022 %>%
#   filter(Compound == "tamoxifen")

junk <- plot_conctime(PK.fit.table = as.data.table(pk_output_1comp_05092022),
                      data.set = as.data.table(raw_1comp_05092022),
                      model = "1compartment")

### 2 compartment
junk <- plot_conctime(PK.fit.table = as.data.table(pk_output_2comp_05092022),
                      data.set = as.data.table(raw_2comp_05092022),
                      model = "2compartment")

junk <- plot_conctime(PK.fit.table = as.data.table(pk_output_flat_05092022),
                      data.set = as.data.table(raw_flat_05092022),
                      model = "flat")

```

```{r, make_cp_rdata, eval = FALSE}

### this block creates a data.frame of predicted concentrations
# using the estimated parameters and the observed conc-time points

######################

###########
### this is 1-compartment data
###########

###################
###################

### subset to only actual model outputs and discard "Joint Analysis" rows
pk_data1 <- pk_output_1comp %>%
  filter(param.value.type == "Fitted geometric mean" & Data.Analyzed != "Joint Analysis") %>% 
  filter(Species != "hamster")

raw_data1 <- dplyr::select(raw_1comp, -c("kelim", 
                                                  "Vdist", 
                                                  "Fgutabs", 
                                                  "kgutabs")) %>% 
  filter(Species != "hamster")

###################
###################

raw_data1$Reference <- as.character(raw_data1$Reference)

full_data_1 <- left_join(raw_data1, pk_data1, by = c("Compound", "CAS", "Reference"))
full_data_1 <- full_data_1 %>% filter(AIC != Inf)

### function that finds cp
find_cp_1 <- function(full_data_1) {plyr::adply(full_data_1, 1, mutate, cp = cp_1comp(time = Time,
                                                                                      params = full_data_1[1, ],
                                                                                      dose = Dose,
                                                                                      iv.dose = iv))}

### have to split data frame into ones with unique parameter values because cp_1comp takes constants
split_df_1 <- split(full_data_1, list(full_data_1$Compound,
                                      full_data_1$Reference,
                                      full_data_1$CAS,
                                      full_data_1$Species.x),
                    drop = TRUE)

cp_df_1 <- lapply(split_df_1, find_cp_1)

cp_df_1 <- do.call(rbind, cp_df_1)

### nomralize cp by Value
cp_df_1$cp_norm <- cp_df_1$cp / cp_df_1$Value
cp_df_1$type <- "1-Compartment"

###########
### this is 2-compartment data
###########

#######################
#######################

### subset to only actual model outputs and discard "Joint Analysis" rows
pk_data2 <- pk_output_2comp_05092022 %>%
  filter(param.value.type == "Fitted geometric mean" & Data.Analyzed != "Joint Analysis")

raw_data2 <- dplyr::select(raw_2comp_05092022, -c("kelim", 
                                                  "Vdist", 
                                                  "Fgutabs", 
                                                  "kgutabs",
                                                  "V1",
                                                  "Ralphatokelim",
                                                  "Fbetaofalpha"))

########################
########################

raw_data2$Reference <- as.character(raw_data2$Reference)

full_data_2 <- left_join(raw_data2, pk_data2, by = c("Compound", "CAS", "Reference"))
full_data_2 <- full_data_2 %>% filter(AIC != Inf)
split_df_2 <- split(full_data_2, list(full_data_2$Compound,
                                      full_data_2$Reference,
                                      full_data_2$CAS,
                                      full_data_2$Species.x),
                    drop = TRUE)

### function that finds cp
find_cp_2 <- function(full_data_2) {plyr::adply(full_data_2, 1, mutate, cp = cp_2comp(time = Time,
                                                                                      params = full_data_2[1, ],
                                                                                      dose = Dose,
                                                                                      iv.dose = iv))}
cp_df_2 <- lapply(split_df_2, find_cp_2)

cp_df_2 <- do.call(rbind, cp_df_2)

cp_df_2$cp_norm <- cp_df_2$cp / cp_df_2$Value
cp_df_2$type <- "2-Compartment"

# cp_df_2 <- cp_df_2 %>%
#   filter(cp != 1e-20)

###########
### this is flat data
###########

pk_data_flat <- pk_output_flat_05092022 %>%
  filter(param.value.type == "Fitted geometric mean" & Data.Analyzed != "Joint Analysis") %>% 
  filter(Species != "hamster")

raw_data_flat <- dplyr::select(raw_flat_05092022, -c("A")) %>% 
  filter(Species != "hamster")

raw_data_flat$Reference <- as.character(raw_data_flat$Reference)

full_data_flat <- left_join(raw_data_flat, pk_data_flat, by = c("Compound", "CAS", "Reference"))
full_data_flat <- full_data_flat %>% filter(AIC != Inf)
split_df_flat <- split(full_data_flat, list(full_data_flat$Compound,
                                            full_data_flat$Reference,
                                            full_data_flat$CAS,
                                            full_data_flat$Species.x),
                       drop = TRUE)

### function that finds cp
find_cp_flat <- function(full_data_flat) {plyr::adply(full_data_flat, 1, mutate, cp = cp_flat(time = Time,
                                                                                              params = full_data_flat[1, ],
                                                                                              dose = Dose,
                                                                                              iv.dose = iv))}
cp_df_flat <- lapply(split_df_flat, find_cp_flat)

cp_df_flat <- do.call(rbind, cp_df_flat)

cp_df_flat$cp_norm <- cp_df_flat$cp / cp_df_flat$Value
cp_df_flat$type <- "Flat"

#######################
#######################

### make all data.frames have the same column names so they are rbind-able
cp_df_1 <- cp_df_1[, intersect(colnames(cp_df_2), colnames(cp_df_1))]
cp_df_2 <- cp_df_2[, intersect(colnames(cp_df_1), colnames(cp_df_2))]

cp_df <- rbind(cp_df_1, cp_df_2)

cp_df <- cp_df[, intersect(colnames(cp_df_flat), colnames(cp_df))]
cp_df_flat <- cp_df_flat[, intersect(colnames(cp_df), colnames(cp_df))]

cp_df <- rbind(cp_df, cp_df_flat)

# cp_df$cp_pred_error <- (cp_df$cp - cp_df$Value) / cp_df$Value
# 
# cp_df <- cp_df %>% group_by(Species.x, 
#                             Compound,
#                             CAS,
#                             type) %>%
#   mutate(cp_pred_error_range = max(cp_pred_error, na.rm = TRUE) - min(cp_pred_error, na.rm = TRUE))

### here is where you save the Rdata file
save(cp_df, file = paste(figures_rdata/cp_df_",THIS.DATE,".RData",sep=""))
```

```{r, load_data}

### this block loads the Rdata files,
# as well as some other files written in below code blocks
# this saves time from having to run those below code blocks every time
# these additional files include 'cp_df' and 'cvt_variability'

### load all of this wonderful data
load(paste("figures_rdata/pk_output_1comp_",THIS.DATE,".RData",sep=""))
load(paste("figures_rdata/pk_output_2comp_",THIS.DATE,".RData",sep=""))
load(paste("figures_rdata/pk_output_flat_",THIS.DATE,".RData",sep=""))
load(paste("figures_rdata/raw_1comp_",THIS.DATE,".RData",sep=""))
load(paste("figures_rdata/raw_2comp_",THIS.DATE,".RData",sep=""))
load(paste("figures_rdata/raw_flat_",THIS.DATE,".RData",sep=""))
load(paste("figures_rdata/cp_df_",THIS.DATE,".RData",sep=""))
load(paste("figures_rdata/cvt_variability_df_",THIS.DATE,".RData",sep=""))

load(paste("figures_rdata/cp_df_",THIS.DATE,".RData",sep=""))

load("figures_rdata/pk_lombardo_02072022.Rdata")
```

```{r, make_cp_histogram}

### this block makes a histogram of the data output
# from make_cp_rdata
# there's a separate facet for each model type
# only data that were best fitted by that model
# are shown in that facet

##################
##################

pk_data <- cp_df_05092022

# cvt_data <- cvt_variability_df_04282022

##################
##################

row.names(pk_data) <- c()
full_data <- pk_data

# cvt_data <- dplyr::select(cvt_data, -c("kelim",
#                                        "Vdist",
#                                        "Fgutabs",
#                                        "kgutabs"))
# 
# ### split pk_data by model so you can properly join with raw data
# model_split <- split(pk_data, pk_data$model)
# 
# cvt_data$Reference <- as.character(cvt_data$Reference)
# 
# cvt_data$type <- "1-Compartment"
# full_data_1 <- left_join(cvt_data, model_split$`1compartment`)
# 
# cvt_data$type <- "2-Compartment"
# full_data_2 <- left_join(cvt_data, model_split$`2compartment`)
# 
# cvt_data$type <- "Flat"
# full_data_flat <- left_join(cvt_data, model_split$`flat`)
# 
# ### put data back together
# full_data <- rbind(full_data_1, full_data_2, full_data_flat)
# 
# full_data <- full_data %>%
#   filter(cp != 0)

full_data$AIC <- round(full_data$AIC, digits = 4)

### only take necessary columns to keep things clean
full_data <- full_data %>%
  distinct(Compound, Reference, Species.x, model, AIC, cp_norm)

### this pivots the data so we have AICs and cp_norms
# for each model row-wise
full_data <- full_data %>%
  group_by(Species.x, Compound, Reference, model) %>%
  mutate(row = row_number()) %>%
  pivot_wider(names_from = model,
              values_from = c(AIC, cp_norm)) %>%
  dplyr::select(-row)

full_data <- as.data.table(full_data)

### make model column equal a specific model based on AIC values
full_data <- full_data %>% mutate(model = case_when(AIC_1compartment < AIC_2compartment &
                                                      AIC_1compartment < AIC_flat ~ "1-Compartment",
                                                    AIC_2compartment < AIC_1compartment &
                                                      AIC_2compartment < AIC_flat ~ "2-Compartment",
                                                    AIC_flat < AIC_1compartment &
                                                      AIC_flat < AIC_2compartment ~ "Null",

                                                    !is.na(AIC_1compartment) & is.na(AIC_2compartment) &
                                                      is.na(AIC_flat) ~ "1-Compartment",
                                                    !is.na(AIC_2compartment) & is.na(AIC_1compartment) &
                                                      is.na(AIC_flat) ~ "2-Compartment",
                                                    !is.na(AIC_flat) & is.na(AIC_1compartment) &
                                                      is.na(AIC_2compartment) ~ "Null",

                                                    is.na(AIC_1compartment) &
                                                      AIC_flat < AIC_2compartment ~ "Null",
                                                    is.na(AIC_1compartment) &
                                                      AIC_2compartment < AIC_flat ~ "2-Compartment",
                                                    is.na(AIC_2compartment) &
                                                      AIC_flat < AIC_1compartment ~ "Null",
                                                    is.na(AIC_2compartment) &
                                                      AIC_1compartment < AIC_flat ~ "1-Compartment",
                                                    is.na(AIC_flat) &
                                                      AIC_1compartment < AIC_2compartment ~"1-Compartment",
                                                    is.na(AIC_flat) &
                                                      AIC_2compartment <AIC_1compartment ~"2-Compartment"))

### make cp_norm equal to the cp_norm that corresponds to the best fitting model
full_data <- full_data %>%
  mutate(cp_norm = case_when(model == "1-Compartment" ~ cp_norm_1compartment,
                             model == "2-Compartment" ~ cp_norm_2compartment,
                             model == "Null" ~ cp_norm_flat))

full_data[, model := paste("Model:", model)]

### this creates a new column that says how many unique sets of Compound, Reference,
# and Species correspond to each model
full_data[, chem_ref_spec_count := uniqueN(.SD), .SDcols = c("Compound", 
                                                             "Reference",
                                                             "Species.x"), 
          by = "model"]

full_data[, chem_ref_spec_count := paste("Chemical-Reference-Species:", chem_ref_spec_count)]

### make plot, facet by model and chem_ref_spec_count
p <- ggplot(data = full_data,
            mapping = aes(x = cp_norm)) +
  geom_histogram(binwidth = 0.325) +
  geom_vline(xintercept = c(0.5, 2),
             color = "Red",
             linetype = "dashed") +
  # xlim(0, 1e42) +
  # ylim(0, 400) +
  # scale_y_log10()+
  theme(title = element_text(size = 16),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14)) +
  # scale_x_log10() +
  scale_x_log10(limits = c(1e-3, 1e3), labels = scientific) +
  scale_y_continuous(labels = scientific) +
  theme_bw() +
  labs(x = "Predicted/Observed",
       y = "Number of Observations") +
  facet_wrap(. ~ model + chem_ref_spec_count)

p
```

```{r, make_cp_error_histogram}

### I think this is little outdated but whatever I'm gonna keep it here
# Think is was just plotting Cpred - Cobs

##################
##################

# data <- cp_df_12212021
# data <- cp_df_03012022
data <- cp_df_04282022

##################
##################

data <- data %>%
  filter(cp != 1e-20)

data <- data %>%
  distinct(Species.x, Compound, CAS, cp_pred_error_range)

p <- ggplot(data = data,
            mapping = aes(x = cp_pred_error_range)) +
  geom_histogram() +
  scale_x_log10() +
  facet_wrap(. ~type)

p
```

```{r, make_cp_error_versus_dose}

### I think this is little outdated but whatever I'm gonna keep it here

##################
##################

data <- cp_df_04282022

##################
##################

data <- data %>%
  filter(type == "1-Compartment")

data <- data %>%
  group_by(Compound, CAS, Dose) %>%
  mutate(average = mean(cp_pred_error))

p <- ggplot(data = data, mapping = aes(x = Dose, y = average)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10()

p

```

```{r, make_cvt_variability_rdata, eval = FALSE}

### this block creates a data.frame that shows the 
# spread of CvT data via across replicate timepoints 
# for a single experiment
# if the data isn't within 2-fold of itself
# is it reasonable to say that a model should fit within 2-fold?

##################
##################

### all the raw data for each model SHOULD be the same
# so only need to use one
data <- as.data.table(raw_1comp_05092022)

##################
##################

### function that finds mean conc of each replicate timepoint
find_mean_conc <- function(data) {

  ### only use data with replicate timepoints
  data <- data[ ave(1:nrow(data), data$Time, FUN = length) > 1 , ]

  if(nrow(data) == 0) {
    return(NULL)
  } else {

    time_split <- split(data, data$Time)

    ### discard replicate timepoints where all concentrations are NA
    time_split_list <- purrr::discard(time_split, ~all(is.na(.$Value)))

    if(length(time_split_list) == 0) {
      return(NULL)
    } else {

      data <- do.call(rbind, time_split_list)

      ### where some replicate concentration data are NA, set them equal to the LOQ
      data$Value[is.na(data$Value)] <- data$LOQ[is.na(data$Value)]

      ### take the average Value for each set of timepoints
      data <- data %>%
        group_by(Time) %>%
        do(mutate(., mean_conc = mean(.$Value, na.rm = TRUE)))

      data <- data %>%
        group_by(Time) %>%
        do(mutate(., range_conc = max(.$Value) - min(.$Value)))

      ### create normalized conc values by dividing conc by mean_conc
      data$conc_norm <- data$Value/data$mean_conc

      data <- data[order(data$Time), ]

      return(data)
    }
  }
}

### this for loop applies the function 
# at data subsets of compound, data, dose, route, media
data.list <- list()
for(this.compound in unique(data[, Compound])) {
  this.compound.data <- data[Compound == this.compound]

  ### list of columns by which to split data.set
  col_list <- list(this.compound.data$Reference,
                   this.compound.data$Dose,
                   this.compound.data$Route,
                   this.compound.data$Media)

  ### split data.set into list of data.tables
  split_list <- split(this.compound.data, col_list)

  ### remove data.tables with 0 rows
  split_list <- keep(split_list, ~nrow(.) > 0)

  ### use find_mean_conc fn
  mean_conc <- lapply(split_list, find_mean_conc)

  test <- do.call(rbind, mean_conc)

  data.list[[this.compound]] <- test
}

### combine data.tables back into one large data.set
cvt_variability_df_05092022 <- do.call(rbind, data.list)

### SAVE the data
save(cvt_variability_df_05092022, file = "figures_rdata/cvt_variability_df_05092022.Rdata")
```

```{r, make_cvt_variability_histogram}

### this block makes the histogram
# of cvt variability using the data.frame
# made from make_cvt_variability_rdata

#################
#################

data <- cvt_variability_df_05092022
# data <- cvt_variability_df_12212021

#################
#################

p <- ggplot(data, aes(x = conc_norm)) +
  geom_histogram(binwidth = 0.25) +
  xlim(c(-0.5, 4)) +
  geom_vline(xintercept = c(0.5, 2),
             color = "Red",
             linetype = "dashed") +
  theme_bw() +
  labs(x = "Normalized concentration",
       y = "Number of observations")
p
```

```{r, make_density_plot}

### this block makes a density plot with cvt variabity
# and normalized predicted concentrations
# it's basically a combination of the other two histograms
# expect here we limit cp observations those that have replicate timepoint
# data, so it matches the cvt variability data, thus making it 
# comparable

##################
##################

# pk_data <- cp_df_03012022
pk_data <- cp_df_05092022

cvt_data <- cvt_variability_df_05092022

##################
##################

### i hate row names
row.names(pk_data) <- c()

### making these null so I don't have to deal
# with duplicate .x and .y's
# only need the ones from pk_data
cvt_data <- dplyr::select(cvt_data, -c("kelim", 
                                       "Vdist", 
                                       "Fgutabs", 
                                       "kgutabs"))

### split pk_data by model so you can properly join with raw data
model_split <- split(pk_data, pk_data$model)

### References were of different classes for some reason
# so had to change
cvt_data$Reference <- as.character(cvt_data$Reference)

cvt_data$type <- "1-Compartment"
full_data_1 <- left_join(cvt_data, model_split$`1compartment`)

cvt_data$type <- "2-Compartment"
full_data_2 <- left_join(cvt_data, model_split$`2compartment`)

cvt_data$type <- "Flat"
full_data_flat <- left_join(cvt_data, model_split$`flat`)

### put data back together
full_data <- rbind(full_data_1, full_data_2, full_data_flat)

### not sure why there were 0s,
# but they were screwing things up
full_data <- full_data %>%
  filter(cp != 0)

full_data$AIC <- round(full_data$AIC, digits = 4)

full_data <- full_data %>%
  distinct(Compound, Reference, model, Species.x, conc_norm, AIC, cp_norm)

### 
### 
full_data <- full_data %>%
  group_by(Species.x, Compound, Reference, model) %>%
  mutate(row = row_number()) %>%
  pivot_wider(names_from = model,
              values_from = c(AIC, cp_norm)) %>%
  dplyr::select(-row)

full_data <- full_data %>% mutate(model = case_when(AIC_1compartment < AIC_2compartment & 
                                                      AIC_1compartment < AIC_flat ~ "1-Compartment",
                                                    AIC_2compartment < AIC_1compartment &
                                                      AIC_2compartment < AIC_flat ~ "2-Compartment",
                                                    AIC_flat < AIC_1compartment &
                                                      AIC_flat < AIC_2compartment ~ "Null",

                                                    !is.na(AIC_1compartment) & is.na(AIC_2compartment) &
                                                      is.na(AIC_flat) ~ "1-Compartment",
                                                    !is.na(AIC_2compartment) & is.na(AIC_1compartment) & 
                                                      is.na(AIC_flat) ~ "2-Compartment",
                                                    !is.na(AIC_flat) & is.na(AIC_1compartment) & 
                                                      is.na(AIC_2compartment) ~ "Null",

                                                    is.na(AIC_1compartment) &
                                                      AIC_flat < AIC_2compartment ~ "Null",
                                                    is.na(AIC_1compartment) &
                                                      AIC_2compartment < AIC_flat ~ "2-Compartment",
                                                    is.na(AIC_2compartment) &
                                                      AIC_flat < AIC_1compartment ~ "Null",
                                                    is.na(AIC_2compartment) &
                                                      AIC_1compartment < AIC_flat ~ "1-Compartment",
                                                    is.na(AIC_flat) &
                                                      AIC_1compartment < AIC_2compartment ~"1-Compartment",
                                                    is.na(AIC_flat) &
                                                      AIC_2compartment <AIC_1compartment ~"2-Compartment"))

full_data <- full_data %>%
  mutate(cp_norm = case_when(model == "1-Compartment" ~ cp_norm_1compartment,
                             model == "2-Compartment" ~ cp_norm_2compartment,
                             model == "Null" ~ cp_norm_flat))

### remove super outlier data from triclosan, for visualization purposes
full_data <- full_data %>%
  filter(cp_norm > 1e-10 & cp_norm < 1e8) # also high one for the heck of it

length(unique(full_data$Compound[full_data$model == "1-Compartment"]))
length(unique(full_data$Compound[full_data$model == "2-Compartment"]))
length(unique(full_data$Compound[full_data$model == "Null"]))

null_vec <- as.vector(unique(full_data$Compound[full_data$model == "Null"]))

full_data <- as.data.table(full_data)

full_data[, model := paste("Model:", model)]

full_data[, chem_ref_spec_count := uniqueN(.SD), .SDcols = c("Compound", 
                                                             "Reference",
                                                             "Species.x"), 
          by = "model"]

full_data[, chem_ref_spec_count := paste("Chemical-Reference-Species:", chem_ref_spec_count)]

# full_data <- full_data %>%
#   mutate(cp_norm = case_when(model == "1comparment" ~ cp_norm_1compartment,
#                              model == "2compartment" ~ cp_norm_2compartment))

# full_data$cp_norm[full_data$model == "2compartment"] <- full_data$cp_norm_2compartment
# full_data$cp_norm[full_data$model == "1compartment"] <- full_data$cp_norm_1compartment

# full_data <- full_data %>%
# mutate(across(everything(), ~ifelse(. == "", NA, as.character(.))))

# filter(Compound != "triclosan")
# filter(Compound != "tamoxifen")

###########
###########
###########

split_full_data <- split(full_data, full_data$model)

full_data_1 <- split_full_data$`Model: 1-Compartment`

### Find number of points in each plot panel for 1-compartment
full_data_1_center <- full_data_1 %>% filter(conc_norm >= 0.5 & conc_norm <= 2)
full_data_1_center <- full_data_1_center %>% filter(cp_norm >= 0.5 & cp_norm <= 2)
ratio_1_center <- nrow(full_data_1_center) / nrow(full_data_1 %>% filter(!is.na(cp_norm)))

full_data_1_green <- full_data_1 %>% filter(conc_norm < 0.5 | conc_norm > 2)
full_data_1_green <- full_data_1_green %>% filter(cp_norm < 0.5 | cp_norm > 2)
ratio_1_green <- nrow(full_data_1_green) / nrow(full_data_1 %>% filter(!is.na(cp_norm)))

full_data_1_gray <- full_data_1 %>% filter(conc_norm < 0.5 | conc_norm > 2)
full_data_1_gray <- full_data_1_gray %>% filter(cp_norm >= 0.5 & cp_norm <= 2)
ratio_1_gray <- nrow(full_data_1_gray) / nrow(full_data_1 %>% filter(!is.na(cp_norm)))

full_data_1_tan <- full_data_1 %>% filter(conc_norm >= 0.5 & conc_norm <= 2)
full_data_1_tan <- full_data_1_tan %>% filter(cp_norm < 0.5 | cp_norm > 2)
ratio_1_tan <- nrow(full_data_1_tan) / nrow(full_data_1 %>% filter(!is.na(cp_norm)))

###########
###########

full_data_2 <- split_full_data$`Model: 2-Compartment`

### Find number of points in each plot panel for 2-compartment
full_data_2_center <- full_data_2 %>% filter(conc_norm >= 0.5 & conc_norm <= 2)
full_data_2_center <- full_data_2_center %>% filter(cp_norm >= 0.5 & cp_norm <= 2)
ratio_2_center <- nrow(full_data_2_center) / nrow(full_data_2 %>% filter(!is.na(cp_norm)))

full_data_2_green <- full_data_2 %>% filter(conc_norm < 0.5 | conc_norm > 2)
full_data_2_green <- full_data_2_green %>% filter(cp_norm < 0.5 | cp_norm > 2)
ratio_2_green <- nrow(full_data_2_green) / nrow(full_data_2 %>% filter(!is.na(cp_norm)))

full_data_2_gray <- full_data_2 %>% filter(conc_norm < 0.5 | conc_norm > 2)
full_data_2_gray <- full_data_2_gray %>% filter(cp_norm >= 0.5 & cp_norm <= 2)
ratio_2_gray <- nrow(full_data_2_gray) / nrow(full_data_2 %>% filter(!is.na(cp_norm)))

full_data_2_tan <- full_data_2 %>% filter(conc_norm >= 0.5 & conc_norm <= 2)
full_data_2_tan <- full_data_2_tan %>% filter(cp_norm < 0.5 | cp_norm > 2)
ratio_2_tan <- nrow(full_data_2_tan) / nrow(full_data_2 %>% filter(!is.na(cp_norm)))

###########
###########

full_data_flat <- split_full_data$`Model: Null`

### Find number of points in each plot panel for 2-compartment
full_data_flat_center <- full_data_flat %>% filter(conc_norm >= 0.5 & conc_norm <= 2)
full_data_flat_center <- full_data_flat_center %>% filter(cp_norm >= 0.5 & cp_norm <= 2)
ratio_flat_center <- nrow(full_data_flat_center) / nrow(full_data_flat %>% filter(!is.na(cp_norm)))

full_data_flat_green <- full_data_flat %>% filter(conc_norm < 0.5 | conc_norm > 2)
full_data_flat_green <- full_data_flat_green %>% filter(cp_norm < 0.5 | cp_norm > 2)
ratio_flat_green <- nrow(full_data_flat_green) / nrow(full_data_flat %>% filter(!is.na(cp_norm)))

full_data_flat_gray <- full_data_flat %>% filter(conc_norm < 0.5 | conc_norm > 2)
full_data_flat_gray <- full_data_flat_gray %>% filter(cp_norm >= 0.5 & cp_norm <= 2)
ratio_flat_gray <- nrow(full_data_flat_gray) / nrow(full_data_flat %>% filter(!is.na(cp_norm)))

full_data_flat_tan <- full_data_flat %>% filter(conc_norm >= 0.5 & conc_norm <= 2)
full_data_flat_tan <- full_data_flat_tan %>% filter(cp_norm < 0.5 | cp_norm > 2)
ratio_flat_tan <- nrow(full_data_flat_tan) / nrow(full_data_flat %>% filter(!is.na(cp_norm)))

###########
###########
###########

# rect_df1 <- data.frame(xmin = c(0, 0, 2, 2, 0, 2, 0.5, 0.5),
#                       xmax = c(0.5, 0.5, Inf, Inf, 0.5, Inf, 2, 2),
#                       ymin = c(2, 0, 0, 2, 0.5, 0.5, 2, 0),
#                       ymax = c(Inf, 0.5, 0.5, Inf, 2, 2, Inf, 0.5),
#                       panel_key = c(paste("Poor Fit, Outlier Data", 
#                                           paste("(", round(100 * ratio_1_green, digits = 1), "%", ")", sep = ""), 
#                                           sep = " "),
#                                     paste("Poor Fit, Outlier Data", 
#                                           paste("(", round(100 * ratio_1_green, digits = 1), "%", ")", sep = ""),
#                                           sep = " "),
#                                     paste("Poor Fit, Outlier Data", 
#                                           paste("(", round(100 * ratio_1_green, digits = 1), "%", ")", sep = ""), 
#                                           sep = " "),
#                                     paste("Poor Fit, Outlier Data", 
#                                           paste("(", round(100 * ratio_1_green, digits = 1), "%", ")", sep = ""),
#                                           sep = " "),
#                                     paste("Good Fit, Outlier Data", 
#                                           paste("(", round(100 * ratio_1_tan, digits = 1), "%", ")", sep = ""), 
#                                           sep = " "),
#                                     paste("Good Fit, Outlier Data", 
#                                           paste("(", round(100 * ratio_1_tan, digits = 1), "%", ")", sep = ""), 
#                                           sep = " "),
#                                    paste("Poor Fit, Typical Data", 
#                                           paste("(", round(100 * ratio_1_gray, digits = 1), "%", ")", sep = ""), 
#                                           sep = " "),
#                                    paste("Poor Fit, Typical Data", 
#                                           paste("(", round(100 * ratio_1_gray, digits = 1), "%", ")", sep = ""), 
#                                           sep = " ")))
# 
# rect_df2 <- data.frame(xmin = c(0, 0, 2, 2, 0, 2, 0.5, 0.5),
#                       xmax = c(0.5, 0.5, Inf, Inf, 0.5, Inf, 2, 2),
#                       ymin = c(2, 0, 0, 2, 0.5, 0.5, 2, 0),
#                       ymax = c(Inf, 0.5, 0.5, Inf, 2, 2, Inf, 0.5),
#                       panel_key = c(paste("Poor Fit, Outlier Data", 
#                                           paste("(", round(100 * ratio_2_green, digits = 1), "%", ")", sep = ""), 
#                                           sep = " "),
#                                     paste("Poor Fit, Outlier Data", 
#                                           paste("(", round(100 * ratio_2_green, digits = 1), "%", ")", sep = ""),
#                                           sep = " "),
#                                     paste("Poor Fit, Outlier Data", 
#                                           paste("(", round(100 * ratio_2_green, digits = 1), "%", ")", sep = ""), 
#                                           sep = " "),
#                                     paste("Poor Fit, Outlier Data", 
#                                           paste("(", round(100 * ratio_2_green, digits = 1), "%", ")", sep = ""),
#                                           sep = " "),
#                                     paste("Good Fit, Outlier Data", 
#                                           paste("(", round(100 * ratio_2_tan, digits = 1), "%", ")", sep = ""), 
#                                           sep = " "),
#                                     paste("Good Fit, Outlier Data", 
#                                           paste("(", round(100 * ratio_2_tan, digits = 1), "%", ")", sep = ""), 
#                                           sep = " "),
#                                    paste("Poor Fit, Typical Data", 
#                                           paste("(", round(100 * ratio_2_gray, digits = 1), "%", ")", sep = ""), 
#                                           sep = " "),
#                                    paste("Poor Fit, Typical Data", 
#                                           paste("(", round(100 * ratio_2_gray, digits = 1), "%", ")", sep = ""), 
#                                           sep = " ")))
# 
# 
# rect_p1 <- ggplot() +
#   geom_rect(data = rect_df1, aes(xmin = xmin,
#                                 ymin = ymin,
#                                 xmax = xmax,
#                                 ymax = ymax,
#                                 fill = panel_key),
#             alpha = 0.3, inherit.aes = FALSE) +
#   scale_fill_manual(name = "Panel Key",
#                     values = c("#361D2E",
#                                "#86A397",
#                                "#E1B07E")) +
#   theme_bw() +
#   theme(panel.spacing = unit(2, "lines")) +
#   geom_vline(xintercept = c(0.5, 2),
#              linetype = "dashed") +
#   geom_hline(yintercept = c(0.5, 2),
#              linetype = "dashed") +
#   scale_x_log10() +
#   scale_y_log10() +
#   labs(x = "Data Variability",
#        y = "Model Error") +
#   guides(fill = guide_legend(override.aes = list(size = 0.1)))
# 
# rect_p2 <- ggplot() +
#   geom_rect(data = rect_df2, aes(xmin = xmin,
#                                 ymin = ymin,
#                                 xmax = xmax,
#                                 ymax = ymax,
#                                 fill = panel_key),
#             alpha = 0.3, inherit.aes = FALSE) +
#   scale_fill_manual(name = "Panel Key",
#                     values = c("#361D2E",
#                                "#86A397",
#                                "#E1B07E")) +
#   theme_bw() +
#   theme(panel.spacing = unit(2, "lines")) +
#   geom_vline(xintercept = c(0.5, 2),
#              linetype = "dashed") +
#   geom_hline(yintercept = c(0.5, 2),
#              linetype = "dashed") +
#   scale_x_log10() +
#   scale_y_log10() +
#   labs(x = "Data Variability",
#        y = "Model Error") +
#   guides(fill = guide_legend(override.aes = list(size = 0.1)))
# 
# ###################
# fill_legend1 <- get_legend(rect_p1)
# fill_legend2 <- get_legend(rect_p2)
# ##################

p <- ggplot(data = full_data,
            mapping = aes(x = conc_norm, y = cp_norm)) +
  annotate(geom = "rect", xmin = 0, xmax = 0.5, ymin = 2, ymax = Inf, alpha = 0.3, fill = "#86A397") +
  annotate(geom = "rect", xmin = 0, xmax = 0.5, ymin = 0, ymax = 0.5, alpha = 0.3, fill = "#86A397") +
  annotate(geom = "rect", xmin = 2, xmax = Inf, ymin = 0, ymax = 0.5, alpha = 0.3, fill = "#86A397") +
  annotate(geom = "rect", xmin = 2, xmax = Inf, ymin = 2, ymax = Inf, alpha = 0.3, fill = "#86A397") +
  annotate(geom = "rect", xmin = 0, xmax = 0.5, ymin = 0.5, ymax = 2, alpha = 0.3, fill = "#361D2E") +
  annotate(geom = "rect", xmin = 2, xmax = Inf, ymin = 0.5, ymax = 2, alpha = 0.3, fill = "#361D2E") +
  annotate(geom = "rect", xmin = 0.5, xmax = 2, ymin = 2, ymax = Inf, alpha = 0.3, fill = "#E1B07E") +
  annotate(geom = "rect", xmin = 0.5, xmax = 2, ymin = 0, ymax = 0.5, alpha = 0.3, fill = "#E1B07E") +
  annotate(geom = "rect", xmin = 0.5, xmax = 2, ymin = 0.5, ymax = 2.0, alpha = 0.5, fill = "black") +
  theme_bw() +
  theme(panel.spacing = unit(2, "lines")) +
  geom_bin2d(bins = 100) +
  scale_fill_distiller(palette = "Blues") +
  geom_vline(xintercept = c(0.5, 2),
             linetype = "dashed") +
  geom_hline(yintercept = c(0.5, 2),
             linetype = "dashed") +
  scale_x_log10(n.breaks = 4, labels = scientific) +
  scale_y_log10() +
  labs(x = "Data Variability",
       y = "Model Performance") +
  facet_wrap(. ~ model + chem_ref_spec_count) +
  theme(strip.text.x = element_text(size = 7))

p

# ##################
# 
# data_p_final <- data_p +
#   # inset_element(fill_legend1, 0.1, 1, 0.5, 1) +
#   inset_element(fill_legend2, 0.6, 0.5, 0.7, 0.2, clip = TRUE)
#  
# data_p_final
```

```{r, make_sigma_distribution_histogram}

# pk_data <- cp_df_03012022
pk_data <- cp_df_04282022

cvt_data <- cvt_variability_df_04282022

##################
##################

row.names(pk_data) <- c()

cvt_data <- dplyr::select(cvt_data, -c("kelim", 
                                       "Vdist", 
                                       "Fgutabs", 
                                       "kgutabs"))

### split pk_data by model so you can properly join with raw data
model_split <- split(pk_data, pk_data$model)

cvt_data$Reference <- as.character(cvt_data$Reference)

cvt_data$type <- "1-Compartment"
full_data_1 <- left_join(cvt_data, model_split$`1compartment`)

cvt_data$type <- "2-Compartment"
full_data_2 <- left_join(cvt_data, model_split$`2compartment`)

cvt_data$type <- "Flat"
full_data_flat <- left_join(cvt_data, model_split$`flat`)

### put data back together
full_data <- rbind(full_data_1, full_data_2, full_data_flat)

full_data <- full_data %>%
  filter(cp != 0)

# full_data$AIC <- round(full_data$AIC, digits = 4)

full_data <- full_data %>%
  pivot_wider(names_from = model,
              values_from = c(AIC, sigma_value)) %>%
  unchop(everything())

full_data <- full_data %>% mutate(model = case_when(AIC_1compartment < AIC_2compartment & 
                                                      AIC_1compartment < AIC_flat ~ "1-Compartment",
                                                    AIC_2compartment < AIC_1compartment &
                                                      AIC_2compartment < AIC_flat ~ "2-Compartment",
                                                    AIC_flat < AIC_1compartment &
                                                      AIC_flat < AIC_2compartment ~ "Null",

                                                    !is.na(AIC_1compartment) & is.na(AIC_2compartment) &
                                                      is.na(AIC_flat) ~ "1-Compartment",
                                                    !is.na(AIC_2compartment) & is.na(AIC_1compartment) & 
                                                      is.na(AIC_flat) ~ "2-Compartment",
                                                    !is.na(AIC_flat) & is.na(AIC_1compartment) & 
                                                      is.na(AIC_2compartment) ~ "Null",

                                                    is.na(AIC_1compartment) &
                                                      AIC_flat < AIC_2compartment ~ "Null",
                                                    is.na(AIC_1compartment) &
                                                      AIC_2compartment < AIC_flat ~ "2-Compartment",
                                                    is.na(AIC_2compartment) &
                                                      AIC_flat < AIC_1compartment ~ "Null",
                                                    is.na(AIC_2compartment) &
                                                      AIC_1compartment < AIC_flat ~ "1-Compartment",
                                                    is.na(AIC_flat) &
                                                      AIC_1compartment < AIC_2compartment ~"1-Compartment",
                                                    is.na(AIC_flat) &
                                                      AIC_2compartment <AIC_1compartment ~"2-Compartment"))

full_data <- full_data %>%
  mutate(AIC = case_when(model == "1-Compartment" ~ AIC_1compartment,
                         model == "2-Compartment" ~ AIC_2compartment,
                         model == "Null" ~ AIC_flat))

full_data <- full_data %>%
  mutate(sigma_value = case_when(model == "1-Compartment" ~ sigma_value_1compartment,
                                 model == "2-Compartment" ~ sigma_value_2compartment,
                                 model == "Null" ~ sigma_value_flat))

full_data <- full_data %>%
  distinct(Compound, 
           Reference, 
           Route, 
           Species,
           Media,
           model, 
           sigma_id,
           sigma_value,
           Value,
           AIC)

full_data <- as.data.table(full_data)

full_data[, sigma_norm := sigma_value / nrow(.SD), by = .(Compound, 
                                                          Reference,
                                                          model,
                                                          sigma_id,
                                                          sigma_value)]

full_data <- as.data.frame(full_data)

full_data <- full_data %>% distinct(Compound,
                                    Reference,
                                    model,
                                    AIC,
                                    sigma_id,
                                    sigma_value,
                                    sigma_norm)

full_data <- full_data %>%
  group_by(Compound, Reference, sigma_id) %>%
  slice_min(AIC)

# ##################
# ##################
# 
# data1 <- pk_output_1comp_03012022 %>%
#   filter(param.value.type == "Fitted geometric mean" & AIC != Inf & Data.Analyzed != "Joint Analysis")
# 
# data2 <- pk_output_2comp_03012022 %>%
#   filter(param.value.type == "Fitted geometric mean" & AIC != Inf & Data.Analyzed != "Joint Analysis")
# 
# ##################
# ##################
# 
# ### assign new column, with either "1-Compartment" or "2-Compartment"
# data1 <- subset(data1, select = sigma_value)
# data1$model <- "1-Compartment"
# 
# data2 <- subset(data2, select = sigma_value)
# data2$model <- "2-Compartment"
# 
# data <- rbind(data1, data2)

p <- ggplot(data = full_data,
            mapping = aes(x = sigma_norm)) +
  geom_histogram(bins = 12) +
  # geom_density() +
  # scale_y_log10()+
  scale_x_log10() +
  theme_bw() +
  labs(x = "Sigma",
       y = "Number of observations") +
  facet_wrap(.~ model)
# geom_vline(xintercept = 0.001476315)
# stat_function(fun = dlnorm, args = list(mean = mean(log(data$sigma_value)), sd = sd(log(data$sigma_value))))

p
```

```{r make_magnitude_histogram}

##################
##################

data <- raw_1comp_04282022

##################
##################

data$X <- NULL

split_df <- split(data, list(data$Compound,
                             data$Reference,
                             data$CAS,
                             data$Dose,
                             data$Species,
                             data$Route,
                             data$Media),
                  drop = TRUE)

### find number of orders of magnitude
find_mag <- function(data) {

  data$Value <- abs(data$Value)

  diff <- log10(max(data$Value, na.rm = TRUE)) - log10(min(data$Value, na.rm = TRUE))

  diff <- round(diff)

  return(diff)

}

data <- lapply(split_df, find_mag)

data <- pivot_longer(as.data.frame(data), cols = everything())

p <- ggplot(data = data, mapping = aes(x = value)) +
  geom_histogram(bins = 5, binwidth = 0.5) +
  xlab("Orders of magnitude") + theme_bw()

p
```

```{r make_joint_sigma_plot, eval = FALSE}

##################
##################

data1 <- as.data.table(pk_output_1comp_04282022 %>%
                         filter(param.value.type == "Fitted geometric mean"))

data2 <- as.data.table(pk_output_2comp_04282022 %>%
                         filter(param.value.type == "Fitted geometric mean"))

##################
##################

### find data with repeated sigma_id values
find_rep_sigma <- function(data) {

  data <- data[ ave(1:nrow(data), data$sigma_id, FUN=length) > 1 , ]

  if(nrow(data) == 0) {
    return(NULL)
  } else {

    return(data)
  }
}

split_list_1 <- split(data1, data1$Compound)
rep_sigma_1 <- lapply(split_list_1, find_rep_sigma)
data1 <- do.call(rbind, rep_sigma_1)
data1 <- subset(data1, select = c(param.value.type, sigma_id, sigma_value, Data.Analyzed, Compound))
data1$Data.Analyzed[data1$Data.Analyzed != "Joint Analysis"] <- "Ind Analysis"
data1 <- data1 %>% pivot_wider(names_from = Data.Analyzed, values_from = sigma_value)
data1 <- data1 %>% filter(sigma_id != "sigma2.14")
data1$`Joint Analysis` <- as.numeric(data1$`Joint Analysis`)
data1$`Ind Analysis` <- as.numeric(data1$`Ind Analysis`)
data1$model <- "1-Compartment"

split_list_2 <- split(data2, data2$Compound)
rep_sigma_2 <- lapply(split_list_2, find_rep_sigma)
data2 <- do.call(rbind, rep_sigma_2)
data2 <- subset(data2, select = c(param.value.type, sigma_id, sigma_value, Data.Analyzed, Compound))
data2$Data.Analyzed[data2$Data.Analyzed != "Joint Analysis"] <- "Ind Analysis"
data2 <- data2 %>% pivot_wider(names_from = Data.Analyzed, values_from = sigma_value)
data2 <- data2 %>% filter(sigma_id != "sigma2.14")
data2$`Joint Analysis` <- as.numeric(data2$`Joint Analysis`)
data2$`Ind Analysis` <- as.numeric(data2$`Ind Analysis`)
data2$model <- "2-Compartment"

data <- rbind(data1, data2)

p <- ggplot(data, aes(x = `Ind Analysis`, y = `Joint Analysis`)) +
  geom_point() +
  scale_y_log10() +
  scale_x_log10() +
  geom_abline(slope = 1) +
  facet_wrap(. ~ model)
p
```

```{r, make_halflife_count_histogram}

###########
### one compartment
###########

##################
##################

raw_data1 <- raw_1comp_04282022

##################
##################

pk_data1 <- pk_output_1comp_04282022 %>% 
  filter(param.value.type == "Fitted geometric mean" & Data.Analyzed != "Joint Analysis")

pk_data1$X <- NULL

raw_data1$Reference <- as.character(raw_data1$Reference)

raw_data1 <- dplyr::select(raw_data1, -c("X",
                                         "kelim",
                                         "Vdist",
                                         "Fgutabs",
                                         "kgutabs"))

# pk_data2 <- pk_output_2comp_12212021

full_data_1 <- left_join(raw_data1, pk_data1, by = c("Compound", "CAS", "Reference"))

full_data_1 <- full_data_1 %>% filter(AIC != Inf)

find_halflife_count <- function(data) {

  data_subset <- data %>% filter(!is.na(Value))

  max_time <- max(unique(data_subset$Time))

  data$halflife_count <- max_time/data$halflife

  return(data)

}

split_full_data_1 <- split(full_data_1, list(full_data_1$Compound,
                                             full_data_1$Reference,
                                             full_data_1$CAS,
                                             full_data_1$Dose,
                                             full_data_1$Species.x,
                                             full_data_1$Route,
                                             full_data_1$Media),
                           drop = TRUE)

full_data_1_hl <- lapply(split_full_data_1, find_halflife_count)

full_data_1_hl <- do.call(rbind, full_data_1_hl)

full_data_1_hl <- full_data_1_hl %>%
  distinct(Compound,
           Reference,
           CAS,
           Dose,
           Species.x,
           Route,
           Media,
           halflife_count)

full_data_1_hl$model <- "1-Compartment"

###########
### two compartment
###########

##################
##################

raw_data2 <- raw_2comp_04282022

pk_data2 <- pk_output_2comp_04282022 %>% 
  filter(param.value.type == "Fitted geometric mean" & Data.Analyzed != "Joint Analysis")

##################
##################

pk_data2$X <- NULL

raw_data2$Reference <- as.character(raw_data2$Reference)

raw_data2 <- dplyr::select(raw_data2, -c("kelim", 
                                         "Vdist", 
                                         "Fgutabs", 
                                         "kgutabs",
                                         "V1",
                                         "Ralphatokelim",
                                         "Fbetaofalpha"))

# pk_data2 <- pk_output_2comp_12212021

full_data_2 <- left_join(raw_data2, pk_data2, by = c("Compound", "CAS", "Reference"))

full_data_2 <- full_data_2 %>% filter(AIC != Inf)

find_halflife_count <- function(data) {

  data_subset <- data %>% filter(!is.na(Value))

  max_time <- max(unique(data_subset$Time))

  data$halflife_count <- max_time/data$halflife

  return(data)

}

split_full_data_2 <- split(full_data_2, list(full_data_2$Compound,
                                             full_data_2$Reference,
                                             full_data_2$CAS,
                                             full_data_2$Dose,
                                             full_data_2$Species.x,
                                             full_data_2$Route,
                                             full_data_2$Media),
                           drop = TRUE)

full_data_2_hl <- lapply(split_full_data_2, find_halflife_count)

full_data_2_hl <- do.call(rbind, full_data_2_hl)

full_data_2_hl <- full_data_2_hl %>%
  distinct(Compound,
           Reference,
           CAS,
           Dose,
           Species.x,
           Route,
           Media,
           halflife_count)

full_data_2_hl$model <- "2-Compartment"

###########
###########
###########

full_data_hl <- rbind(full_data_1_hl, full_data_2_hl)

###########

p <- ggplot(data = full_data_hl, mapping = aes(x = halflife_count)) +
  geom_histogram() +
  scale_x_log10() +
  xlab("halflife count") + 
  theme_bw() +
  # geom_vline(xintercept = 330.7598,
  #            color = "red",
  #            linetype = "dashed") + 
  facet_wrap(. ~ model)
p
```

```{r, make_max_loq_histogram}

###############
###############

raw_data1 <- raw_1comp_04282022

###############
###############

find_normalized_max <- function(data) {

  data$normalized_max <- max(data$Value, na.rm = TRUE)/data$LOQ

  return(data)

}

split_df_1 <- split(raw_data1, list(raw_data1$Compound,
                                    raw_data1$Reference,
                                    raw_data1$CAS,
                                    raw_data1$Dose,
                                    raw_data1$Species,
                                    raw_data1$Route,
                                    raw_data1$Media),
                    drop = TRUE)

raw_data1_normalized_max <- lapply(split_df_1, find_normalized_max)

raw_data1_normalized_max <- do.call(rbind, raw_data1_normalized_max)

raw_data1_normalized_max <- raw_data1_normalized_max %>%
  distinct(Compound,
           Reference,
           CAS,
           Dose,
           Species,
           Route,
           Media,
           normalized_max)

p <- ggplot(data = raw_data1_normalized_max, mapping = aes(x = normalized_max)) +
  geom_histogram(bins = 15) +
  scale_x_log10() +
  xlab("Max Concentration / LOQ") + 
  theme_bw()
# geom_vline(xintercept = 24.8482,
#            color = "red",
#            linetype = "dashed")
p
```

```{r, make_mean_loq_histogram}

###############
###############

raw_data1 <- raw_1comp_04282022

###############
###############

find_normalized_mean <- function(data) {

  data$normalized_mean <- mean(data$Value, na.rm = TRUE)/data$LOQ

  return(data)

}

split_df_1 <- split(raw_data1, list(raw_data1$Compound,
                                    raw_data1$Reference,
                                    raw_data1$CAS,
                                    raw_data1$Dose,
                                    raw_data1$Species,
                                    raw_data1$Route,
                                    raw_data1$Media),
                    drop = TRUE)

raw_data1_normalized_mean <- lapply(split_df_1, find_normalized_mean)

raw_data1_normalized_mean <- do.call(rbind, raw_data1_normalized_mean)

raw_data1_normalized_mean <- raw_data1_normalized_mean %>%
  distinct(Compound,
           Reference,
           CAS,
           Dose,
           Species,
           Route,
           Media,
           normalized_mean)

p <- ggplot(data = raw_data1_normalized_mean, mapping = aes(x = normalized_mean)) +
  geom_histogram(bins = 15) +
  scale_x_log10() +
  xlab("Mean Concentration / LOQ") + 
  theme_bw()
# geom_vline(xintercept = 24.8482,
#            color = "red",
#            linetype = "dashed")
p
```

```{r, make_model_params_table, eval = FALSE}

### this code almost entirely is imported from John's bake-off RMarkdown file
### what this produces is a table of compounds that were fit by at least one of the models
### AICs for both models are reported, and subsequent parameter values correspond to the model with the lower AIC

###################
###################

pk_data1 <- pk_output_1comp_04282022 
# pk_data1 <- pk_output_1comp_12212021

pk_data2 <- pk_output_2comp_04282022 
# pk_data2 <- pk_output_2comp_12212021

###################
###################

pk_data1$sigma_id <- NULL
pk_data1$sigma_value <- NULL
pk_data2$sigma_id <- NULL
pk_data2$sigma_value <- NULL

pk_data1 <- distinct(pk_data1)
pk_data2 <- distinct(pk_data2)

fits1comp <- subset(pk_data1, is.finite(AIC) & AIC < 10^4)
fits2comp <- subset(pk_data2, is.finite(AIC) & AIC < 10^4)

#How many chemicals have 1 or 2 compartment fits:
length(unique(c(fits1comp$CAS,fits2comp$CAS)))

fittable <- NULL
# test <- function(fits1comp, fits2comp) {
for (this.cas in unique(c(fits1comp$CAS,fits2comp$CAS)))
{ 
  # print(this.cas)
  # if (this.cas == "123-91-1") browser()
  if (this.cas %in% fits1comp$CAS) this.compound <- fits1comp[
    fits1comp$CAS==this.cas,"Compound"][1]
  else this.compound <- fits2comp[
    fits2comp$CAS==this.cas,"Compound"][1]

  this.data.1comp <- subset(fits1comp,CAS==this.cas)
  this.data.2comp <- subset(fits2comp,CAS==this.cas)

  for (this.species in unique(tolower(c(
    this.data.1comp$Species,
    this.data.2comp$Species))))
  {
    Vdist.1comp <- NA
    kelim.1comp <- NA
    Vdist.2comp <- NA
    kelim.2comp <- NA

    this.data.1comp.species <- subset(this.data.1comp,tolower(Species)==this.species)
    this.data.2comp.species <- subset(this.data.2comp,tolower(Species)==this.species)

    this.row <- data.frame(
      Compound=this.compound,
      CAS=this.cas,
      Species=this.species,
      Reference=NA,
      AIC.1comp=NA,
      AIC.2comp=NA,
      Model=NA,
      Vdist=NA,
      kelim=NA)

    if (this.cas %in% fits1comp$CAS)
    {
      this.data.1comp.species <- subset(this.data.1comp.species,
                                        param.value.type=="Fitted geometric mean")

      # If more than one source, require that the joint analysis worked:
      if (dim(this.data.1comp.species)[1]>1)
      {
        this.data.1comp.species<- subset(this.data.1comp.species, 
                                         Data.Analyzed=="Joint Analysis")
      }
      if (dim(this.data.1comp.species)[1]>0) 
      {
        if (nrow(this.data.1comp.species) > 1) {
          this.data.1comp.species <- this.data.1comp.species[1, ]
        }

        this.row$AIC.1comp <- this.data.1comp.species$AIC
        Vdist.1comp <- this.data.1comp.species$Vdist
        kelim.1comp <- this.data.1comp.species$kelim
        this.row$Reference <- this.data.1comp.species$Reference
      }
    }

    if (this.cas %in% fits2comp$CAS)
    {
      this.data.2comp.species <- subset(this.data.2comp.species,
                                        param.value.type=="Fitted geometric mean")

      # If more than one source, require that the joint analysis worked:
      if (dim(this.data.2comp.species)[1]>1)
      {
        this.data.2comp.species <- subset(this.data.2comp.species, 
                                          Data.Analyzed=="Joint Analysis")
      }
      if (dim(this.data.2comp.species)[1]>0) 
      {
        if(nrow(this.data.2comp.species) > 1 ) {
          this.data.2comp.species <- this.data.2comp.species[1, ]
        }
        this.row$AIC.2comp <- this.data.2comp.species$AIC
        Vdist.2comp <- this.data.2comp.species$Vss
        kelim.2comp <- this.data.2comp.species$kelim
        this.row$Reference <- this.data.2comp.species$Reference
      }
    }

    if (!is.na(this.row$AIC.1comp))
    {
      if (!is.na(this.row$AIC.2comp))
      {
        if (this.row$AIC.1comp < this.row$AIC.2comp)
        {
          this.row$Model <- "1Comp"
        } else this.row$Model <- "2Comp"
      } else this.row$Model <- "1Comp"
    } else if (!is.na(this.row$AIC.2comp))
    {
      this.row$Model <- "2Comp"
    }

    if (!is.na(this.row$Model))
      if (this.row$Model == "1Comp")
      {
        this.row$Vdist <- Vdist.1comp
        this.row$kelim <- kelim.1comp
      } else 
      {
        this.row$Vdist <- Vdist.2comp
        this.row$kelim <- kelim.2comp
      }

    fittable <- rbind(fittable,this.row)
  }
}

# debug(test)
# test(fits1comp, fits2comp)
fittable$halflife <- log(2)/fittable$kelim

#Lets only keep chemicals where we can fit a PK model:
fittable <- subset(fittable, !is.na(Model))

for (this.col in c("AIC.1comp","AIC.2comp","Vdist","kelim","halflife"))
  fittable[,this.col] <- signif(as.numeric(fittable[,this.col]),4)

# We have parameters estimated for 86 chemicals:
dim(fittable)[1]

### compounds fit better by flat model should be removed
# if present in fittable
### some are only fit by the flat model,
# but some are fit just fit better by the flat model

# these are
# hexachlorobenzene -- nevermind, for the joint 1comp does best
# phenolphthalein glucuronide
# tamoxifen
# 2,3,7,8-tetrachlorodibenzo-p-dioxin
# imidacloprid", # well this one is only fit by the flat,
# so isn't present anyway

### when I take these out, 48 1 compartment best fits, 59 2 compartment, 107 total

### I also manually checked all Joint Analysis compounds 
# to make sure no joint flat models performed better
# none did

# write.csv(fittable, file = "inst/ext/model_params_table_04182022.csv", row.names = FALSE)

```

```{r, make_no_fit_table}

##################
##################

pk_data1 <- as.data.table(pk_output_1comp_04282022)
pk_data2 <- as.data.table(pk_output_2comp_04282022)

##################
##################

pk_data1 <- pk_data1 %>%
  filter(param.value.type == "Fitted geometric mean") %>%
  dplyr::select(c("Compound", "CAS", "Reference", "AIC")) %>%
  mutate(model = "1-Compartment")

pk_data2 <- pk_data2 %>%
  filter(param.value.type == "Fitted geometric mean") %>%
  dplyr::select(c("Compound", "CAS", "Reference", "AIC")) %>%
  mutate(model = "2-Compartment")

pk_all <- rbind(pk_data1, pk_data2)

### loop that says if a compound has any references that do not have AIC of Inf, it was a fit compound
this.compound.list <- list()
for (this.compound in unique(pk_all[, Compound])) {

  this.compound.data <- pk_all[Compound == this.compound]

  if(any(this.compound.data$AIC != Inf)) {
    this.compound.data$status <- "good fit"
  } else {
    this.compound.data$status <- "bad fit"
  }

  this.compound.list[[this.compound]] <- this.compound.data

}

pk_all <- do.call(rbind, this.compound.list)

pk_all <- pk_all %>%
  filter(status == "bad fit") %>% ### only want bad fit
  mutate(status = NULL) %>% ### get rid of status column
  group_by(Compound) %>%  
  filter(n() == 2) %>% ### group_by compound and only keep compounds with two observations, aka neither model worked
  mutate(model = NULL) %>% ### get rid of model column
  distinct() ### discard repeated compound names

# write.csv(pk_all, file = "inst/ext/no_fit_table_03012022.csv", row.names = FALSE)
```

```{r, compare_AIC}

##################
##################

pk_data1 <- pk_output_1comp_04282022

pk_data2 <- pk_output_2comp_04282022

pk_data_flat <- pk_output_flat_04282022 %>%
  filter(param.value.type == "Predicted" | param.value.type == "Fitted geometric mean" | param.value.type == "Fitted geometric std dev")

##################
##################

### make both data.frames have the same column names so they are rbind-able
pk_data1 <- pk_data1[, intersect(intersect(colnames(pk_data2), colnames(pk_data_flat)), colnames(pk_data1))]
pk_data2 <- pk_data2[, intersect(intersect(colnames(pk_data1), colnames(pk_data_flat)), colnames(pk_data2))]
pk_data_flat <- pk_data_flat[, intersect(intersect(colnames(pk_data1), colnames(pk_data2)), colnames(pk_data_flat))]

pk_data1 <- pk_data1 %>%
  rename(LogLikelihood_1 = LogLikelihood,
         AIC_1 = AIC,
         sigma_value_1 = sigma_value) %>%
  mutate(model = NULL) %>%
  mutate(model.type = NULL)

pk_data2 <- pk_data2 %>%
  rename(LogLikelihood_2 = LogLikelihood,
         AIC_2 = AIC,
         sigma_value_2 = sigma_value) %>%
  mutate(model = NULL) %>%
  mutate(model.type = NULL)

pk_data_flat <- pk_data_flat %>%
  rename(LogLikelihood_flat = LogLikelihood,
         AIC_flat = AIC,
         sigma_value_flat = sigma_value) %>%
  mutate(model = NULL) %>%
  mutate(model.type = NULL)

pk_all <- left_join(pk_data1, pk_data2, by = c("CAS",
                                               "Species",
                                               "param.value.type",
                                               "sigma_id",
                                               "Reference",
                                               "Data.Analyzed",
                                               "Compound"))

pk_all <- left_join(pk_all, pk_data_flat, by = c("CAS",
                                                 "Species",
                                                 "param.value.type",
                                                 "sigma_id",
                                                 "Reference",
                                                 "Data.Analyzed",
                                                 "Compound"))

pk_all <- subset(pk_all, select = c(CAS, Species, Compound, Reference, AIC_1, AIC_2, AIC_flat))

pk_all <- distinct(pk_all)

pk_all$AIC_1[pk_all$AIC_1 == Inf] <- NA
pk_all$AIC_2[pk_all$AIC_2 == Inf] <- NA
pk_all$AIC_flat[pk_all$AIC_flat == Inf] <- NA

pk_all <- pk_all %>%
  mutate(flat_lowest = ifelse(AIC_flat < AIC_1 & AIC_flat < AIC_2, 1, 0)) %>%
  mutate(flat_better_than_1 = ifelse(AIC_flat < AIC_1 & AIC_flat > AIC_2, 1, 0)) %>%
  mutate(flat_better_than_2 = ifelse(AIC_flat > AIC_1 & AIC_flat < AIC_2, 1, 0))

pk_just_bad <- pk_all %>%
  filter(flat_lowest == 1 | flat_better_than_1 == 1 | flat_better_than_2 == 1)
```

```{r, lombardo_comparisons}

##################
##################

pk_data1 <- pk_output_1comp_05092022
# pk_data1 <- pk_output_1comp_04282022

# pk_data1$Vdist <- log(pk_data1$Vdist)

pk_lombardo <- pk_lombardo

##################
##################

pk_data1 <- pk_data1 %>%
  filter(param.value.type == "Fitted geometric mean")
         # , CAS != "50-49-7", CAS != "106266-06-2")
# , Species == "human")

pk_all <- left_join(pk_data1, pk_lombardo, by = c("CAS" = "CAS #"))

pk_all <- pk_all %>%
  filter(!is.na(`human VDss (L/kg)`), 
         !is.na(`human CL (mL/min/kg)`))

pk_all$`human CL (L/hr/kg)` <- pk_all$`human CL (mL/min/kg)` * 60 / 1000

pk_all$Vdist_norm <- abs(pk_all$Vdist / pk_all$`human VDss (L/kg)`)
pk_all$CLtot_norm <- abs(pk_all$CLtot / pk_all$`human CL (L/hr/kg)`)
pk_all$halflife_norm <- abs(pk_all$halflife / pk_all$`terminal  t1/2 (h)`)

pk_all$mae <- rowMeans(pk_all[, c("Vdist_norm", "CLtot_norm", "halflife_norm")])

### volume of distribution comparison
p1 <- ggplot(pk_all, aes(x = Vdist, y = `human VDss (L/kg)`)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  geom_abline(slope = 1, intercept = 0)

p1

### half-life comparision
p2 <- ggplot(pk_all, aes(x = halflife, y = `terminal  t1/2 (h)`)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  geom_abline(slope = 1, intercept = 0)

p2

### clearance comparison

### convert to same units

p3 <- ggplot(data = pk_all, mapping = aes(x = CLtot, y = `human CL (L/hr/kg)`)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  geom_abline(slope = 1, intercept = 0)

p3

```

<!-- # ```{r, make_model_params_table, eval = FALSE} -->
<!-- #  -->
<!-- # ### this code almost entirely is imported from John's bake-off RMarkdown file -->
<!-- # ### what this produces is a table of compounds that were fit by at least one of the models -->
<!-- # ### AICs for both models are reported, and subsequent parameter values correspond to the model with the lower AIC -->
<!-- #  -->
<!-- # ################### -->
<!-- # ################### -->
<!-- #  -->
<!-- # pk_data1 <- pk_output_1comp_04142022  -->
<!-- # # pk_data1 <- pk_output_1comp_12212021  -->
<!-- #  -->
<!-- # pk_data2 <- pk_output_2comp_04142022  -->
<!-- # # pk_data2 <- pk_output_2comp_12212021  -->
<!-- #  -->
<!-- # pk_data_flat <- pk_output_flat_04142022 -->
<!-- #  -->
<!-- # ################### -->
<!-- # ################### -->
<!-- #  -->
<!-- # pk_data1$sigma_id <- NULL -->
<!-- # pk_data1$sigma_value <- NULL -->
<!-- # pk_data2$sigma_id <- NULL -->
<!-- # pk_data2$sigma_value <- NULL -->
<!-- # pk_data_flat$sigma_id <- NULL -->
<!-- # pk_data_flat$sigma_value <- NULL -->
<!-- #  -->
<!-- # pk_data1 <- distinct(pk_data1) -->
<!-- # pk_data2 <- distinct(pk_data2) -->
<!-- # pk_data_flat <- distinct(pk_data_flat) -->
<!-- #  -->
<!-- # fits1comp <- subset(pk_data1, is.finite(AIC) & AIC < 10^4) -->
<!-- # fits2comp <- subset(pk_data2, is.finite(AIC) & AIC < 10^4) -->
<!-- # fitsflat <- subset(pk_data_flat, is.finite(AIC) & AIC < 10^4) -->
<!-- #  -->
<!-- # #How many chemicals have 1 or 2 compartment fits: -->
<!-- # length(unique(c(fits1comp$CAS,fits2comp$CAS,fitsflat$CAS))) -->
<!-- #  -->
<!-- # fittable <- NULL -->
<!-- # # test <- function(fits1comp, fits2comp, fitsflat) { -->
<!-- #   for (this.cas in unique(c(fits1comp$CAS,fits2comp$CAS, fitsflat$CAS))) -->
<!-- #   { print(this.cas) -->
<!-- #     # if (this.cas == "96-47-9") browser() -->
<!-- #     if (this.cas %in% fits1comp$CAS) this.compound <- fits1comp[ -->
<!-- #       fits1comp$CAS==this.cas,"Compound"][1] -->
<!-- #     else if (this.cas %in% fits2comp$CAS) this.compound <- fits2comp[ -->
<!-- #       fits2comp$CAS==this.cas,"Compound"][1] -->
<!-- #     else if (this.cas %in% fitsflat$CAS) this.compound <- fitsflat[ -->
<!-- #       fitsflat$CAS==this.cas,"Compound"][1] -->
<!-- #      -->
<!-- #     this.data.1comp <- subset(fits1comp,CAS==this.cas) -->
<!-- #     this.data.2comp <- subset(fits2comp,CAS==this.cas) -->
<!-- #     this.data.flat <- subset(fitsflat,CAS==this.cas) -->
<!-- #      -->
<!-- #     for (this.species in unique(tolower(c( -->
<!-- #       this.data.1comp$Species, -->
<!-- #       this.data.2comp$Species, -->
<!-- #       this.data.flat$Species)))) -->
<!-- #     { -->
<!-- #       Vdist.1comp <- NA -->
<!-- #       kelim.1comp <- NA -->
<!-- #       Vdist.2comp <- NA -->
<!-- #       kelim.2comp <- NA -->
<!-- #       A.flat <- NA -->
<!-- #        -->
<!-- #       this.data.1comp.species <- subset(this.data.1comp,tolower(Species)==this.species) -->
<!-- #       this.data.2comp.species <- subset(this.data.2comp,tolower(Species)==this.species) -->
<!-- #       this.data.flat.species <- subset(this.data.flat, tolower(Species)==this.species) -->
<!-- #        -->
<!-- #       this.row <- data.frame( -->
<!-- #         Compound=this.compound, -->
<!-- #         CAS=this.cas, -->
<!-- #         Species=this.species, -->
<!-- #         Reference=NA, -->
<!-- #         AIC.1comp=NA, -->
<!-- #         AIC.2comp=NA, -->
<!-- #         Model=NA, -->
<!-- #         Vdist=NA, -->
<!-- #         kelim=NA) -->
<!-- #        -->
<!-- #       if (this.cas %in% fits1comp$CAS) -->
<!-- #       { -->
<!-- #         this.data.1comp.species <- subset(this.data.1comp.species, -->
<!-- #                                           param.value.type=="Fitted geometric mean") -->
<!-- #          -->
<!-- #         # If more than one source, require that the joint analysis worked: -->
<!-- #         if (dim(this.data.1comp.species)[1]>1) -->
<!-- #         { -->
<!-- #           this.data.1comp.species<- subset(this.data.1comp.species,  -->
<!-- #                                            Data.Analyzed=="Joint Analysis") -->
<!-- #         } -->
<!-- #         if (dim(this.data.1comp.species)[1]>0)  -->
<!-- #         { -->
<!-- #           if (nrow(this.data.1comp.species) > 1) { -->
<!-- #             this.data.1comp.species <- this.data.1comp.species[1, ] -->
<!-- #           } -->
<!-- #            -->
<!-- #           this.row$AIC.1comp <- this.data.1comp.species$AIC -->
<!-- #           Vdist.1comp <- this.data.1comp.species$Vdist -->
<!-- #           kelim.1comp <- this.data.1comp.species$kelim -->
<!-- #           this.row$Reference <- this.data.1comp.species$Reference -->
<!-- #         } -->
<!-- #       } -->
<!-- #        -->
<!-- #       if (this.cas %in% fits2comp$CAS) -->
<!-- #       { -->
<!-- #         this.data.2comp.species <- subset(this.data.2comp.species, -->
<!-- #                                           param.value.type=="Fitted geometric mean") -->
<!-- #          -->
<!-- #         # If more than one source, require that the joint analysis worked: -->
<!-- #         if (dim(this.data.2comp.species)[1]>1) -->
<!-- #         { -->
<!-- #           this.data.2comp.species <- subset(this.data.2comp.species,  -->
<!-- #                                             Data.Analyzed=="Joint Analysis") -->
<!-- #         } -->
<!-- #         if (dim(this.data.2comp.species)[1]>0)  -->
<!-- #         { -->
<!-- #           if(nrow(this.data.2comp.species) > 1 ) { -->
<!-- #             this.data.2comp.species <- this.data.2comp.species[1, ] -->
<!-- #           } -->
<!-- #           this.row$AIC.2comp <- this.data.2comp.species$AIC -->
<!-- #           Vdist.2comp <- this.data.2comp.species$Vss -->
<!-- #           kelim.2comp <- this.data.2comp.species$kelim -->
<!-- #           this.row$Reference <- this.data.2comp.species$Reference -->
<!-- #         } -->
<!-- #       } -->
<!-- #        -->
<!-- #       if (this.cas %in% fitsflat$CAS) -->
<!-- #       { -->
<!-- #         this.data.flat.species <- subset(this.data.flat.species, -->
<!-- #                                          param.value.type=="Fitted geometric mean") -->
<!-- #          -->
<!-- #         # If more than one source, require that the joint analysis worked: -->
<!-- #         if (dim(this.data.flat.species)[1]>1) -->
<!-- #         { -->
<!-- #           this.data.flat.species <- subset(this.data.flat.species,  -->
<!-- #                                            Data.Analyzed=="Joint Analysis") -->
<!-- #         } -->
<!-- #         if (dim(this.data.flat.species)[1]>0)  -->
<!-- #         { -->
<!-- #           if(nrow(this.data.flat.species) > 1 ) { -->
<!-- #             this.data.flat.species <- this.data.flat.species[1, ] -->
<!-- #           } -->
<!-- #           this.row$AIC.flat <- this.data.flat.species$AIC -->
<!-- #           A.flat <- this.data.flat.species$A -->
<!-- #           this.row$Reference <- this.data.flat.species$Reference -->
<!-- #         } -->
<!-- #       } -->
<!-- #       ######################## -->
<!-- #       ######################### -->
<!-- #        -->
<!-- #       if(!"AIC.flat" %in% names(this.row)) this.row <- this.row %>%  -->
<!-- #         add_column(AIC.flat = NA) -->
<!-- #        -->
<!-- #       if(!"AIC.1comp" %in% names(this.row)) this.row <- this.row %>%  -->
<!-- #         add_column(AIC.1comp = NA) -->
<!-- #        -->
<!-- #       if(!"AIC.2comp" %in% names(this.row)) this.row <- this.row %>%  -->
<!-- #         add_column(AIC.2comp = NA) -->
<!-- #        -->
<!-- #       this.row <- this.row %>% mutate(Model = case_when(AIC.1comp < AIC.2comp &  -->
<!-- #                                                           AIC.1comp < AIC.flat ~ "1Comp", -->
<!-- #                                                         AIC.2comp < AIC.1comp & -->
<!-- #                                                           AIC.2comp < AIC.flat ~ "2Comp", -->
<!-- #                                                         AIC.flat < AIC.1comp & -->
<!-- #                                                           AIC.flat < AIC.2comp ~ "Null", -->
<!-- #                                                          -->
<!-- #                                                         !is.na(AIC.1comp) & is.na(AIC.2comp) & -->
<!-- #                                                           is.na(AIC.flat) ~ "1Comp", -->
<!-- #                                                         !is.na(AIC.2comp) & is.na(AIC.1comp) &  -->
<!-- #                                                           is.na(AIC.flat) ~ "2Comp", -->
<!-- #                                                         !is.na(AIC.flat) & is.na(AIC.1comp) &  -->
<!-- #                                                           is.na(AIC.2comp) ~ "Null", -->
<!-- #                                                          -->
<!-- #                                                         is.na(AIC.1comp) & -->
<!-- #                                                           AIC.flat < AIC.2comp ~ "Null", -->
<!-- #                                                         is.na(AIC.1comp) & -->
<!-- #                                                           AIC.2comp < AIC.flat ~ "2Comp", -->
<!-- #                                                         is.na(AIC.2comp) & -->
<!-- #                                                           AIC.flat < AIC.1comp ~ "Null", -->
<!-- #                                                         is.na(AIC.2comp) & -->
<!-- #                                                           AIC.1comp < AIC.flat ~ "1Comp", -->
<!-- #                                                         is.na(AIC.flat) & -->
<!-- #                                                           AIC.1comp < AIC.2comp ~"1Comp", -->
<!-- #                                                         is.na(AIC.flat) & -->
<!-- #                                                           AIC.2comp <AIC.1comp ~"2Comp")) -->
<!-- #        -->
<!-- #        -->
<!-- #       fittable <- rbind(fittable,this.row) -->
<!-- #     } -->
<!-- #   } -->
<!-- # # } -->
<!-- # # debug(test) -->
<!-- # # test(fits1comp, fits2comp, fitsflat) -->
<!-- # fittable$halflife <- log(2)/fittable$kelim -->
<!-- #  -->
<!-- # #Lets only keep chemicals where we can fit a PK model: -->
<!-- # fittable <- subset(fittable, !is.na(Model)) -->
<!-- #  -->
<!-- # # for (this.col in c("AIC.1comp","AIC.2comp","Vdist","kelim","halflife")) -->
<!-- # #   fittable[,this.c=-] <- signif(as.numeric(fittable[,this.col]),4) -->
<!-- #  -->
<!-- # # We have parameters estimated for 86 chemicals: -->
<!-- # dim(fittable)[1] -->
<!-- #  -->
<!-- # # write.csv(fittable, file = "inst/ext/model_params_table_04182022.csv", row.names = FALSE) -->
<!-- #  -->
<!-- # ``` -->