### 2023-01-06 Working Rmds/impute_missing_sds.Rmd
Fix error in computing sd vs. mean relation, and implement the fixed result when imputing missing SDs.

---
title: "Imputing missing SDs"
author: "Caroline Ring"
date: "2023-01-04"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(results = TRUE)
knitr::opts_chunk$set(message = FALSE)
```

```{r, include = FALSE}
library(data.table)
devtools::load_all("../invivopkfit")
```

```{r}
timestamp <- "2023_01_05"
```

Read in CvTdb data pulled in `pulling_oral_iv_data_cvtdb.Rmd`.

```{r, load_cvt, eval = TRUE}
cvt_data <- fread(paste0("../inst/ext/cvt_data_",
timestamp,
".csv"))
```
List of new_name = old_name to supply for preprocessing data

```{r}
names_list <- list(
  "Compound_Dosed" = "studies.test_substance_name_original",
  "DTXSID_Dosed" = "chemicals_dosed.dsstox_substance_id",
  "CAS_Dosed" = "chemicals_dosed.dsstox_casrn",
  "Compound_Analyzed" = "series.analyte_name_original",
  "DTXSID_Analyzed" = "chemicals_analyzed.dsstox_substance_id",
  "CAS_Analyzed" = "chemicals_analyzed.dsstox_casrn",
  "Reference" = "documents_reference.id",
  "Extraction" = "documents_extraction.id",
  "Species" = "subjects.species",
  "Weight" ="subjects.weight_kg",
  "Weight.Units" = NULL,
  "Dose" = "studies.dose_level_normalized_corrected", #note that dose levels were corrected in pulling_oral_iv_data_cvtdb.Rmd
  "Dose.Units" = NULL,
  "Time" = "conc_time_values.time_hr",
  "Time.Units" = NULL,
  "Media" = "series.conc_medium_normalized",
  "Value" = "conc_time_values.conc", #already normalized to mg/L units in CvTdb
  "Value.Units" = NULL,
  "Route" = "studies.administration_route_normalized",
  "LOQ" = "series.loq_normalized", #note that LOQs were normalized to mg/L units in pulling_oral_iv_data_cvtdb.Rmd
  "Subject" = "subjects.id",
  "N_Subjects" = "series.n_subjects_in_series",
  "Value_SD" = "conc_time_values.conc_sd_normalized" #note that SDs were normalized to mg/L units in pulling_oral_iv_data_cvtdb.Rmd
  )
```

List of default values for new columns added in preprocessing data

```{r}
defaults_list <- list(
                      "Weight.Units" = "kg",
                      "Dose.Units" = "mg/kg",
                      "Time.Units" = "h",
                      "Value.Units" = "mg/L")
```

For ease of plotting, get the pre-processed, harmonized data:

```{r}
cvt <- preprocess_data(data.set = cvt_data,
                       names_list =names_list,
                       defaults_list =   defaults_list,
                       ratio_conc_to_dose = 1,
                       calc_loq_factor = 0.45,
                       routes_keep = c("po", "iv"),
                       media_keep = c("blood", "plasma"),
                       impute_loq = TRUE,
                       impute_sd = FALSE,
                       suppress.messages = FALSE)
cvt <- as.data.table(cvt)
```

There are many, many cases where N > 1 but no SD is reported. Can we impute these based on other data?

# Individual subject data

By chemical, species, reference, medium, dose group, route, and time point: calculate mean and standard deviation.

```{r}

cvt_summarized <- cvt[N_Subjects==1,
                      .(meanval = mean(Value, na.rm = TRUE),
                        sdvals = sd(Value, na.rm = TRUE),
                        n = .N),
                      by = .(Reference,
                             DTXSID,
                             Species,
                             Media,
                             Route,
                             Dose,
                             Time)]
```

Plot SD vs. mean across all chemicals, studies, media, doses, and timepoints. Is there an overall relation between SD and mean?

```{r}

ggplot(cvt_summarized,
       aes(x = meanval,
           y = sdvals)) +
  geom_point(aes(color = n)) +
  geom_abline(intercept = 0, slope =1 , linetype = 2) +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  scale_x_log10() +
  scale_y_log10() +
  scale_color_viridis_c(trans = "log10",
                        name = "# subjects") +
  annotation_logticks(sides = "bl") +
  xlab("Mean conc., mg/kg") +
  ylab("Sample std dev conc., mg/kg") +
  ggtitle("Mean vs. SD for N=1 data") 
```

Linear regression of the relation:

```{r}
cvt_summarized[, log10_sdvals := log10(sdvals)]
cvt_summarized[, log10_meanval := log10(meanval)]

cvt_summarized[!is.finite(log10_sdvals), log10_sdvals := NA_real_]
cvt_summarized[!is.finite(log10_meanval), log10_meanval := NA_real_]

summary(lm(log10_sdvals ~ log10_meanval,
           data = cvt_summarized,
           na.action = na.omit))
```

What about the multi-subject cases with reported SD?

```{r}
ggplot(cvt[N_Subjects > 1],
       aes(x = Value,
           y = Value_SD)) +
  geom_point(aes(color = N_Subjects)) +
  geom_abline(intercept = 0, slope =1 , linetype = 2) +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  scale_x_log10() +
  scale_y_log10() +
  scale_color_viridis_c(trans = "log10",
                        name = "# subjects") +
  annotation_logticks(sides = "bl") +
  xlab("Mean conc., mg/kg") +
  ylab("Sample std dev conc., mg/kg") +
  ggtitle("Mean vs. SD for N>1 data") 
```

```{r}

summary(lm(log10(Value_SD) ~ log10(Value),
           data = cvt[N_Subjects > 1]))
```

Studies with entirely missing SDs are different from the studies where only a few SDs are missing. When only a few are missing, it was usually because the data were pulled from a plot where the error bars were smaller thant the plotting symbol on a few points. Instead of recording the size of the symbol as the error bars, the error was recorded as NA. But when all SDs are missing, it is usually because the study did not report the SDs in the first place.

I suggest using the relation from the summarized individual-animal data to impute standard deviations. That is

$$ \log \textrm{SD} = -0.555 + 0.939 \log \textrm{Mean} $$

When translated to the natural scale (raising 10 to the power of both sides), this becomes

$$ \log_{10} \textrm{SD} = -0.555 + \log_{10} \textrm{Mean}^{0.939} $$
$$ \log_{10} \textrm{SD} = \log_{10}{ \left( 10^{-0.555} \times \textrm{Mean}^{0.939} \right)}  $$

Raising 10 to the power of both sides,

$$ \textrm{SD} = 10^{-0.555} \times \textrm{Mean}^{0.939} $$

or approximately

$$ \textrm{SD} = 0.278 \times \textrm{Mean}^{0.939} $$

Very roughly, this is about a 30% coefficient of variation.