### 2023-01-10 Working Rmds/plot_fits.Rmd
Read in pre-processed data.

---
title: "Plot curve fits"
author: "Caroline Ring"
date: "2023-01-04"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(results = TRUE)
knitr::opts_chunk$set(message = FALSE)
```

```{r, include = FALSE}
library(tidyverse)
library(data.table)
library(readxl)
devtools::load_all("../invivopkfit")
```

Timestamp

```{r}
timestamp <- "2023_01_06"
```

Read in CvTdb data pulled in `pulling_oral_iv_data_cvtdb.Rmd`.

```{r, load_cvt, eval = TRUE}
cvt <- fread("../inst/ext/cvt_preprocessed_2023_01_06.csv")
cvt[, Reference:=as.character(Reference)]
cvt[, Reference_orig:=as.character(Reference_orig)]
```

Read in the fitted model parameter values.

```{r, read_data}

PK.fit.table.1comp <- fread(paste0("../inst/ext/", "PK_fit_table_1comp_", timestamp, ".csv",sep=""))
PK.fit.table.2comp <- fread(paste0("../inst/ext/", "PK_fit_table_2comp_", timestamp, ".csv",sep=""))
PK.fit.table.flat <- fread(paste0("../inst/ext/", "PK_fit_table_flat_", timestamp, ".csv",sep=""))

#pk_lombardo <- read_xlsx("inst/ext/Supplemental_82966_revised_corrected.xlsx")
```

Rowbind together all of the fitted parameter tables into one big one.
```{r}
pk_fit <- rbindlist(list("1comp" = PK.fit.table.1comp,
                         "2comp" = PK.fit.table.2comp,
                         "flat" = PK.fit.table.flat),
                    use.names = TRUE,
                    fill = TRUE)

rm(PK.fit.table.1comp,
   PK.fit.table.2comp,
   PK.fit.table.flat)
```

Get a table of the winning model for each dataset (the one with minimum AIC).

```{r}
tmp <- unique(pk_fit[, .(DTXSID, Species, Analysis_Type, References.Analyzed, model, AIC)])

#sort AIC from smallest to largest within each dataset, putting NAs last
setorder(tmp, DTXSID, Species, Analysis_Type, References.Analyzed, AIC, na.last = TRUE)
#the winning model will now be listed first within each dataset (smallest AIC)
model_win <- tmp[, .(winmodel = model[1]), by = .(DTXSID, Species, Analysis_Type, References.Analyzed)]
pk_fit <- model_win[pk_fit, on = c("DTXSID", "Species", "Analysis_Type", "References.Analyzed")]
pk_fit[, winning:=(model == winmodel)]

```

Plot concentration vs time with the best-fit params by data subset (DTXSID and species).

Indicate the following:

- One line for each model
- Which is the winning model
- Dose group
- Blood vs. plasma
- Below LOQ vs. above LOQ (plot below-LOQ points at LOQ)
- IV vs. oral administration

Would also like to plot uncertainty in predictions based on uncertainty in parameter estimates.

Plot versions:
- Winning model only
- One model at a time
- All three models on same plot

Get unique combinations of DTXSID and species -- each of these will be one plot.

```{r}
plot_table <- unique(pk_fit[, .(DTXSID, Species)])
```

Plot all models for the joint analysis.
```{r}

#plot all models
tmp <- mapply(plot_fit,
              DTXSID_in = plot_table$DTXSID,
                     Species_in = plot_table$Species,
              MoreArgs = list(
                     model_in = "all",
                     Analysis_Type_in = "Joint",
                     DF = cvt,
                     pk_fit = pk_fit,
                     n_interp_time = 10,
                     limit_y_axis = FALSE,
                     log10_scale_y = FALSE,
                     plot_errbars = TRUE,
                     log10_scale_time = FALSE,
                     file_path = "../inst/ext/plots/Joint Analysis/model_all/errbars_naturaly")
                     )

```

```{r}

#plot all models
tmp <- mapply(plot_fit,
              DTXSID_in = plot_table$DTXSID,
                     Species_in = plot_table$Species,
              MoreArgs = list(
                     model_in = "all",
                     Analysis_Type_in = "Joint",
                     DF = cvt,
                     pk_fit = pk_fit,
                     n = 101,
                     limit_y_axis = FALSE,
                     log10_scale_y = TRUE,
                     plot_errbars = FALSE,
                     log10_scale_time = FALSE,
                     file_path = "../inst/ext/plots/Joint Analysis/model_all/noerrbars_logy")
                     )

```

Problem: DTXSID0022985_rat_Joint_all does not plot any model lines. Why?

```{r}
pk_fit[DTXSID %in% "DTXSID0022985"]
```

Apparently there were no detected data points?

```{r}
cvt[DTXSID %in% "DTXSID0022985", .(Reference, Value, LOQ)]
```

Hmm. Something is very weird here. I definitely see some detected points. They are all the same reference, so it isn't that a single-reference fit did not work but a joint fit did work.

Plot only the winning model

```{r}
#plot all models
tmp <- mapply(plot_fit,
              DTXSID_in = plot_table$DTXSID,
                     Species_in = plot_table$Species,
              MoreArgs = list(
                Analysis_Type_in = c("Joint", "Separate"),
                     model_in = "winning",
                     fitdata = DF,
                     pk_fit = pk_fit,
                     file_path = "inst/ext/plots/analysis_all_model_winning/",
                log10_scale_time = TRUE)
                     )
```