### 2023-01-12 Working Rmds/MainAnalysis.Rmd
note on weird fit

---
title: "invivoPKfit analysis of CvTdb data"
author: "Christopher Cook, John Wambaugh, Caroline Ring"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document: default
  word_document: default
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(results = TRUE)
knitr::opts_chunk$set(message = FALSE)
```

```{r, include = FALSE}
library(tidyverse)
library(data.table)
library(readxl)
devtools::load_all("../invivopkfit")
```

# Read in CvTdb data

Read in CvTdb data that was pulled and normalized in `pulling_oral_iv_data_cvtdb.Rmd`.

```{r, load_cvt, eval = TRUE}
cvt_data <- fread("../inst/ext/cvt_data_2023_01_05.csv")
```

# Setup for data processing and model fitting

## Timestamp
Timestamp the file name with the current date and time (at the beginning of the runs) in format YYYY_MM_DD_hh_mm.

```{r}
timestamp <- format(Sys.time(), "%Y_%m_%d")
```

## Settings for pre-processing data

List of new_name = old_name to supply for preprocessing data

```{r}
names_list <- list(
  "Compound_Dosed" = "studies.test_substance_name_original",
  "DTXSID_Dosed" = "chemicals_dosed.dsstox_substance_id",
  "CAS_Dosed" = "chemicals_dosed.dsstox_casrn",
  "Compound_Analyzed" = "series.analyte_name_original",
  "DTXSID_Analyzed" = "chemicals_analyzed.dsstox_substance_id",
  "CAS_Analyzed" = "chemicals_analyzed.dsstox_casrn",
  "Reference" = "documents_reference.id",
  "Extraction" = "documents_extraction.id",
  "Species" = "subjects.species",
  "Weight" ="subjects.weight_kg",
  "Weight.Units" = NULL,
  "Dose" = "studies.dose_level_normalized_corrected", #note that dose levels were corrected in pulling_oral_iv_data_cvtdb.Rmd
  "Dose.Units" = NULL,
  "Time" = "conc_time_values.time_hr",
  "Time.Units" = NULL,
  "Media" = "series.conc_medium_normalized",
  "Value" = "conc_time_values.conc", #already normalized to mg/L units in CvTdb
  "Value.Units" = NULL,
  "Route" = "studies.administration_route_normalized",
  "LOQ" = "series.loq_normalized", #note that LOQs were normalized to mg/L units in pulling_oral_iv_data_cvtdb.Rmd
  "Subject" = "subjects.id",
  "N_Subjects" = "series.n_subjects_in_series",
  "Value_SD" = "conc_time_values.conc_sd_normalized" #note that SDs were normalized to mg/L units in pulling_oral_iv_data_cvtdb.Rmd
  )
```

List of default values for new columns added in preprocessing data

```{r}
defaults_list <- list(
                      "Weight.Units" = "kg",
                      "Dose.Units" = "mg/kg",
                      "Time.Units" = "hours",
                      "Value.Units" = "mg/L")
```

# Preprocess data

```{r}
cvt <- preprocess_data(data.set = cvt_data,
                       names_list =names_list,
                       defaults_list =   defaults_list,
                       ratio_conc_to_dose = 1,
                       calc_loq_factor = 0.45,
                       routes_keep = c("po", "iv"),
                       media_keep = c("blood", "plasma"),
                       impute_loq = TRUE,
                       impute_sd = TRUE,
                       suppress.messages = FALSE)

#save pre-processed data
write.csv(cvt,
          paste0("../inst/ext/cvt_preprocessed_",
                 timestamp,
          ".csv"),
          row.names = FALSE)
```

# Fit TK models to data

## Flat model

```{r, run_invivoPKfit_flat}

system.time(
  PK.fit.table.flat <- fit_all(
    data.set = cvt,
    model = "flat",
    modelfun = "analytic",
    preprocess = FALSE,
    optimx_args = list(
      "method" = "bobyqa",
      "itnmax" = 1e6,
      "control" = list("kkt" = FALSE)
    ),
    suppress.messages = TRUE
  )
)

write.csv(PK.fit.table.flat, paste("../inst/ext/PK_fit_table_flat_",
                                    timestamp,
                                    ".csv",
                                    sep=""), row.names = FALSE)

post_flat <- postprocess_data(PK_fit = PK.fit.table.flat,
                               model = "flat")

write.csv(post_flat,
          paste("../inst/ext/PK_fit_postprocess_flat_",
                                    timestamp,
                                    ".csv",
                                    sep=""),
          row.names = FALSE)

rm(PK.fit.table.flat, post_flat)
```

## 1-compartment model

```{r, run_invivoPKfit_1comp}
### do 1-Compartment fit
system.time(
    PK.fit.table.1comp <- fit_all(
      data.set = cvt,
      model = "1compartment",
      modelfun = "analytic",
      preprocess = FALSE,
      optimx_args = list(
                      "method" = "bobyqa",
                      "itnmax" = 1e6,
                      "control" = list("kkt" = FALSE)
                    ),
      suppress.messages = TRUE
  )
)

write.csv(PK.fit.table.1comp,
          paste("../inst/ext/PK_fit_table_1comp_",
                                    timestamp,
                                    ".csv",
                                    sep=""),
          row.names = FALSE)

post_1comp <- postprocess_data(PK_fit = PK.fit.table.1comp,
                               model = "1compartment")

write.csv(post_1comp,
          paste("../inst/ext/PK_fit_postprocess_1comp_",
                                    timestamp,
                                    ".csv",
                                    sep=""),
          row.names = FALSE)

rm(PK.fit.table.1comp, post_1comp)

```

## 2-compartment model

```{r, run_invivoPKfit_2comp}
### do 2-Compartment fit
system.time(
    PK.fit.table.2comp <- fit_all(
      data.set = cvt,
      model = "2compartment",
      modelfun = "analytic",
      preprocess = FALSE,
      optimx_args = list(
                      "method" = "bobyqa",
                      "itnmax" = 1e6,
                      "control" = list("kkt" = FALSE)
                    ),
      suppress.messages = TRUE
    )
)

write.csv(PK.fit.table.2comp, paste("../inst/ext/PK_fit_table_2comp_",
                                    timestamp,
                                    ".csv",
                                    sep=""), row.names = FALSE)

post_2comp <- postprocess_data(PK_fit = PK.fit.table.2comp,
                               model = "2compartment")

write.csv(post_2comp,
          paste("../inst/ext/PK_fit_postprocess_2comp_",
                                    timestamp,
                                    ".csv",
                                    sep=""),
          row.names = FALSE)

rm(PK.fit.table.2comp, post_2comp)
```

# Test case

DTXSID1022265 (Alachlor) is an example of one chemical where there is only one Reference ID, but the PO and IV data are on totally different scales -- IV concentration/dose ranges from 1e-3 to 1e-1, but PO concentration/dose ranges from 1e-5 to 1e-4. Now that we are fitting on natural scale, this means that with reference-specific sigma, the curve really only fits the IV data, because sigma is so large compared to the PO data scale. This results in a too-small Fgutabs estimate that does not fit the peak very well. But with "study" specific sigma -- chemical + species + reference + route + media -- the PO sigma can be on a different order than the IV sigma, and is.

Another possible thing to try is to go back to fitting on the log scale. I had switched to the natural scale because I wasn't sure how to derive the "summary" parameterization of the residual distribution for the log scale. But I could try again.