### 2023-01-17 Working Rmds/plot_fits.Rmd
Plot fits both natural and log scale, both concentration/dose and dose-faceted concentrations.

---
title: "Plot curve fits"
author: "Caroline Ring"
date: "2023-01-04"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(results = TRUE)
knitr::opts_chunk$set(message = FALSE)
```

```{r, include = FALSE}
library(tidyverse)
library(data.table)
library(readxl)
devtools::load_all("../invivopkfit")
```

Timestamp

```{r}
timestamp <- "2023_01_13"
```

Read in CvTdb data pulled in `pulling_oral_iv_data_cvtdb.Rmd`.

```{r, load_cvt, eval = TRUE}
cvt <- fread(paste0("../inst/ext/cvt_preprocessed_",
"2023_01_13",
".csv"))
cvt[, Reference:=as.character(Reference)]
cvt[, Reference_orig:=as.character(Reference_orig)]
```

Read in the fitted model parameter values.

```{r, read_data}

DT_1comp <- fread(paste0("../inst/ext/", "PK_fit_postprocess_1comp_", timestamp, ".csv",sep=""))
DT_2comp <- fread(paste0("../inst/ext/", "PK_fit_postprocess_2comp_", timestamp, ".csv",sep=""))
DT_flat <- fread(paste0("../inst/ext/", "PK_fit_postprocess_flat_", timestamp, ".csv",sep=""))

idcols <- c( "Analysis_Type",
                                 "DTXSID",
                                "Species",
                                "References.Analyzed",
                               "Studies.Analyzed",
                               "N_Routes",
                               "N_Media",
              "time_units_reported")

setnames(DT_1comp,
         setdiff(names(DT_1comp),
                 idcols),
         paste(setdiff(names(DT_1comp),
                 idcols),
               "1compartment",
               sep = "."))

setnames(DT_2comp,
         setdiff(names(DT_2comp),
                 idcols),
         paste(setdiff(names(DT_2comp),
                 idcols),
               "2compartment",
               sep = "."))

setnames(DT_flat,
         setdiff(names(DT_flat),
                 idcols),
         paste(setdiff(names(DT_flat),
                 idcols),
               "flat",
               sep = "."))

#pk_lombardo <- read_xlsx("inst/ext/Supplemental_82966_revised_corrected.xlsx")
```

Merge these together.

```{r}
pk_fit <- merge(DT_2comp,
               merge(DT_flat, DT_1comp, by = idcols),
               by = idcols)
```

Find the winning model for each dataset

```{r}
pk_AIC <- melt(pk_fit,
     id.vars = idcols,
     measure.vars = c("AIC.flat",
                      "AIC.1compartment",
                      "AIC.2compartment"),
     value.name = "AIC",
     variable.name = "model")

pk_AIC[, model := gsub(x = model,
                       pattern = "AIC.",
                       replacement = "",
                       fixed = TRUE)]

#sort AIC from smallest to largest within each dataset, putting NAs last
setorderv(pk_AIC,
        c(idcols, "AIC"),
         na.last = TRUE)
#the winning model will now be listed first within each dataset (smallest AIC)
model_win <- pk_AIC[,
                    .(winmodel = model[1]),
                    by = idcols]
pk_fit <- model_win[pk_fit,
                    on = idcols]
```

Keep parameters only -- that is, drop the calculated quantities like tpeak, Cpeak, Css, total clearance, etc.

```{r}
all_params <- unlist(sapply(c("flat",
         "1compartment",
         "2compartment"),
       function(model){
         tmp <- get_model_paramnames(model)
         return(paste(tmp,
                      model,
                      sep= "."))
       }))

pk_params <- pk_fit[,
                    .SD,
                    .SDcols = c(idcols,
                                "winmodel",
                                all_params)]

#Ensure comma-separated lists of studies analyzed are properly sorted
pk_params[, Studies.Analyzed := {
  tmp <- strsplit(x = Studies.Analyzed,
           split = ", ")
  sapply(tmp,
         function(x) paste(sort(x), collapse = ", "))
}]

# pk_params[is.na(Fgutabs_Vdist.flat),
#           Fgutabs_Vdist.flat := Fgutabs.flat / Vdist.flat]
# 
# pk_params[is.na(Rblood2plasma.flat), Rblood2plasma.flat := 1]
# pk_params[is.na(Rblood2plasma.1compartment), Rblood2plasma.1compartment := 1]
# pk_params[is.na(Rblood2plasma.2compartment), Rblood2plasma.2compartment := 1]
```

Merge the parameter information with the original concentration-time data.

```{r}
cvt[, Study_pooled :=
  paste(
    sort(unique(Study)),
    collapse = ", "
  ),
  by = .(DTXSID, Species)]

  #Melt to longer format. The effect will be as though we row-bound two versions
  #of DF together: one with the original single references, and one with
  #comma-separated combined References. There is one column "Study_type" that
  #contains either "Study" or "Study_pooled", and then one column "Study" that
  #contains the studyID or the comma-separated list of study IDs.
cvt[, Study_orig := Study]
  cvt1 <- melt(cvt,
              measure.vars = c("Study", "Study_pooled"),
              variable.name = "Study_type",
              value.name = "Studies.Analyzed")

  #There will be duplicated rows for single-study analyses, because
  #"Study_pooled" (the comma-separated list of study IDs) is the same as "Study"
  #(the individual study ID) for these analyses. Remove the duplicated rows.
  cvt1 <- unique(cvt1[, .(DTXSID, Compound, Species, Studies.Analyzed, Reference,
                        Route, Dose, iv, Media, Study_orig, Time,
                        Value, LOQ, Value_SD, N_Subjects,
                        Conc, Conc_Dose, Detect, Value_Dose,
                        Value_SD_Dose, LOQ_Dose)])

  cvt_fit <- pk_params[cvt1, 
                    on = c("DTXSID" = "DTXSID",
                            "Species" = "Species",
                            "Studies.Analyzed" = "Studies.Analyzed"),
                     allow.cartesian = TRUE]
```

Also prepare a table of study-specific sigmas.

```{r}
DT_1comp <- fread(paste0("../inst/ext/", "PK_fit_table_1comp_", timestamp, ".csv",sep=""))
DT_2comp <- fread(paste0("../inst/ext/", "PK_fit_table_2comp_", timestamp, ".csv",sep=""))
DT_flat <- fread(paste0("../inst/ext/", "PK_fit_table_flat_", timestamp, ".csv",sep=""))

idcols <- c( "Analysis_Type",
                                 "DTXSID",
                                "Species",
                                "References.Analyzed",
                               "Studies.Analyzed")

DT_all <- rbindlist(list(DT_1comp, DT_2comp, DT_flat),
                    use.names = TRUE,
                    fill = TRUE)
  DT_all <- DT_all[grepl(x = param_name,
                 pattern = "sigma"),
           .SD, .SDcols= c(idcols, "model", "param_name", "Fitted mean", "Fitted std dev")]
  DT_melt <- dcast(DT_all,
                   formula = Analysis_Type + DTXSID + Species + References.Analyzed + Studies.Analyzed + param_name ~ model,
                   value.var = "Fitted mean")
  setnames(DT_melt,
           unique(DT_all$model),
           paste0("sigma.", unique(DT_all$model)))
  DT_melt[param_name %in% "sigma", sigma_studyID := Studies.Analyzed]
  DT_melt[!(param_name %in% "sigma"), sigma_studyID := gsub(x = param_name,
                                                            pattern = "sigma_study_",
                                                            replacement = "",
                                                            fixed = TRUE)]
  DT_melt_sd <- dcast(DT_all,
                   formula = Analysis_Type + DTXSID + Species + References.Analyzed + Studies.Analyzed + param_name ~ model,
                   value.var = "Fitted std dev")
   setnames(DT_melt_sd,
           unique(DT_all$model),
           paste0("sigma_sd.", unique(DT_all$model)))
 DT_melt_sd[param_name %in% "sigma", sigma_studyID := Studies.Analyzed]
  DT_melt_sd[!(param_name %in% "sigma"), sigma_studyID := gsub(x = param_name,
                                                            pattern = "sigma_study_",
                                                            replacement = "",
                                                            fixed = TRUE)]

  DT_melt_all <- merge(DT_melt, DT_melt_sd, by = intersect(names(DT_melt),
                                                           names(DT_melt_sd)))

```

Get SD of data for each study

```{r}
cvt_sd <- cvt_fit[, .(sd_Value_Dose = sd(Value_Dose, na.rm = TRUE)),
        by = Study_orig]
```

Merge SD of data with fitted SDs and compare

```{r}
sd_table <- cvt_sd[DT_melt_all,
      on = c("Study_orig" = "sigma_studyID")]
```

Plot concentration vs time with the best-fit params by data subset (DTXSID and species).

Indicate the following:

- One line for each model
- Which is the winning model
- Dose group
- Blood vs. plasma
- Below LOQ vs. above LOQ (plot below-LOQ points at LOQ)
- IV vs. oral administration

Would also like to plot uncertainty in predictions based on uncertainty in parameter estimates.

Plot versions:
- Winning model only
- One model at a time
- All three models on same plot

Plot all models for the joint analysis.
```{r}
cvt_fit[,
        plot_fit(DTXSID = unique(DTXSID),
                 Species_in = unique(Species),
                 DFsub = .SD,
                 model_in = "all",
                 Analysis_Type_in = "Joint",
                 plot_dose_norm = TRUE,
                 split_dose = FALSE,
                 n_interp_time = 10,
                     limit_y_axis = FALSE,
                     log10_scale_y = FALSE,
                     plot_errbars = TRUE,
                     log10_scale_time = FALSE,
                 file_suffix = "errbars_naturaly_dose_norm",
                     file_path = "../inst/ext/plots/Joint Analysis/model_all/errbars_naturaly/dose_norm",
                 verbose = TRUE),
        by = .(DTXSID,
               Species),
        .SDcols = names(cvt_fit)]

```
Do the same, but faceting by dose.

```{r}
cvt_fit[,
        plot_fit(DTXSID = unique(DTXSID),
                 Species_in = unique(Species),
                 DFsub = .SD,
                 model_in = "all",
                 Analysis_Type_in = "Joint",
                 plot_dose_norm = FALSE,
                 split_dose = TRUE,
                 n_interp_time = 10,
                     limit_y_axis = FALSE,
                     log10_scale_y = FALSE,
                     plot_errbars = TRUE,
                     log10_scale_time = FALSE,
                 file_suffix = "errbars_naturaly_dose_norm",
                     file_path = "../inst/ext/plots/Joint Analysis/model_all/errbars_naturaly/dose_split",
                 verbose = TRUE),
        by = .(DTXSID,
               Species),
        .SDcols = names(cvt_fit)]

```

Do non-dose-faceted, but log scale.

```{r}
cvt_fit[,
        plot_fit(DTXSID = unique(DTXSID),
                 Species_in = unique(Species),
                 DFsub = .SD,
                 model_in = "all",
                 Analysis_Type_in = "Joint",
                 plot_dose_norm = TRUE,
                 split_dose = FALSE,
                 n_interp_time = 10,
                     limit_y_axis = FALSE,
                     log10_scale_y = TRUE,
                     plot_errbars = FALSE,
                     log10_scale_time = FALSE,
                 file_suffix = "errbars_naturaly_dose_norm",
                     file_path = "../inst/ext/plots/Joint Analysis/model_all/noerrbars_logy/dose_norm",
                 verbose = TRUE),
        by = .(DTXSID,
               Species),
        .SDcols = names(cvt_fit)]

```

Do dose-faceted, but log scale.

```{r}
cvt_fit[,
        plot_fit(DTXSID = unique(DTXSID),
                 Species_in = unique(Species),
                 DFsub = .SD,
                 model_in = "all",
                 Analysis_Type_in = "Joint",
                 plot_dose_norm = FALSE,
                 split_dose = TRUE,
                 n_interp_time = 10,
                     limit_y_axis = FALSE,
                     log10_scale_y = TRUE,
                     plot_errbars = FALSE,
                     log10_scale_time = FALSE,
                 file_suffix = "errbars_naturaly_dose_norm",
                     file_path = "../inst/ext/plots/Joint Analysis/model_all/noerrbars_logy/dose_split",
                 verbose = TRUE),
        by = .(DTXSID,
               Species),
        .SDcols = names(cvt_fit)]

```

```{r}

#plot all models
tmp <- mapply(plot_fit,
              DTXSID_in = plot_table$DTXSID,
                     Species_in = plot_table$Species,
              MoreArgs = list(
                     model_in = "all",
                     Analysis_Type_in = "Joint",
                     DF = cvt,
                     pk_fit = pk_fit,
                     n = 101,
                     limit_y_axis = FALSE,
                     log10_scale_y = TRUE,
                     plot_errbars = FALSE,
                     log10_scale_time = FALSE,
                     file_path = "../inst/ext/plots/Joint Analysis/model_all/noerrbars_logy")
                     )

```

Problem: DTXSID0022985_rat_Joint_all does not plot any model lines. Why?

```{r}
pk_fit[DTXSID %in% "DTXSID0022985"]
```

Apparently there were no detected data points?

```{r}
cvt[DTXSID %in% "DTXSID0022985", .(Reference, Value, LOQ)]
```

Hmm. Something is very weird here. I definitely see some detected points. They are all the same reference, so it isn't that a single-reference fit did not work but a joint fit did work.

Plot only the winning model

```{r}
#plot all models
tmp <- mapply(plot_fit,
              DTXSID_in = plot_table$DTXSID,
                     Species_in = plot_table$Species,
              MoreArgs = list(
                Analysis_Type_in = c("Joint", "Separate"),
                     model_in = "winning",
                     fitdata = DF,
                     pk_fit = pk_fit,
                     file_path = "inst/ext/plots/analysis_all_model_winning/",
                log10_scale_time = TRUE)
                     )
```