### 2023-01-20 Working Rmds/summary_loglike_lognormal.Rmd

---
title: "Log-likelihood for summary data"
author: "Caroline Ring"
date: "2022-12-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data

Imagine we have $N$ observations of a dependent (response) variable, $\boldsymbol{z_1}, \boldsymbol{z_2}, \ldots, \boldsymbol{z_i), \ldots, \boldsymbol{z_N}$. Each of these corresponds to one or more independent predictor variables, $\boldsymbol{x_1}, \ldots, \boldsymbol{x_i}, \ldots, \boldsymbol{x_N}$.

The $i^{th}$ observation is a summary composed of three values: the sample mean, sample standard deviation, and count of a sample of {M_i} underlying individual-subject observations.

$$ \boldsymbol{z_i} =  \left(
\begin{array}{c}
\bar{y}_i  \\
s_i \\
M_i\\
\end{array} 
\right) $$

where

$$ \boldsymbol{y_i} = \left(
\begin{array}{c}
y_{i1} \\
\cdots \\
y_{ij}\\
\cdots \\
y_{iM_i} \\
\end{array} 
\right) $$

For example, this might be a toxicokinetic study where, say, 36 rats were divided into 6 groups, and each group was dosed at a different level, resulting in 6 dose groups. Blood was drawn from the animals in each dose group at a series of time points, and the concentration of chemical X was analyzed. For each dose group and each time point, the mean and standard deviation of concentrations was reported. In this case, group $i$ represents one dose and one time point.

# Definitions of sample mean and sample variance

The sample mean and sample variance (or SD) were computed from the original individual-level data using the following definitions.

## Sample mean

$$ \bar{y}_i = \frac{1}{M_i} \sum_{j=1}^{M_i} y_{ij} $$
## Sample variance

$$ s_i^2 = \frac{1}{M_i - 1} \sum_{j=1}^{M_i} \left( y_{ij} - \bar{y}_i \right) ^2$$

# Normal log-likelihood

Assume we have a model that predicts a value $\mu_i$ for the $i^{th}$ observation, given the independent (predictor) variables $\boldsymbol{x_i}$ and some vector of parameters $\boldsymbol{\theta}$.

$$ \mu_i = f(\boldsymbol{x_i}; \boldsymbol{\theta}) $$

In the example of a TK study, $f$ might be a TK model -- e.g. a 1-compartment or 2-compartment model.

Our goal is to get the maximum-likelihood estimates of the parameters $\boldsymbol{\theta}$ -- in other words, the parameter values that make it most likely that the model would produce the data actually observed.

To do that we need to formulate the likelihood of the data given the parameters.

In this case, we assume a zero-mean normal distribution of the residual errors. This is equivalent to assuming the observed values are normally distributed around a mean given by the predicted values.

Write this likelihood function for individual $j$ in group $i$. This is just the normal probability density function.

$$ L_{ij} = \frac{1}{\sigma \sqrt{2 \pi}} \exp \left[ \frac{-1}{2} \left( \frac{\log y_{ij} -  \mu_i}{\sigma} \right) ^2 \right] $$

## Log-likelihood for individual subjects in group i

It's usually more convenient to work with the log-likelihood, so log-transform it.

$$LL_{ij} = \log \left( \frac{1}{\sigma \sqrt{2 \pi}} \right) + \frac{-1}{2} \left( \frac{\log y_{ij} -  \mu_i}{\sigma} \right) ^2 $$
## Total log-likelihood of group i

The joint log-likelihood over all subjects in group $i$ is just the sum of individual log-likelihoods.

$$ LL_i = \sum_{j=1}^{M_i} LL_{ij} =  \sum_{j=1}^{M_i} \left[ \log \left( \frac{1}{\sigma \sqrt{2 \pi}} \right) + \frac{-1}{2} \left( \frac{\log y_{ij} - \mu_i}{\sigma} \right) ^2 \right]$$
Breaking apart the two terms inside the summation (square brackets), and expanding the square in the second term, this becomes

$$ LL_i = M_i \log \left(\frac{1}{\sigma \sqrt{2 \pi}} \right) + \frac{-1}{2 \sigma^2} \left[\sum_{j=1}^{M_i} \left( \left( \log y_{ij} \right) ^2 - 2  \mu_i \log y_{ij} + \mu_i^2  \right) \right] $$
Further splitting apart the terms inside the summation (square brackets), this becomes

$$ LL_i = M_i \log \left(\frac{1}{\sigma \sqrt{2 \pi}} \right) + \\
\frac{-1}{2 \sigma^2} \sum_{j=1}^{M_i} \left( \log y_{ij} \right)^2 + \\
\frac{2 \mu_i}{2 \sigma^2} \sum_{j=1}^{M_i} \log y_{ij} + \\
\frac{-\mu_i^2 M_i}{2 \sigma^2} $$

Here is where things get complicated with a log-normal distribution. With a normal distribution, we could substitute in the definition of the sample mean for $\sum y_{ij}$. With a log-normal distribution, we instead have $\sum \log y_{ij}$, which does not correspond to the definition of the sample mean. At best we could do the log of the sample mean -- but that doesn't give us what we are looking for.

$$\log \bar{y}_i = \log \sum_{j=1}^{M_i} y_{ij} \neq \sum_{j=1}^{M_i} \log y_{ij}$$
If log-scale population mean is $m_i$ and log-scale standard deviation is $s_i$, and sample mean is $\bar{y}_i$ and sample standard deviation is $S_i$, then

$$ m_i = \log \frac{\bar{y}_i^2}{\sqrt{S_i^2 + \bar{y}_i^2}} = \frac{1}{M_{i}} \sum_{j=1}^{M_i} \log y_{ij}$$
and

$$ s_i = \sqrt{ \log \left( 1 + \frac{S_i^2}{\bar{y}_i^2} \right)} = \sqrt{ \frac{1}{M_i - 1} \sum_{i=1}^{M_i} (\log y_{ij} - m_i)^2}$$

Substitute in the definition of the log-scale mean for $\sum_{j=1}^{M_i} \log y_{ij}$

$$ LL_i = M_i \log \left(\frac{1}{\sigma \sqrt{2 \pi}} \right) + \\
\frac{-1}{2 \sigma^2} \sum_{j=1}^{M_i} \log y_{ij}^2 + \\
\frac{2 \mu_i}{2 \sigma^2} M_i \log \frac{\bar{y}_i^2}{\sqrt{S_i^2 + \bar{y}_i^2}} + \\
\frac{-\mu_i^2 M_i}{2 \sigma^2} $$

Using the definitions of the log-scale sample variance and log-scale sample mean, it can be shown that

$$ \sum_{j=1}^{M_i}  \left( \log y_{ij} \right)^2 = \left(M_i - 1 \right)  s_i^2 + M_i m_i^2 = (M_i -1) \log \left( 1 + \frac{S_i^2}{\bar{y}_i^2} \right) + M_i  \left(\log \frac{\bar{y}_i^2}{\sqrt{S_i^2 + \bar{y}_i^2}}\right)^2 $$
Substitute this back into the log-likelihood:

$$ LL_i = M_i \log \left(\frac{1}{\sigma \sqrt{2 \pi}} \right) + \\
\frac{-1}{2 \sigma^2} \left( (M_i -1) \log \left( 1 + \frac{S_i^2}{\bar{y}_i^2} \right) + M_i  \left(\log \frac{\bar{y}_i^2}{\sqrt{S_i^2 + \bar{y}_i^2}}\right)^2 \right) + \\
\frac{2  \mu_i}{2 \sigma^2} M_i \log \frac{\bar{y}_i^2}{\sqrt{S_i^2 + \bar{y}_i^2}} + \\
\frac{-\mu_i^2 M_i}{2 \sigma^2} $$

This looks horrible, but I think it is correct.

# Appendix

How to use the definition of the sample variance and smaple mean to show that

$$ s_i^2 = \left( \frac{1}{M_i - 1} \sum_{j=1}^{M_i}  y_{ij}^2 \right) - \frac{M_i}{M_i - 1} \bar{y}_i^2 $$

Start with the definition of the sample variance

$$ s_i^2 = \frac{1}{M_i - 1} \sum_{j=1}^{M_i} \left( y_{ij} - \bar{y}_i \right) ^2$$
Expand the square

$$ s_i^2 = \frac{1}{M_i - 1} \sum_{j=1}^{M_i} \left( y_{ij}^2 - 2 y_{ij} \bar{y}_i + \bar{y}_i^2 \right) $$

Split apart the terms in the summation

$$ s_i^2 = \frac{1}{M_i - 1} \left( \sum_{j=1}^{M_i} y_{ij}^2  - 2 \bar{y}_i  \sum_{j=1}^{M_i} y_{ij} + M_i \bar{y}_i^2 \right)$$
In the second term, substitute from the definition of the sample mean, $\sum_{j=1}^{M_i} y_{ij} = M_i \bar{y}_i $

$$ s_i^2 = \frac{1}{M_i - 1} \left( \left[ \sum_{j=1}^{M_i} y_{ij}^2 \right]  - 2 M_i \bar{y}_i^2  + M_i \bar{y}_i^2 \right) $$
This then simplifies to

$$ s_i^2 = \frac{1}{M_i - 1} \left( \left[ \sum_{j=1}^{M_i} y_{ij}^2 \right]  -  M_i \bar{y}_i^2 \right) $$
We canthen solve for $\sum_{j=1}^{M_i} y_{ij}^2$

$$ \sum_{j=1}^{M_i} y_{ij}^2 = (M_i - 1) s_i^2 + M_i \bar{y}_i^2 $$