### 2023-01-23 Working Rmds/MainAnalysis.Rmd
Added several functions to evaluate fits.
preprocess_data() now retains CvTdb series and study IDs.

Updated documentation for several functions.

---
title: "invivoPKfit analysis of CvTdb data"
author: "Christopher Cook, John Wambaugh, Caroline Ring"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document: default
  word_document: default
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(results = TRUE)
knitr::opts_chunk$set(message = FALSE)
```

```{r, include = FALSE}
library(tidyverse)
library(data.table)
library(readxl)
devtools::load_all("../invivopkfit")
```

# Read in CvTdb data

Read in CvTdb data that was pulled and normalized in `pulling_oral_iv_data_cvtdb.Rmd`.

```{r, load_cvt, eval = TRUE}
cvt_data <- fread("../inst/ext/cvt_data_2023_01_05.csv")
```

# Setup for data processing and model fitting

## Timestamp
Timestamp the file name with the current date and time (at the beginning of the runs) in format YYYY_MM_DD_hh_mm.

```{r}
timestamp <- format(Sys.time(), "%Y_%m_%d")
```

## Settings for pre-processing data

List of new_name = old_name to supply for preprocessing data

```{r}
names_list <- list(
  "Compound_Dosed" = "studies.test_substance_name_original",
  "DTXSID_Dosed" = "chemicals_dosed.dsstox_substance_id",
  "CAS_Dosed" = "chemicals_dosed.dsstox_casrn",
  "Compound_Analyzed" = "series.analyte_name_original",
  "DTXSID_Analyzed" = "chemicals_analyzed.dsstox_substance_id",
  "CAS_Analyzed" = "chemicals_analyzed.dsstox_casrn",
  "Reference" = "documents_reference.id",
  "Extraction" = "documents_extraction.id",
  "Species" = "subjects.species",
  "Weight" ="subjects.weight_kg",
  "Weight.Units" = NULL,
  "Dose" = "studies.dose_level_normalized_corrected", #note that dose levels were corrected in pulling_oral_iv_data_cvtdb.Rmd
  "Dose.Units" = NULL,
  "Time" = "conc_time_values.time_hr",
  "Time.Units" = NULL,
  "Media" = "series.conc_medium_normalized",
  "Value" = "conc_time_values.conc", #already normalized to mg/L units in CvTdb
  "Value.Units" = NULL,
  "Route" = "studies.administration_route_normalized",
  "LOQ" = "series.loq_normalized", #note that LOQs were normalized to mg/L units in pulling_oral_iv_data_cvtdb.Rmd
  "Subject" = "subjects.id",
  "N_Subjects" = "series.n_subjects_in_series",
  "Value_SD" = "conc_time_values.conc_sd_normalized", #note that SDs were normalized to mg/L units in pulling_oral_iv_data_cvtdb.Rmd
  "Study_ID" = "studies.id",
  "Series_ID" = "series.id"
  )
```

List of default values for new columns added in preprocessing data

```{r}
defaults_list <- list(
                      "Weight.Units" = "kg",
                      "Dose.Units" = "mg/kg",
                      "Time.Units" = "hours",
                      "Value.Units" = "mg/L")
```

# Preprocess data

```{r}
cvt <- preprocess_data(data.set = cvt_data,
                       names_list =names_list,
                       defaults_list =   defaults_list,
                       ratio_conc_to_dose = 1,
                       calc_loq_factor = 0.45,
                       routes_keep = c("po", "iv"),
                       media_keep = c("blood", "plasma"),
                       impute_loq = TRUE,
                       impute_sd = TRUE,
                       study_def = c("DTXSID", "Species", "Reference", "Route", "Media"),
                       suppress.messages = TRUE)

#save pre-processed data
write.csv(cvt,
          paste0("../inst/ext/cvt_preprocessed_",
                 timestamp,
          ".csv"),
          row.names = FALSE)
```

# Fit TK models to data

```{r}
for(this_rescale_time in c(FALSE, TRUE)){
  for(this_fit_conc_dose in c(FALSE, TRUE)){
    for(this_fit_log_conc in c(FALSE, TRUE)){
      if(!(this_rescale_time %in% FALSE &
           this_fit_conc_dose %in% FALSE &
           this_fit_log_conc %in% FALSE)){ #already did the all FALSE case
      for (this_model in c("flat",
                           "1compartment",
                           "2compartment")){
        print(paste0("model = ", this_model, "\n",
                     "fit_log_conc = ", this_fit_log_conc, "\n",
                     "fit_conc_dose = ", this_fit_conc_dose, "\n",
                     "rescale_time = ", this_rescale_time))
        system.time(
          PK.fit.table <- fit_all(
            data.set = cvt,
            model = this_model,
            modelfun = "analytic",
            preprocess = FALSE,
            fit_conc_dose = this_fit_conc_dose,
            fit_log_conc = this_fit_log_conc,
            rescale_time = this_rescale_time,
            optimx_args = list(
              "method" = "bobyqa",
              "itnmax" = 1e6,
              "control" = list("kkt" = FALSE)
            ),
            suppress.messages = FALSE
          )
        )

        write.csv(PK.fit.table,
                  paste("../inst/ext/PK_fit_table_",
                        this_model,
                        "_fit_log_conc_",
                        this_fit_log_conc, "_",
                        "_fit_conc_dose_",
                        this_fit_conc_dose, "_",
                        "_rescale_time_",
                        this_rescale_time, "_",
                        timestamp,
                        ".csv",
                        sep=""), row.names = FALSE)

        post_table <- postprocess_data(PK_fit = PK.fit.table,
                                       model= this_model)

        rm(PK,fit.table)

        write.csv(post_table,
                  paste("../inst/ext/PK_fit_postprocess_",
                        this_model,
                        "_fit_log_conc_",
                        this_fit_log_conc, "_",
                        "_fit_conc_dose_",
                        this_fit_conc_dose, "_",
                        "_rescale_time_",
                        this_rescale_time, "_",
                        timestamp,
                        ".csv",
                        sep=""), row.names = FALSE)

        rm(post_table)
      } #end loop over models
      } #end if not all FALSE
    } #end loop over this_fit_log_conc
  } #end loop over this_fit_conc_dose
} #end loop over this_rescale_time
```