### 2023-01-23 Working Rmds/summary_loglike_lognormal.Rmd
Correctly log-transform predictions if log = TRUE specified when computing RMSE, R-squared, etc.
Update documentation.

---
title: "Log-likelihood for summary data"
author: "Caroline Ring"
date: "2022-12-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data

Imagine we have $N$ observations of a dependent (response) variable, $\boldsymbol{z_1}, \boldsymbol{z_2}, \ldots, \boldsymbol{z_i), \ldots, \boldsymbol{z_N}$. Each of these corresponds to one or more independent predictor variables, $\boldsymbol{x_1}, \ldots, \boldsymbol{x_i}, \ldots, \boldsymbol{x_N}$.

The $i^{th}$ observation is a summary composed of three values: the sample mean, sample standard deviation, and count of a sample of {M_i} underlying individual-subject observations.

$$ \boldsymbol{z_i} =  \left(
\begin{array}{c}
\bar{y}_i  \\
s_i \\
M_i\\
\end{array} 
\right) $$

where

$$ \boldsymbol{y_i} = \left(
\begin{array}{c}
y_{i1} \\
\cdots \\
y_{ij}\\
\cdots \\
y_{iM_i} \\
\end{array} 
\right) $$

For example, this might be a toxicokinetic study where, say, 36 rats were divided into 6 groups, and each group was dosed at a different level, resulting in 6 dose groups. Blood was drawn from the animals in each dose group at a series of time points, and the concentration of chemical X was analyzed. For each dose group and each time point, the mean and standard deviation of concentrations was reported. In this case, group $i$ represents one dose and one time point.

# Definitions of sample mean and sample variance

The sample mean and sample variance (or SD) were computed from the original individual-level data using the following definitions.

## Sample mean

$$ \bar{y}_i = \frac{1}{M_i} \sum_{j=1}^{M_i} y_{ij} $$
## Sample variance

$$ s_i^2 = \frac{1}{M_i - 1} \sum_{j=1}^{M_i} \left( y_{ij} - \bar{y}_i \right) ^2$$

# Normal log-likelihood

Assume we have a model that predicts a value $\mu_i$ for the $i^{th}$ observation, given the independent (predictor) variables $\boldsymbol{x_i}$ and some vector of parameters $\boldsymbol{\theta}$.

$$ \mu_i = f(\boldsymbol{x_i}; \boldsymbol{\theta}) $$

In the example of a TK study, $f$ might be a TK model -- e.g. a 1-compartment or 2-compartment model.

Our goal is to get the maximum-likelihood estimates of the parameters $\boldsymbol{\theta}$ -- in other words, the parameter values that make it most likely that the model would produce the data actually observed.

To do that we need to formulate the likelihood of the data given the parameters.

In this case, we assume a zero-mean normal distribution of the residual errors. This is equivalent to assuming the observed values are normally distributed around a mean given by the predicted values.

Write this likelihood function for individual $j$ in group $i$. This is just the normal probability density function.

$$ L_{ij} = \frac{1}{\sigma \sqrt{2 \pi}} \exp \left[ \frac{-1}{2} \left( \frac{\log y_{ij} -  \mu_i}{\sigma} \right) ^2 \right] $$

## Log-likelihood for individual subjects in group i

It's usually more convenient to work with the log-likelihood, so log-transform it.

$$LL_{ij} = \log \left( \frac{1}{\sigma \sqrt{2 \pi}} \right) + \frac{-1}{2} \left( \frac{\log y_{ij} -  \mu_i}{\sigma} \right) ^2 $$
## Total log-likelihood of group i

The joint log-likelihood over all subjects in group $i$ is just the sum of individual log-likelihoods.

$$ LL_i = \sum_{j=1}^{M_i} LL_{ij} =  \sum_{j=1}^{M_i} \left[ \log \left( \frac{1}{\sigma \sqrt{2 \pi}} \right) + \frac{-1}{2} \left( \frac{\log y_{ij} - \mu_i}{\sigma} \right) ^2 \right]$$
Breaking apart the two terms inside the summation (square brackets), and expanding the square in the second term, this becomes

$$ LL_i = M_i \log \left(\frac{1}{\sigma \sqrt{2 \pi}} \right) + \frac{-1}{2 \sigma^2} \left[\sum_{j=1}^{M_i} \left( \left( \log y_{ij} \right) ^2 - 2  \mu_i \log y_{ij} + \mu_i^2  \right) \right] $$
Further splitting apart the terms inside the summation (square brackets), this becomes

$$ LL_i = M_i \log \left(\frac{1}{\sigma \sqrt{2 \pi}} \right) + \\
\frac{-1}{2 \sigma^2} \sum_{j=1}^{M_i} \left( \log y_{ij} \right)^2 + \\
\frac{2 \mu_i}{2 \sigma^2} \sum_{j=1}^{M_i} \log y_{ij} + \\
\frac{-\mu_i^2 M_i}{2 \sigma^2} $$

Here is where things get complicated with a log-normal distribution. With a normal distribution, we could substitute in the definition of the sample mean for $\sum y_{ij}$. With a log-normal distribution, we instead have $\sum \log y_{ij}$, which does not correspond to the definition of the sample mean. At best we could do the log of the sample mean -- but that doesn't give us what we are looking for.

$$\log \bar{y}_i = \log \sum_{j=1}^{M_i} y_{ij} \neq \sum_{j=1}^{M_i} \log y_{ij}$$
If log-scale population mean is $m_i$ and log-scale standard deviation is $s_i$, and sample mean is $\bar{y}_i$ and sample standard deviation is $S_i$, then

$$ m_i = \log \frac{\bar{y}_i^2}{\sqrt{S_i^2 + \bar{y}_i^2}} = \frac{1}{M_{i}} \sum_{j=1}^{M_i} \log y_{ij}$$
and

$$ s_i = \sqrt{ \log \left( 1 + \frac{S_i^2}{\bar{y}_i^2} \right)} = \sqrt{ \frac{1}{M_i - 1} \sum_{i=1}^{M_i} (\log y_{ij} - m_i)^2}$$

Substitute in the definition of the log-scale mean for $\sum_{j=1}^{M_i} \log y_{ij}$

$$ LL_i = M_i \log \left(\frac{1}{\sigma \sqrt{2 \pi}} \right) + \\
\frac{-1}{2 \sigma^2} \sum_{j=1}^{M_i} \log y_{ij}^2 + \\
\frac{2 \mu_i}{2 \sigma^2} M_i \log \frac{\bar{y}_i^2}{\sqrt{S_i^2 + \bar{y}_i^2}} + \\
\frac{-\mu_i^2 M_i}{2 \sigma^2} $$

Using the definitions of the log-scale sample variance and log-scale sample mean, it can be shown that

$$ \sum_{j=1}^{M_i}  \left( \log y_{ij} \right)^2 = \left(M_i - 1 \right)  s_i^2 + M_i m_i^2 = (M_i -1) \log \left( 1 + \frac{S_i^2}{\bar{y}_i^2} \right) + M_i  \left(\log \frac{\bar{y}_i^2}{\sqrt{S_i^2 + \bar{y}_i^2}}\right)^2 $$
Substitute this back into the log-likelihood:

$$ LL_i = M_i \log \left(\frac{1}{\sigma \sqrt{2 \pi}} \right) + \\
\frac{-1}{2 \sigma^2} \left( (M_i -1) \log \left( 1 + \frac{S_i^2}{\bar{y}_i^2} \right) + M_i  \left(\log \frac{\bar{y}_i^2}{\sqrt{S_i^2 + \bar{y}_i^2}}\right)^2 \right) + \\
\frac{2  \mu_i}{2 \sigma^2} M_i \log \frac{\bar{y}_i^2}{\sqrt{S_i^2 + \bar{y}_i^2}} + \\
\frac{-\mu_i^2 M_i}{2 \sigma^2} $$

This looks horrible, but I think it is correct.

# Appendix: Error in log-scale moments from arithmetic moments, a simulation study

How much error is there in estimating the log-scale moments (mean and SD of log-transformed observations) from the arithmetic moments (mean and SD of natural-scale observations)?

We can get an idea using a simulation study.

Take a log-normal distribution with known log-scale mean and SD.

Draw some number of samples from this distribution.

Calculate the sample mean and SD of the log-scaled samples, and of the natural-scale samples.

Repeat these steps some large number of times.

Compare the log-scale sample means and SDs to the natural-scale sample means and SDs.

Write a function to draw the samples and calculate the sample moments:

```{r}
get_sample_moments <- function(meanlog, sdlog, n_samp){
  #samples
  y <- rlnorm(n = n_samp, meanlog = meanlog, sdlog = sdlog)
  #log-transformed samples
  ylog <- log(y)
  return(c("sample_mean" = mean(y),
           "sample_sd" = sd(y),
           "sample_meanlog" = mean(ylog),
           "sample_sdlog" = sd(ylog),
           "sample_n" = n_samp))
}
```

Choose values for meanlog, sdlog, and n_samp
```{r}
meanlog <- 0
sdlog <- 1
n_samp <- 7
```

Set a seed for reproducibility, and generate the values

```{r}
set.seed(42)
n_rep <- 1000
samp <- replicate(n = n_rep,
          expr = get_sample_moments(meanlog = meanlog,
                                    sdlog = sdlog,
                                    n_samp = n_samp))
samp <- t(samp)
samp <- as.data.frame(samp)
```

Calculate the estimated log-scale sample moments from the arithmetic sample moments

```{r}
samp[c("est_sample_meanlog",
       "est_sample_sdlog")] <- convert_summary_to_log(sample_mean = samp$sample_mean,
                                                                   sample_SD = samp$sample_sd)
```

Now, plot the estimated log-scale moments vs. the "actual" log-scale moments.

Log-scale sample mean. The "true" population log-scale mean (0.0) is indicated by vertical and horizontal dashed lines.

```{r}
ggplot(data = samp) +
  geom_point(aes(x = sample_meanlog,
                 y = est_sample_meanlog),
             color = "gray50") +
  geom_abline(intercept = 0, slope = 1, color = "black") +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_vline(xintercept = 0, linetype = 2) +
  xlab("Actual log-scale sample mean") +
  ylab("Estimated log-scale sample mean") +
  ggtitle("Estimated vs. actual log-scale sample mean, n = 7") +
  theme_bw()
```

The estimated log-scale sample mean usually overestimates the "true" log-scale sample mean (most points are above the identity line).

Log-scale sample SD. The "true" population log-scale SD (1.0) is indicated by vertical and horizontal dashed lines.

```{r}
ggplot(data = samp) +
  geom_point(aes(x = sample_sdlog,
                 y = est_sample_sdlog),
             color = "gray50") +
  geom_abline(intercept = 0, slope = 1, color = "black") +
    geom_hline(yintercept = 1, linetype = 2) +
  geom_vline(xintercept = 1, linetype = 2) +
  xlab("Actual log-scale sample SD") +
  ylab("Estimated log-scale sample SD") +
  ggtitle("Estimated vs. actual log-scale sample SD, n = 7") +
  theme_bw()
```

The estimated log-scale sample SD usually underestimates the actual log-scale sample SD (most points are below the identity line).

```{r}
summary(lm(est_sample_sdlog ~ sample_sdlog, data = samp))
```