### 2023-02-01 Working Rmds/MainAnalysis.Rmd
Adding upper bounds because L-BFGS-B sometimes fails without them (returning NA for optimized param values, but not failing with an error). Adding more detail in output data.frame (including number of points in "empirical" absorption and elimination phases).

---
title: "invivoPKfit analysis of CvTdb data"
author: "Christopher Cook, John Wambaugh, Caroline Ring"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document: default
  word_document: default
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(results = TRUE)
knitr::opts_chunk$set(message = FALSE)
```

```{r, include = FALSE}
library(tidyverse)
library(data.table)
library(readxl)
devtools::load_all("../invivopkfit")
```

# Read in CvTdb data

Read in CvTdb data that was pulled and normalized in `pulling_oral_iv_data_cvtdb.Rmd`.

```{r, load_cvt, eval = TRUE}
cvt_data <- fread("../inst/ext/cvt_data_2023_01_05.csv")
```

# Setup for data processing and model fitting

## Timestamp
Timestamp the file name with the current date and time (at the beginning of the runs) in format YYYY_MM_DD_hh_mm.

```{r}
timestamp <- format(Sys.time(), "%Y_%m_%d")
```

## Settings for pre-processing data

List of new_name = old_name to supply for preprocessing data

```{r}
names_list <- list(
  "Compound_Dosed" = "studies.test_substance_name_original",
  "DTXSID_Dosed" = "chemicals_dosed.dsstox_substance_id",
  "CAS_Dosed" = "chemicals_dosed.dsstox_casrn",
  "Compound_Analyzed" = "series.analyte_name_original",
  "DTXSID_Analyzed" = "chemicals_analyzed.dsstox_substance_id",
  "CAS_Analyzed" = "chemicals_analyzed.dsstox_casrn",
  "Reference" = "documents_reference.id",
  "Extraction" = "documents_extraction.id",
  "Species" = "subjects.species",
  "Weight" ="subjects.weight_kg",
  "Weight.Units" = NULL,
  "Dose" = "studies.dose_level_normalized_corrected", #note that dose levels were corrected in pulling_oral_iv_data_cvtdb.Rmd
  "Dose.Units" = NULL,
  "Time" = "conc_time_values.time_hr",
  "Time.Units" = NULL,
  "Media" = "series.conc_medium_normalized",
  "Value" = "conc_time_values.conc", #already normalized to mg/L units in CvTdb
  "Value.Units" = NULL,
  "Route" = "studies.administration_route_normalized",
  "LOQ" = "series.loq_normalized", #note that LOQs were normalized to mg/L units in pulling_oral_iv_data_cvtdb.Rmd
  "Subject" = "subjects.id",
  "N_Subjects" = "series.n_subjects_in_series",
  "Value_SD" = "conc_time_values.conc_sd_normalized", #note that SDs were normalized to mg/L units in pulling_oral_iv_data_cvtdb.Rmd
  "Study_ID" = "studies.id",
  "Series_ID" = "series.id"
  )
```

List of default values for new columns added in preprocessing data

```{r}
defaults_list <- list(
                      "Weight.Units" = "kg",
                      "Dose.Units" = "mg/kg",
                      "Time.Units" = "hours",
                      "Value.Units" = "mg/L")
```

# Preprocess data

```{r}
cvt <- preprocess_data(data.set = cvt_data,
                       names_list =names_list,
                       defaults_list =   defaults_list,
                       ratio_conc_to_dose = 1,
                       calc_loq_factor = 0.45,
                       routes_keep = c("po", "iv"),
                       media_keep = c("blood", "plasma"),
                       impute_loq = TRUE,
                       impute_sd = TRUE,
                       study_def = c("DTXSID", "Species", "Reference"),
                       suppress.messages = TRUE)

#save pre-processed data
write.csv(cvt,
          paste0("../inst/ext/cvt_preprocessed_",
                 timestamp,
          ".csv"),
          row.names = FALSE)
```

# Fit TK models to data

```{r}
for(this_method in c("bobyqa", "L-BFGS-B")){
  for(this_rescale_time in c(FALSE, TRUE)){
    for(this_fit_conc_dose in c(FALSE, TRUE)){
      for(this_fit_log_conc in c(FALSE, TRUE)){
          for (this_model in c("flat",
                               "1compartment",
                               "2compartment")){
            cat(paste0("method = ", this_method, "\n",
                         "fit_log_conc = ", this_fit_log_conc, "\n",
                         "fit_conc_dose = ", this_fit_conc_dose, "\n",
                         "rescale_time = ", this_rescale_time, "\n",
                                     "model = ", this_model, "\n"))
            system.time(
              PK.fit.table <- fit_all(
                data.set = cvt,
                model = this_model,
                modelfun = "analytic",
                preprocess = FALSE,
                fit_conc_dose = this_fit_conc_dose,
                fit_log_conc = this_fit_log_conc,
                rescale_time = this_rescale_time,
                optimx_args = list(
                  method = this_method,
                  itnmax = 1e6,
                  control = list(kkt = FALSE,
                                 factr = 1e7)
                ),
                suppress.messages = FALSE
              )
            )

            write.csv(PK.fit.table,
                      paste0("../inst/ext/PK_fit_table_",
                            this_model, "_",
                            "fit_log_conc_",
                            this_fit_log_conc, "_",
                            "fit_conc_dose_",
                            this_fit_conc_dose, "_",
                            "rescale_time_",
                            this_rescale_time, "_",
                            "method_", this_method, "_",
                            timestamp,
                            ".csv"), row.names = FALSE)

          } #end loop over models
      } #end loop over this_fit_log_conc
    } #end loop over this_fit_conc_dose
  } #end loop over this_rescale_time
}#end loop over this_method
```