### 2023-09-25 data-raw/batchMode_trial_gpm.Rmd
Updating vignette to prepare for publication, included a TOC for figure's produced.

---
title: "PK object batch mode trial"
author: "Gilberto Padilla Mercado"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
set.seed(2023)
library(tidyverse, quietly = TRUE)
devtools::load_all() # Loads the invivopkfit package in feature/pkS3obj_facet branch
library(DBI)
tidyverse_conflicts()
```

# TOC

1. Pulling Data from CvTdb  
  - Minimal `pk` object  
2. Running all possible fitting options  
  - Evaluating fitting options  
  - Head-to-head comparison of pooled vs joint models  
3. Running dose-normalized joint fits  
4. Analysis Plots  
  - CvTdb replicate variation (Figure 5)  
  - CvTdb replicate variation over ADME-normalized time (Figure 5B)  
  - CvTdb concentration and final time-point distributions (Supplementary Figure 1)  
  - invivoPKfit fit option selection (Figure 6)  
  - invivoPKfit model performance (Supplementary Figure 2)  
  - Model performance vs Data variability (Figure 7A)  
  - Goodness of fit plots (Figure 7B)  
  - Fits that may be improved (Supplementary Figure 3)  
5. Lombardo Analysis  
  - Chemicals included  
  - Comparison of derived TK stats  (Figure 8)
6. Benchmarking invivoPKfit
  - Parallelization speeds up the process of fitting (Supplementary Figure 4)

## Pulling data from CvTdb
This is largely taken from a `Querying CvTdb` vignette from Taylor Wall.

```{r query_functions}

postgreSQL_connection <- function() {
  return(
    dbConnect(RPostgreSQL::PostgreSQL(),
              user = Sys.getenv("CVT_USER"),
              password = Sys.getenv("CVT_PASS"),
              host = Sys.getenv("CVT_HOST"),
              dbname = Sys.getenv("CVT_DB_NAME"))
  )
}

query_db <- function(query = NULL, schema) {
  if (is.null(query)) return(cat("\nMust provide a query to send"))
  con <- postgreSQL_connection()

  query_result <- tryCatch(
    return(dbGetQuery(con, query %>%
                        gsub("FROM ", paste0("FROM ", schema, "."), .) %>%
                        gsub("JOIN ", paste0("JOIN ", schema, "."), .))),
  error = function(cond){ 
    cat(paste0("\nError message: ", cond))
    return(NULL)
  },
  finally = { 
    dbDisconnect(con)
    }
    )
  return(query_result)
}

```

The following is the new SQL query, this will eventually be moved another vignette.

```{r SQL_queries}
# Sample query of CvTdb Documents SQL
# query <- "SELECT pmid, year, first_author, title FROM documents LIMIT 10"
# query <- "SELECT count(*) FROM studies"
# query <- "SELECT DISTINCT administration_route_normalized FROM studies"

# Need to make this in a specified ordered
# First all IDs (so I can check them quickly)
# Note that this uses an ALIAS for each table which is later specified in
# the FROM and LEFT JOIN statments
query <- paste0(
  "SELECT distinct ",
  "a.id, a.time_original,  a.time_hr, a.conc_original, a.conc_sd_original, a.conc, a.conc_sd, ",
  "b.id, b.analyte_name_original, b.analyte_dtxsid, b.analyte_casrn, ",
  "b.fk_analyzed_chemical_id, b.fk_test_chemical_id, ",
  "b.test_substance_dtxsid, b.analyte_name_secondary_original, ",
  "b.conc_medium_normalized, b.time_units_original, b.conc_units_original, ",
  "b.loq, b.loq_units, b.n_subjects_in_series, ",
  "c.id, c.administration_route_normalized, c.fk_dosed_chemical_id, c.dose_volume, c.dose_vehicle, ",
  "c.dose_duration, c.dose_duration_units, c.dose_frequency, c.fasting_period, ",
  "c.dose_level_normalized, c.dose_level_original, c.dose_level_original_units, ",
  "d.id, d.weight_kg, d.species, d.sex, d.age, d.age_units, d.age_category, ",
  "e.id, e.pmid, e.year, e.other_study_identifier, e.url, e.doi, e.extracted ",

  # Main Table
  "FROM conc_time_values a ",

  # Join with series table by series ID
  "LEFT JOIN series b ON b.id = a.fk_series_id ",

  # Join to studies table by study ID
  "LEFT JOIN studies c ON c.id = b.fk_study_id ",

  # Join to subjects table by subject ID
  "LEFT JOIN subjects d ON d.id = b.fk_subject_id ",

  # Join to documents table by extraction document ID
  "LEFT JOIN documents e ON e.id = c.fk_extraction_document_id ",

  # Filtering steps
  "WHERE b.conc_medium_normalized IN ('blood', 'plasma') AND ",
  "c.administration_route_normalized IN ('iv', 'oral')"
)

sql_query <- query_db(query = query,
         schema = "cvt")

names(sql_query)
# Need to change column names for this
{
  names(sql_query)[1] <- "conc_time_id"
  names(sql_query)[8] <- "series_id"
  names(sql_query)[22] <- "study_id"
  names(sql_query)[34] <- "subject_id"
  names(sql_query)[41] <- "document_id"
}

# Harmonize values for easier analysis
sql_query <- sql_query %>%
  mutate(
    species = tolower(species),
    sex = tolower(sex),
    loq = as.numeric(loq),
    n_subjects_in_series = as.numeric(n_subjects_in_series),
    dose_level_normalized = as.numeric(dose_level_normalized),
    dose_level_original = as.numeric(dose_level_original),
    dose_volume_units = str_extract(dose_volume,
                                    pattern = "(?<=[:space:]).+"),
    dose_volume = as.numeric(str_extract(dose_volume,
                                         pattern = "^(\\d|\\.)*")),
    dose_duration_units = str_extract(dose_duration,
                                      pattern = "(?<=[space]).+"),
    dose_duration = as.numeric(str_extract(dose_duration,
                                pattern = "^\\d"))
  )
sql_query <- sql_query %>%
  rename_with(~gsub(pattern = "fk_", replacement = "",
                         .x), .cols = contains("fk_"))

# Copying
cvt_df <- sql_query # Copy so I don't have to do query again to reset any errors

# Check other_study_identifier == S0916 so that it has dose_level_original_units == ug/mL
cvt_df %>% filter(other_study_identifier == "S0916") %>%
  pull(dose_level_original_units) %>% unique()
# There are still columns that are mg/kg... changing these to ug/mL
cvt_df <- cvt_df %>% mutate(
  dose_level_original_units = ifelse(
    other_study_identifier == "S0916",
    "ug/mL",
    dose_level_original_units
  )
)

# Need to convert the units to mg/kg
# In pulling_oral_iv_data_cvtdb.Rmd, it is noted that oral was 5mL/kg and iv 2mL/kg
# So I need to do ug/mL * mL/kg * 1mg/1000ug = mg/kg
cvt_df <- cvt_df %>%
  mutate(
    dose_level_corrected = ifelse(
      other_study_identifier %in% "S0916",
      ifelse(administration_route_normalized %in% "oral",
             dose_level_original * 5/1000,
             dose_level_original * 2/1000),
      dose_level_normalized),
    dose_level_corrected_units = "mg/kg",
    .after = dose_level_original_units)

# For this study, it was in ng/kg not mg/kg
cvt_df <- cvt_df %>%
  mutate(
    dose_level_corrected = ifelse(
      other_study_identifier %in% "S0624",
      dose_level_original * 1E-6,
      dose_level_normalized),
    .after = dose_level_original_units)

# The following measured C-14/H-3 but are indicators for the following Chemicals
# document_id = 65, DTXSID = DTXSID7031290, name = Isocyclemone E or beta-OTNE
# document_id = 61, DTXSID = DTXSID4022527, DTXSID3020209,
### name(s) = propylparaben, butylparaben, missing...methylparaben
# Propylparaben SERIES_ID = 110, Butylparaben SERIES_ID = 112
# For above, HPLC reviealed a single peak corresponding with p-hydroxybenzoic acid (PHBA, DTXSID3026647)
# document_id = 46, DTXSID = DTXSID2020139, name = benzo[a]pyrene
# document_id = 31, DTXSID = DTXSID2020139, name = benzo[a]pyrene

# However, now would be good time to eliminate any ones where they analyte != test chemical
# Some studies may have dosed with multiple chemicals... important to note
# cvt_df <- cvt_df %>% filter(analyzed_chemical_id == test_chemical_id)
# For now, I will change the chemical name and dtxsid
cvt_df <- cvt_df %>%
  mutate(analyte_dtxsid = ifelse(document_id %in% 65, "DTXSID7031290",
                                 analyte_dtxsid),
         analyte_name_original = ifelse(document_id %in% 65, "beta-OTNE",
                                        analyte_name_original)) %>%
  mutate(analyte_dtxsid = ifelse(document_id %in% 61,
                                 case_when(
                                   series_id %in% 110 ~ "DTXSID4022527",
                                   series_id %in% 112 ~ "DTXSID3020209",
                                   .default = analyte_dtxsid
                                 ),
                                 analyte_dtxsid),
         analyte_name_original = ifelse(document_id %in% 61,
                                 case_when(
                                   series_id %in% 110 ~ "Propylparaben",
                                   series_id %in% 112 ~ "Butylparaben",
                                   .default = analyte_name_original
                                 ),
                                 analyte_name_original)
         ) %>%
  mutate(analyte_dtxsid = ifelse(document_id %in% c(46, 31),
                                 "DTXSID2020139",
                                 analyte_dtxsid),
         analyte_name_original = ifelse(document_id %in% c(46, 31),
                                        "benzo[a]pyrene",
                                        analyte_name_original))

# For study S0976, Caffeine, L-ephedrine, and psuedoephedrine were dosed
cvt_df <- cvt_df %>%
  mutate(loq = ifelse(
    other_study_identifier %in% "S0976",
    case_when(
      analyte_name_original %in% "Caffeine" ~ 4.00,
      analyte_name_original %in% "L-Ephedrine" ~ 4.44,
      analyte_name_original %in% "Pseudoephedrine" ~ 4.43
    ),
    loq),
    loq_units = ifelse(
      other_study_identifier %in% "S0976",
      "ng/mL",
      loq_units
    ))

# Some pentachloranisole studies
# S0361 seems to have some data that has not been extracted (also for mouse)
cvt_df <- cvt_df %>%
  mutate(loq = case_when(
    other_study_identifier %in% "S0361" ~ 0.05,
    other_study_identifier %in% "S0328" ~ 1, # This is inferred
    # Highest value under loq is 0.9
    # Lowest value above loq is 1.263, but there are differences in GC-EC and HPLC methods
    .default = loq
  ),
  loq_units = case_when(
    other_study_identifier %in% "S0361" ~ "ug/mL",
    other_study_identifier %in% "S0328" ~ "ug/mL",
    .default = loq_units
  ))

# Checking if entire groups of data are NA for conc and loq
# Really only eliminates Tamoxifen
# cvt_df %>%
#   # filter(analyte_dtxsid %in% "DTXSID1034187") %>%
#   group_by(analyte_dtxsid, document_id, 
#                     administration_route_normalized,
#                     conc_medium_normalized,
#                     species, dose_level_corrected) %>%
#   filter(all(is.na(conc) & is.na(loq))) %>% 
#   distinct(analyte_dtxsid, document_id, analyte_name_original,
#                     administration_route_normalized,
#                     conc_medium_normalized,
#                     species, dose_level_corrected) %>% View()

# Checking individual chemicals
# cvt_df %>%
#   filter(analyte_dtxsid %in% "DTXSID2020682") %>%
#   distinct(analyte_dtxsid, document_id, 
#                     administration_route_normalized,
#                     conc_medium_normalized,
#                     species, dose_level_corrected)

# PMID 6310077 is a tritium tracing of B[a]P and these were given as % of Dose
# The initial dose was 117 nmol, 2 mCi/kg
# Molecular Mass of benzo[a]pyrene (C20H12) is 252.316 g/mol
# The initial dose *I think* is in dose_level_normalized
cvt_df <- cvt_df %>%
  mutate(conc = ifelse(document_id %in% 38,
                       conc_original*dose_level_corrected,
                       conc))

# NTP-CEB C96019 just has a lot of data below LOQ
# For S0916, there is no > LOQ data for anything but Tamoxifen
# N-di-butyl phthalate is only good at the iv dose
# DTXSID2021315 is only good at 1E-4 Dose
# S0623 Doesn't have any plasma data
# There is a maybe control sample for hexachlorobenzene (Dose = 0.0 ) ?

# Filter by whether at each dose, there are no LOQ or Concentration value reported
# This is the threshold for Minimally Informative Data in our use case.
cvt_df %>%
  group_by(analyte_dtxsid, document_id, 
                    administration_route_normalized,
                    conc_medium_normalized,
                    species, dose_level_corrected) %>%
  filter(!all(is.na(conc) & is.na(loq))) %>%
  distinct(analyte_dtxsid, species,) %>% nrow() # Says there are 433 distinct

cvt_df <- cvt_df %>%
  group_by(analyte_dtxsid, document_id, 
                    administration_route_normalized,
                    conc_medium_normalized,
                    species, dose_level_corrected) %>%
  filter(!all(is.na(conc) & is.na(loq)))

# Give N-desmethyltamoxifen it's due DTXSID501315924
cvt_df <- cvt_df %>%
  mutate(analyte_dtxsid = ifelse(
    analyte_name_original %in% "N-desmethyltamoxifen",
    "DTXSID501315924",
    analyte_dtxsid
  ))

cvt_df <- cvt_df %>%
  mutate(analyte_dtxsid = ifelse(
    analyte_name_original %in% "Methemoglobin",
    "DTXSID301056085",
    analyte_dtxsid
  ))

# Standardizing the names (example: glucuronide had two entries)
cvt_df <- cvt_df %>%
  mutate(analyte_name_original = tolower(analyte_name_original))

# Eliminate study that gave chemical each day, not single dose
# This may be used in another type of model, but right now we deal with single bolus
# Only include those that have NA or 1 for dose frequency
cvt_df <- cvt_df %>%
  filter(dose_frequency %in% c("1", NA))

# Normalize LOQ units
cvt_df <- cvt_df %>%
  mutate(loq = ifelse(
    loq_units %in% c("ug/L", "ng/mL"), loq * 1E-3, loq)) # mg/L and ug/mL are same scale
# Normalize conc_sd and conc units manually
cvt_df <- cvt_df %>%
  mutate(conc_sd_normalized = case_when(
    conc_units_original %in% c("ug/L", "ng/mL", "ng-Eq/mL", "ng/g tissue") ~ conc_sd_original*1E-3,
    conc_units_original %in% c("g/dL") ~ conc_sd_original*1E-4,
    conc_units_original %in% c("pg/mL") ~ conc_sd_original*1E-6,
    .default = conc_sd_original
  ))
# Need to normalize the ng-eq/mL sample, conc is generally already normalized
cvt_df <- cvt_df %>%
  mutate(conc = case_when(
    conc_units_original %in% c("ng-eq/mL") ~ conc_original*1E-3,
    .default = conc
  ))

cvt_df <- cvt_df %>%
  dplyr::filter(!(analyte_dtxsid %in% c("DTXSID501315924",
                                      "DTXSID10169854",
                                      "DTXSID2021103") )) # These only have 1 or 2 values

# More notes:
# DTXSID6020438 & DTXSID0020868 rat is from water (top) and oil (bottom curve) gavage

```

```{r pk_obj_fromSQL}

### PK OBJECT ###----
# Minimum pk object, add options later
minimal_pk <- pk(data = cvt_df,
                 mapping = ggplot2::aes(
                   Chemical = analyte_dtxsid,
                   Chemical_Name = analyte_name_original,
                   DTXSID = analyte_dtxsid,
                   CASRN = analyte_casrn,
                   Species = species,
                   Reference = document_id,
                   Media = conc_medium_normalized,
                   Route = administration_route_normalized,
                   Dose = dose_level_normalized,
                   Dose.Units = "mg/kg",
                   Subject_ID = subject_id,
                   Series_ID = series_id,
                   Study_ID = study_id,
                   ConcTime_ID = conc_time_id,
                   N_Subjects = n_subjects_in_series,
                   Weight = weight_kg,
                   Weight.Units = "kg",
                   Time = time_hr,
                   Time.Units = "hours",
                   Value = conc,
                   Value.Units = "mg/L",
                   Value_SD = conc_sd_normalized,
                   LOQ = loq
                 ))
```

## Running all possible fitting options

This chunk is to run all the options through a
standard pipeline that calculates RMSE for Cmax and AUC_inf
for the Chemical Species Route Media grouping...

```{r pooled_fitting_choices}
# Types of Choices:
## Dose Normalized
## Log10 transform
## Scale time

# Pooled
my_pk_000p <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = FALSE, log10_trans = FALSE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species))

my_pk_100p <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = FALSE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species))

my_pk_110p <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = TRUE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species))

my_pk_111p <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = TRUE) +
  scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species))

my_pk_010p <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = FALSE, log10_trans = TRUE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species))

my_pk_011p <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = FALSE, log10_trans = TRUE) +
  scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species))

my_pk_001p <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = FALSE, log10_trans = FALSE) +
  scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species))

my_pk_101p <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = FALSE) +
  scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species))

```

```{r joint_fitting_choices}
# Joint
my_pk_000j <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = FALSE, log10_trans = FALSE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species, Reference))

my_pk_100j <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = FALSE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species, Reference))

my_pk_110j <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = TRUE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species, Reference))

my_pk_111j <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = TRUE) +
  scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species, Reference))

my_pk_010j <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = FALSE, log10_trans = TRUE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species, Reference))

my_pk_011j <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = FALSE, log10_trans = TRUE) +
  scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species, Reference))

my_pk_001j <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = FALSE, log10_trans = FALSE) +
  scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species, Reference))

my_pk_101j <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = FALSE) +
  scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species, Reference)) 
```

### Evaluating fitting options

```{r Evaluating_Choices}
# Evaluation
# Current: my_pk_111p

get_choices <- function(obj, PJ = "Pooled") {
  dose_norm <- obj$scales$conc$dose_norm
  log10_norm <- obj$scales$conc$log10_trans
  time_scale <- ifelse(obj$scales$time$new_units == "identity", FALSE, TRUE)

  my_choices <- paste0(dose_norm+0,
                       log10_norm+0,
                       time_scale+0,
                       " ", PJ)
  return(my_choices)
}

# Write into package
evaluate_choices <- function(obj, PJ = "Pooled") {
  pk_name <- get_choices(obj, PJ = PJ)

  cli::cli_progress_step("Winning Model...")
  # Wide winning model
  suppressMessages({
    winning_model <- get_winning_model(obj = obj) 
    wide_winning_model <- winning_model %>%
      group_by(method, model) %>%
      count() %>%
      pivot_wider(names_from = model, values_from = n) %>%
      mutate(Options = pk_name)
  })

  cli::cli_progress_step("Cmax and AUC_inf RMSEs...")
  # Cmax and AUC_inf RMSEs
  suppressMessages({
    RMSE_eval <- eval_tkstats(obj = obj) %>%
    mutate(Options = pk_name) %>%
    dplyr::select(Options,
                Chemical, Species,
                Route, Media, Reference,
                method, model,
                starts_with("Cmax"),
                starts_with("AUC_")) %>%
      filter(abs(AUC_inf.nca / AUC_tlast.nca) < 10,
             is.finite(AUC_inf.nca), is.finite(AUC_inf.tkstats),
             AUC_inf.nca > 0, AUC_inf.tkstats > 0) %>%
      rowwise() %>%
      mutate(SE_Cmax = (Cmax.tkstats - Cmax.nca)^2,
             SE_AUC_inf = (AUC_inf.tkstats - AUC_inf.nca)^2) %>%
      filter(!is.na(SE_Cmax), !is.na(SE_AUC_inf)) %>%
      group_by(Options, method) %>%
      summarize(RMSE_Cmax = sqrt(mean(SE_Cmax,
                                      na.rm = FALSE)) %>% signif(digits = 4),
                RMSE_AUC = sqrt(mean(SE_AUC_inf,
                                     na.rm = FALSE)) %>% signif(digits = 4))
  })

  cli::cli_progress_step("More detailed metrics...")
  # Other metrics (detailed)
  suppressMessages({
    r2_df <- rsq(obj, use_scale_conc = FALSE) %>%
    inner_join(winning_model)
  myAIC <- AIC(obj) %>%
    inner_join(winning_model)
  fold_errors_all <- fold_errors(obj) %>%
    inner_join(winning_model) %>%
    dplyr::filter(Detect %in% TRUE) %>%
    dplyr::group_by(!!!obj$data_group, model, method) %>%
    dplyr::summarize(avg_FoldErr = mean(Fold_Error, na.rm = TRUE) %>%
                       signif(digits = 4),
                     median_FoldErr = median(Fold_Error, na.rm = TRUE) %>%
                       signif(digits = 4),
                     within_2fold = sum(Fold_Error >= 0.5 & Fold_Error <= 2)/length(Fold_Error) %>%
                       signif(digits = 4))

  myRMSE_all <- rmse(obj,
                     rmse_group = vars(Route, Media, Dose),
                     use_scale_conc = FALSE) %>% 
    inner_join(winning_model) %>%
    dplyr::group_by(!!!obj$data_group, model, method) %>%
    dplyr::summarize(avg_MSE = mean(RMSE^2, na.rm = TRUE) %>%
                       signif(digits = 4),
                     median_MSE = median(RMSE^2, na.rm = TRUE) %>%
                       signif(digits = 4))

  my_df <- left_join(fold_errors_all, r2_df) %>%
    inner_join(myAIC) %>%
    inner_join(myRMSE_all) %>%
    mutate(Options = pk_name, .before = Chemical)

  full_list <- list(wide_winning_model, RMSE_eval, my_df)
  })

  return(full_list)
}
gc()

### Evaluations-----
my_pk_000p <- do_fit(my_pk_000p, n_cores = 14) 
my_pk_000p <- my_pk_000p %>% evaluate_choices() # Done
my_pk_100p <- do_fit(my_pk_100p, n_cores = 8) %>% evaluate_choices() # Done
my_pk_110p <- do_fit(my_pk_110p, n_cores = 8) %>% evaluate_choices() # Done
my_pk_111p <- do_fit(my_pk_111p, n_cores = 8) %>% evaluate_choices() # Done
rm(my_pk_100p,my_pk_110p,my_pk_111p)
gc()
my_pk_010p <- do_fit(my_pk_010p, n_cores = 8) %>% evaluate_choices() # Done
my_pk_011p <- do_fit(my_pk_011p, n_cores = 8) %>% evaluate_choices() # Done 
my_pk_001p <- do_fit(my_pk_001p, n_cores = 8) %>% evaluate_choices() # Done
my_pk_101p <- do_fit(my_pk_101p, n_cores = 8) %>% evaluate_choices() # Done

my_pk_000j <- do_fit(my_pk_000j, n_cores = 8) %>% evaluate_choices(PJ = "Joint") # Done
my_pk_100j <- do_fit(my_pk_100j, n_cores = 8) %>% evaluate_choices(PJ = "Joint") # Done
my_pk_110j <- do_fit(my_pk_110j, n_cores = 8) %>% evaluate_choices(PJ = "Joint") # Done
my_pk_111j <- do_fit(my_pk_111j, n_cores = 8) %>% evaluate_choices(PJ = "Joint") # Done
rm(my_pk_000j, my_pk_100j, my_pk_110j, my_pk_111j)
gc()
my_pk_010j <- do_fit(my_pk_010j, n_cores = 8) %>% evaluate_choices(PJ = "Joint") # Done
my_pk_011j <- do_fit(my_pk_011j, n_cores = 8) %>% evaluate_choices(PJ = "Joint") # Done
my_pk_001j <- do_fit(my_pk_001j, n_cores = 8) %>% evaluate_choices(PJ = "Joint") # Done
my_pk_101j <- do_fit(my_pk_101j, n_cores = 8) %>% evaluate_choices(PJ = "Joint") # Done

# Run after each evaluation
# initialization (only run once)
{
winning_model_final <- NULL
rmse_summary <- NULL
detail_stats <- NULL
}

# Run all at once
{
current_pk <- my_pk_101j # Define current evaluated pk object here
winning_model_final <- bind_rows(winning_model_final, current_pk[[1]])
rmse_summary <- bind_rows(rmse_summary, current_pk[[2]])
detail_stats <- bind_rows(detail_stats, current_pk[[3]])
rm(current_pk)
gc()
}

# For copy + pasting into Excel workbook
winning_model_final %>% clipr::write_clip(col.names = FALSE)
clipr::clear_clip()
rmse_summary %>% clipr::write_clip(col.names = FALSE)
clipr::clear_clip()
detail_stats %>% clipr::write_clip(col.names = FALSE)
clipr::clear_clip()
gc()

```

## Head-to-head comparison of pooled vs joint models

Now I need to go back and re-analyze the data only for 100, 101, and 000
Pooled and Joint, L-BFGS-B only. I need to filter the cvt_df to only include Chem + Species combinations
with multiple references. Then I will join them with the excel sheet of already collected data.

```{r load_multirefs}
# Reference Multiples
multiRef_ChemSpec <- cvt_df %>%
  ungroup() %>%
  distinct(analyte_dtxsid, species, document_id) %>%
  group_by(analyte_dtxsid, species) %>%
  count() %>%
  arrange(desc(n)) %>%
  filter(n > 1) %>%
  dplyr::select(!n)
multiRef_ChemSpec %>% nrow() # Surprisingly only 13 Chem + Species have multiple references

```

```{r Plotting_Summary}
finalist_options <- c("000 Pooled", "000 Joint",
                      "100 Pooled", "100 Joint",
                      "101 Pooled", "101 Joint")

readxl::excel_sheets(paste(Sys.getenv("OD_DIR"),
                           "DataSummary_9_21_2023.xlsx",
                           sep = "/")) # Predictions Evaluation

preval <- readxl::read_xlsx(paste(Sys.getenv("OD_DIR"),
                                  "DataSummary_9_21_2023.xlsx",
                                  sep = "/"),
                            sheet = "Predictions Evaluation") %>%
  left_join(multiRef_ChemSpec,
            by = join_by(Chemical == analyte_dtxsid,
                         Species == species)) %>%
  filter(method %in% "L-BFGS-B", 
         Options %in% finalist_options) %>%
  distinct()

preval %>%
  separate(Options, sep = " ",
           into = c("Options", "Error_Model")) %>%
  ggplot(aes(x = avg_FoldErr,
             color = Options)) +
  geom_freqpoly(bins = 10) +
  facet_grid(cols = vars(Error_Model))

preval %>% 
  separate(Options, sep = " ",
           into = c("Options", "Error_Model")) %>%
  group_by(Options, Error_Model) %>%
  summarize(totalFoldError = sum(avg_FoldErr))

readxl::read_xlsx(paste(Sys.getenv("OD_DIR"),
                                  "DataSummary_9_21_2023.xlsx",
                                  sep = "/"),
                            sheet = "Fit_Counts") %>%
  filter(Options %in% finalist_options, method %in% "L-BFGS-B")

readxl::read_xlsx(paste(Sys.getenv("OD_DIR"),
                                  "DataSummary_9_21_2023.xlsx",
                                  sep = "/"),
                            sheet = "Cmax and AUC_inf Evaluation") %>%
  filter(Options %in% finalist_options, method %in% "L-BFGS-B")

library(broom)

filtered_preval <- preval %>%
  filter(if_all(where(is.numeric), is.finite))

filtered_preval %>%
  dplyr::select(Chemical, Species, model, method, Options, Rsq) %>%
  mutate(time_scaled = ifelse(str_detect(Options,
                                         pattern = "^\\d\\d1"),
                              "Time_Scaled",
                              "notTime_Scaled"),
         Options = str_remove(Options, pattern = "\\d(?=[:blank:])")) %>%
  # dplyr::select(-Options) %>%
  pivot_wider(names_from = time_scaled, values_from = Rsq) %>% 
  mutate(differ = Time_Scaled - notTime_Scaled) %>%
  View()

fit <- lm(within_2fold ~ avg_FoldErr, data = filtered_preval)
summary(fit)
tidy(fit)
head(augment(fit))
glance(fit)

filtered_preval %>% group_by(Options) %>%
  nest() %>%
  mutate(fit = map(data, \(x) {
    lm(within_2fold ~ avg_FoldErr, data = x)
    }),
    tidy_fit = map(fit, tidy)) %>%
  unnest(tidy_fit) %>%
  dplyr::select(Options, term, estimate, std.error, statistic) %>%
  View()

Cmax_AUC <- readxl::read_xlsx(paste("C:/Users/GPADILLA",
                                  "OneDrive - Environmental Protection Agency (EPA)",
                                  "DataSummary_8_18_2023.xlsx",
                                  sep = "/"),
                            sheet = "Cmax and AUC_inf Evaluation") %>%
  dplyr::select(1:4) %>%
  distinct()

Cmax_AUC %>%
  mutate(DFO = rank(sqrt(RMSE_Cmax^2 + signif(RMSE_AUC, 5)^2)),
         OLS = rank(abs(RMSE_Cmax + signif(RMSE_AUC, 5))/sqrt(2))) %>%
  #filter(DFO < 11 | OLS < 11) %>%
  arrange(DFO, OLS) %>%
  dplyr::select(-RMSE_Cmax, -RMSE_AUC) %>% View()
  clipr::write_clip()

Cmax_AUC %>%
  mutate(DFO = rank(sqrt(RMSE_Cmax^2 + RMSE_AUC^2)),
         OLS = rank(abs(RMSE_Cmax + RMSE_AUC)/sqrt(2))) %>%
  ggplot(aes(
    x = DFO,
    y = OLS,
    fill = Options
  )) + geom_point(shape = 21)

```

## Running dose-normalized joint fits

Dose-normalization and a hierarchical model were the options that produced the least
amount of error in predictions and derived TK stats.

```{r Standard_Settings}
# Full option setting
my_pk <- pk(data = cvt_df,
            mapping = ggplot2::aes(
              Chemical = analyte_dtxsid,
              Chemical_Name = analyte_name_original,
              DTXSID = analyte_dtxsid,
              CASRN = analyte_casrn,
              Species = species,
              Reference = document_id,
              Media = conc_medium_normalized,
              Route = administration_route_normalized,
              Dose = dose_level_normalized,
              Dose.Units = "mg/kg",
              Subject_ID = subject_id,
              Series_ID = series_id,
              Study_ID = study_id,
              ConcTime_ID = conc_time_id,
              N_Subjects = n_subjects_in_series,
              Weight = weight_kg,
              Weight.Units = "kg",
              Time = time_hr,
              Time.Units = "hours",
              Value = conc,
              Value.Units = "mg/L",
              Value_SD = conc_sd_normalized,
              LOQ = loq
            )) +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = TRUE,
                      suppress.messages = FALSE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = FALSE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species, Reference)) 
my_pk <- do_preprocess(my_pk)
my_data <- get_data(my_pk)

```

An aside: working with one DTXSID + Species combination

The first step is understanding how to work with a single DTXSID + Species combination.
Once we know how to do this, it will be easier to apply it to (potentially)
a list of these objects in a vectorized fashion.
Here I subset the `cvt` object and use that as my working data.

```{r data_loading}
# There are 103 unique chemicals in this object
cvt %>% pull(chemicals_analyzed.dsstox_substance_id) %>%
  unique() %>%
  length()

# easy fit
my_predata <- cvt %>%
  dplyr::filter(chemicals_analyzed.dsstox_substance_id %in% "DTXSID5020732")

```

Next, I want to follow the steps required for the particular analysis I would like.
In this instance, I set the optimizer to "L-BFGS-B" and dose-normalize the output.
I do not scale the time or log-transform the concentration values.
Since I would like to do a "Pooled" analysis,
I set the `error_group = vars(Chemical, Species)`.

```{r single_fit}
my_pk_original <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = FALSE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = FALSE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species, Reference)) 
# Add Reference for Joint Analysis

my_pk_original <- do_preprocess(my_pk_original)

my_pk <- do_data_info(my_pk)

my_pk <- do_prefit(my_pk) # needs reframe somewhere
# Timing for normal fit sapply
my_pk <- do_fit(my_pk, 
                n_cores = 12)
my_pk <- do_fit(my_pk)

```

## Analysis Plots

### Figure 5A: Replicated timepoint variation in CvTdb

### Figure 5B: Variation over time in CvTdb data

### Supplementary Figure 1: Concentration and final time-points have wide range

### Figure 6: Selecting best fitting options for CvTdb data

### Supp Figure 2: invivoPKfit model performance

### Figure 7A: Model performance vs Data Variability

### Figure 7B: Multiple goodness-of-fit metrics validate model performance

### Supp. Fig. 3: Fits that may be improved with other settings

### Figure 8: Comparing derived TK stats with human TK stats from Lombardo et al.

### Supp. Fig. 4: Parallelization decreases runtime of invivoPKfit

```{r cvt_analysis}
my_df %>%
  ggplot(aes(x = maxT/(24)
             )) +
  geom_histogram(fill = "grey5") +
  scale_x_log10() +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 30)) +
  expand_limits(y = 0) +
  labs(x = "Day of final Timepoint",
       y = "Count") +
  theme(aspect.ratio = 1,
        panel.border = element_rect(color = "black", fill = NA, size = 1.5),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(face = "bold"),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        axis.text = element_text(face = "bold"))

```

The purpose of the following code is to generate data.frames with information for
all 103 chemicals in the `cvt` dataset that comes with `invivopkfit`.
I will need to do append each dataframe to the last for the outputs of the following
methods:

- `AIC()`
- `BIC()` left-joined with `AIC` output
- `get_tkstats()`

I will also have to fit these iteratively and manage the memory as part of the loop.

```{r allDTXSIDs_analysis}

# after fitting pk object for all cvt objects
# saveRDS(my_pk_100p, file = "data-raw/all_cvt_pk.rds")
# saveRDS(my_pk, file = "data-raw/all_cvt_pk_JOINT.rds")
my_pk <- readRDS("data-raw/all_cvt_pk.rds")
my_coefs <- coef(my_pk, drop_sigma = FALSE)
get_winning_model(my_pk) %>%
  filter(model %in% c("model_2comp"),
         method == "L-BFGS-B") %>%
  dplyr::select(Chemical, Species) %>% distinct() %>% nrow()
all_my_data  %>% dplyr::select(Chemical, Species) %>% distinct() %>% nrow()
my_AIC <- AIC(my_pk)
my_tkstats <- get_tkstats(my_pk)
my_IC <- left_join(AIC(my_pk), BIC(my_pk))

my_IC %>% group_by(Chemical, Species, method) %>%
  filter(AIC == min(AIC),
         method == "bobyqa") %>%
  ungroup() %>%
  count(model)
# model_1comp 59
# model_2comp 50
# model_flat   2

my_IC %>% group_by(Chemical, Species, method) %>%
  filter(AIC == min(AIC),
         method == "L-BFGS-B") %>%
  ungroup() %>%
  # filter(model == "model_flat") %>% pull(Chemical) %>% unique()
  count(model)
# model_1comp 57
# model_2comp 50
# model_flat   4

my_preds <- predict(my_pk,
                    use_scale_conc = TRUE)

winmodels <- my_IC %>% group_by(Chemical, Species, method) %>%
  filter(AIC == min(AIC),
         method == "L-BFGS-B") %>%
  ungroup() %>%
  dplyr::select(Chemical, Species,
                method, model)

my_preds <- inner_join(winmodels, my_preds)
my_tkstats <- inner_join(winmodels, my_tkstats) %>%
  dplyr::select(-TKstats) %>% distinct()
my_nca <- get_nca(my_pk,
                  nca_group = vars(Chemical, Species,))
my_nca <- my_nca %>% dplyr::select(-param_sd_z, -param_units) %>%
  pivot_wider(names_from = param_name,
              values_from = param_value)

all_my_data <- get_data(my_pk)
all_my_data %>%
  filter(Detect %in% TRUE,
         exclude %in% FALSE,
         Dose > 0) %>%
  group_by(Chemical,
           Species,
           Reference,
           Route, 
           Media,
           Dose, Time) %>%
  filter(n() > 1,
         sum(is.na(Conc)) != n(),
         LOQ < Conc) %>%
  mutate(meanConc = mean(Conc, na.rm = TRUE)) %>%
  ungroup() %>%
  rowwise() %>%
  mutate(foldConc = Conc / meanConc) %>%
  filter(Route == "iv") %>%
  count(foldConc < 0.5) %>%
  {.[[2,2]]/(.[[1,2]] + .[[2,2]])}
# 88.9% inside twofold ## New CvT extraction: 91.3%
# 2.2%  above twofold  ## 1.7%
# 8.8%  below twofold  ## 7.0%
# 
#

all_my_data %>%
  filter(Detect %in% TRUE,
         exclude %in% FALSE,
         Dose > 0) %>%
  group_by(Chemical,
           Species,
           Reference,
           Route, 
           Media,
           Dose, Time) %>%
  filter(n() > 1,
         sum(is.na(Conc)) != n(),
         LOQ < Conc) %>%
  mutate(meanConc = mean(Conc, na.rm = TRUE)) %>%
  ungroup() %>%
  rowwise() %>%
  mutate(foldConc = Conc / meanConc) %>%
  ggplot(aes(x = foldConc)) +
  geom_histogram(binwidth = 0.1,
                 fill = "grey5") +
  coord_cartesian(xlim =  c(0, 4),
                  ylim = c(0, 2000)) +
  scale_y_continuous(expand = c(0, 0)) +
  expand_limits(y = 0) +
  geom_vline(xintercept = c(0.5, 2), color = "red3", linetype = "dashed") +
  # annotate(geom= "text",
  #          x = 1.25, y = 1800,
  #          label = "88.9% of Observations\n within a factor of 2") +
  theme_classic() +
  labs(x = "Concentration Normalized to Average",
                y = "Number of Observations") +
  theme(panel.border = element_rect(color = "black", fill = NA),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.4))

ggsave(paste0(Sys.getenv("FIG_DIR"),
              "NormalizedConcentrations_2023-09-21.png"),
       height = 3.5,
       width = 5.5,
       units = "in")

all_my_data %>%
  group_by(Chemical,
           Species,
           Reference,
           Route,
           Dose, Time) %>%
  filter(n() > 1,
         sum(is.na(Conc_trans)) != n(),
         exclude == FALSE) %>%
  mutate(meanConc = mean(Conc_trans, na.rm = TRUE)) %>%
  ungroup() %>%
  rowwise() %>%
  mutate(foldConc = Conc_trans / meanConc) %>%
  inner_join(my_tkstats %>%
              dplyr::select(Chemical, Species,
                            Route, Media,
                            tmax) %>%
              filter(!is.na(tmax))) %>%
  group_by(Chemical,
           Species,
           Reference,
           Route,
           Media,
           Dose) %>%
  mutate(
         normTime = ifelse(
           Route == "iv",
           1 + Time/max(Time),
           ifelse(
             Time > tmax,
             ((Time - tmax)/max(Time)) + 1,
             0.5 + Time/(2*tmax)
           )
         )) %>%
  ggplot(aes(
    x = normTime,
    y = foldConc
  )) +
  geom_point(alpha = 0.1, size = 0.7) +
  geom_vline(xintercept = 1, color = "blue",
             size = 1, linetype = "dashed") +
  geom_vline(xintercept = 2, color = "green4",
             size = 1, linetype = "dashed") +
  geom_hline(yintercept = c(0.5, 2), color = "red3",
             size = 1, linetype = "dashed") +
  facet_grid(cols = vars(Route),
             rows = vars(Species),
             scales = "free_x") +
  labs(x = "ADME-Normalized Time",
       y = "Mean-Normalized Concentration") +
  theme(aspect.ratio = 1,
        panel.border = element_rect(color = "black", fill = NA, size = 1.5),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(face = "bold"),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        axis.text.x = element_blank(),
        legend.position = "none")

ggsave(paste0(Sys.getenv("FIG_DIR"),
              "ADME_NormalizedTime-NormalizedConcentrations_onlyDetects_2023-09-21.png"),
       height = 4,
       width = 6,
       units = "in",
       dpi = 300)

# Per chemical + Species average mean normalized concentration variability over time
all_my_data %>%
  group_by(Chemical,
           Species,
           Reference,
           Route,
           Dose, Time) %>%
  filter(n() > 1,
         sum(is.na(Conc)) != n(),
         LOQ < Conc,
         exclude == FALSE) %>%
  mutate(meanConc = mean(Conc, na.rm = TRUE)) %>%
  ungroup() %>%
  rowwise() %>%
  mutate(foldConc = Conc / meanConc) %>%
  inner_join(my_tkstats %>%
              dplyr::select(Chemical, Species,
                            Route, Media,
                            tmax) %>%
              filter(!is.na(tmax))) %>%
  group_by(Chemical, Species, Route, Media, Dose) %>%
  summarize(sdFC = sd(foldConc)) %>%
  mutate(maxDose = max(Dose),
         minDose = min(Dose)) %>%
  ungroup() %>%
  add_count(Dose, name = "Unique_Doses") %>%
  filter(Unique_Doses == 1) %>%
  arrange(desc(sdFC))

my_tkstats %>%
              dplyr::select(Chemical, Species,
                            Route, Media, model, method,
                            tmax) %>%
              filter(!is.na(tmax))

all_my_data %>%
              dplyr::select(Chemical, Species,
                            Route, Media, Dose,
                            Reference) %>%
  distinct() %>% nrow()

my_preds <- predict(my_pk,
                    use_scale_conc = TRUE)

winmodels <- get_winning_model(my_pk)

my_preds <- inner_join(winmodels, my_preds)
my_tkstats <- inner_join(winmodels, my_tkstats) %>%
  dplyr::select(-TKstats) %>% distinct()
my_nca <- get_nca(my_pk,
                  nca_group = vars(Chemical, Species,))
my_nca <- my_nca %>% dplyr::select(-param_sd_z, -param_units) %>%
  pivot_wider(names_from = param_name,
              values_from = param_value)

my_preds %>%
  group_by(Chemical,
           Species,
           Route, Media,
           Dose, Time) %>%
  filter(n() > 1,
         sum(is.na(Conc_trans)) != n(),
         # model %in% c("model_flat"),
         exclude %in% FALSE,
         Detect %in% TRUE) %>%
  mutate(meanConc = mean(Conc_trans, na.rm = TRUE)) %>%
  ungroup() %>%
  rowwise() %>%
  mutate(foldConc = Conc_trans / meanConc,
         foldPred = Conc_est / Conc_trans) %>% 
  # For percentages
  # ungroup() %>%
  # count(!(foldPred >= 0.5 & foldPred <= 2) &
  #         !(foldConc >= 0.5 & foldConc <= 2)) %>%
  # {.[[2,2]]/(.[[1,2]] + .[[2,2]])}
  # For Plotting
  ggplot(aes(
    y = foldPred,
    x = foldConc
  )) +
  geom_bin2d(bins = 40, color = NA) +
  geom_hline(yintercept = c(0.5, 2), linetype = "dashed") +
  geom_vline(xintercept = c(0.5, 2), linetype = "dashed") +
  facet_grid(cols = vars(model)) +
  scale_x_log10(limits = c(0.0001, 100)) +
  scale_y_log10(limits = c(0.001, 1000)) +
  scale_fill_viridis_c(option = "cividis",
                       limits = c(1, 100),
                       oob = oob_squish) +
  labs(y = "Model Performance",
       x = "Data Variability") +
  theme(aspect.ratio = 1,
        panel.border = element_rect(color = "black", fill = NA, size = 1.5),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(face = "bold"),
        axis.ticks = element_blank(),
        axis.line = element_blank())

ggsave(paste0(Sys.getenv("FIG_DIR"),
              "ModelPerformance_vDataVariability_2023-09-21.png"),
       height = 3.5,
       width = 7,
       units = "in")

# Why are 38-39% of adequate data have models that are too variable?
# 
my_preds %>%
  group_by(Chemical,
           Species,
           Route,
           Dose, Time) %>%
  filter(n() > 1,
         sum(is.na(Conc_trans)) != n(),
         exclude == FALSE,
         Detect == TRUE) %>%
  mutate(meanConc = mean(Conc_trans, na.rm = TRUE)) %>%
  ungroup() %>%
  rowwise() %>%
  mutate(foldConc = Conc_trans / meanConc,
         foldPred = Conc_est / Conc_trans) %>% 
  # filter(model == "model_2comp") %>%
  mutate(VAR_DATA = ifelse(
    (foldPred > 2 | foldPred < 0.5) & (foldConc <= 2 & foldConc >= 0.5),
    TRUE, FALSE
  )) %>%
  group_by(Chemical, Species) %>%
  mutate(VAR_TRUE = sum(VAR_DATA %in% TRUE),
         VAR_FALSE = sum(VAR_DATA %in% FALSE)) %>%
  dplyr::select(-VAR_DATA) %>% distinct() %>%
  mutate(PERCENT_TRUE = VAR_TRUE/(VAR_TRUE + VAR_FALSE),
         MAJOR_TRUE = ifelse(PERCENT_TRUE > 0.6, TRUE, FALSE)) %>%
  # View("Count")
  ggplot(aes(
    x = Time,
    y = Conc_est,
    color = paste(Chemical, Species),
    group = paste(Chemical, Species)
  )) +
  geom_point() +
  geom_line() +
  facet_grid(cols = vars(MAJOR_TRUE),
             rows = vars(model)) +
  labs(y = "Predicted Concentrations",
       x = "Time") +
  # xlim(0, 2000) +
  theme(aspect.ratio = 1,
        panel.border = element_rect(color = "black", fill = NA, size = 1.5),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(face = "bold"),
        legend.position = "none",
        axis.ticks = element_blank(),
        axis.line = element_blank())

# Saving some CSVs
# my_tkstats %>% write_csv(file = "data-raw/Pooled_TKstats.csv")
# my_nca %>% write_csv(file = "data-raw/NonCompartmentalAnalysis.csv")
# Loading CSVs
my_nca <- read_csv(file = "data-raw/NonCompartmentalAnalysis.csv")

my_preds %>%
  rowwise() %>%
  mutate(Pred_Obs = Conc_trans/Conc_est) %>%
  filter(is.finite(Pred_Obs),
         model %in% c("model_1comp", "model_2comp")) %>%
  # mutate(pass = ifelse(Pred_Obs < 0.5, TRUE, FALSE)) %>%
  # pull(pass) %>% table()
  ggplot(aes(x = Pred_Obs)) +
  geom_histogram(binwidth = 0.4,
                 fill = "grey5") +
  scale_x_log10(limits = c(NA, 1E6)) +
  expand_limits(y = 0) +
  geom_vline(xintercept = c(0.5, 2), color = "red3", linetype = "dashed") +
  facet_grid(cols = vars(model),
             rows = vars(Route)) +
  labs(x = "Predicted/Observed",
       y = "Number of Observations") +
  theme_classic() +
  theme(aspect.ratio = 1,
        panel.border = element_rect(color = "black", fill = NA),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.4))

# 91.1% of Pred_Obs are finite
# 16.6% Are below two-fold
# 33.2%  Are above two-fold
# 50.2% Are within twofold

ggsave(paste0(Sys.getenv("FIG_DIR"),
              "PredictedObserved_compartmentModels_2023-07-11.png"),
       height = 6, width = 7)

# Pooled analysis
all_my_data %>% group_by(Chemical, Species, Reference, Media, Route) %>%
  filter(is.finite(Conc)) %>%
  summarise(OOM = log10(max(Conc) / min(Conc))) %>% 
  ggplot(aes(x = OOM)) +
  geom_histogram(bins = 15,
                 color = NA,
                 fill = "grey5") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  scale_x_continuous(breaks = c(0, 1, 2, 3, 4, 5)) +
  labs(x = "Orders of Magnitude per Series\n log10(max/min)",
       y = "Count") +
  theme(aspect.ratio = 1,
        panel.border = element_rect(color = "black", fill = NA, size = 1.5),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
        axis.ticks = element_blank(),
        axis.line = element_blank())

ggsave(paste0(Sys.getenv("FIG_DIR"),
              "OrdersOfMagnitude_2023-07-12.png"),
       height = 4, width = 4)

my_pk$prefit$fit_check %>% filter(fit_decision == "abort") %>%
  dplyr::select(model, Chemical) %>% distinct()

all_my_data %>%
  filter(Chemical == "DTXSID2021781")

my_coefs %>% inner_join(get_winning_model(my_pk)) -> my_coefs
my_coefs %>% inner_join(get_data_summary(my_pk)) -> my_coefs

my_coefs %>%
  mutate(model = 
           case_when(
             model == "model_1comp" ~ "1-compartment",
             model == "model_2comp" ~ "2-compartment",
             model == "model_flat"  ~ "Null"
           )) %>%
  ggplot(aes(x = sigma_value/n_obs)) +
  geom_histogram(bins = 15,
                 color = NA, fill = "grey5") +
  scale_x_log10(labels = label_log()) +
  facet_grid(cols = vars(model)) +
  labs(x = "Size-normalized Sigma Value",
       y = "Number of Observations",
       title = "Distribution of Sigmas in CvTdb data") +
  theme(panel.border = element_rect(color = "black", fill = NA, size = 1.5),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(face = "bold"),
        plot.title = element_text(hjust = 0.5),
        axis.ticks = element_blank(),
        axis.line = element_blank())

ggsave(paste0(Sys.getenv("FIG_DIR"),
              "NormalizedSigma_Distributions_2023-07.png"),
       height = 3, width = 5)
```

This next set of code is a preliminary attempt at summarizing the TK statistics
for Pooled and Joint analyses.

```{r update_eval_tkstats}
devtools::load_all()
my_pk <- readRDS("data-raw/all_cvt_pk.rds") # This no longer works somehow
my_eval <- eval_tkstats(my_pk)

# To give summary RMSEs of Cmax and AUC...
my_eval <- my_eval %>%
  group_by(method, model, .add = TRUE) %>%
  ungroup(Reference) %>%
  summarize(AUC_inf_RMSE = sqrt((1/n())*(sum(AUC_inf.tkstats - AUC_inf.nca,
                                             na.rm = TRUE)^2)),
            Cmax_RMSE = sqrt((1/n())*(sum(Cmax.tkstats - Cmax.nca,
                                          na.rm = TRUE)^2)),
            .groups = "keep")

my_eval <- my_eval %>%
  mutate(dose_norm = TRUE,
         log10_norm = FALSE,
         scale_time = FALSE,
         Analysis_Type = "Pooled")

ggplot(my_eval,
       aes(x = AUC_inf_RMSE,
           y = Cmax_RMSE)) +
  geom_point() +
  geom_abline(slope = 1) +
  scale_y_log10() +
  scale_x_log10()

ggplot(my_eval,
       aes(x = -log10(AUC_inf_RMSE))) +
  geom_histogram(fill = "grey5") +
  facet_grid(cols = vars(method)) +
  labs(x = "Accuracy of AUC",
       y = "Observations") +
  theme(panel.border = element_rect(color = "black", fill = NA, size = 1.5),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
        strip.background = element_rect(fill = "white"),
        axis.ticks = element_blank(),
        axis.line = element_blank())

```

```{r eval_Joint}
my_pk2 <- readRDS("data-raw/all_cvt_pk_JOINT.rds")
my_eval_joint <- eval_tkstats(my_pk2)

# To give summary RMSEs of Cmax and AUC...
my_eval_joint <- my_eval_joint %>%
  group_by(method, model, .add = TRUE) %>%
  ungroup(Reference) %>%
  summarize(AUC_inf_RMSE = sqrt((1/n())*(sum(AUC_inf.tkstats - AUC_inf.nca,
                                             na.rm = TRUE)^2)),
            Cmax_RMSE = sqrt((1/n())*(sum(Cmax.tkstats - Cmax.nca,
                                          na.rm = TRUE)^2)),
            .groups = "keep")

my_eval_joint <- my_eval_joint %>%
  mutate(dose_norm = TRUE,
         log10_norm = FALSE,
         scale_time = FALSE,
         Analysis_Type = "Joint")

ggplot(my_eval_joint,
       aes(x = AUC_inf_RMSE,
           y = Cmax_RMSE)) +
  geom_point()

ggplot(my_eval_joint,
       aes(x = -log10(AUC_inf_RMSE))) +
  geom_histogram(fill = "grey5") +
  facet_grid(cols = vars(method)) +
  labs(x = "Accuracy of AUC",
       y = "Observations") +
  theme(panel.border = element_rect(color = "black", fill = NA, size = 1.5),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
        strip.background = element_rect(fill = "white"),
        axis.ticks = element_blank(),
        axis.line = element_blank())

total_eval <- bind_rows(my_eval, my_eval_joint)

ggplot(total_eval,
       aes(x = AUC_inf_RMSE,
           y = Cmax_RMSE)) +
  geom_point() +
  facet_grid(rows = vars(Analysis_Type))

total_eval %>%
  dplyr::select(!Cmax_RMSE) %>%
  pivot_wider(names_from = Analysis_Type,
              values_from = AUC_inf_RMSE) %>%
  ggplot(aes(x = log10(Pooled),
             y = log10(Joint),
             color = Media)) +
  geom_point(alpha = 0.3) +
  geom_abline(slope = 1, linetype = "dashed") +
  facet_grid(rows = vars(model),
             cols = vars(method)) +
  theme(panel.border = element_rect(color = "black", fill = NA, size = 1.5),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
        strip.background = element_rect(fill = "white"),
        axis.ticks = element_blank(),
        axis.line = element_blank())

total_eval %>%
  dplyr::select(!Cmax_RMSE) %>%
  pivot_wider(names_from = Analysis_Type,
              values_from = AUC_inf_RMSE) %>%
  mutate(ratio = Joint/Pooled) %>%
  # filter(!(ratio == 1)) %>% 
  # View()
  pull(Chemical) %>%
  # unique()
  length()

```

Here I introduce the notion whether the data is variable compared across time.

```{r PredictionVariability_overtime}
my_pk <- readRDS("data-raw/all_cvt_pk.rds")
winmodels <- AIC(my_pk) %>% group_by(Chemical, Species, method) %>%
  filter(AIC == min(AIC),
         method == "L-BFGS-B") %>%
  ungroup() %>%
  dplyr::select(Chemical, Species,
                method, model)
my_preds <- inner_join(winmodels,
                       predict(my_pk,
                               use_scale_conc = TRUE))

my_tkstats <- inner_join(winmodels, get_tkstats(my_pk))
my_residuals <- residuals(my_pk, use_scale_conc = FALSE)

my_residuals %>%
  filter(exclude == FALSE) %>%
  inner_join(winmodels) %>%
  inner_join(my_tkstats %>%
              dplyr::select(Chemical, Species,
                            Route, Media,
                            tmax) %>%
              filter(!is.na(tmax))) %>%
  group_by(Chemical,
           Species,
           Route,
           Media,
           Dose) %>%
  mutate(
         normTime = ifelse(
           Route == "iv",
           1 + Time/max(Time),
           ifelse(
             Time > tmax,
             ((Time - tmax)/max(Time)) + 1,
             0.5 + Time/(2*tmax)
           )
         )) %>%
  distinct() %>%
  filter(Detect %in% TRUE) -> my_residuals2
my_residuals2 %>%
  # filter(Route == "oral") %>% View()
  ggplot(aes(
    x = normTime,
    y = Residuals
  )) +
  geom_point(alpha = 0.2, size = 0.5) +
  geom_point(data = filter(my_residuals2,
                           Chemical %in% "DTXSID4020533",
                           Species %in% "rat"),
             color = "orange2", size = 0.5) +
  geom_point(data = filter(my_residuals2,
                           Chemical %in% "DTXSID8031865",
                           Species %in% "rat"),
             color = "magenta4", size = 0.5) +
  geom_point(data = filter(my_residuals2,
                           Chemical %in% "DTXSID8025337",
                           Species %in% "rat"),
             color = "green2", size = 0.5) +
  geom_vline(xintercept = 1, color = "blue", linetype = "dashed") +
  geom_vline(xintercept = 2, color = "green4", linetype = "dashed") +
  facet_grid(cols = vars(Route),
             # rows = vars(model),
             scales = "free_x") +
  labs(x = "ADME-Normalized Time",
       y = "Winning Model Residuals") +
  theme(aspect.ratio = 1,
        panel.border = element_rect(color = "black", fill = NA, size = 1.5),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(face = "bold"),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        axis.text.x = element_blank())

ggsave(paste0(Sys.getenv("FIG_DIR"),
              "ADME_NormalizedTime-WinModelResiduals_onlyDetects_Highlighted_2023-07-19.png"),
       height = 4,
       width = 6,
       units = "in")
```

This is another preliminary attempt to summarize RMSEs of each fit
(Predictions - Observed Values)

```{r boilerplate_RMSE_analysis}
my_pk <- readRDS("data-raw/all_cvt_pk.rds")
my_AIC <- AIC(my_pk)

my_AIC %>% group_by(Chemical, Species, method) %>%
  filter(AIC == min(AIC),
         method == "bobyqa") %>%
  ungroup() %>%
  count(model)

my_AIC %>% group_by(Chemical, Species, method) %>%
  filter(AIC == min(AIC),
         method == "L-BFGS-B") %>%
  ungroup() %>%
  count(model)

my_eval <- eval_tkstats(my_pk)

# To give summary RMSEs of Cmax and AUC...
my_eval <- my_eval %>%
  group_by(method, model, .add = TRUE) %>%
  ungroup(Reference) %>%
  summarize(AUC_inf_RMSE = sqrt((1/n())*(sum(AUC_inf.tkstats - AUC_inf.nca,
                                             na.rm = TRUE)^2)),
            Cmax_RMSE = sqrt((1/n())*(sum(Cmax.tkstats - Cmax.nca,
                                          na.rm = TRUE)^2)),
            .groups = "keep")

my_eval %>%
  group_by(method) %>%
  summarize(AUC_meanRMSE = mean(AUC_inf_RMSE),
            AUC_mediRMSE = median(AUC_inf_RMSE),
            AUC_sdRMSE = sd(AUC_inf_RMSE),
            Cmax_meanRMSE = mean(Cmax_RMSE),
            Cmax_mediRMSE = median(Cmax_RMSE),
            Cmax_sdRMSE = sd(Cmax_RMSE)) %>%
  as.data.frame()
```

Here I explore some bad-fitting chemicals using the dose-normalized pooled analysis.

```{r bad_fits}
predata_chemical <- my_pk$prefit$fit_check %>%
  filter(fit_decision == "abort") %>%
  pull(Chemical) %>%
  unique()

my_pk <- pk(data = cvt %>%
  dplyr::filter(chemicals_analyzed.dsstox_substance_id %in% predata_chemical)) +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = FALSE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species)) 
# Add Reference for Joint Analysis

my_pk <- do_preprocess(my_pk)

get_data(my_pk) %>% View()

plot(my_pk, print_out = TRUE)[[1]] +
  labs(subtitle = "Dibutyl phthalate (DBP)") +
  theme(aspect.ratio = 1,
        plot.subtitle = element_text(hjust = 0.5))
ggsave(paste0(Sys.getenv("FIG_DIR"),
              "DTXSID20211781_rat_badFit.png"),
       width = 6,
       height = 4,
       units = "in")

my_pk <- pk(data = cvt %>%
  dplyr::filter(chemicals_analyzed.dsstox_substance_id %in% "DTXSID2021315")) +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = FALSE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species))

my_pk <- do_fit(my_pk)

plot(my_pk,
     method = "L-BFGS-B",
     n_interp = 10,
     print_out = TRUE)[[1]] +
  labs(subtitle = "2,3,7,8-tetrachlorodibenzo-p-dioxin") +
  scale_x_continuous(breaks = c(0, (1:4)*(24*7*4)),
                     labels = c(0, 1:4),
                     name = "Time (months)") +
  guides(color = "none",
         shape = "none") +
  theme(aspect.ratio = 1,
        plot.subtitle = element_text(hjust = 0.5))
ggsave(paste0(Sys.getenv("FIG_DIR"),
              "DTXSID2021315_rat_badFit.png"),
       width = 8,
       height = 5,
       units = "in")

my_pk <- pk(data = cvt %>%
  dplyr::filter(chemicals_analyzed.dsstox_substance_id %in% "DTXSID3032179")) +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = FALSE, log10_trans = FALSE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species))

my_pk <- do_fit(my_pk)

plot(my_pk,
     method = "L-BFGS-B",
     use_scale_conc = FALSE,
     log10_C = FALSE,
     n_interp = 50,
     print_out = TRUE)[[1]] +
  labs(subtitle = "3,3',4,4',5-pentachlorobiphenyl (without Dose-Normalized Fit)") +
  scale_x_continuous(breaks = c(0, (1:14)*(24*7*4)),
                     labels = c(0, 1:14),
                     name = "Time (months)") +
  guides(color = "none",
         shape = "none") +
  theme(aspect.ratio = 1,
        plot.subtitle = element_text(hjust = 0.5))
ggsave(paste0(Sys.getenv("FIG_DIR"),
              "DTXSID3032179_rat_badFit_noDoseNorm.png"),
       width = 7.5,
       height = 5,
       units = "in")

my_pk <- pk(data = cvt %>%
  dplyr::filter(chemicals_analyzed.dsstox_substance_id %in% "DTXSID3032179")) +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = FALSE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species))

my_pk <- do_fit(my_pk)

plot(my_pk,
     method = "L-BFGS-B",
     use_scale_conc = FALSE,
     log10_C = FALSE,
     n_interp = 50,
     print_out = TRUE)[[1]] +
  labs(subtitle = "3,3',4,4',5-pentachlorobiphenyl (Dose-Normalized Fit)") +
  scale_x_continuous(breaks = c(0, (1:14)*(24*7*4)),
                     labels = c(0, 1:14),
                     name = "Time (months)") +
  guides(color = "none",
         shape = "none") +
  theme(aspect.ratio = 1,
        plot.subtitle = element_text(hjust = 0.5))
ggsave(paste0(Sys.getenv("FIG_DIR"),
              "DTXSID3032179_rat_badFit_DoseNorm.png"),
       width = 7.5,
       height = 5,
       units = "in")

my_pk <- pk(data = cvt %>%
  dplyr::filter(chemicals_analyzed.dsstox_substance_id %in% "DTXSID1034187")) +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = FALSE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species))

my_pk <- do_fit(my_pk)

get_data(my_pk) %>% View()
plot(my_pk,
     method = "L-BFGS-B",
     use_scale_conc = TRUE,
     log10_C = FALSE,
     n_interp = 100,
     print_out = TRUE)[[1]] +
  labs(subtitle = "Tamoxifen") +
  theme(aspect.ratio = 1,
        plot.subtitle = element_text(hjust = 0.5))
ggsave(paste0(Sys.getenv("FIG_DIR"),
              "DTXSID1034187_rat_badFit_.png"),
       width = 7,
       height = 7,
       units = "in")
# Fit is actually ok, but there are lot's of non-Detects in the data
# There are many doses that are fully non-Detects.

problem_dose <- all_my_data %>%
  group_by(Chemical, Species, Route, Media, Dose) %>%
  filter(sum(Detect) == 0) %>%
  dplyr::select(Chemical, Species, Route, Media, Dose) %>%
  rename(chemicals_analyzed.dsstox_substance_id = Chemical,
         subjects.species_harmonized = Species,
         series.conc_medium_normalized = Media,
         studies.administration_route_normalized = Route,
         studies.dose_level_normalized_corrected = Dose) %>%
  distinct()

problem_dose_DTX <- all_my_data %>%
  group_by(Chemical, Species, Route, Media, Dose) %>%
  filter(sum(Detect) == 0) %>%
  ungroup() %>%
  pull(1) %>% unique()

# Figure out if the fits improve if we exclude Dose groups that do not have
# any Detects

my_pk_full <- pk(data = cvt %>%
                   dplyr::filter(chemicals_analyzed.dsstox_substance_id %in% 
                                   "DTXSID1034187")) +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = FALSE) +
  stat_error_model(error_group = vars(Chemical, Species))

my_pk_full <- do_preprocess(my_pk_full)
my_pk_full <- do_fit(my_pk_full)

plot(my_pk_full,
     n_interp = 50,
     use_scale_conc = TRUE,
     print_out = TRUE)

my_pk_trimmed <- pk(data = cvt %>%
                      dplyr::filter(chemicals_analyzed.dsstox_substance_id %in%
                                      "DTXSID1034187") %>%
                      anti_join(problem_dose)) +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = FALSE) +
  stat_error_model(error_group = vars(Chemical, Species))

my_pk_trimmed <- do_preprocess(my_pk_trimmed)
my_pk_trimmed <- do_fit(my_pk_trimmed)

AIC(my_pk_full) %>% View("full")
AIC(my_pk_trimmed) %>% View("trimmed")

get_winning_model(my_pk_full)
get_winning_model(my_pk_trimmed)

eval_tkstats(my_pk_full) %>% View("full_tkstats")
eval_tkstats(my_pk_trimmed) %>% View("trimmed_tkstats")
get_tkstats(my_pk_trimmed) %>% View()

plot(my_pk_trimmed,
     n_interp = 50,
     method = "L-BFGS-B",
     print_out = TRUE)

# PFAS Chemical
my_predata <- cvt %>%
  dplyr::filter(chemicals_analyzed.dsstox_substance_id %in% "DTXSID80108992")
my_pk <- pk(data = my_predata) +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = TRUE,
                      suppress.messages = FALSE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = FALSE, log10_trans = FALSE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species)) 
my_pk <- do_fit(my_pk)
plot(my_pk, n_interp = 20,
     method = "L-BFGS-B", print_out = TRUE)
ggsave(paste0(Sys.getenv("FIG_DIR"),
              "DTXSID80108992_noTimeScaling.png"),
       height = 3,
       width = 5.5,
       units = "in")
# Time-scaled!
my_pk <- pk(data = my_predata) +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = FALSE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = FALSE, log10_trans = FALSE) +
  scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species)) 
my_pk <- do_data_info(my_pk)
my_pk <- do_fit(my_pk)
dplot(my_pk, n_interp = 20,
     method = "L-BFGS-B", print_out = TRUE)
ggsave(paste0(Sys.getenv("FIG_DIR"),
              "PerluorodeconoicAcid_noTimeScaling.png"),
       height = 4,
       width = 5.5,
       units = "in")

```

I try to find the connections between functions in _invivoPKfit_ to best understand
and represent the structure of the package.

```{r function_connections}
fw <- mvbutils::foodweb(where = asNamespace("invivoPKfit"),
                        border = TRUE, cex = 0.6, lwd = 1.5)
fun_mat <- fw$funmat

mvbutils::foodweb(where = asNamespace("invivoPKfit"),
                  prune = "convert_time",
                  # descendents = FALSE,
                  # ancestors = FALSE,
                  border = TRUE, cex = 1, lwd = 1.5)

# Predict dependents: rmse, residuals, rsq, calc_auc, fold_errors, plot
# Coef dependents: calc_auc, predict, get_tkstats, summary,log_Lik, coef_sd

fun_mat2 <- fun_mat[rowSums(fun_mat) > 0, ]

library(ComplexHeatmap)
ComplexHeatmap::Heatmap(fun_mat2, row_names_side = "left", column_names_side = "top",
                        row_names_gp = gpar(fontsize = 10),
                        column_names_gp = gpar(fontsize = 10),
                        col = circlize::colorRamp2(c(0, 1), colors = c("white", "green4")),
                        rect_gp = gpar(col = "black", lwd = 1),
                        show_row_dend = FALSE,
                        show_column_dend = FALSE)

fun_mat %>%
  as_tibble(rownames = NA) %>%
  rownames_to_column(var = "from") %>%
  pivot_longer(cols = !from,
               names_to = "to",
               values_to = "Call") %>%
   mutate(
    from = str_remove(from, pattern = ".pk"),
    to = str_remove(to, pattern = ".pk")
  ) %>%
  filter(from %in% "residuals") %>% View()

fun_mat2 %>%
  as_tibble(rownames = NA) %>%
  rownames_to_column(var = "from") %>%
  pivot_longer(cols = !from,
               names_to = "to",
               values_to = "Call") %>%
  filter(Call != 0) %>% 
  transmute(
    from = str_remove(from, pattern = ".pk"),
    to = str_remove(to, pattern = ".pk")
  ) %>%
  filter(from %in% c("residuals"),
         !(to %in% c("check_model",
                     "check_method",
                     "check_newdata",
                     "check_required_status"))) %>%
  ggflowchart::ggflowchart(horizontal = TRUE)

```

## Q & A

Here I try asking the remaining questions for the paper/presentation.

```{r questions}
my_pk <- readRDS("data-raw/all_cvt_pk.rds")
pk_summary <- summary(my_pk) # Testing the summary functionality

# Q: What do "bad" fits have in common?
gof_df <- pk_summary$goodness_of_fit %>%
  inner_join(suppressMessages(get_winning_model(my_pk))) %>%
  filter(method == "L-BFGS-B") %>%
  ungroup()

## Find the within_2fold for each data_group
gof_df %>%
  dplyr::select(Chemical, Species, full_fold_errors) %>%
  unnest(full_fold_errors) %>%
  group_by(Chemical, Species) %>%
  summarize(within_2fold = sum(
    Fold_Error >= 0.5 & Fold_Error <= 2, na.rm = TRUE
    ) / length(na.omit(Fold_Error))) %>%
  filter(within_2fold == 0) %>% pull(Chemical) %>% unique()
  arrange(within_2fold) %>%
  View()

gof_df %>%
  ggplot(aes(
    x = log2(avg_FoldErr),
    y = log10(avg_rmse))) +
  geom_point() +
  geom_vline(color = "red3", xintercept = c(-1, 1))

winning_model <- get_winning_model(my_pk, method = "L-BFGS-B")
fe_df <- fold_errors(my_pk, method = "L-BFGS-B") %>% inner_join(winning_model)
# May need to change the use_scale_conc defaults
r2_df <- rsq(my_pk, method = "L-BFGS-B", use_scale_conc = FALSE) %>%
  inner_join(winning_model)
myAIC <- AIC(my_pk, method = "L-BFGS-B") %>%
  inner_join(winning_model)
myRMSE <- rmse(my_pk, method = "L-BFGS-B",
               rmse_group = vars(Route, Media, Dose),
               use_scale_conc = FALSE) %>% 
  inner_join(winning_model)

# Something wrong with the summary method... maybe only use detects?
fold_errors_all <- fe_df %>%
  dplyr::filter(Detect %in% TRUE) %>%
  dplyr::group_by(!!!my_pk$data_group, model, method) %>%
  dplyr::summarize(avg_FoldErr = mean(Fold_Error, na.rm = TRUE),
                median_FoldErr = median(Fold_Error, na.rm = TRUE),
                within_2fold = sum(Fold_Error >= 0.5 & Fold_Error <= 2)/length(Fold_Error))

myRMSE_all <- myRMSE %>%
  dplyr::group_by(!!!my_pk$data_group, model, method) %>%
  dplyr::summarize(avg_MSE = mean(RMSE^2, na.rm = TRUE),
                median_MSE = median(RMSE^2, na.rm = TRUE))

my_df <- left_join(fold_errors_all, r2_df) %>% inner_join(myAIC) %>% inner_join(myRMSE_all)
mypl <- ggplot(data = my_df,
               aes(
                 x = within_2fold,
                 y = Rsq,
                 color = model
               )) +
  geom_point() +
  geom_abline(slope = 1, color = "red4", linetype = "longdash") +
  scale_color_manual(values = c("#0398FC", "#D68E09", "black")) +
  labs(x = "% Fold Error within 2-fold",
       y = "R-squared value") +
  theme(aspect.ratio = 1,
        panel.border = element_rect(color = "black", fill = NA, size = 1.5),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.key = element_blank())
panelA_plot <- ggExtra::ggMarginal(mypl, groupFill = TRUE,
                    type = "histogram",
                    xparams = list(binwidth = 0.05),
                    yparams = list(binwidth = 0.05))
ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "R2_within2fold_comparison.png"),
       plot = ggExtra::ggMarginal(mypl, groupFill = TRUE,
                    type = "histogram",
                    xparams = list(binwidth = 0.05),
                    yparams = list(binwidth = 0.05)),
       height = 5,
       width = 5,
       units = "in")

panelB_plot <- ggplot(data = myRMSE %>%
         inner_join(winning_model) %>%
         mutate(RMSE = ifelse(RMSE == 0, NA, RMSE)),
               aes(
                 x = log10(RMSE),
                 color = model,
                 fill = model
               )) + 
  geom_histogram(bins = 100) +
  labs(y = "Count") +
  scale_color_manual(values = c("#0398FC", "#D68E09", "grey10")) +
  scale_fill_manual(values = c("#0398FC", "#D68E09", "grey10")) +
  scale_y_continuous(breaks = scales::pretty_breaks()) +
  facet_grid(rows = vars(model),
             cols = vars(Route)) +
  theme(aspect.ratio = 0.5,
        panel.border = element_rect(color = "black", fill = NA, size = 1.5),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        strip.background = element_blank(),
        legend.position = "none",
        legend.title = element_blank())
ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RMSEs_winningmodel_LBFGSB_zero1E-6_RouteFacet_notNormalized.png"),
       height = 4,
       width = 6,
       units = "in")

cowplot::plot_grid(panelA_plot, panelB_plot, labels = c("A", "B"))
ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "CombinedPlot_Rsquared_RMSE.png"),
       height = 5,
       width = 7.45,
       bg = "white",
       units = "in")

 rmse(my_pk, method = "L-BFGS-B",
      rmse_group = vars(Route, Media, Dose)) %>%
   inner_join(get_winning_model(my_pk, method = "L-BFGS-B")) %>%
   count(model)

rmse(my_pk, method = "L-BFGS-B",
                   rmse_group = vars(Route, Media, Dose)) %>%
  inner_join(winning_model) %>%
  filter(RMSE < 0.01, RMSE > 0.0001, Route == "oral", model == "model_1comp") %>%
  group_by(Chemical, Species, Route, Media) %>%
  mutate(group_size = n(),
         logRMSE = log10(RMSE)) %>%
  arrange(Chemical, Species, Route, Media, RMSE) %>%
  View()

fold_errors_all %>% 
  ggplot(aes(x = within_2fold)) +
  geom_histogram(binwidth = 0.1) +
  facet_grid(cols = vars(model))

ggplot() +
  geom_histogram(data = fold_errors_all, 
                 aes(x = within_2fold),
                 bins = 100) +
  geom_vline(xintercept = fold_errors_all %>%
               filter(Chemical == "DTXSID8025337") %>%
               pull(within_2fold)) +
  scale_x_log10()

fold_errors_all %>% ungroup() %>%
  mutate(per_rank = rank(within_2fold)/n()) %>%
  filter(Chemical == "DTXSID8025337") %>%
  dplyr::select(-full_fold_errors)

fe_df %>%
  inner_join(suppressMessages(get_winning_model(my_pk))) %>%
  dplyr::filter(Detect %in% TRUE) %>%
  ggplot(aes(x = log2(Fold_Error))) +
  xlim(-4, 4) +
  geom_histogram(binwidth = 0.1) +
  facet_grid(cols = vars(model))

# Evaluate at level of Cmax and AUC RMSE
my_tkstats <- eval_tkstats(my_pk, method = "L-BFGS-B")
my_tkstats %>%
  dplyr::select(Chemical, Species, Route, Media, model, method,
                starts_with("Cmax"),
                starts_with("AUC")) %>%
  distinct() %>% 
  filter(is.finite(Cmax.tkstats), is.finite(AUC_inf.tkstats),
         is.finite(Cmax.nca), is.finite(AUC_inf.nca)) %>%
  ggplot(aes(x = Cmax.tkstats/Cmax.nca,
             y = AUC_inf.tkstats/AUC_inf.nca)) +
  geom_point() +
  geom_abline(slope = 1, color = "red4")

# Q: What do "great" fits have in common?

# Q: Does the dataset have a bias?
```

Here I do a bit of benchmarking for the parallel-ized version of do_fit

```{r do_fit-benchmarking}

# Set to 
single_core <- microbenchmark::microbenchmark(do_fit(my_pk, n_cores = 1), times = 3)
four_core <- microbenchmark::microbenchmark(do_fit(my_pk, n_cores = 4), times = 3)
# eight_core <- microbenchmark::microbenchmark(do_fit(my_pk, n_cores = 8), times = 3) # Only for four_DG
ten_core <- microbenchmark::microbenchmark(do_fit(my_pk, n_cores = 10), times = 3)
fourteen_core <- microbenchmark::microbenchmark(do_fit(my_pk, n_cores = 14), times = 3)

single_core
four_core
# eight_core
ten_core
fourteen_core

four_DG <- bind_rows(summary(single_core),
                     summary(four_core),
                     summary(eight_core),
                     summary(ten_core),
                     summary(fourteen_core)) # Done ~ 140s

eight_DG <- bind_rows(summary(single_core),
                      summary(four_core),
                      summary(ten_core),
                      summary(fourteen_core)) # Done

sixteen_DG <- bind_rows(summary(single_core),
                        summary(four_core),
                        summary(ten_core),
                        summary(fourteen_core)) # Done

my_pk %>% get_data() %>% dplyr::select(Chemical, Species) %>%
  distinct() %>% nrow()
my_pk %>% get_data() %>% pull(Chemical) %>% unique() -> used_chemicals

unused_cvt <- cvt %>%
  dplyr::filter(!(chemicals_analyzed.dsstox_substance_id %in% used_chemicals))

new_chemicals <- sample(unique(unused_cvt$chemicals_analyzed.dsstox_substance_id),
                       size = 8)

my_predata <- cvt %>%
  dplyr::filter(chemicals_analyzed.dsstox_substance_id %in% union(used_chemicals,
                                                                  new_chemicals))
my_predata %>%
  dplyr::select(chemicals_analyzed.dsstox_substance_id, subjects.species_harmonized) %>%
  distinct() %>% nrow()

my_pk <- pk(data = my_predata) +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = TRUE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = FALSE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species))
my_pk <- do_prefit(my_pk) # Then run again

##
##
# After getting all of them
full_df <- bind_rows(
  four_DG %>% mutate(data_group_N = 4),
  eight_DG %>% mutate(data_group_N = 8),
  sixteen_DG %>% mutate(data_group_N = 16)
)

full_df <- full_df %>%
  mutate(N_cores = as.numeric(str_extract(expr, pattern = "[:digit:]+")))

full_long <- full_df %>% 
         dplyr::select(expr, N_cores, data_group_N, min, median, max) %>%
         pivot_longer(cols = c(min, median, max)) %>%
  dplyr::select(-name)

df_long <- full_df %>% 
         dplyr::select(expr, N_cores, data_group_N, min, median, max) %>%
         pivot_longer(cols = c(min, median, max)) %>%
  dplyr::select(-name) %>% group_by(N_cores, data_group_N) %>%
  summarize(avg_val = mean(value),
            sd_val = sd(value))

ggplot(full_long,
       aes(x = N_cores,
           y = value,
           color = factor(data_group_N))) +

  geom_point() +
  geom_line(data = df_long,
            aes(x = N_cores,
                y = avg_val)) +
  scale_x_continuous(breaks = c(0, 1, 4, 8, 10, 14)) +
  labs(x = "Number of Cores",
       y = "Runtime (seconds)",
       color = "Data Groups",
       title = "Parallelization diminishes runtime at scale") +
  theme(panel.border = element_rect(color = "black", fill = NA, size = 1.5),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
        legend.key = element_rect(fill = "white"),
        axis.ticks = element_blank(),
        axis.line = element_blank())
ggsave(paste0(Sys.getenv("FIG_DIR"),
              "Parallelization_Benchmarking.png"),
       width = 4.5,
       height = 3.5,
       units = "in")

```

### Time table (refactoring convert_time)

```{r convert_time_table}

expand_grid(TimeA = time_units, TimeB = time_units) %>%
  mutate(A_fun = paste0("d", TimeA), B_fun = paste0("d", TimeB)) %>%
  rowwise() %>%
  mutate(conversion = 
           do.call(A_fun, list(1))/do.call(B_fun, list(1))) %>%
  dplyr::select(!c(A_fun, B_fun))

time_table <- convert_time_table()

time_table$TimeA %>% clipr::write_clip()
time_table$TimeB %>% clipr::write_clip()
time_table$conversion %>% MASS::fractions(max.denominator = 1000000000000000)

### Plotting for all (2.5MB file)
pl <- plot(my_pk, n_interp = 8, use_scale_conc = TRUE)
pdf(file = paste0(Sys.getenv("FIG_DIR"),
              "DoseNormalized_Pooled_AllFits.pdf"),
    height = 6, width = 10)
for (i in 1:nrow(pl)) {
  print(pl$final_plot[[i]])
}
dev.off()

# From my new SQL pull
pl <- plot(my_pk, n_interp = 8, use_scale_conc = TRUE)
pdf(file = paste0(Sys.getenv("FIG_DIR"),
              "NormalizedUpdate_NewCvT_SQLquery.pdf"),
    height = 6, width = 10)
for (i in 1:nrow(pl)) {
  print(pl$final_plot[[i]])
}
dev.off()

ggsave(paste0(Sys.getenv("FIG_DIR"),
              "DoseNormalized_DTXSID8025337.png"),
       units = "in",
       height = 5.05,
       width = 5.55,
       dpi = 200)

```