### 2023-10-05 data-raw/batchMode_trial_gpm.Rmd
Major reorganization of vignette to specify how each figure in manuscript was made.

---
title: "PK object batch mode trial"
author: "Gilberto Padilla Mercado"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
set.seed(2023)
library(tidyverse, quietly = TRUE)
devtools::load_all() # Loads the invivopkfit package in feature/pkS3obj_facet branch
library(cowplot)
library(RColorBrewer)
tidyverse_conflicts()
# Note I have now saved the RDS file for JOINT model with L-BFGS-B
my_pk <- readRDS("data-raw/all_cvt_pk_JOINT.rds")
```

# TOC

1. Pulling Data from CvTdb  
  - Minimal `pk` object  
2. Running all possible fitting options  
  - Evaluating fitting options  
  - Head-to-head comparison of pooled vs joint models  
3. Running dose-normalized joint fits  
4. Analysis Plots  
  - CvTdb replicate variation (Figure 4A)  
  - CvTdb replicate variation over ADME-normalized time (Figure 4B)  
  - CvTdb concentration and final time-point distributions (Supplementary Figure 1)  
  - invivoPKfit fit option selection (Supplementary Table 1)  
  - invivoPKfit model performance (Supplementary Figure 2)  
  - Model performance vs Data variability (Figure 5)  
  - Goodness of fit plots (Figure 6)  
  - Fits that may be improved (Supplementary Figure 3)  
5. Lombardo Analysis  
  - Chemicals included  
  - Comparison of derived TK stats  (Figure 7A)  
  - Histograms for Steady-state Volume of Dist. and Total Clearance (Supplemental Figure 4)
6. Benchmarking invivoPKfit
  - Parallelization speeds up the process of fitting (Supplementary Figure 5)

## Pulling data from CvTdb

See `pulling_iv_oral_cvtdb.Rmd`.

```{r pk_obj_fromSQL}

### Minimal PK Object ###----
# Minimum pk object, add options later
# Note that these mappings are now a default mappings, just being verbose here.
minimal_pk <- pk(data = cvt_df,
                 mapping = ggplot2::aes(
                   Chemical = analyte_dtxsid,
                   Chemical_Name = analyte_name_original,
                   DTXSID = analyte_dtxsid,
                   CASRN = analyte_casrn,
                   Species = species,
                   Reference = document_id,
                   Media = conc_medium_normalized,
                   Route = administration_route_normalized,
                   Dose = dose_level_normalized,
                   Dose.Units = "mg/kg",
                   Subject_ID = subject_id,
                   Series_ID = series_id,
                   Study_ID = study_id,
                   ConcTime_ID = conc_time_id,
                   N_Subjects = n_subjects_in_series,
                   Weight = weight_kg,
                   Weight.Units = "kg",
                   Time = time_hr,
                   Time.Units = "hours",
                   Value = conc,
                   Value.Units = "mg/L",
                   Value_SD = conc_sd_normalized,
                   LOQ = loq
                 ))
```

## Running all possible fitting options

This chunk is to run all the options through a
standard pipeline that calculates RMSE for Cmax and AUC_inf
for the Chemical Species Route Media grouping...

```{r pooled_fitting_choices}
# Types of Choices:
## Dose Normalized
## Log10 transform
## Scale time

# Pooled
my_pk_000p <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = FALSE, log10_trans = FALSE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species))

my_pk_100p <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = FALSE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species))

my_pk_110p <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = TRUE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species))

my_pk_111p <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = TRUE) +
  scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species))

my_pk_010p <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = FALSE, log10_trans = TRUE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species))

my_pk_011p <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = FALSE, log10_trans = TRUE) +
  scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species))

my_pk_001p <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = FALSE, log10_trans = FALSE) +
  scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species))

my_pk_101p <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = FALSE) +
  scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species))

```

```{r joint_fitting_choices}
# Joint
my_pk_000j <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = FALSE, log10_trans = FALSE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species, Reference))

my_pk_100j <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = FALSE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species, Reference))

my_pk_110j <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = TRUE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species, Reference))

my_pk_111j <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = TRUE) +
  scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species, Reference))

my_pk_010j <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = FALSE, log10_trans = TRUE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species, Reference))

my_pk_011j <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = FALSE, log10_trans = TRUE) +
  scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species, Reference))

my_pk_001j <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = FALSE, log10_trans = FALSE) +
  scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species, Reference))

my_pk_101j <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = FALSE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = FALSE) +
  scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species, Reference)) 
```

### Evaluating fitting options

```{r Evaluating_Choices}
# Evaluation
# Current: my_pk_111p

get_choices <- function(obj, PJ = "Pooled") {
  dose_norm <- obj$scales$conc$dose_norm
  log10_norm <- obj$scales$conc$log10_trans
  time_scale <- ifelse(obj$scales$time$new_units == "identity", FALSE, TRUE)

  my_choices <- paste0(dose_norm+0,
                       log10_norm+0,
                       time_scale+0,
                       " ", PJ)
  return(my_choices)
}

# Write into package
evaluate_choices <- function(obj, PJ = "Pooled") {
  pk_name <- get_choices(obj, PJ = PJ)

  cli::cli_progress_step("Winning Model...")
  # Wide winning model
  suppressMessages({
    winning_model <- get_winning_model(obj = obj) 
    wide_winning_model <- winning_model %>%
      group_by(method, model) %>%
      count() %>%
      pivot_wider(names_from = model, values_from = n) %>%
      mutate(Options = pk_name)
  })

  cli::cli_progress_step("Cmax and AUC_inf RMSEs...")
  # Cmax and AUC_inf RMSEs
  suppressMessages({
    RMSE_eval <- eval_tkstats(obj = obj) %>%
    mutate(Options = pk_name) %>%
    dplyr::select(Options,
                Chemical, Species,
                Route, Media, Reference,
                method, model,
                starts_with("Cmax"),
                starts_with("AUC_")) %>%
      filter(abs(AUC_inf.nca / AUC_tlast.nca) < 10,
             is.finite(AUC_inf.nca), is.finite(AUC_inf.tkstats),
             AUC_inf.nca > 0, AUC_inf.tkstats > 0) %>%
      rowwise() %>%
      mutate(SE_Cmax = (Cmax.tkstats - Cmax.nca)^2,
             SE_AUC_inf = (AUC_inf.tkstats - AUC_inf.nca)^2) %>%
      filter(!is.na(SE_Cmax), !is.na(SE_AUC_inf)) %>%
      group_by(Options, method) %>%
      summarize(RMSE_Cmax = sqrt(mean(SE_Cmax,
                                      na.rm = FALSE)) %>% signif(digits = 4),
                RMSE_AUC = sqrt(mean(SE_AUC_inf,
                                     na.rm = FALSE)) %>% signif(digits = 4))
  })

  cli::cli_progress_step("More detailed metrics...")
  # Other metrics (detailed)
  suppressMessages({
    r2_df <- rsq(obj, use_scale_conc = FALSE) %>%
    inner_join(winning_model)
  myAIC <- AIC(obj) %>%
    inner_join(winning_model)
  fold_errors_all <- fold_errors(obj) %>%
    inner_join(winning_model) %>%
    dplyr::filter(Detect %in% TRUE) %>%
    dplyr::group_by(!!!obj$data_group, model, method) %>%
    dplyr::summarize(avg_FoldErr = mean(Fold_Error, na.rm = TRUE) %>%
                       signif(digits = 4),
                     median_FoldErr = median(Fold_Error, na.rm = TRUE) %>%
                       signif(digits = 4),
                     within_2fold = sum(Fold_Error >= 0.5 & Fold_Error <= 2)/length(Fold_Error) %>%
                       signif(digits = 4))

  myRMSE_all <- rmse(obj,
                     rmse_group = vars(Route, Media, Dose),
                     use_scale_conc = FALSE) %>% 
    inner_join(winning_model) %>%
    dplyr::group_by(!!!obj$data_group, model, method) %>%
    dplyr::summarize(avg_MSE = mean(RMSE^2, na.rm = TRUE) %>%
                       signif(digits = 4),
                     median_MSE = median(RMSE^2, na.rm = TRUE) %>%
                       signif(digits = 4))

  my_df <- left_join(fold_errors_all, r2_df) %>%
    inner_join(myAIC) %>%
    inner_join(myRMSE_all) %>%
    mutate(Options = pk_name, .before = Chemical)

  full_list <- list(wide_winning_model, RMSE_eval, my_df)
  })

  return(full_list)
}
gc()

### Evaluations-----
my_pk_000p <- do_fit(my_pk_000p, n_cores = 14)  %>% evaluate_choices() # Done
my_pk_100p <- do_fit(my_pk_100p, n_cores = 8) %>% evaluate_choices() # Done
my_pk_110p <- do_fit(my_pk_110p, n_cores = 8) %>% evaluate_choices() # Done
my_pk_111p <- do_fit(my_pk_111p, n_cores = 8) %>% evaluate_choices() # Done
rm(my_pk_100p,my_pk_110p,my_pk_111p)
gc()
my_pk_010p <- do_fit(my_pk_010p, n_cores = 8) %>% evaluate_choices() # Done
my_pk_011p <- do_fit(my_pk_011p, n_cores = 8) %>% evaluate_choices() # Done 
my_pk_001p <- do_fit(my_pk_001p, n_cores = 8) %>% evaluate_choices() # Done
my_pk_101p <- do_fit(my_pk_101p, n_cores = 8) %>% evaluate_choices() # Done

my_pk_000j <- do_fit(my_pk_000j, n_cores = 8) %>% evaluate_choices(PJ = "Joint") # Done
my_pk_100j <- do_fit(my_pk_100j, n_cores = 8) %>% evaluate_choices(PJ = "Joint") # Done
my_pk_110j <- do_fit(my_pk_110j, n_cores = 8) %>% evaluate_choices(PJ = "Joint") # Done
my_pk_111j <- do_fit(my_pk_111j, n_cores = 8) %>% evaluate_choices(PJ = "Joint") # Done
rm(my_pk_000j, my_pk_100j, my_pk_110j, my_pk_111j)
gc()
my_pk_010j <- do_fit(my_pk_010j, n_cores = 8) %>% evaluate_choices(PJ = "Joint") # Done
my_pk_011j <- do_fit(my_pk_011j, n_cores = 8) %>% evaluate_choices(PJ = "Joint") # Done
my_pk_001j <- do_fit(my_pk_001j, n_cores = 8) %>% evaluate_choices(PJ = "Joint") # Done
my_pk_101j <- do_fit(my_pk_101j, n_cores = 8) %>% evaluate_choices(PJ = "Joint") # Done

# Run after each evaluation
# initialization (only run once)
{
winning_model_final <- NULL
rmse_summary <- NULL
detail_stats <- NULL
}

# Run all at once
{
current_pk <- my_pk_101j # Define current evaluated pk object here
winning_model_final <- bind_rows(winning_model_final, current_pk[[1]])
rmse_summary <- bind_rows(rmse_summary, current_pk[[2]])
detail_stats <- bind_rows(detail_stats, current_pk[[3]])
rm(current_pk)
gc()
}

# For copy + pasting into Excel workbook
winning_model_final %>% clipr::write_clip(col.names = FALSE)
clipr::clear_clip()
rmse_summary %>% clipr::write_clip(col.names = FALSE)
clipr::clear_clip()
detail_stats %>% clipr::write_clip(col.names = FALSE)
clipr::clear_clip()
gc()

```

## Head-to-head comparison of pooled vs joint models

Now I need to go back and re-analyze the data only for 100, 101, and 000
Pooled and Joint, L-BFGS-B only. I need to filter the cvt_df to only include Chem + Species combinations
with multiple references. Then I will join them with the excel sheet of already collected data.

```{r load_multirefs}
# Reference Multiples
multiRef_ChemSpec <- cvt_df %>%
  ungroup() %>%
  distinct(analyte_dtxsid, species, document_id) %>%
  group_by(analyte_dtxsid, species) %>%
  count() %>%
  arrange(desc(n)) %>%
  filter(n > 1) %>%
  dplyr::select(!n)
multiRef_ChemSpec %>% nrow() # Surprisingly only 13 Chem + Species have multiple references

```

```{r Plotting_Summary}
finalist_options <- c("000 Pooled", "000 Joint",
                      "100 Pooled", "100 Joint",
                      "101 Pooled", "101 Joint")

readxl::excel_sheets(paste(Sys.getenv("OD_DIR"),
                           "DataSummary_9_21_2023.xlsx",
                           sep = "/")) # Predictions Evaluation

preval <- readxl::read_xlsx(paste(Sys.getenv("OD_DIR"),
                                  "DataSummary_9_21_2023.xlsx",
                                  sep = "/"),
                            sheet = "Predictions Evaluation") %>%
  left_join(multiRef_ChemSpec,
            by = join_by(Chemical == analyte_dtxsid,
                         Species == species)) %>%
  filter(method %in% "L-BFGS-B", 
         Options %in% finalist_options) %>%
  distinct()

preval %>%
  separate(Options, sep = " ",
           into = c("Options", "Error_Model")) %>%
  ggplot(aes(x = avg_FoldErr,
             color = Options)) +
  geom_freqpoly(bins = 10) +
  facet_grid(cols = vars(Error_Model))

preval %>% 
  separate(Options, sep = " ",
           into = c("Options", "Error_Model")) %>%
  group_by(Options, Error_Model) %>%
  summarize(totalFoldError = sum(avg_FoldErr))

readxl::read_xlsx(paste(Sys.getenv("OD_DIR"),
                                  "DataSummary_9_21_2023.xlsx",
                                  sep = "/"),
                            sheet = "Fit_Counts") %>%
  filter(Options %in% finalist_options, method %in% "L-BFGS-B")

readxl::read_xlsx(paste(Sys.getenv("OD_DIR"),
                                  "DataSummary_9_21_2023.xlsx",
                                  sep = "/"),
                            sheet = "Cmax and AUC_inf Evaluation") %>%
  filter(Options %in% finalist_options, method %in% "L-BFGS-B")

library(broom)

filtered_preval <- preval %>%
  filter(if_all(where(is.numeric), is.finite))

filtered_preval %>%
  dplyr::select(Chemical, Species, model, method, Options, Rsq) %>%
  mutate(time_scaled = ifelse(str_detect(Options,
                                         pattern = "^\\d\\d1"),
                              "Time_Scaled",
                              "notTime_Scaled"),
         Options = str_remove(Options, pattern = "\\d(?=[:blank:])")) %>%
  # dplyr::select(-Options) %>%
  pivot_wider(names_from = time_scaled, values_from = Rsq) %>% 
  mutate(differ = Time_Scaled - notTime_Scaled) %>%
  View()

fit <- lm(within_2fold ~ avg_FoldErr, data = filtered_preval)
summary(fit)
tidy(fit)
head(augment(fit))
glance(fit)

filtered_preval %>% group_by(Options) %>%
  nest() %>%
  mutate(fit = map(data, \(x) {
    lm(within_2fold ~ avg_FoldErr, data = x)
    }),
    tidy_fit = map(fit, tidy)) %>%
  unnest(tidy_fit) %>%
  dplyr::select(Options, term, estimate, std.error, statistic) %>%
  View()

Cmax_AUC <- readxl::read_xlsx(paste("C:/Users/GPADILLA",
                                  "OneDrive - Environmental Protection Agency (EPA)",
                                  "DataSummary_8_18_2023.xlsx",
                                  sep = "/"),
                            sheet = "Cmax and AUC_inf Evaluation") %>%
  dplyr::select(1:4) %>%
  distinct()

Cmax_AUC %>%
  mutate(DFO = rank(sqrt(RMSE_Cmax^2 + signif(RMSE_AUC, 5)^2)),
         OLS = rank(abs(RMSE_Cmax + signif(RMSE_AUC, 5))/sqrt(2))) %>%
  #filter(DFO < 11 | OLS < 11) %>%
  arrange(DFO, OLS) %>%
  dplyr::select(-RMSE_Cmax, -RMSE_AUC) %>% View()
  clipr::write_clip()

Cmax_AUC %>%
  mutate(DFO = rank(sqrt(RMSE_Cmax^2 + RMSE_AUC^2)),
         OLS = rank(abs(RMSE_Cmax + RMSE_AUC)/sqrt(2))) %>%
  ggplot(aes(
    x = DFO,
    y = OLS,
    fill = Options
  )) + geom_point(shape = 21)

```

## Running dose-normalized joint fits

Dose-normalization and a hierarchical model were the options that produced the least
amount of error in predictions and derived TK stats.
In addition, _L-GFGS-B_ was the optimizer of choice

```{r Final_Options}
my_pk <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = TRUE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B")) +
  scale_conc(dose_norm = TRUE, log10_trans = FALSE) +
  stat_error_model(error_group = vars(Chemical, Species, Reference))

my_pk <- do_preprocess(my_pk)
cvt_DTXSIDs <- unique(my_pk$data$Chemical)

my_pk <- do_data_info(my_pk)

my_pk <- do_prefit(my_pk) # needs reframe somewhere

{
  start <- Sys.time()
  my_pk <- do_fit(my_pk, 
                  n_cores = 12)
  round(Sys.time() - start, 4)
}

```

## Analysis Plots

```{r data_aggregation}
# Need the following data for any of the subsequent plots

# after fitting pk object for all cvt objects or reading the fitted pk object in `setup`
winmodel <- get_winning_model(my_pk)
my_preds <- inner_join(predict(my_pk, use_scale_conc = TRUE), winmodel)
my_residuals <- residuals(my_pk, use_scale_conc = FALSE) %>%
  inner_join(winmodel)
my_tkstats <- eval_tkstats(my_pk)
my_nca <- get_nca(my_pk)
all_my_data <- get_data(my_pk)

fe_df <- fold_errors(my_pk, method = "L-BFGS-B") %>% inner_join(winmodel)
# May need to change the use_scale_conc defaults
r2_df <- rsq(my_pk, method = "L-BFGS-B", use_scale_conc = FALSE) %>%
  inner_join(winmodel)
myAIC <- AIC(my_pk, method = "L-BFGS-B") %>%
  inner_join(winmodel)
myRMSE <- rmse(my_pk, method = "L-BFGS-B",
               rmse_group = vars(Route, Media, Dose),
               use_scale_conc = FALSE) %>% 
  inner_join(winmodel)

# Something wrong with the summary method... maybe only use detects?
fold_errors_all <- fe_df %>%
  dplyr::filter(Detect %in% TRUE) %>%
  dplyr::group_by(!!!my_pk$data_group, model, method) %>%
  dplyr::summarize(avg_FoldErr = mean(Fold_Error, na.rm = TRUE),
                median_FoldErr = median(Fold_Error, na.rm = TRUE),
                within_2fold = sum(Fold_Error >= 0.5 & Fold_Error <= 2)/length(Fold_Error))

myRMSE_all <- myRMSE %>%
  dplyr::group_by(!!!my_pk$data_group, model, method) %>%
  dplyr::summarize(avg_MSE = mean(RMSE^2, na.rm = TRUE),
                median_MSE = median(RMSE^2, na.rm = TRUE))

```

### Figure 4A: Replicated timepoint variation in CvTdb
```{r fig4A}

all_my_data %>%
  filter(Detect %in% TRUE,
         exclude %in% FALSE,
         Dose > 0) %>%
  group_by(Chemical,
           Species,
           Reference,
           Route, 
           Media,
           Dose, Time) %>%
  filter(n() > 1,
         sum(is.na(Conc)) != n(),
         LOQ < Conc) %>%
  mutate(meanConc = mean(Conc, na.rm = TRUE)) %>%
  ungroup() %>%
  rowwise() %>%
  mutate(foldConc = Conc / meanConc) %>%
  filter(Route == "iv") %>%
  count(foldConc < 0.5) %>%
  {.[[2,2]]/(.[[1,2]] + .[[2,2]])}
# Stats for ALL | ORAL | IV
# 88.64% inside twofold | 87.20%  | 91.23%
# 2.23%  above twofold  | 2.52%   | 1.70%
# 9.11%  below twofold  | 10.25%  | 7.04%

pl5A <- all_my_data %>%
  filter(Detect %in% TRUE,
         exclude %in% FALSE,
         Dose > 0) %>%
  group_by(Chemical,
           Species,
           Reference,
           Route, 
           Media,
           Dose, Time) %>%
  filter(n() > 1,
         sum(is.na(Conc)) != n(),
         LOQ < Conc) %>%
  mutate(meanConc = mean(Conc, na.rm = TRUE)) %>%
  ungroup() %>%
  rowwise() %>%
  mutate(foldConc = Conc / meanConc) %>%
  ggplot(aes(x = foldConc)) +
  geom_histogram(binwidth = 0.2,
                 fill = "grey5",
                 color = "grey5",
                 linewidth = 0.4) +
  coord_cartesian(xlim =  c(0, 4),
                  ylim = c(0, 4000)) +
  scale_y_continuous(expand = c(0, 0),
                     breaks = c(0, 1000, 2000, 3000),
                     labels = c("0", "1", "2", "3")) +
  expand_limits(y = 0) +
  geom_vline(xintercept = c(0.5, 2), color = "red3", linetype = "dashed") +
  theme_classic() +
  labs(x = "Mean-normalized\nconcentration",
                y = "Observations\n(thousands)") +
  theme(text = element_text(size = 10),
        panel.border = element_rect(color = "black", linewidth = 1, fill = NA),
        panel.background = element_blank(),
        panel.grid.major.y = element_line(color = "grey90", linewidth = 0.4),
        axis.title = element_text(face = "bold"),
        axis.line = element_blank(),
        plot.background = element_blank(),
        axis.ticks.y = element_blank())

ggsave(paste0(Sys.getenv("FIG_DIR"),
              "NormalizedConcentrations_2023-09-29.png"),
       plot = pl5A,
       height = 2.2,
       width = 2.5,
       units = "in",
       device = "png",
       dpi = 300)

```

### Figure 4B: Variation over time in CvTdb data

```{r fig4B-cowplot}

pl5B <- all_my_data %>%
  group_by(Chemical,
           Species,
           Reference,
           Route,
           Dose, Time) %>%
  filter(n() > 1,
         sum(is.na(Conc_trans)) != n(),
         exclude == FALSE) %>%
  mutate(meanConc = mean(Conc_trans, na.rm = TRUE)) %>%
  ungroup() %>%
  rowwise() %>%
  mutate(foldConc = Conc_trans / meanConc) %>%
  inner_join(my_tkstats %>%
              dplyr::select(Chemical, Species,
                            Route, Media,
                            tmax = tmax.nca) %>%
              filter(!is.na(tmax))) %>%
  group_by(Chemical,
           Species,
           Reference,
           Route,
           Media,
           Dose) %>%
  mutate(
         normTime = ifelse(
           Route == "iv",
           1 + Time/max(Time),
           ifelse(
             Time > tmax,
             ((Time - tmax)/max(Time)) + 1,
             0.5 + Time/(2*tmax)
           )
         )) %>%
  ggplot(aes(
    x = normTime,
    y = foldConc
  )) +
  geom_point(alpha = 0.1, size = 0.7) +
  geom_vline(xintercept = 1, color = "blue",
             linewidth = 0.8, linetype = "dashed") +
  geom_vline(xintercept = 2, color = "green4",
             linewidth = 0.8, linetype = "dashed") +
  geom_hline(yintercept = c(0.5, 2), color = "red3",
             linewidth = 0.8, linetype = "dashed") +
  facet_grid(cols = vars(Route),
             scales = "free_x") +
  labs(x = "ADME-Normalized Time",
       y = "Mean-Normalized\nConcentration") +
  theme(text = element_text(size = 10),
        panel.border = element_rect(color = "black", fill = NA, size = 1),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(face = "bold"),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        axis.text.x = element_blank(),
        axis.title = element_text(face = "bold"),
        panel.spacing = unit(0.125, units = "in"),
        plot.background = element_blank(),
        legend.position = "none")

ggsave(paste0(Sys.getenv("FIG_DIR"),
              "ADME_NormTime-NormConcs_onlyDetects_2023-09-29.png"),
       height = 4,
       width = 5,
       units = "in",
       dpi = 300)

# Make the Combination Plot----

plot_grid(pl5A, pl5B, labels = c("A", "B"),
                     align = "h", axis = "bt", rel_widths = c(1, 1.5))

figure5 <- plot_grid(pl5A, pl5B, labels = c("A", "B"),
                     align = "h", axis = "bt", rel_widths = c(1, 1.5))

ggsave(paste0(Sys.getenv("FIG_DIR"),
              "Figure5_onlyDetects_2023-10-02.png"),
       plot = figure5,
       bg = "white",
       height = 3.25,
       width = 6.5,
       units = "in",
       dpi = 300)

```

### Supp. Figure 1: Concentration and final time-points have wide range

```{r cvt_analysis}
data_range <- all_my_data %>%
  filter(Detect) %>%
  group_by(Chemical, Species, Reference, Media, Route, Dose, Dose.Units) %>%
  summarize(maxT = max(Time)/24,
         concRange = log10(max(Conc)/min(Conc))) %>%
  ungroup()

sf_2B <- data_range %>%
  ggplot(aes(x = maxT)) +
  geom_histogram(fill = "grey5",
                 bins = 40,
                 color = "grey5") +
  scale_x_log10(labels = scales::number,
                breaks = c(0.1, 1, 7, 30, 365)) +
  scale_y_continuous(expand = c(0, 0),
                     limits = c(0, 60),
                     breaks = c(0, 10, 20, 30, 40, 50)) +
  labs(x = "Day of final Timepoint",
       y = "Count") +
  theme(aspect.ratio = 1,
        text = element_text(size = 10),
        panel.border = element_rect(color = "black", fill = NA, size = 1),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
        axis.ticks = element_blank(),
        axis.line = element_blank())

ggsave(paste0(Sys.getenv("FIG_DIR"),
              "SFig2B_TimePoints_2023-10-02.png"),
       height = 4,
       width = 4,
       units = "in",
       dpi = 300)

sf_2A <- data_range %>%
  ggplot(aes(x = concRange)) +
  geom_histogram(fill = "grey5",
                 bins = 40,
                 color = "grey5") +
  scale_y_continuous(expand = c(0, 0),
                     limits = c(0, 25),
                     breaks = c(0, 5, 10, 15, 20)) +
  labs(x = "log10(Concentration Range)",
       y = "Count") +
  theme(aspect.ratio = 1,
        panel.border = element_rect(color = "black", fill = NA, size = 1),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(face = "bold"),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 11))

ggsave(paste0(Sys.getenv("FIG_DIR"),
              "SFig2A_ConcRange_2023-09-29.png"),
       height = 4,
       width = 4,
       units = "in",
       dpi = 300)

plot_grid(sf_2A, sf_2B, labels = "AUTO", align = "h", axis = "bt")

ggsave(paste0(Sys.getenv("FIG_DIR"),
              "SFig2_ConcTimeRanges_2023-10-02.png"),
       height = 3.25,
       width = 6,
       units = "in",
       dpi = 300)

```

### Supp Figure 2: invivoPKfit model performance

```{r sFig2-model_performance}
my_preds %>%
  right_join(winmodel) %>%
  rowwise() %>%
  mutate(Pred_Obs = Conc_trans/Conc_est) %>%
  filter(is.finite(Pred_Obs),
         model %in% c("model_1comp", "model_2comp")) %>%
  # mutate(pass = ifelse(Pred_Obs < 0.5, TRUE, FALSE)) %>%
  # pull(pass) %>% table()
  ggplot(aes(x = Pred_Obs)) +
  geom_histogram(binwidth = 0.5,
                 fill = "grey5") +
  scale_x_log10(limits = c(NA, 1E6)) +
  expand_limits(y = 0) +
  geom_vline(xintercept = c(0.5, 2), color = "red3", linetype = "dashed") +
  facet_grid(cols = vars(model),
             rows = vars(Route)) +
  labs(x = "Dose-normalized\nPredicted/Observed",
       y = "Number of Observations") +
  theme_classic() +
  theme(aspect.ratio = 1,
        panel.border = element_rect(color = "black", fill = NA),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.4))

# 91.1% of Pred_Obs are finite
# 16.6% Are below two-fold
# 33.2%  Are above two-fold
# 50.2% Are within twofold

ggsave(paste0(Sys.getenv("FIG_DIR"),
              "PredictedObserved_compartmentModels_2023-10-02.png"),
       height = 6, width = 6)

```

### Figure 5: Model performance vs Data Variability

```{r fig5-modelPerform-v-dataVar}

pl_7A <- my_preds %>%
  group_by(Chemical,
           Species,
           Route, Media,
           Dose, Time) %>%
  filter(n() > 1,
         sum(is.na(Conc_trans)) != n(),
         # model %in% c("model_flat"),
         exclude %in% FALSE,
         Detect %in% TRUE) %>%
  mutate(meanConc = mean(Conc_trans, na.rm = TRUE)) %>%
  ungroup() %>%
  rowwise() %>%
  mutate(foldConc = Conc_trans / meanConc,
         foldPred = Conc_est / Conc_trans) %>% 
  # For percentages
  # ungroup() %>%
  # filter(model %in% c("model_flat")) %>%
  # count((foldPred >= 0.5 & foldPred <= 2) & !(foldConc >= 0.5 & foldConc <= 2)) %>%
  # {100*.[[2,2]]/(.[[1,2]] + .[[2,2]])}
  # For Plotting
  ggplot(aes(
    y = foldPred,
    x = foldConc
  )) +
  geom_bin2d(bins = 40, color = NA) +
  geom_hline(yintercept = c(0.5, 2), linetype = "dashed") +
  geom_vline(xintercept = c(0.5, 2), linetype = "dashed") +
  facet_grid(cols = vars(model)) +
  scale_x_log10(limits = c(0.0001, 100)) +
  scale_y_log10(limits = c(0.001, 1000)) +
  scale_fill_viridis_c(option = "cividis",
                       limits = c(1, 100),
                       oob = oob_squish) +
  labs(y = "Model Error",
       x = "Data Variability") +
  theme(aspect.ratio = 1,
        panel.border = element_rect(color = "black", fill = NA, size = 1.5),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(face = "bold"),
        axis.ticks = element_blank(),
        axis.line = element_blank())

ggsave(paste0(Sys.getenv("FIG_DIR"),
              "ModelPerformance_vDataVariability_2023-10-02.png"),
       height = 3,
       width = 6.5,
       device = "png",
       dpi = 300,
       units = "in")

my_table <- data.frame(
  c("Both within a factor of 2",
    "Model too variable",
    "Data too variable",
    "Both outside a factor of 2"),
  c(46.3, 39.5, 2.3, 12),
  c(46, 44, 1.8, 7.4),
  c(26, 60, 5, 8.5),
  c(45, 42.7, 2.3, 10))

names(my_table) <- c(" ","1-compartment (%)", "2-compartment (%)", "Flat (%)", "Overall (%)")

library(flextable)
library(officer)

flextable(my_table) %>%
  border_inner() %>%
  fontsize(part = "all", size = 11) %>%
  bold(part = "all", j = 5)

table_plot <- gen_grob(flextable(my_table) %>%
                         border_inner() %>%
                         fontsize(part = "all", size = 10) %>%
                         bold(part = "all", j = 5) %>%
                         autofit())

dual_plot <- plot_grid(pl_7A, table_plot, ncol = 1, align = "v",
          rel_heights = c(1, 0.5))

ggsave(paste0(Sys.getenv("FIG_DIR"),
              "ModelPerformance_vDataVariability_wTable_2023-10-02.png"),
       plot = dual_plot,
       height = 4.5,
       width = 6.5,
       device = "png",
       bg = "white",
       dpi = 300,
       units = "in")  
```

### Figure 6: Multiple goodness-of-fit metrics validate model performance
This should be it's own figure

```{r fig6-goodness-of-fit}

my_df <- left_join(fold_errors_all, r2_df) %>% inner_join(myAIC) %>% inner_join(myRMSE_all)

mypl <- ggplot(data = my_df,
               aes(
                 x = within_2fold,
                 y = Rsq,
                 color = model
               )) +
  geom_point() +
  geom_abline(slope = 1, color = "red4", linetype = "longdash") +
  scale_color_manual(values = c("#0398FC", "#D68E09", "black")) +
  labs(x = "% Fold Error within 2-fold",
       y = "R-squared value") +
  theme(aspect.ratio = 1,
        panel.border = element_rect(color = "black", fill = NA, size = 1),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        legend.position = "none",
        legend.title = element_blank(),
        legend.key = element_blank())
panelA_plot <- ggExtra::ggMarginal(mypl, groupFill = TRUE,
                    type = "histogram",
                    xparams = list(binwidth = 0.05),
                    yparams = list(binwidth = 0.05))
ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "R2_within2fold_comparison.png"),
       plot = ggExtra::ggMarginal(mypl, groupFill = TRUE,
                    type = "histogram",
                    xparams = list(binwidth = 0.05),
                    yparams = list(binwidth = 0.05)),
       height = 5,
       width = 5,
       units = "in")

panelB_plot <- ggplot(data = myRMSE %>%
         mutate(RMSE = ifelse(RMSE == 0, NA, RMSE)),
               aes(
                 x = log10(RMSE),
                 color = model,
                 fill = model
               )) + 
  geom_histogram(bins = 100) +
  labs(y = "Count") +
  scale_color_manual(values = c("#0398FC", "#D68E09", "grey10")) +
  scale_fill_manual(values = c("#0398FC", "#D68E09", "grey10")) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 4),
                     expand = c(0,NA)) +
  facet_grid(rows = vars(model),
             cols = vars(Route)) +
  theme(text = element_text(size = 10),
        aspect.ratio = 0.6,
    panel.border = element_rect(color = "black", fill = NA, size = 1),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        strip.background = element_blank(),
    panel.spacing.y = unit(0.125, units = "in"),
        legend.position = "bottom",
        legend.title = element_blank())

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RMSEs_winningmodel_LBFGSB_zero1E-6_RouteFacet_notNormalized.png"),
       height = 4,
       width = 6,
       units = "in")

legend <- get_legend(panelB_plot)

plot_grid(plot_grid(panelA_plot,
                   panelB_plot + theme(legend.position = "none"), labels = c("A", "B"),
                   axis = "tb", align = "h", rel_heights = c(1,1)), legend,
          ncol = 1,
          rel_heights = c(1, 0.1))
ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "CombinedPlot_Rsquared_RMSE.png"),
       height = 4,
       width = 6.5,
       bg = "white",
       units = "in")
```

### Supp. Fig. 3: Distribution of Sigmas in CvTdb data

```{r sFig3-sigmaDistribution}

my_coefs <- inner_join(get_winning_model(my_pk), coef(my_pk))
my_coefs %>% inner_join(get_data_summary(my_pk)) -> my_coefs

my_coefs %>%
  mutate(model = 
           case_when(
             model == "model_1comp" ~ "1-compartment",
             model == "model_2comp" ~ "2-compartment",
             model == "model_flat"  ~ "Null"
           )) %>%
  ggplot(aes(x = sigma_value/n_obs)) +
  geom_histogram(bins = 15,
                 color = NA, fill = "grey5") +
  scale_x_log10(labels = label_log()) +
  facet_grid(cols = vars(model)) +
  labs(x = "Size-normalized Sigma Value",
       y = "Number of Observations") +
  theme(panel.border = element_rect(color = "black", fill = NA, size = 1.5),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(face = "bold"),
        plot.title = element_text(hjust = 0.5),
        axis.ticks = element_blank(),
        axis.line = element_blank())

ggsave(paste0(Sys.getenv("FIG_DIR"),
              "NormalizedSigma_Distributions_2023-07.png"),
       height = 3, width = 5)

```

### Figure 7: Comparing derived TK stats with human TK stats from Lombardo et al.

```{r Lombardo_Comparison}
cvt_DTXSIDs <- unique(my_pk$data$Chemical)
# First things first, let's load the data from Lombardo et al.
lombardo <- readxl::read_xlsx(
  paste0(Sys.getenv("FIG_DIR"),
         "Lombardo2018-Supplemental_82966_revised_corrected.xlsx"),
  skip = 8)

l_batch_search <- readxl::read_xlsx(
  paste0(Sys.getenv("FIG_DIR"),
         "LombardoBatchSearch_Name.xlsx"),
  sheet = "Main Data"
)

l_batch_search <- l_batch_search[c("INPUT", "CASRN","DTXSID")] %>%
  distinct() %>%
  right_join(lombardo[c("Name", "CAS #")], by = c("CASRN" = "CAS #", "INPUT" = "Name"))

l_noNamesCASRN <- readxl::read_xlsx(
  paste0(Sys.getenv("FIG_DIR"),
         "LombardoBatchSearch_CASRNnoName.xlsx"),
  sheet = "Main Data"
) %>%
  dplyr::select(CASRN = INPUT, DTXSID)

l_curatedDTX <- read_tsv(paste0(Sys.getenv("FIG_DIR"),
                         "CuratedIDs_lombardo.txt")) %>%
  rename(INPUT = Name)

l_batch_search <- l_batch_search %>%
  left_join(l_noNamesCASRN, by = "CASRN") %>%
  mutate(DTXSID = coalesce(DTXSID.x, DTXSID.y)) %>%
  dplyr::select(-DTXSID.x, -DTXSID.y) %>%
  left_join(l_curatedDTX, by = "INPUT") %>%
  mutate(DTXSID = coalesce(DTXSID.x, DTXSID.y)) %>%
  dplyr::select(-DTXSID.x, -DTXSID.y)
# Now all but two of the IDs are unidentified

lombardo <- lombardo %>%
  right_join(l_batch_search, by = c("Name" = "INPUT"))

lombard_abbr <- lombardo %>% dplyr::select(DTXSID,
                                           hVss = `human VDss (L/kg)`,
                                           hCltot = `human CL (mL/min/kg)`,
                                           halflife = `terminal  t1/2 (h)` ) %>%
  filter(DTXSID %in% cvt_DTXSIDs)

# 15 chemicals in common with values for both
lombardo_comparison <- my_tkstats %>%
  dplyr::select(DTXSID = Chemical, Species,
                model, Vss.tkstats, CLtot.tkstats) %>%
  distinct() %>%
  inner_join(lombard_abbr, by = "DTXSID") %>%
  filter(!is.na(Vss.tkstats)) %>%
  arrange(halflife)

comparable_ids <- lombardo_comparison %>% pull(DTXSID)

lombard_abbr <- lombard_abbr %>%
  mutate(Species = "human", model = "lombardo") %>%
  rename(Vss = hVss, CLtot = hCltot) %>%
  mutate(CLtot = CLtot*60/1000) %>% # Lombardo reports this as mL/min/kg, need L/h/kg
  filter(DTXSID %in% comparable_ids)

lombardo_comparison <- my_tkstats %>%
  dplyr::select(DTXSID = Chemical, Species,
                model, Vss.tkstats, CLtot.tkstats, halflife.tkstats) %>%
  filter(DTXSID %in% comparable_ids) %>%
  distinct() %>%
  rename(Vss = Vss.tkstats, CLtot = CLtot.tkstats, halflife = halflife.tkstats) %>%
  bind_rows(lombard_abbr) %>%
  mutate(model = ifelse(str_detect(model, "^model"), "invivoPKfit", model),
         DTXSID = factor(DTXSID, levels = unique(comparable_ids)))

lombardo_comparison %>%
  pivot_longer(cols = c("Vss", "CLtot", "halflife")) %>%
  mutate(name = factor(name, levels = c("halflife", "Vss", "CLtot"))) %>%
  ggplot(mapping = aes(
    x = value,
    y = DTXSID
  )) +
  geom_point(aes(color = model,
                 shape = Species),
             position = position_dodge(0.7),
             size = 5/.pt, stroke = 2.5/.pt) +
  facet_grid(cols = vars(name),
             scales = "free_x") +
  scale_shape_manual(values = c(21, 22, 24),
                     breaks = c("human", "dog", "rat")) +
  scale_color_manual(values = c("black", "#1064c9")) +
  guides(color = guide_legend(override.aes = list(shape = 21),
                              nrow = 2,
                              title.position = "top",
                              title.hjust = 0.5),
         shape = guide_legend(nrow = 1,
                              title.position = "top",
                              title.hjust = 0.5)) +
  theme(panel.border = element_rect(color = "black", fill = NA, linewidth = 1.5),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(face = "bold", size = 12),
        text = element_text(size = 10),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        axis.text = element_text(face = "bold"),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "bottom",
        legend.key = element_blank(),
        legend.title = element_text(face = "bold"),
        legend.text = element_text())

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "lombardoComparison_Figure8.png"),
       height = 4,
       width = 6.5,
       device = "png", dpi = 300)

```

### Supp. Fig. 4 Histogram of Lombardo Vss and InvivoPKfit & another meta-analysis

```{r otherMeta-analysis}

# Histograms
lombardo_hist <- data.frame(Chemical = c(lombardo$DTXSID,
                              my_tkstats$Chemical),
                            Vss = c(lombardo$`human VDss (L/kg)`,
                              my_tkstats$Vss.tkstats),
                            halflife = c(lombardo$`terminal  t1/2 (h)`,
                              my_tkstats$halflife.tkstats),
                            Source = c(rep_len("lombardo",
                                    length.out = nrow(lombardo)),
                              rep_len("invivoPKfit",
                                      length.out = nrow(my_tkstats))))

sfig4A <- lombardo_hist %>% ggplot(aes(x = log2(Vss), color = Source,
                             group = Source)) +
  geom_freqpoly(aes(y = after_stat(ndensity)),
                linewidth = 2/.pt,
                bins = 20) +
  labs(y = "Scaled Frequency") +
  scale_color_manual(values = c("black", "#1064c9")) +
  guides(color = guide_legend(nrow = 2, linewidth = 1)) +
  theme(panel.border = element_rect(color = "black", fill = NA, linewidth = 1.5),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(face = "bold", size = 12),
        text = element_text(size = 10),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        axis.text = element_text(face = "bold"),
        legend.position = "bottom",
        legend.key = element_blank(),
        legend.title = element_text(face = "bold"),
        legend.text = element_text())

other_meta <- read_csv(
  file = paste0(Sys.getenv("FIG_DIR"), "SupTableInVivoDatandPreds.csv")) %>%
  dplyr::select(DTXSID, fbioh, fbior) %>%
  filter(DTXSID %in% my_tkstats$Chemical)

sfig4B <- other_meta %>% left_join(my_tkstats, by = c("DTXSID" = "Chemical")) %>%
  filter(Species %in% c("rat")) %>%
  filter(!is.na(Fgutabs.tkstats),
         !is.na(fbior)) %>%
  distinct(DTXSID, Species,
           Fgutabs.tkstats, fbior) %>%
  pivot_longer(cols = c(Fgutabs.tkstats, fbior),
               names_to = "Source",
               values_to = "Fgutabs") %>%
    mutate(Source = ifelse(str_detect(Source, "tkstats"), "invivoPKfit",
                           "Caco-2 Paper")) %>%
    ggplot(mapping = aes(
      x = Fgutabs,
      y = reorder(DTXSID, Fgutabs)
    )) +
    geom_point(aes(color = Source),
               shape = 24,
               position = position_dodge(0.7),
               size = 5/.pt, stroke = 2.5/.pt) +
    scale_color_manual(values = c("black", "green4"),
                       breaks = c("invivoPKfit", "Caco-2 Paper")) +
  guides(color = guide_legend(nrow = 2)) +
    theme(panel.border = element_rect(color = "black", fill = NA, linewidth = 1.5),
          panel.background = element_blank(),
          panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
          strip.background = element_rect(fill = "white"),
          strip.text = element_text(face = "bold", size = 12),
          text = element_text(size = 10),
          axis.ticks = element_blank(),
          axis.line = element_blank(),
          axis.text = element_text(face = "bold"),
          axis.title.y = element_blank(),
          legend.position = "bottom",
          legend.key = element_blank(),
          legend.title = element_text(face = "bold"),
          legend.text = element_text())

plot_grid(sfig4A, sfig4B, align = "h", axis = "bt")
ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "metaAnalysis_Comparisons_SuppFigure4.png"),
       height = 3,
       width = 5,
       device = "png", dpi = 300)

```

### Supp. Fig. 5: Parallelization decreases runtime of invivoPKfit

Here I do a bit of benchmarking for the parallel-ized version of do_fit

```{r do_fit-benchmarking}

# Set to 
single_core <- microbenchmark::microbenchmark(do_fit(my_pk, n_cores = 1), times = 3)
four_core <- microbenchmark::microbenchmark(do_fit(my_pk, n_cores = 4), times = 3)
# eight_core <- microbenchmark::microbenchmark(do_fit(my_pk, n_cores = 8), times = 3) # Only for four_DG
ten_core <- microbenchmark::microbenchmark(do_fit(my_pk, n_cores = 10), times = 3)
fourteen_core <- microbenchmark::microbenchmark(do_fit(my_pk, n_cores = 14), times = 3)

single_core
four_core
# eight_core
ten_core
fourteen_core

four_DG <- bind_rows(summary(single_core),
                     summary(four_core),
                     summary(eight_core),
                     summary(ten_core),
                     summary(fourteen_core)) # Done ~ 140s

eight_DG <- bind_rows(summary(single_core),
                      summary(four_core),
                      summary(ten_core),
                      summary(fourteen_core)) # Done

sixteen_DG <- bind_rows(summary(single_core),
                        summary(four_core),
                        summary(ten_core),
                        summary(fourteen_core)) # Done

my_pk %>% get_data() %>% dplyr::select(Chemical, Species) %>%
  distinct() %>% nrow()
my_pk %>% get_data() %>% pull(Chemical) %>% unique() -> used_chemicals

unused_cvt <- cvt %>%
  dplyr::filter(!(chemicals_analyzed.dsstox_substance_id %in% used_chemicals))

new_chemicals <- sample(unique(unused_cvt$chemicals_analyzed.dsstox_substance_id),
                       size = 8)

my_predata <- cvt %>%
  dplyr::filter(chemicals_analyzed.dsstox_substance_id %in% union(used_chemicals,
                                                                  new_chemicals))
my_predata %>%
  dplyr::select(chemicals_analyzed.dsstox_substance_id, subjects.species_harmonized) %>%
  distinct() %>% nrow()

my_pk <- pk(data = my_predata) +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = TRUE,
                      suppress.messages = TRUE) +
  settings_optimx(method = c("L-BFGS-B", "bobyqa")) +
  scale_conc(dose_norm = TRUE, log10_trans = FALSE) +
  # scale_time(new_units = "auto") +
  # stat_model() + 
  stat_error_model(error_group = vars(Chemical, Species))
my_pk <- do_prefit(my_pk) # Then run again

##
##
# After getting all of them
full_df <- bind_rows(
  four_DG %>% mutate(data_group_N = 4),
  eight_DG %>% mutate(data_group_N = 8),
  sixteen_DG %>% mutate(data_group_N = 16)
)

full_df <- full_df %>%
  mutate(N_cores = as.numeric(str_extract(expr, pattern = "[:digit:]+")))

full_long <- full_df %>% 
         dplyr::select(expr, N_cores, data_group_N, min, median, max) %>%
         pivot_longer(cols = c(min, median, max)) %>%
  dplyr::select(-name)

df_long <- full_df %>% 
         dplyr::select(expr, N_cores, data_group_N, min, median, max) %>%
         pivot_longer(cols = c(min, median, max)) %>%
  dplyr::select(-name) %>% group_by(N_cores, data_group_N) %>%
  summarize(avg_val = mean(value),
            sd_val = sd(value))

ggplot(full_long,
       aes(x = N_cores,
           y = value,
           color = factor(data_group_N))) +

  geom_point() +
  geom_line(data = df_long,
            aes(x = N_cores,
                y = avg_val)) +
  scale_x_continuous(breaks = c(0, 1, 4, 8, 10, 14)) +
  labs(x = "Number of Cores",
       y = "Runtime (seconds)",
       color = "Data Groups",
       title = "Parallelization diminishes runtime at scale") +
  theme(panel.border = element_rect(color = "black", fill = NA, size = 1.5),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
        legend.key = element_rect(fill = "white"),
        axis.ticks = element_blank(),
        axis.line = element_blank())
ggsave(paste0(Sys.getenv("FIG_DIR"),
              "Parallelization_Benchmarking.png"),
       width = 4.5,
       height = 3.5,
       units = "in")

```

### Supp. Fig. 6: Example plots for data fits

```{r sfig6-examples-of-fits}
pl <- plot(my_pk, use_scale_conc = TRUE, n_interp = 10)
names(pl) # Chemical Species observations oservation_plot predicted predicted_plot final_plot

s6_plA <- pl %>% filter(Chemical %in% "DTXSID90382620") %>%
  pull(observation_plot) %>% .[[1]]
s6_plB_mouse <- pl %>% filter(Chemical %in% "DTXSID4020533",
                              Species %in% "mouse") %>% pull(final_plot) %>% .[[1]]
s6_plB_rat <- pl %>% filter(Chemical %in% "DTXSID4020533",
                              Species %in% "rat") %>% pull(final_plot) %>% .[[1]]
s6_plC <- pl %>% filter(Chemical %in% "DTXSID8021359") %>%
  pull(final_plot) %>% .[[1]]
s6_plD <- pl %>% filter(Chemical %in% "DTXSID8025337") %>%
  pull(final_plot) %>% .[[1]]

# Set colors palletes for each
# panel A : Oranges
panelA_cols <- brewer.pal(8, "Oranges")[3:8]
# panel B : Greens
panelB_cols <- brewer.pal(9, "Greens")[2:9]
# panel C : Blue (only 2 values) 
# panel D : Purple (6 values)
panelD_cols <- brewer.pal(9, "Purples")[4:9]

s6_plA <- s6_plA + 
  labs(title = "fluorotelomer acid 7+3") +
  scale_color_manual(values = panelA_cols, aesthetics = c("color", "fill")) +
  theme(text = element_text(size = 20),
        legend.position = "none",
        strip.background = element_rect(fill = "white"))

s6_plB_mouse <- s6_plB_mouse + 
  labs(title = "1,4-dioxane\nMouse") +
  scale_color_manual(values = panelB_cols[c(5,8)],
                     aesthetics = c("color", "fill")) +
  theme(text = element_text(size = 20),
        legend.position = "none",
        strip.background = element_rect(fill = "white"))

s6_plB_rat <- s6_plB_rat + 
  labs(title = "1,4-dioxane\nRat") +
  scale_color_manual(values = panelB_cols[c(1, 2, 3, 4, 6,8)],
                     aesthetics = c("color", "fill")) +
  theme(text = element_text(size = 20),
        legend.position = "none",
        strip.background = element_rect(fill = "white"))

s6_plC <- s6_plC +
  labs(title = "tolbutamide") +
  scale_color_manual(values = c("#6BAED6", "#08519C"),
                     aesthetics = c("color", "fill")) +
  theme(text = element_text(size = 20),
        legend.position = "none",
        strip.background = element_rect(fill = "white"))

s6_plD <- s6_plD +
  labs(title = "formamide") +
  scale_color_manual(values = panelD_cols,
                     aesthetics = c("color", "fill")) +
  theme(text = element_text(size = 20),
        legend.position = "none",
        strip.background = element_rect(fill = "white"))

# Put it all together
s6_plB <- plot_grid(s6_plB_mouse, s6_plB_rat,
            align = "h", axis = "bt",
           rel_widths = c(1.2,2))

s6_plCD <- plot_grid(s6_plC, s6_plD, labels = c("C", "D"), label_size = 24)

plot_grid(
  s6_plA,
  s6_plB,
  s6_plCD,
  align = "v",
  ncol = 1,
  rel_heights = c(1.25, 1, 1.25),
  labels = c("A", "B", ""),
  label_size = 24
)

# Am making this 2X the size I need it to be 
# so the points are a reasonable size when I scale it down
ggsave(paste0(Sys.getenv("FIG_DIR"),
              "ExampleFitsPlots.png"),
       width = 13,
       height = 14,
       units = "in")

```

The purpose of the following code is to generate data.frames with information for
all 103 chemicals in the `cvt` dataset that comes with `invivopkfit`.
I will need to do append each dataframe to the last for the outputs of the following
methods:

- `AIC()`
- `BIC()` left-joined with `AIC` output
- `get_tkstats()`

I will also have to fit these iteratively and manage the memory as part of the loop.

```{r more_analysis}

my_pk$prefit$fit_check %>% filter(fit_decision == "abort") %>%
  dplyr::select(model, Chemical) %>% distinct()

all_my_data %>%
  filter(Chemical == "DTXSID2021315")

# On average, how many time-points before tmax (empirical)?
my_nca %>% filter(param_name %in% "tmax") %>%
  dplyr::select(Chemical, Species, Route, Media, tmax.nca = param_value) %>%
  left_join(all_my_data) %>%
  filter(Route %in% "oral", Time < tmax.nca) %>%
  distinct(Chemical, Chemical_Name, Species, Route, Media, Detect, Time) %>%
  group_by(Chemical, Chemical_Name, Species, Route, Media, Detect) %>%
  count(name = "tp_before_tmax") %>%
  ungroup() %>% 
  filter(Detect == TRUE) %>%
  dplyr::select(!Detect) %>%
  arrange(tp_before_tmax) %>%
  count(tp_before_tmax > 1)

```

Here I introduce the notion whether the data is variable compared across time.

```{r PredictionVariability_overtime}
my_residuals %>%
  filter(exclude == FALSE) %>%
  inner_join(winmodel) %>%
  inner_join(my_tkstats %>%
              dplyr::select(Chemical, Species,
                            Route, Media,
                            tmax = tmax.tkstats) %>%
              filter(!is.na(tmax))) %>%
  group_by(Chemical,
           Species,
           Route,
           Media,
           Dose) %>%
  mutate(
         normTime = ifelse(
           Route == "iv",
           1 + Time/max(Time),
           ifelse(
             Time > tmax,
             ((Time - tmax)/max(Time)) + 1,
             0.5 + Time/(2*tmax)
           )
         )) %>%
  distinct() %>%
  filter(Detect %in% TRUE) -> my_residuals2
my_residuals2 %>%
  # filter(Route == "oral") %>% View()
  ggplot(aes(
    x = normTime,
    y = Residuals
  )) +
  geom_point(alpha = 0.2, size = 0.5) +
  geom_point(data = filter(my_residuals2,
                           Chemical %in% "DTXSID4020533",
                           Species %in% "rat"),
             color = "orange2", size = 0.5) +
  geom_point(data = filter(my_residuals2,
                           Chemical %in% "DTXSID8031865",
                           Species %in% "rat"),
             color = "magenta4", size = 0.5) +
  geom_point(data = filter(my_residuals2,
                           Chemical %in% "DTXSID8025337",
                           Species %in% "rat"),
             color = "green2", size = 0.5) +
  geom_vline(xintercept = 1, color = "blue", linetype = "dashed") +
  geom_vline(xintercept = 2, color = "green4", linetype = "dashed") +
  facet_grid(cols = vars(Route),
             # rows = vars(model),
             scales = "free_x") +
  labs(x = "ADME-Normalized Time",
       y = "Winning Model Residuals") +
  theme(aspect.ratio = 1,
        panel.border = element_rect(color = "black", fill = NA, size = 1.5),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(face = "bold"),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        axis.text.x = element_blank())

ggsave(paste0(Sys.getenv("FIG_DIR"),
              "ADME_NormalizedTime-WinModelResiduals_onlyDetects_Highlighted_2023-07-19.png"),
       height = 4,
       width = 6,
       units = "in")
```

I try to find the connections between functions in _invivoPKfit_ to best understand
and represent the structure of the package.

```{r function_connections}
fw <- mvbutils::foodweb(where = asNamespace("invivoPKfit"),
                        border = TRUE, cex = 0.6, lwd = 1.5)
fun_mat <- fw$funmat

mvbutils::foodweb(where = asNamespace("invivoPKfit"),
                  prune = "convert_time",
                  # descendents = FALSE,
                  # ancestors = FALSE,
                  border = TRUE, cex = 1, lwd = 1.5)

# Predict dependents: rmse, residuals, rsq, calc_auc, fold_errors, plot
# Coef dependents: calc_auc, predict, get_tkstats, summary,log_Lik, coef_sd

```

## Other Plots

### Plotting ALL fits for Dose-normalized Joint fits (L-BFGS-B)

```{r plotting_all_Fits}
### Plotting for all (2.5MB file)
pl <- plot(my_pk, n_interp = 8, use_scale_conc = TRUE)
pdf(file = paste0(Sys.getenv("FIG_DIR"),
              "DoseNormalized_Joint_AllFits_Final.pdf"),
    height = 6, width = 10)
for (i in 1:nrow(pl)) {
  print(pl$final_plot[[i]])
}
dev.off()

# From my new SQL pull
pl <- plot(my_pk, n_interp = 8, use_scale_conc = TRUE)
pdf(file = paste0(Sys.getenv("FIG_DIR"),
              "NormalizedUpdate_NewCvT_SQLquery.pdf"),
    height = 6, width = 10)
for (i in 1:nrow(pl)) {
  print(pl$final_plot[[i]])
}
dev.off()

```

### Figure 2
```{r Explain-Heirarchical-Model}
# Note that the data will run one "Series" of time point at a time
my_1comp <- function(t, F_V, kelim, kabs, sd = 0) {
  sd + (F_V * (kabs * (exp(-kelim * t) - exp(-kabs * t))) / (kabs - kelim))
}

my_cdata <- data.frame(
  Reference = rep(rep(c("X", "O"), each = 3), 6),
  Time = rep(c(0, 2, 4, 6, 8, 12), each = 6),
  Concentration = c(
    4, 5, 6.5,
    4.5, 5, 6,
    8, 9, 9.5,
    7.5, 8, 8.5,
    3, 4.5, 6,
    3.5, 4, 5,
    2, 2.5, 4,
    1, 1.5, 2.25,
    0.4, 1, 2,
    0.4, 0.8, 1,
    0.2, 0.5, 1,
    0.25, 0.4, 0.5
  ) 
)

pl_pool <- ggplot() +
  geom_point(data = my_cdata,
             mapping = aes(
               x = Time,
               y = Concentration,
               color = Reference
             ),
             position = position_dodge(width = 0.4),
             size = 2) +
  stat_function(
    fun = my_1comp,
    args = list(F_V = 20, kelim = 0.6, kabs = 0.8),
    linewidth = 1) +
  stat_function(
    fun = my_1comp,
    args = list(F_V = 20, kelim = 0.6, kabs = 0.8, sd = 1),
    linewidth = 0.5, linetype = "dashed") +
  stat_function(
    fun = my_1comp,
    args = list(F_V = 20, kelim = 0.6, kabs = 0.8, sd = -1),
    linewidth = 0.5, linetype = "dashed") +
  scale_y_continuous(breaks = scales::pretty_breaks(10)) +
  scale_x_continuous(breaks = scales::pretty_breaks(13)) +
  scale_color_manual(values = c("black", "black")) +
  theme_classic() +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_text(size = 12, face = "bold"),
        axis.line = element_line(linewidth = 1, lineend = "square"),
        legend.position = "none")
ggsave(paste0(Sys.getenv("FIG_DIR"), "PooledExampleFigure.png"),
       units = "in",
       height = 4,
       width = 5,
       device = "png",
       dpi = 300)

# Change shape(s) from circles
pl_joint <- ggplot() +
  geom_point(data = my_cdata,
             mapping = aes(
               x = Time,
               y = Concentration,
               color = Reference,
               shape = Reference
             ),
             position = position_dodge(width = 0.4),
             size = 2) +
  stat_function(
    fun = my_1comp,
    args = list(F_V = 20, kelim = 0.6, kabs = 0.8),
    linewidth = 1) +
  stat_function(
    fun = my_1comp,
    args = list(F_V = 20, kelim = 0.6, kabs = 0.8, sd = 1.3),
    linewidth = 0.6, linetype = "dashed", color = "green4") +
  stat_function(
    fun = my_1comp,
    args = list(F_V = 20, kelim = 0.6, kabs = 0.8, sd = -1.3),
    linewidth = 0.6, linetype = "dashed", color = "green4") +
  stat_function(
    fun = my_1comp,
    args = list(F_V = 20, kelim = 0.6, kabs = 0.8, sd = 0.7),
    linewidth = 0.6, linetype = "dashed", color = "magenta3") +
  stat_function(
    fun = my_1comp,
    args = list(F_V = 20, kelim = 0.6, kabs = 0.8, sd = -0.7),
    linewidth = 0.6, linetype = "dashed", color = "magenta3") +
  scale_y_continuous(breaks = scales::pretty_breaks(10)) +
  scale_x_continuous(breaks = scales::pretty_breaks(13)) +
  scale_color_manual(values = c("magenta3", "green4")) +
  scale_shape_manual(values = c(15, 17)) +
  theme_classic() +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_text(size = 12, face = "bold"),
        axis.line = element_line(linewidth = 1, lineend = "square"),
        axis.title.y = element_blank(),
        legend.position = "none")

pl_joint
ggsave(paste0(Sys.getenv("FIG_DIR"), "JointExampleFigure.png"),
       units = "in",
       height = 4,
       width = 5,
       device = "png",
       dpi = 300)

plot_grid(pl_pool, pl_joint,
          align = "h", axis = "bt")

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "Pooled_Joint_ExampleFigure.png"),
       units = "in",
       height = 3.4,
       width = 6.5,
       device = "png",
       dpi = 300)

ggplot(data.frame(point_size = seq(0.2, 3, by = 0.2)),
       aes(x = as.factor(point_size), size = point_size)) +
  geom_point(y = 0) +
  scale_y_continuous(limits = c(-1, 1)) +
  theme_bw() +
  theme(legend.position = "none")

ggsave(paste0(Sys.getenv("FIG_DIR"), "Point_Reference_3x5.png"),
       units = "in",
       height = 3,
       width = 5,
       device = "png",
       dpi = 300)

```

# Notes

Specify that favorable gains in Cmax RMSE not accompanied by gains in AUC RMSE.
Histogram of Volume of Distributions and half-lifes... other TKstats... Lombardo comparison
- Stratified on species... then chemical class (pharma vs non-pharma, Pharmapendium)
- Classyfier for chemical superclass (talk to Paul) Ranges and Mean for each class...