### 2023-11-30 data-raw/TKstatsForCompTox_vignette.Rmd
Vignette for CompTox TK stats predictions needed

---
title: "Updated invivoPKfit predictions"
author: "Gilberto Padilla Mercado"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
set.seed(2023)
library(tidyverse, quietly = TRUE)
devtools::load_all() # Loads the invivopkfit package in feature/pkS3obj_facet branch
library(cowplot)
library(RColorBrewer)
tidyverse_conflicts()
# Note I have now saved the RDS file for JOINT model with L-BFGS-B optimization
my_pk <- readRDS("data-raw/all_cvt_pk_JOINT.rds")
```

First we explore the different _settings_ used. These fits were generated using a 
hierarchical error model. The data group was Chemical and Species and the error
group additionally included Reference. The optimization algorithm used was **L-BFGS-B**.

```{r settings_output}
my_pk$settings_preprocess
my_pk$settings_data_info
my_pk$settings_optimx
my_pk$stat_error_model
my_pk$stat_model
my_pk$data_group
```

The following chunk pulls the *winning model* (with the lowest AIC)
and pulls the predicted TKstats from this for each chemical and species listed.
This results in 109 unique observations.

```{r tkstats_predictions}
winmodel <- get_winning_model(my_pk)
all_tkstats <- get_tkstats(my_pk) %>% inner_join(winmodel)

all_tkstats <- all_tkstats %>% ungroup() %>%
  dplyr::select(Chemical, Species, Vdist, V1, Vss, halflife, model) %>%
  distinct()

```