### 2024-10-15 data-raw/testing_summary.Rmd

---
title: "testing_summary"
author: "Gilberto Padilla Mercado"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
today <- format(Sys.Date(), "%d%B%Y")
set.seed(as.numeric(Sys.Date()))
options(digits = 10, warnPartialMatchDollar = TRUE)
```

## Purpose

The pupose of this document is to test all functions in invivoPKfit and
assist with debugging. No function will remain unturned. Perhaps some of these
examples will be formalized into unit tests included in the package.

The first step is loading some basic packages: _magrittr_ and _devtools_

```{r libraries}
library(devtools)
load_all()
library(magrittr)
library(ggplot2)
```

Now the first thing we could do is get a list of all the functions.
Since we are in a project-based development environment, we are already in the
necessary working directory. We can inquire all the _.R_ files in the Terminal as so:

```
ls R/*.R
ls -l R/*.R | wc -l
```

We have 111 functions and methods in this package! This will be a lot of work to get
through, hence the inspiration for this document. First things first, make a 
"small" version of `cvt` that we can quickly fit. For the purposes of this
document, I will take 20% of CVT_Legacy data.
I will also add a few chemicals + species' data that I know will fail or
be skipped, so that these features may be tested.

```{r tiny-cvt}
tiny_cvt <- cvt %>%
  dplyr::filter(curation_set_tag == "CVT_Legacy") %>%
  dplyr::distinct(analyte_dtxsid, species) %>%
  dplyr::slice_sample(prop = 0.20) %>%
  dplyr::bind_rows(
    data.frame(
      analyte_dtxsid = c("DTXSID00861411", "DTXSID6023577", 
                         "DTXSID1043881", "DTXSID3061635",
                         "DTXSID1021116", "DTXSID2020139",
                         "DTXSID3031862"),
      species = rep("rat", 7)
    )) %>%
  dplyr::distinct() %>%
  dplyr::left_join(cvt, by = dplyr::join_by(analyte_dtxsid, species))

length(unique(tiny_cvt$analyte_dtxsid))
length(unique(tiny_cvt$administration_route_normalized)) # Should be two
length(unique(tiny_cvt$conc_medium_normalized)) # Should be two
tiny_cvt %>% 
  dplyr::distinct(analyte_dtxsid, species, n_subjects_normalized, curation_set_tag)
```

The next step is to create a minimal PK object like in the main vignette.

```{r minimal-pk}
flat_chems <- paste0("DTXSID",
                     c(2020139,
                       5037523,
                       8020541,
                       8025337,
                       5034773,
                       4021268))

minimal_pk <- pk(data = tiny_cvt %>%
                   # filter(analyte_dtxsid %in% flat_chems) %>%
                   mutate(n_subjects_normalized = if_else(
                     n_subjects_normalized > 1 & is.na(invivPK_conc_sd),
                     min(n_subjects_normalized, 6), n_subjects_normalized)),
                 mapping = ggplot2::aes(
                   Chemical = analyte_dtxsid,
                   Chemical_Name = analyte_name_original,
                   DTXSID = analyte_dtxsid,
                   CASRN = analyte_casrn,
                   Species = species,
                   Reference = fk_extraction_document_id,
                   Media = conc_medium_normalized,
                   Route = administration_route_normalized,
                   Dose = invivPK_dose_level,
                   Dose.Units = "mg/kg",
                   Subject_ID = fk_subject_id,
                   Series_ID = fk_series_id,
                   Study_ID = fk_study_id,
                   ConcTime_ID = conc_time_id,
                   N_Subjects = n_subjects_normalized,
                   Weight = weight_kg,
                   Weight.Units = "kg",
                   Time = time_hr,
                   Time.Units = "hours",
                   Value = invivPK_conc,
                   Value.Units = "mg/L",
                   Value_SD = invivPK_conc_sd,
                   LOQ = invivPK_loq
                 ))

testthat::expect_identical(length(minimal_pk), 10L)
```

From here I need to set up the pk object settings.
I will be using all settings as TRUE.

```{r pk-settings}
my_pk <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(keep_data_original = TRUE,
                      suppress.messages = FALSE) +
  settings_optimx(method = c(
    "bobyqa",
    "L-BFGS-B")) +
  scale_conc(dose_norm = TRUE, log10_trans = TRUE) +
  # scale_time(new_units = "auto") +
  stat_error_model(error_group = vars(Chemical, Species, Reference))
```

## Setting Getters Test

Below are just a few checks for the settings I changed above.

```{r setting-getters}
get_data_group(my_pk) # Modified to follow output from get_error_group
get_error_group(my_pk)

get_mapping(my_pk)
get_settings_data_info(my_pk) # Removed nca_group
get_settings_preprocess(my_pk)
get_scale_time(my_pk)
get_scale_conc(my_pk)
get_stat_error_model(my_pk) # Removed Chemical <- NULL
get_stat_model(my_pk)

get_status(my_pk)
get_status(my_pk, suppress.messages = TRUE)

testthat::expect_identical(get_status(my_pk), 1L)
```

Now we can do the pre-processing steps. Important to note that there is actually
an impute_sd setting that is on by default but has been commented out.

I need to figure out what the effect of the 30% is on calculating sigma.

```{r pre-process}
my_pk <- do_preprocess(my_pk)
my_data <- get_data(my_pk)
my_data %>% dplyr::distinct(Chemical, Species) %>% NROW()
```

Next I will go through the data info function. Implicitly, this tests
the `nca` and `calc_nca`

```{r data-info}
my_pk <- do_data_info(my_pk)
get_status(my_pk)
```

Next is pre-fitting. Here, the upper/lower bounds & starts for the parameters.
Here is also where sigma values are initialized.

```{r prefit}
my_pk <- do_prefit(my_pk)
get_prefit(my_pk) # This has some odd outputs :/
get_status(my_pk)
get_data_summary(my_pk)
get_settings_optimx(my_pk)
```

Finally do the fitting, here is where most of the functions come to play,
including the model functions. Walk through the main function and set `browser()`
markers in the concentration prediction functions.

```{r fit}
thispk_profvis <- profvis::profvis(do_fit(my_pk))
system.time(my_pk <- do_fit(my_pk))
system.time(my_pk <- do_fit(my_pk, n_cores = 10))
my_pk <- do_fit(my_pk)
# 100 seconds, 11 data_groups
# 270 seconds, 25 data_groups
# 12 cores, full dataset:  489 seconds

get_winning_model(my_pk) %>% View()
```

After some optimization of the fitting functions... now it's time
to optimize some of the downstream analysis functions and make sure they work
as intended! Firstly, I changed the way fit object _look_ so I need to change how
it interfaces with the coef method.

I have also found a _safer_ way to select the model functions in `predict.pk()`
that is based on joins and creating a data.frame of model names and functions
ahead of time.

```{r coef-preds}
my_coefs <- coef(my_pk)
my_preds <- predict(my_pk)
my_logLik <- logLik(my_pk)
```

There are a bunch of methods that call these three...  
For coef.pk() :

+ coef_sd.pk()  
+ get_tkstats.pk()  
+ _logLik.pk()_  
+ _predict.pk()_  
+ summary.pk()

For predict.pk() :

+ fold_errors.pk()  
+ residuals.pk()  
+ rsq.pk()  
+ rmse.pk()

For logLik.pk() :

+ AIC.pk()  
+ BIC.pk()  
+ get_winning_model.pk()

I will start with adjusting the `predict.pk()`-dependent methods.
Since most of these simply use the results from `predict.pk()`, it should
be simple.

```{r}
my_residuals <- residuals(my_pk)
my_fes <- fold_errors(my_pk)
my_rmse <- rmse(my_pk)
my_rmsle <- rmse(my_pk, use_scale_conc = list(dose_norm = FALSE,
                                              log10_trans = TRUE))

my_rsq <- rsq(my_pk)

plot(my_rsq$Rsq, my_rmsle$RMSLE)
```

Next lets address `logLik.pk()`-dependent methods. Generally, what I did here
is let `logLik.pk()` take the responsibility of checking that the methods
and models were alright. This was done more for efficiency in reading the code.
One important consideration for `get_winning_model.pk()` is what happens
when there are ties? This is probably unlikely to be the case, but we
need to account for it nonetheless.

```{r}
my_AIC <- AIC(my_pk)
my_BIC <- BIC(my_pk)
my_wm <- get_winning_model(my_pk)
table(my_wm$model)

```

Lastly, let's take a look at the remaining `coef.pk()`-dependent methods.
To make more efficient use of the results, I modified the output of
`coef_sd.pk` to look like `coef`, removing the table_format option.
`get_tkstats(my_pk)` has some nested operations that I think can be optimized
further.

```{r}
my_coefsd <- coef_sd(my_pk)
my_tkstats <- get_tkstats(my_pk)
best_tkstats <- eval_tkstats(my_pk)
```

Plotting is of the important outputs of _invivoPKfit_, as it allows
people to visualize the predictions and form a second opinion on the
fits.

```{r}
mypl <- plot(my_pk, log10_C = TRUE)
mypl %>% pull(final_plot) %>% .[[13]]
```