### 2025-01-23 data-raw/RestrictiveClearance_comparison.Rmd
Re-structured analysis to fit new httk release and ensured that restrictive and non-restrictive estimates & parameterization are being appropriately handled.

---
title: "Restrictive Clearance Comparison"
author: "Gilberto Padilla Mercado"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

This vignette describes comparisons between invivoPKfit predicted and those from httk.

```{r library_loading}
today <- format(Sys.Date(), "%d%b%Y")
devtools::load_all()
library(tidyverse)
library(ctxR)
library(ggvenn)
library(ggalluvial)
library(flextable)
library(reticulate)
source("data-raw/RC_comparison_bkgd.R") # httk & dawson_2021 loaded here
```

I select which parameterization (based on species) that I will use.

```{r selecting_chemical_params}
chosen_list <- rat_pars_human
# The following are derived from the above selected list (see environment)
this_chem_v <- unique(chosen_list$param_df$Chemical)
this_param_rest_l <- chosen_list$named_params_restrictive
this_param_nrest_l <- chosen_list$named_params_nonrestrictive
this_param_df <- chosen_list$param_df
this_httk_clearance_df <- chosen_list$clearance_df
this_species_v <- c("rat", "human")
# Testing to make sure there are some chemicals that have different clearances
count(this_httk_clearance_df, restrictive_clearance == nonrestrictive_clearance)
# The majority (140) do!
```

Here I do a quick fit of the CvT data only for those that have plasma measurements.

```{r quick-fit}
# Only use plasma for simplicity
this_pk <- pk(cvt %>% filter(conc_medium_normalized == "plasma")) +
  settings_optimx(method = "bobyqa") +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(
    keep_data_original = TRUE,
    suppress.messages = FALSE) +
  scale_conc(dose_norm = TRUE, log10_trans = TRUE) +
  stat_error_model(error_group = vars(Chemical, Species))

this_pk <- do_prefit(this_pk)
this_pk <- do_fit(this_pk, n_cores = 14)
```

Relationships in the mathematical PK models:

- When there is only oral data available, `Fgutabs_Vdist` and `CLtot_Fgutabs`
will be calculated instead of the numerator alone.  
- $C_{ss} \cdot CL_{tot} = C_{H,u}\cdot CL{int}$ (Bennet & Sodhi 2024)  
- [Add More Here]

Steps in the analysis:

1. Get chemicals + species combinations in CvTdb that have _httk_ predictions.
2. Safely get parameters, clearance, and solutions for each of these groups.
3. Compare clearances from _httk_ with those from _invivoPKfit_.
4. Compare $f_{up}$ for subset of data with total values available via _httk_.
5. Use new _clearance-aware_ optimization of 1-compartment parameters to
benchmark _httk_ solutions.
6. Estimate the fraction unbound in plasma for chemicals with rat data,
for those same chemicals with human data for Clint and Fup. 
Begin with the assumption that there is restrictive clearance, those chemicals
that do not change when optimized (and produce reasonable fits) are likely to
be _restrictively cleared_, whereas those that increase Fup during maximum
likelihood estimation are _possibly non-restrictively cleared_. 
7. Other factors such as renal clearance may play a role in those cases.
Validate the clearance estimate by measuring changes in predicted clearance.
Important to identify whether there seems to be a "fraction missing".
8. Wrap the previous steps into attempts to identify true positive
categorizations for Restrictive and Potentially Non-Restrictively metabolized
chemicals (include Lipinski's Rule of 5 too).
9. Implement a _slow_ PBPK model that optimizes $f_{up}$ and $CL_{int}$ under
restrictive and non-restrictive clearance conditions. Test whether these have a
substantial difference from initial values. Use `coef_sd()` to evaluate
uncertainty of these parameter estimates for a larger share of chemicals.
10. Compare the results of 9 with the putative/predicted clearance from 1-6.

The first thing I need to do is get the chemicals included in both the CvTdb subset
and the httk empirical data for pbpk.

Next I need to get the clearances using the `calc_total_clearance(dtxsid, restrictive.clearance)`
function from httk. Also need the clearances from invivoPKfit's winning models.

Note: There is about 50 chemicals with rat data for both, but 3X that when
we allow substitution of human numbers. However, this does not necessarily mean
that we will get appropriate fits with the human numbers.
The aforementioned set of rat data, for example, 33 of 51 chemicals find
appropriate fits with invivoPKfit + httk.

We first want to see how many chemicals are in either _httk_ or _invivoPKfit_

```{r venn_diagrams}
member_list <- list(
  httk_gas_pbtk = unique(get_cheminfo(info = "DTXSID",
                                species = "human",
                                model = "gas_pbtk",
                                median.only = TRUE,
                                physchem.exclude = TRUE,
                                suppress.messages = TRUE)) %>% 
    suppressWarnings(),
  cvt = cvt %>% 
    filter(species %in% c("human", "rat"),
           conc_medium_normalized == "plasma") %>%
    pull(analyte_dtxsid)
)

ggvenn(member_list, fill_color = c("white", "white"))

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "httk_vs_CvT_VennDiagram_human.png"),
       bg = "white",
       height = 3, width = 5
)

target_chems <- intersect(member_list$cvt, member_list$httk_gas_pbtk)
```

## Lipinski's Rules

Here I will look at the whether chemicals are "drug-like" by evaluating if:

1. Molecular weight less than 500 Da.  
2. logP (octanol-water partitioning coefficient) no greater than 5.  
3. Less than 5 hydrogen bond donors and less than 10 acceptors.

```{r lipinksi_test_data}
target_chem_info <- get_chemical_details_batch(DTXSID = target_chems,
                                               rate_limit = 1L)
target_chem_properties <- get_chem_info_batch(DTXSID = target_chems, 
                                              rate_limit = 1L) %>%
  filter(name == "LogKow: Octanol-Water") %>% distinct()

target_chem_properties %>%
  ggplot(aes(x = source, y = value)) +
  geom_point() +
  geom_line(aes(group = dtxsid), color = "grey10", alpha = 0.1) +
  geom_hline(yintercept = 5, linetype = "longdash") +
  facet_grid(cols = vars(propType), scales = "free_x") +
  theme_bw() +
  theme(axis.text.x = element_text(hjust = 1, angle = 45))

target_chem_properties %>%
  count(source)

target_chem_info %>%
  ggplot(aes(x = monoisotopicMass)) +
  geom_histogram(binwidth = 50, fill = "grey5",
                 boundary = 0) +
  geom_vline(xintercept = 500, linetype = "longdash") +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 100)) +
  theme_bw()

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "target_chems_molmass.png"),
       bg = "white",
       height = 3, width = 5
)

target_chem_info %>%
  count(monoisotopicMass <= 500)

```

```{r reticulate-prep}
smiles_list <- target_chem_info %>% pull(smiles) %>% unique()
mykit <- import("rdkit")
```

```{python get_H_bonds}
import rdkit
from rdkit import Chem
from rdkit.Chem import Descriptors

py_smiles_list = r.smiles_list
mol_from_smiles = [Chem.MolFromSmiles(x) for x in py_smiles_list]

py_h_donors = [Chem.Descriptors.NOCount(x) for x in mol_from_smiles]
py_h_acceptors = [Chem.Descriptors.NHOHCount(x) for x in mol_from_smiles]
```

```{r lipinksi_alluvial}
target_chem_summary <- target_chem_info %>%
  distinct(dtxsid, monoisotopicMass, smiles) %>%
  right_join(data.frame(smiles = smiles_list)) %>%
  bind_cols(data.frame(H_donors = py_to_r(py$py_h_donors),
                       H_acceptors = py_to_r(py$py_h_acceptors))
            ) %>%
  left_join(
    target_chem_properties %>%
      filter(source == "ACD/Labs Consensus") %>%
      distinct(dtxsid, source, value)
  ) %>%
  rename(LogKow = value)

target_chem_lipinski <- target_chem_summary %>%
  select(-c(smiles, source)) %>%
  group_by(dtxsid) %>%
  mutate("Mass <= 500 Da" = (monoisotopicMass <= 500),
          "H-donors < 5" = (H_donors < 5),
          "H-acceptors < 10" = (H_acceptors < 10),
          "logP < 5" = (LogKow < 5),
          .keep = "none") %>%
  ungroup() %>%
  mutate(Lipinski = rowSums(dplyr::select(., where(is.logical))) >= 3)

target_chem_lipinski %>%
  group_by(Lipinski) %>%
  count(`Mass <= 500 Da`, `H-donors < 5`, `H-acceptors < 10`, `logP < 5`,
        name = "Count") %>%
  ggplot(aes(
    axis3 = `Mass <= 500 Da`,
    axis4 = `logP < 5`,
    axis2 = `H-donors < 5`,
    axis1 = `H-acceptors < 10`,
    y = Count
  )) +
  geom_alluvium(aes(fill = Lipinski)) +
  geom_stratum() +
  scale_x_discrete(limits = c( 
    "H-acceptors < 10",
    "H-donors < 5",
    "<= 500 Da",
    "logP < 5"
  ),
  position = "top",
  expand = c(.2, .05)) +
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  labs(title = "Overwhelmingly, these chemicals passes Lipinksi's Rules of 5",
       fill = "Passes Lipinksi's Rules?") +
  scale_fill_manual(values = c("grey20", "green4")) +
  theme_minimal() +
  theme(axis.text.x = element_text(face = "bold", color = "black", size = 10),
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5),
        legend.position = "bottom")

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "Alluvium_Lipinski.png"),
       bg = "white",
       height = 4, width = 6
)
```

Overwhelmingly, we see that chemicals (by Lipinski's rules of 5) are "drug-like".
While this may not matter effectively for clearance, it may be nice to look 
at whether chemicals deemed not "drug-like" (11 total) are more likely to be
restrictively or non-restrictively cleared.

We also want to assess whether the "winning" _invivoPKfit_ fits are better than
the chosen _httk_ fit. This is a bit of a pre-requisite to our analysis of the
clearances and whether they are more or less reasonable in their reflection of
the reality. The metric used here will be RMSLE.

```{r httk_predictions}
restrictive_preds <- get_httk_preds(
  parameters = this_param_rest_l,
  pk_obj = this_pk,
  species = c("rat", "human")
) %>% rename_with(
  .fn = ~paste0(.x, "_restrictive"),
  .cols = starts_with("httk_")
)

nonrestrictive_preds <- get_httk_preds(
  parameters = this_param_nrest_l,
  pk_obj = this_pk,
  species = c("rat", "human")
) %>% rename_with(
  .fn = ~paste0(.x, "_nonrestrictive"),
  .cols = starts_with("httk_")
)

all_httk_preds <- inner_join(restrictive_preds, nonrestrictive_preds)

this_nonflat_winmodel <- get_winning_model(this_pk) %>%
  filter(model %in% c("model_1comp", "model_2comp"))

winning_preds <- predict(this_pk) %>%
  semi_join(this_nonflat_winmodel) %>%
  distinct(Chemical, Species, Chemical_Name,
           model, method, Dose, Media, Route, Reference, pLOQ, Time,
           N_Subjects, Conc, Conc_est)

# RMSLE formula
# LE = log1p(Conc_est) - log1p(Conc)
# SLE = LE^2
# RMSLE = sqrt(mean(SLE, na.rm = TRUE))

combined_preds <- all_httk_preds %>%
  left_join(winning_preds) %>%
  filter(!is.na(model))

combined_RMSLE <- combined_preds %>%
  ungroup() %>%
  mutate(
    restrictive_SLE = (log1p(httk_Preds_restrictive) - log1p(Conc))^2,
    nonrestrictive_SLE = (log1p(httk_Preds_nonrestrictive) - log1p(Conc))^2,
    invivoPK_SLE =  (log1p(Conc_est) - log1p(Conc))^2 
  ) %>%
  group_by(Chemical, Species) %>%
  summarize(RMSLE_restrictive = sqrt(mean(restrictive_SLE, na.rm = TRUE)),
            RMSLE_nonrestrictive = sqrt(mean(nonrestrictive_SLE, na.rm = TRUE)),
            RMSLE_invivoPKfit = sqrt(mean(invivoPK_SLE, na.rm = TRUE)))

combined_RMSLE %>%
  ungroup() %>%
  pivot_longer(cols = starts_with("RMSLE_"),
               names_to = "Source",
               values_to = "RMSLE") %>%
  mutate(Source = factor(Source,
                         levels = c("RMSLE_nonrestrictive",
                                    "RMSLE_invivoPKfit",
                                    "RMSLE_restrictive")
  )) %>%
  ggplot(aes(x = Source, y = RMSLE, group = Chemical)) +
  geom_point(alpha = 0.5) +
  geom_line(alpha = 0.5) +
  theme_bw()

combined_RMSLE %>%
  ungroup() %>%
  mutate(comparison_group = case_when(
    between(RMSLE_invivoPKfit, RMSLE_nonrestrictive, RMSLE_restrictive) |
     between(RMSLE_invivoPKfit, RMSLE_restrictive, RMSLE_nonrestrictive) ~ "Between",
    RMSLE_invivoPKfit < RMSLE_nonrestrictive &
      RMSLE_invivoPKfit < RMSLE_restrictive ~ "Less than",
    RMSLE_invivoPKfit > RMSLE_restrictive &
      RMSLE_invivoPKfit > RMSLE_nonrestrictive~ "More than"
  )) %>% 
  ggplot(aes(x = Chemical, y = RMSLE_invivoPKfit)) +
  geom_point(color = "green4") +
  geom_point(aes(y = RMSLE_nonrestrictive), shape = 4, size = 0.5) +
  geom_point(aes(y = RMSLE_restrictive), shape = 4, size = 0.5) +
  geom_linerange(aes(ymin = RMSLE_nonrestrictive, ymax = RMSLE_restrictive)) +
  facet_grid(~factor(comparison_group,
                     levels = c("Less than", "Between", "More than")),
             scales = "free_x", space = "free_x") +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid = element_blank())

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "RMSLE_comparison_fullChart.png"),
       bg = "white",
       height = 4, width = 8
)

```

Another weight that we may consider is relative log-likelihood. This can be
calculated using AIC, compared to the the flat model.

```{r relative-likelihood}
this_rll_probs <- AIC(this_pk) %>%
  distinct(Chemical, Species, model, AIC) %>%
  pivot_wider(names_from = model, values_from = AIC) %>%
  rowwise() %>% 
  mutate(minAIC = min(c_across(starts_with("model_")), na.rm = TRUE)) %>% 
  ungroup() %>%
  mutate(across(starts_with("model_"), \(x) {
    signif(exp((minAIC - x)/2) * 100, 4)
  })) %>%
  mutate(across(starts_with("model_"), \(x) {
    ifelse(x < 1E-6, 0, x)
  })) %>%
  inner_join(this_nonflat_winmodel %>% distinct(Chemical, Species, model)) %>%
  ungroup() %>%
  filter(model_1comp >= 99 | model_2comp >= 99,
        model_flat < 1, Chemical %in% target_chems)

this_rll_probs %>%
  count(model_1comp >= 99, model_2comp >= 99)

```

The weights to be applied are calculated as such:

$$
\gamma = \frac{100\cdot \rm{RMSLE}}{\theta}
$$
Where $\theta$ is the relative log-likelihood. The smaller number the better.
If the RMSLE is zero, then it means that it is a "perfect" fit and we are highly
confident in using other metrics to gauge clearance. Below I put these together.

```{r model_weighting}
combined_model_weights <- combined_RMSLE %>%
  left_join(this_rll_probs) %>%
  ungroup() %>%
  pivot_longer(cols = starts_with("model_"),
               names_to = "all_models",
               values_to = "relative_likelihood") %>%
  filter(all_models == model) %>%
  ungroup() %>%
  mutate(gamma_weight = 100 * RMSLE_invivoPKfit / relative_likelihood,
         rmsle_comparison_group = case_when(
    between(RMSLE_invivoPKfit, RMSLE_nonrestrictive, RMSLE_restrictive) |
     between(RMSLE_invivoPKfit, RMSLE_restrictive, RMSLE_nonrestrictive) ~ "Between",
    RMSLE_invivoPKfit < RMSLE_nonrestrictive &
      RMSLE_invivoPKfit < RMSLE_restrictive ~ "Less than",
    RMSLE_invivoPKfit > RMSLE_restrictive &
      RMSLE_invivoPKfit > RMSLE_nonrestrictive~ "More than")
  ) %>%
  mutate(rmsle_comparison_group = factor(rmsle_comparison_group,
                                         levels = c("Less than", "Between",
                                                    "More than"))) %>%
  select(-all_models) %>% distinct()

```

Next we need to compare clearances. Sometimes this is difficult as the overlap
with certain calculations will be defined by the type of experimental data
available. For instance, if there is only oral data available, `CLtot.tkstats`
will _not_ be calculated and instead `CLtot/Fgutabs.tkstats` will be available.

```{r combine_tkstats}
# Get the winning model and only include chemicals in target_chems
this_winmodel <- get_winning_model(this_pk) %>%
    filter(Chemical %in% target_chems)
# How many overlapping chemicals have winning models that aren't flat?
count(
  this_winmodel,
  model
) # 153 = 97 1-compartment + 56 2-compartment

this_tkstats <- eval_tkstats(this_pk,
                             model = c("model_1comp", "model_2comp"),
                             finite_only = TRUE,
                             dose_norm = TRUE
) %>%
  rename(CLtot_Fgutabs.tkstats = `CLtot/Fgutabs.tkstats`) %>%
  distinct(Chemical, Species, method, model, Dose, Route, Media,
           CLtot.tkstats, CLtot_Fgutabs.tkstats, halflife.tkstats, Css.tkstats) %>%
  inner_join(this_winmodel)

count(
  this_tkstats,
  # model,
  !is.na(CLtot.tkstats),
  !is.na(CLtot_Fgutabs.tkstats),
  !is.na(halflife.tkstats)
) %>% flextable() %>%
  set_header_labels(
    `!is.na(CLtot.tkstats)` = "Valid Total Clearance",
    `!is.na(CLtot_Fgutabs.tkstats)` = "Valid Total Clearance/Fgutabs",
    `!is.na(halflife.tkstats)` = "Valid halflife",
    n = "Number of Chemicals"
    )
# 46 chemicals have valid total clearance
# 144 chemicals have valid total clearance/Fgutabs
# 5 chemicals have only valid total clearance and no valid total clearance/Fgutabs

count(this_tkstats, is.na(Css.tkstats))

```

Here I will evaluate $C_{SS}$, the steady-state concentration to get some
understanding of whether chemicals may be likely restrictive or non-restrictive.

```{r Css_assessment}

css_httk_batch <- function(param_list,
                           restrictive = TRUE) {
  stopifnot(is.logical(restrictive))

  css_list_r <- vapply(
    param_list,
    \(x) {
      calc_analytic_css_3comp2(parameters = x,
                               restrictive.clearance = restrictive)
    },
    FUN.VALUE = numeric(1L)
  )
  css_list_names <- names(param_list)

  css_df <- data.frame(Chemical = css_list_names,
                       Css_httk = css_list_r)
  if (restrictive) {
    css_df <- rename(css_df, Css_httk_r = Css_httk)
  } else {
    css_df <- rename(css_df, Css_httk_nr = Css_httk)
  }
  return(css_df)
}

this_css_df <- inner_join(
  css_httk_batch(this_param_rest_l, restrictive = TRUE),
  css_httk_batch(this_param_nrest_l, restrictive = FALSE)
) %>%
  inner_join(this_tkstats) %>%
  distinct(Chemical, model, Route, Css_httk_r, Css_httk_nr, Css.tkstats) %>%
  filter(Css_httk_nr > 0, Css_httk_r > 0, Css.tkstats > 0) %>%
  left_join(target_chem_lipinski %>% rename(Chemical = dtxsid)) %>%
  mutate(comparison_group_Css = factor(
    case_when(
      between(Css.tkstats, Css_httk_r, Css_httk_nr) |
        between(Css.tkstats, Css_httk_nr, Css_httk_r) ~ "Between",
      Css.tkstats < Css_httk_r & Css.tkstats < Css_httk_nr ~ "Below",
      Css.tkstats > Css_httk_r & Css.tkstats > Css_httk_nr ~ "Above"),
    levels = c("Above", "Between", "Below")
  ), Chemical = factor(Chemical) %>% fct_reorder(Css_httk_r)
  )

this_css_df %>%
  ggplot(aes(x = Chemical)) +
  geom_point(aes(y = Css_httk_r), shape = 4, size = 0.5) +
  geom_point(aes(y = Css_httk_nr), shape = 4, size = 0.5) +
  geom_linerange(aes(ymin = Css_httk_r,
                     ymax = Css_httk_nr)
  ) +
  geom_point(aes(y = Css.tkstats, shape = Route), color = "turquoise4") +
  facet_grid(rows = vars(comparison_group_Css),
             scales = "free_x",
             space = "free_x") +
  scale_y_log10(name = "Css") +
  scale_shape_manual(values = c(0,1)) +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid = element_blank())

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "Css_comparison_fullChart.png"),
       bg = "white",
       height = 4, width = 8
)

```

```{r}
this_css_df %>%
  distinct(Chemical, model, Route, Css_httk_r, Css_httk_nr, Css.tkstats) %>%
  filter(Css_httk_nr > 0, Css_httk_r > 0, Css.tkstats > 0) %>%
  left_join(data.frame(Chemical = names(this_param_rest_l),
                       CLint = vapply(names(this_param_rest_l),
                                      \(x) {
                                        this_param_rest_l[[x]][["Clint"]]
                                      },
                                      FUN.VALUE = numeric(1L)
                       ))) %>%
  mutate(CLint = log10(CLint + 1)) %>%
  ggplot(aes(x = CLint)) +
  geom_point(aes(y = Css_httk_r), shape = 4, size = 0.5) +
  geom_point(aes(y = Css_httk_nr), shape = 4, size = 0.5) +
  geom_linerange(aes(ymin = Css_httk_r,
                     ymax = Css_httk_nr)
  ) +
  scale_y_log10(name = "C<sub>SS</sub>") + 
  labs(x = "CL<sub>int</sub>") +
  geom_point(aes(y = Css.tkstats, shape = Route), color = "turquoise4") +
  scale_shape_manual(values = c(0,1)) +
  theme_bw() +
  theme(axis.title = ggtext::element_markdown())

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "Css_x_CLint.png"),
       bg = "white",
       height = 4, width = 8
)
```

Now I will try to triangulate the model weights witht the Css estimates.

```{r weighted_Css}
combined_model_weights %>%
  inner_join(this_css_df) %>%
  ggplot(aes(x = fct_reorder(Chemical, Css_httk_r))) +
  geom_point(aes(y = Css_httk_r), shape = 4, size = 0.5) +
  geom_point(aes(y = Css_httk_nr), shape = 4, size = 0.5) +
  geom_linerange(aes(ymin = Css_httk_r,
                     ymax = Css_httk_nr)
  ) +
  geom_point(aes(y = Css.tkstats, shape = Route, color = gamma_weight)) +
  facet_grid(rows = vars(comparison_group_Css),
             cols = vars(rmsle_comparison_group),
             scales = "free_x",
             space = "free_x") +
  scale_y_log10(name = "Css") +
  scale_shape_manual(values = c(0,1)) +
  scale_color_viridis_b(option = "viridis", 
                        guide = guide_colorbar(position = "bottom")) +
  labs(subtitle = "RMSLE for invivoPKfit compared to httk predictions",
       x = "Chemical\n(ordered by increasing restrictive Css prediction)",
       y = "Css",
       color = "Inverse Confidence") +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.subtitle = element_text(hjust = 0.5),
        panel.grid = element_blank(),
        legend.title = element_text(vjust = 0.8))

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "Css_withRMSLEweights.png"),
       bg = "white",
       height = 5, width = 10
)

```

```{r calc_CLtot_Fgutabs}
this_tkstats2 <- this_tkstats %>%
  left_join(
    this_param_df %>%
      distinct(Chemical, Species, Fabsgut.restrictive)
  ) %>%
  mutate(CLtot_calc = CLtot_Fgutabs.tkstats * Fabsgut.restrictive)

ggplot(data = this_httk_clearance_df,
       aes(x = Chemical)) +
  geom_point(aes(y = restrictive_clearance), color = "green4") +
  geom_point(aes(y = nonrestrictive_clearance), color = "orange2") +
  geom_point(data = this_tkstats2,
             aes(y = CLtot_calc)) +
  scale_y_log10() +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

# Combine these two
combined_clearance_df <- this_httk_clearance_df %>%
  inner_join(this_tkstats) %>%
  distinct()

length(unique(combined_clearance_df$Chemical))
```

Finally, let's visualize where the tkstats predictions land.
Because there are only a handful of chemicals which _invivoPKfit_ can calculate
`CLtot` for, I will employ two strategies to gauge how different these estimates
are in reality.

1. Visualize differences in half-life.  
2. Multiply `CLtot_Fgutabs` by the `Fabsgut` value parameterized in _httk_ to
get a clearance value.

```{r half-life}
combined_clearance_df %>%
  ggplot(aes(x = halflife.tkstats)) +
  geom_point(aes(y = restrictive_halflife), color = "magenta2") +
  geom_point(aes(y = nonrestrictive_halflife), color = "green4") +
  coord_equal(xlim = c(0, 200), y = c(0, 200))
```

Notes:
Look at Css values for restrictive/nonrestrictive (httk solutions)
AICs for models

```{r clearance_viz}

this_param_df %>%
  distinct(Chemical, Species, Fabsgut.restrictive, Fabsgut.nonrestrictive) %>%
  inner_join(combined_clearance_df) %>%
  mutate(CLtot.nonrest = CLtot_Fgutabs.tkstats * Fabsgut.nonrestrictive,
         CLtot.rest = CLtot_Fgutabs.tkstats * Fabsgut.restrictive) %>%
  pivot_longer(cols = c(restrictive_clearance, nonrestrictive_clearance,
                        CLtot.nonrest, CLtot.rest),
               names_to = "Source",
               values_to = "Clearance") %>%
  ggplot(aes(y = Chemical, x = Clearance, color = Source)) +
  geom_point() +
  labs(title = "comparison of restrictive and non-restrictive clearance estimates") +
  theme_bw() +
  coord_cartesian(xlim = c(0.0001, 1000)) +
  scale_x_log10()

this_param_df %>%
  distinct(Chemical, Species, Fabsgut.restrictive, Fabsgut.nonrestrictive) %>%
  inner_join(combined_clearance_df) %>%
  mutate(nonrest_fgutabs = nonrestrictive_clearance / Fabsgut.nonrestrictive,
         rest_fgutabs = nonrestrictive_clearance / Fabsgut.restrictive) %>%
  pivot_longer(cols = c(CLtot_Fgutabs.tkstats,
                        rest_fgutabs, nonrest_fgutabs),
               names_to = "Source",
               values_to = "Clearance") %>%
  ggplot(aes(y = Chemical, x = Clearance, color = Source)) +
  geom_point() +
  labs(title = "comparison of restrictive and non-restrictive clearance estimates") +
  theme_bw() +
  coord_cartesian(xlim = c(0.000001, 100000)) +
  scale_x_log10()

this_param_df %>%
  distinct(Chemical, Species, Fabsgut.restrictive, Fabsgut.nonrestrictive) %>%
  inner_join(combined_clearance_df) %>%
  mutate(nonrest_fgutabs = nonrestrictive_clearance / Fabsgut.nonrestrictive,
         rest_fgutabs = nonrestrictive_clearance / Fabsgut.restrictive) %>%
   mutate(AD_restrictive = abs(CLtot_Fgutabs.tkstats - rest_fgutabs),
         AD_nonrestrictive = abs(CLtot_Fgutabs.tkstats - nonrest_fgutabs),
         rough_determination = ifelse(
           min(AD_restrictive, AD_nonrestrictive) == AD_restrictive &
             CLtot.tkstats <= nonrest_fgutabs,
           "Restrictive", "Possibly non-Restrictive"
         )) %>% 
  dplyr::ungroup() %>% View()

clearance_guess <- combined_clearance_df %>%
  mutate(AD_restrictive = abs(CLtot.tkstats - restrictive_clearance),
         AD_nonrestrictive = abs(CLtot.tkstats - nonrestrictive_clearance),
         rough_determination = ifelse(
           min(AD_restrictive, AD_nonrestrictive) == AD_restrictive &
             CLtot.tkstats < nonrestrictive_clearance,
           "Restrictive", "Possibly non-Restrictive"
         )) %>% 
  dplyr::ungroup()

clearance_guess %>%
  dplyr::distinct(Chemical, rough_determination) %>%
  group_by(rough_determination) %>%
  tally()

# Preliminary viz, plot all clearances
ccd <- clearance_guess %>%
  relocate(CLtot.nca_mean, CLtot.nca_sd, .after = tidyselect::last_col()) %>%
  pivot_longer(
    cols = tidyselect::all_of(c("CLtot.tkstats", "CLtot.nca_mean",
                              "restrictive_clearance", "nonrestrictive_clearance")),
    names_to = "Clearance_Source",
    values_to = "Value"
  ) %>%
  mutate(
    Chemical = factor(Chemical),
    Clearance_Source = factor(
      Clearance_Source,
      levels = c(
        "CLtot.tkstats", "CLtot.nca_mean",
        "nonrestrictive_clearance", "restrictive_clearance"
      )
    ),
    label_cols = ifelse(rough_determination == "Restrictive",
                     "green4", "orange3")
  ) %>%
  transform(
    Chemical = fct_reorder(Chemical, Value, max)
  ) %>% 
  as_tibble()

ccd %>%
  ggplot(aes(
    x = Value,
    y = Chemical,
    fill = Clearance_Source
  )) +
  geom_point(shape = 21, alpha = 0.7) +
  scale_x_log10() +
  scale_fill_manual(
    breaks = c(
      "nonrestrictive_clearance", "CLtot.tkstats",
      "CLtot.nca_mean",
      "restrictive_clearance"
    ),
    labels = c(
      "non-restrictive clearance (httk)",
      "predicted clearance (invivoPKfit)",
      "mean non-compartmental total clearance (invivoPKfit)",
      "restrictive clearance (httk)"
    ),
    values = c("orange2", "grey5", "skyblue3", "green4")
  ) +
  facet_grid(rows = vars(Species), scales = "free", space = "free") +
  labs(
    x = "Total Clearance",
    y = "Chemical",
    title = "Comparing clearance estimates of chemicals with rat CvT data"
  ) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1.5),
    panel.background = element_blank(),
    panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold", angle = 0),
    axis.title.y = element_blank(),
    legend.position = "right",
    legend.direction = "vertical",
    axis.ticks = element_blank(),
    axis.line = element_blank()
  )

clear_chem_levels <- levels(ccd$Chemical)

ggsave(
  filename = paste0(
    Sys.getenv("FIG_DIR"),"RestrictiveClearance_Analysis/",
    today,
    "all_tkClearance_comparison_httk_PBPK.png"
  ),
  height = 8,
  width = 8,
  device = "png"
)
```

I would also like to know, how does the $f_{up}$ of this set of chemicals
compare to that of all chemicals in `httk` (Dawson 2021).

```{r fup_comparison}
fup_df <- this_param_df %>%
  distinct(Chemical, Species, forced_human_values,
           Funbound.plasma, Funbound.plasma.adjustment, 
           Funbound.plasma.dist) %>%
  mutate(
    Funbound.plasma.median = as.numeric(str_extract(
      Funbound.plasma.dist, pattern = "^\\d+\\.?\\d+e?-?\\d+(?=,)"
    )),
    Funbound.plasma.lower95 = as.numeric(str_extract(
      Funbound.plasma.dist, pattern = "(?<=,)\\d+\\.?\\d+e?-?\\d+(?=,)"
    )),
    Funbound.plasma.upper95 = as.numeric(str_extract(
      Funbound.plasma.dist, pattern = "(?<=,)\\d+\\.?\\d+e?-?\\d$"
    ))
    )

fup_df %>%
  ggplot(aes(x = Funbound.plasma.adjustment)) +
  geom_histogram(
    data = fup_df %>%
      filter(Chemical %in% this_httk_clearance_df$Chemical),
    # aes(y = after_stat(density)),
    fill = "white",
    color = "darkgreen", binwidth = 0.03
  ) +
  expand_limits(expand = c(0,0)) +
  geom_vline(xintercept = 0.05, linetype = "dashed") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  scale_x_continuous(expand = expansion(mult = c(0.01, 0.01))) +
  labs(
    title = "Distribution of Funbound.plasma values",
    subtitle = "(chemicals fit by invivoPKfit in dark green)"
  ) +
  coord_cartesian(clip = "off") +
  theme_bw()

ggsave(paste0(Sys.getenv("FIG_DIR"), "fup_density_plots.png"),
       height = 5,
       width = 6,
       device = ragg::agg_png,
       units = "in"
)

test_fup <- fup_df %>%
  pull(Funbound.plasma)
for (i in seq(0, 0.5, by = 1e-9)) {
  if (sum(test_fup > i) == sum(test_fup <= i)) {
    print(i)
    break
  }
}

fup_df %>%
  ggplot(aes(x = Funbound.plasma, y = Funbound.plasma.adjustment)) +
  geom_point()

```

## Exploring _httk_ parameters and solved values for `gas_pbtk`

Note that this is only for rats. This is supposed to be the best case,
but of course _httk_ is "made" for predicting **human exposures**. Also,
I will try to order some of this data the same way as the clearance figures.

This will ultimately help understand whether Total Clearance
differences and prediction differences relate. Of course it would
be possible to do a formal test for correlation but that overdoes what we need
to build this relationship for with the rather limited _rat_ data we have
in both _httk_ and _invivoPKfit_.

```{r rat_invivo_comparison}

solve_gas_pbtk(
  dtxsid = common_chems_dawson[10],
  dose = 1, exp.conc = 0,
  iv.dose = FALSE,
  input.units = "mg/kg",
  output.units = "mg/L",
  suppress.messages = TRUE
) %>%
  ggplot(aes(x = time)) +
  geom_line(aes(y = Cplasma)) +
  geom_line(aes(y = Crest), color = "blue4") +
  geom_line(aes(y = Cgut), color = "goldenrod") +
  geom_line(aes(y = Cliver), color = "darkgreen") +
  geom_line(aes(y = Clung), color = "hotpink") +
  geom_line(aes(y = Ckidney), color = "purple4")

# Functions wrapping httk single dose evaluation
httk_bolus_gas_pbtk <- function(dtxsid, species = "Human",
                                dose, times, iv.dose, LOQ) {
  tmp_solution <- solve_gas_pbtk(
    dtxsid = unique(dtxsid),
    dose = unique(dose),
    exp.conc = 0,
    iv.dose = unique(iv.dose),
    times = unique(c(0, times/24)),
    input.units = "mg/kg",
    output.units = "mg/L",
    species = unique(stringr::str_to_title(species)),
    default.to.human = TRUE,
    suppress.messages = TRUE
  )
  # Remove the 0.0001 intermediate time value that is returned
  retdf <- tibble(
    Time = round(tmp_solution[, "time"]*24, digits = 3),
    httk_Preds = tmp_solution[, "Cplasma"]
  )
  retdf[which(retdf$Time %in% unique(round(times, 3))), ]
}

httk_preds_rat <- get_httk_preds(rat_pars_rat, this_pk, species = "rat")

```

We need to put the _in vivo_ data in an amenable format.
For this I will limit the species to "rat", the chemicals to
`common_chems_rat`, and the medium to "plasma".

```{r rat_data}
rat_df <- httk_preds_rat %>%
  left_join(predict(this_pk)) %>%
  semi_join(this_winmodel)

rat_df %>% ggplot(mapping = aes(
    x = Conc_est,
    y = httk_Preds,
    color = interaction(Route)
  )) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "black") +
  scale_x_log10() + scale_y_log10() +
  facet_wrap(~Route)

rat_plots <- rat_df %>%
  distinct(Chemical, Species, Dose, Route, Media, Reference,
           Time, httk_Preds, Conc, Conc_est, pLOQ) %>%
  mutate(Dose = as.factor(Dose)) %>%
  group_by(Chemical, Species, Route, Media, Reference) %>%
  nest(.key = "data_pl")

rat_plots <- rat_plots %>%
  reframe(plots = purrr::map(data_pl,
                             \(x) {
                               cowplot::plot_grid(
                               ggplot(data = x %>%
                                        group_by(Time, Dose) %>%
                                        mutate(meanConc = mean(Conc,
                                                           na.rm = TRUE)) %>%
                                        select(-Conc) %>% distinct() %>%
                                        pivot_longer(cols = c(httk_Preds,
                                                              meanConc,
                                                              Conc_est),
                                                     names_to = "Measurement",
                                                     values_to = "Concentration"
                                                     ), 
                                      mapping = aes(
                                        x = Time,
                                        y = Concentration,
                                        color = Measurement,
                                        shape = Dose
                                      )) +
                                 geom_point() +
                                 geom_line(
                                   aes(group = interaction(Measurement, Dose))
                                   ) +
                                 theme_bw() +
                                 labs(
                                   title = paste(Chemical, Species),
                                   subtitle = paste(Route, Media)
                                 ),
                               ggplot(data = x) +
                                 geom_point(aes(x = Conc_est, httk_Preds,
                                                shape = Dose)) +
                                 geom_abline(slope = 1, intercept = 0,
                                             linetype = "dashed") +
                                 theme_bw()
                               )
                             })
  )

pdf(file = paste0(Sys.getenv("FIG_DIR"), "RC_comparison_CvTplots.pdf"),
    onefile = TRUE,
    height = 7,
    width = 10,
    bg = "white")

for (i in 1L:NROW(rat_plots)) {
  message("Now plotting: ", i)
  print(rat_plots %>% pull(plots) %>% .[[i]])
}

dev.off()

length(unique(rat_df$Chemical))

rat_df$Chemical <- factor(rat_df$Chemical, levels = clear_chem_levels)

rat_out <- rat_df %>%
  filter(Chemical %in% clear_chem_levels) %>%
  group_by(Chemical, Species, Route, Reference, Media) %>%
  mutate(httk_Preds = ifelse(httk_Preds < pLOQ, pLOQ, httk_Preds)) %>% 
  mutate(SLE = (log2(Conc) - log2(httk_Preds))^2,
         AE = abs(Conc - httk_Preds)) %>%
  summarize(
    RMSLE = sqrt(mean(SLE, na.rm = TRUE)),
    MAE = mean(AE, na.rm = TRUE)
   ) %>% pivot_longer(cols = c(RMSLE, MAE),
                      names_to = "metric",
                      values_to = "value")

rat_out %>%
  na.omit() %>%
  ggplot() +
  geom_tile(data = rat_out %>% na.omit() %>%
              filter(metric %in% "RMSLE"),
            aes(x = metric, y = Chemical, fill = value)) +
  scale_fill_gradient(low = "grey5", high = "grey90",
                      name = "RMSLE") +
  ggnewscale::new_scale_fill() +
  geom_tile(data = rat_out %>% na.omit() %>%
              filter(metric %in% "MAE"),
            aes(x = metric, y = Chemical, fill = value)) +
  scale_fill_gradient(low = "darkgreen", high = "seagreen1",
                      limits = c(0, 80),
                      labels = c("0", "20",
                                 "40", "60",
                                 "80+"),
                      oob = scales::squish,
                      name = "MAE") +
  theme_minimal() +
  theme(axis.text.x = element_text(hjust = 1, angle = 50),
        legend.title = element_text(hjust = 0.5)) +
  coord_fixed(ratio = 0.75)

ggsave(
  filename = paste0(
    Sys.getenv("FIG_DIR"),
    "rat_httk_predictionRMSLE.png"
  ),
  height = 8.1,
  width = 4,
  device = "png"
)

rat_out2 <- rat_df %>%
  group_by(Chemical, Species, Route, Dose) %>% nest() %>%
  summarize(plot = map(
    data,
    \(d){
      ggplot(d, aes(x = Time)) +
        geom_point(aes(y = Conc)) +
        geom_line(aes(y = httk_Preds),
                    color = "forestgreen") +
        geom_point(aes(y = httk_Preds),
                   color = "forestgreen",
                   shape = 1) +
        labs(title = Chemical,
             subtitle = paste(Dose, "mg/kg given via",
                               Route, "bolus"),
             y = "Plasma Concentration (mg/L)",
             x = "Time (hrs)") +
        theme_bw() +
        theme(plot.title = element_text(hjust = 0.5),
              plot.subtitle = element_text(hjust = 0.5))
    }
  ))
```

```{r httk_solution}

rat_params %>%
  ggplot(aes(y = fupClint)) + 
  geom_point(aes(x = Clint), color = "darkred", shape = 1) +
  geom_abline(slope = 1, intercept = 0, color = "darkblue",
              linetype = "dashed")
rat_params %>%
  mutate(fup_z = scale(Funbound.plasma),
         clint_z = scale(Clint)) %>%
  dplyr::select(Chemical, fup_z, clint_z) %>%
  pivot_longer(cols = c(fup_z, clint_z),
               names_to = "pars", values_to = "vals") %>%
  ggplot(aes(x = pars, y = vals, group = Chemical)) +
  geom_line()

httk::calc_tkstats(dtxsid = unique(rat_out$Chemical)[[5]])
httk::parameterize_gas_pbtk(
  restrictive.clearance = TRUE,
  dtxsid = unique(rat_out$Chemical)[[5]]
)

httk::solve_gas_pbtk(
  dtxsid = unique(combined_clearance_df$Chemical)[[5]],
  parameters = httk::parameterize_gas_pbtk(
    restrictive.clearance = TRUE,
    dtxsid = unique(combined_clearance_df$Chemical)[[5]]
  ),
  dose = 10,
  exp.conc = 0,
  input.units = "mg",
  output.units = "mg",
  days = 3,
  plots = TRUE
)

# For the invivoPKfit functions, we will hold kelim constant
# This constant will be calculated as CLtot/Vdist (or V1, if 2-compartment)
# This constant is derived by httk in the following way:
calc_total_clearance(
  dtxsid = common_chems[[10]],
  restrictive.clearance = FALSE
) /
  calc_vdist(dtxsid = common_chems[[10]])
```

I have modified the starts of the 1- and 2-compartment models,
as well as adding a restrictive_clearance argument in stat_model.

```{r running-invivoPKfit}
# This is now updated to use dawson chem table
minimal_pk <- pk(data = cvt %>%
                   filter(analyte_dtxsid %in% this_chem_v,
                          species %in% "rat",
                          administration_route_normalized == "iv"))

new_pk <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(
    keep_data_original = TRUE,
    suppress.messages = FALSE
  ) +
  scale_conc(dose_norm = TRUE, log10_trans = TRUE) +
  stat_error_model(error_group = vars(Chemical, Species)) +
  stat_model(model = c(
    "model_1comp", "model_1comp_cl_rest", 
    "model_1comp_cl_nonrest", "model_1comp_fup",
    "model_flat")) +
  settings_optimx(method = c("bobyqa"))

new_pk <- do_data_info(new_pk)

new_pk <- do_prefit(new_pk)
gc()

unique(new_pk$prefit$par_DF$param_name) # Check all parameters are there
length(unique(new_pk$data$Chemical)) 
length(unique(this_chem_v))

new_pk <- do_fit(new_pk, n_cores = 14)
gc()

new_pk$fit %>%
  View()

# saveRDS(new_pk, "data-raw/ratCVT_wHTTK_reformulated.rds")
```

Here we begin the analysis by estimating the relative likelihood of
"clearance aware" models. This provides an additional evaluation of the evidence
that points towards a restrictive or non-restrictively cleared model.

```{r begin_analysis}
# new_pk <- readRDS("data-raw/ratCVT_wHTTK_reformulated.rds")
coef(new_pk) %>% View()
new_winmodel <- get_winning_model(new_pk)
eval_tkstats(new_pk, finite_only = TRUE) %>% 
  semi_join(new_winmodel) %>%
  View()

AIC(new_pk) %>%
  arrange(Chemical, Species) %>%
  group_by(model, method) %>% 
  count(is.finite(AIC))

new_tkstats <- eval_tkstats(new_pk, finite_only = TRUE)

pl2comp <- plot(new_pk,
                use_scale_conc = FALSE,
                n_interp = 12
)

pl2comp %>%
  filter(Chemical %in% "DTXSID4020533") %>%
  pull(final_plot)

rel_ll <- AIC(new_pk) %>% 
  filter(is.finite(AIC)) %>% 
  dplyr::select(-c(log_likelihood, npar)) %>% 
  pivot_wider(names_from = model,
              values_from = AIC) %>% 
  rowwise() %>% 
  mutate(minAIC = min(c_across(starts_with("model_")), na.rm = TRUE)) %>% 
  ungroup() %>%
  mutate(across(starts_with("model_"), \(x) {
    signif(exp((minAIC - x)/2) * 100, 4)
  }))

# Test whether the flat model has lowest relative likelihood
rel_ll %>%
  dplyr::select(-c(Species, method))
# We can inspect this visually
rel_ll %>% arrange(desc(model_flat)) %>% head()
rel_ll %>%
  dplyr::select(-c(Species, method)) %>% 
  rowwise() %>% 
  mutate(across(contains("_1comp"),
                \(x) {
                  ifelse(
                    x < model_flat | x < 1,
                    NA_real_,
                    x
                  )
                })) %>% 
  filter(!if_all(contains("_1comp_"), is.na)) %>%
  clipr::write_clip()

rel_ll %>%
  rowwise() %>% 
  mutate(across(contains("_1comp"),
                \(x) {
                  ifelse(
                    x < model_flat | x < 1,
                    NA_real_,
                    x
                  )
                })) %>% 
  count(is.na(model_1comp_fup))
  # drop_na(model_1comp_fup) %>%
  # filter(Chemical %in% clear_chem_levels) %>% View()

new_pk$fit %>%
  filter(param_name %in% "Fup",
         model %in% "model_1comp_fup") %>%
  ggplot(aes(x = estimate, y = start)) +
  geom_point(size = 1.25) +
  theme_bw() +
  coord_equal(xlim = c(0,1), ylim = c(0,1),
              expand = TRUE) +
  coord_flip()

ggsave(filename = paste0(
  Sys.getenv("FIG_DIR"), "RestrictiveClearance_Analysis/",
  today, "fup_optimization_allowingHuman.png"),
  height = 5, width = 5)

new_pk$fit %>%
  filter(param_name %in% "Fup",
         model %in% "model_1comp_fup") %>%
  distinct(Chemical, Species, method, model,
           param_name, estimate, start) %>%
  mutate(direction = ifelse(abs(estimate - start) < 0.1,
                            "similar (less than 10% difference)",
                            ifelse(estimate > start,
                                   "increase",
                                   "decrease"))) %>%
  # filter(direction %in% "similar (less than 10% difference)") %>%
  pivot_longer(cols = c(estimate, start),
               names_to = "parameterization",
               values_to = "value") %>%
  mutate(parameterization = factor(parameterization,
                                   levels = c("start", "estimate"))) %>%
  ggplot(aes(x = parameterization, y = value)) +
  geom_point() +
  geom_line(aes(group = Chemical)) +
  facet_grid(cols = vars(direction)) +
  labs(title = "Optimization of fraction unbound in plasma") +
  theme_bw()

ggsave(filename = paste0(
  Sys.getenv("FIG_DIR"), "RestrictiveClearance_Analysis/",
  today, "fup_optimization_generalFits_allowingHuman.png"),
  height = 5, width = 7)

```

Why are so many fits failing?
Well, we should take a look at the parameter which we are holding constant,
`kelim` and whether the new starting points vastly differ from those in the
starting points based on the data.

For this I will just use the PK object I have saved `this_pk` and take the
_prefit_ parameter estimations and first compare that with eventual optimized `kelim`.

```{r kelim-questions}
# Original data, filtered for common_chems
# par_DF
this_kelim_prefit <- this_pk$prefit$par_DF %>%
  filter(
    param_name %in% "kelim",
  ) %>%
  distinct(Chemical, Species, param_name, lower_bound, upper_bound, start)

this_kelim_fit <- get_fit(this_pk) %>%
  semi_join(get_winning_model(this_pk)) %>%
  ungroup() %>%
  filter(param_name == "kelim") %>%
  distinct(model, Chemical, Species, param_name, value)

this_kelim_df <- inner_join(this_kelim_fit, this_kelim_prefit)

httk_df <- new_pk$prefit$par_DF %>%
  filter(param_name %in% "kelim") %>%
  distinct(Chemical, Species, param_name, start) %>%
  rename(init_restrictive = "start") %>%
  inner_join(
    nonrest_pk$prefit$par_DF %>%
      filter(param_name %in% "kelim") %>%
      distinct(Chemical, Species, param_name, start) %>%
      rename(init_nonrestrictive = "start")
  )

full_df <- inner_join(this_kelim_df, httk_df) %>%
  relocate(fit_kelim, .after = last_col())

full_df %>%
  ggplot(aes(x = Chemical)) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "grey30") +
  geom_errorbar(aes(ymax = upper_bound, ymin = lower_bound)) +
  geom_point(aes(y = fit_kelim, color = "Original Optimized")) +
  geom_point(aes(y = init_restrictive, color = "Restrictive")) +
  geom_point(aes(y = init_nonrestrictive, color = "Non-restrictive")) +
  geom_point(aes(y = start, shape = "Original Initialized")) +
  facet_grid(rows = vars(Species), scales = "free", space = "free") +
  scale_color_manual(
    name = "kelim",
    breaks = c("Original Optimized", "Restrictive", "Non-restrictive"),
    values = c(
      "Original Optimized" = "skyblue4",
      "Restrictive" = "orange2",
      "Non-restrictive" = "green4"
    )
  ) +
  scale_shape_manual(
    name = "",
    values = 4
  ) +
  coord_flip() +
  scale_y_log10(labels = scales::label_math(format = log10)) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1.5),
    panel.background = element_blank(),
    panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold"),
    legend.direction = "vertical",
    axis.ticks = element_blank(),
    axis.line = element_blank(),
    axis.text.y = element_blank(),
    strip.text.y = element_text(angle = 0)
  )

ggsave(
  filename = paste0(Sys.getenv("FIG_DIR"), "kelim_compare.png"),
  height = 11,
  width = 8.5
)
```

```{r new_analysis}
# Loading the new data

new_wm <- get_winning_model(new_pk)
new_AIC <- AIC(new_pk)
new_rsq <- rsq(new_pk)
new_rmse <- rmse(new_pk, use_scale_conc = FALSE)

new_tkstats <- eval_tkstats(new_pk, finite_only = FALSE)

new_resid <- residuals(new_pk)

this_data <- get_data(new_pk) %>%
  distinct(Chemical, Species, Reference, Route, Media, Dose, Time, Conc) %>%
  right_join(new_resid)

pl <- plot(new_pk, n_interp = 12, use_scale_conc = FALSE)
pdf(
  file = paste0(
    Sys.getenv("FIG_DIR"),
    "AllFits_nonDoseNorm_newPK_reParametizedModels_2024_0412.pdf"
  ),
  height = 6, width = 10
)
for (i in seq_len(nrow(pl))) {
  print(pl$final_plot[[i]])
}
dev.off()
```

Comparing total clearance estimates for those where a restrictive or non-restrictive model
was the "best fit" as determined by AIC.

```{r cltot-comparison}
new_cltot2 <- new_tkstats %>%
  filter(str_detect(model, "rest")) %>%
  select(
    Chemical, Species, Media, Route,
    model, method,
    CLtot.tkstats, CLtot.nca
  ) %>%
  filter(!is.na(CLtot.tkstats)) %>%
  mutate(clearance_type = ifelse(str_detect(model, "nonrest"),
                                 "NonRestrictive",
                                 ifelse(str_detect(model, "rest"),
                                        "Restrictive",
                                        "optimized kelim"
                                 )
  )) %>%
  rowwise() %>%
  mutate(
    CLtot.httk_rest = calc_total_clearance(
      dtxsid = Chemical,
      restrictive.clearance = TRUE
    ),
    CLtot.httk_nonrest = calc_total_clearance(
      dtxsid = Chemical,
      restrictive.clearance = FALSE
    )
  )

new_cltot2 %>%
  ggplot(aes(x = Chemical)) +
  geom_point(aes(y = CLtot.httk_rest),
             shape = 19, size = 2, color = "green4") +
  geom_point(aes(y = CLtot.httk_nonrest),
             shape = 19, size = 2, color = "orange2") +
  geom_point(aes(y = CLtot.tkstats,
                 shape = Species), size = 3, color = "black", stroke = 1) +
  scale_y_log10(labels = scales::label_math(format = log10)) +
  geom_hline(yintercept = c(0.01, 0.1, 1, 10)) +
  scale_shape_manual(values = c(0, 1, 2, 5, 6)) +
  coord_flip() +
  facet_grid(
    rows = vars(clearance_type),
    scales = "free", space = "free"
  ) +
  labs(title = paste(
    "CLtot for constant kelim \"best fits\" (empty)",
    "and httk calculations (filled)",
    sep = "\n"
  )) +
  theme_bw() +
  theme(
    axis.line = element_line(linewidth = 1),
    axis.text = element_text(size = 12),
    axis.title = element_blank(),
    plot.title = element_text(hjust = 0.5),
    plot.background = element_rect(fill = "white"),
    strip.text.y = element_text(angle = 0),
    strip.background = element_blank(),
    legend.position = "none"
  )

ggsave(
  filename = paste0(Sys.getenv("FIG_DIR"),
                    "Cltot_backcompare_wDawson_2024_0416.png"),
  height = 8,
  width = 9
)
```

```{r chemical-correlation}
new_AIC %>%
  ungroup() %>%
  filter(
    model %in% c("model_1comp", "model_2comp"),
    Species %in% "rat"
  ) %>%
  group_by(Chemical) %>%
  filter(AIC == min(AIC)) %>%
  mutate(estimate = "nonConst") %>%
  distinct(Chemical, model, estimate) -> nAIC_nonConst
new_AIC %>%
  ungroup() %>%
  filter(
    model %in% c("model_1comp_rest", "model_2comp_rest"),
    Species %in% "rat"
  ) %>%
  group_by(Chemical) %>%
  filter(AIC == min(AIC)) %>%
  mutate(estimate = "Rest") %>%
  distinct(Chemical, model, estimate) -> nAIC_rest
new_AIC %>%
  ungroup() %>%
  filter(
    model %in% c("model_1comp_nonrest", "model_2comp_nonrest"),
    Species %in% "rat"
  ) %>%
  group_by(Chemical) %>%
  filter(AIC == min(AIC)) %>%
  mutate(estimate = "nonRest") %>%
  distinct(Chemical, model, estimate) -> nAIC_nonrest

full_nAIC <- bind_rows(nAIC_nonConst, nAIC_rest) %>%
  bind_rows(nAIC_nonrest)

new_tkstats_full <- get_tkstats(new_pk)

new_tkstats_full %>%
  filter(Species %in% "rat", !(model %in% "model_flat")) %>%
  distinct(Chemical, model, CLtot) %>%
  inner_join(full_nAIC) %>%
  na.omit() %>%
  distinct(Chemical, estimate, CLtot) %>%
  pivot_wider(
    names_from = estimate,
    values_from = CLtot
  ) %>%
  rowwise() %>%
  mutate(
    CLtot.httk_rest = calc_total_clearance(
      dtxsid = Chemical,
      restrictive.clearance = TRUE
    ),
    CLtot.httk_nonrest = calc_total_clearance(
      dtxsid = Chemical,
      restrictive.clearance = FALSE
    )
  ) %>%
  ungroup() %>%
  mutate(
    A_rest = log(nonConst) - log(CLtot.httk_rest),
    A_nonrest = log(nonConst) - log(CLtot.httk_nonrest),
    B_rest = log(Rest) - log(CLtot.httk_rest),
    B_nonrest = log(Rest) - log(CLtot.httk_nonrest),
    C_rest = log(nonRest) - log(CLtot.httk_rest),
    C_nonrest = log(nonRest) - log(CLtot.httk_nonrest)
  ) %>%
  select(Chemical, A_rest:C_nonrest) -> df_valcomp

df_valcomp %>%
  column_to_rownames(var = "Chemical") %>%
  t() %>%
  as.matrix() %>%
  cor() %>%
  signif() -> val_cormat

hclust(d = as.dist((1 + val_cormat) / 2), method = "average") %>% plot()

hclust(d = as.dist((1 + val_cormat) / 2), method = "average") -> val_hclust

corrplot::corrplot(abs(val_cormat))

# reorder the Chemical names prior to plotting
df_valcomp %>%
  mutate(Chemical = factor(
    x = Chemical,
    levels = val_hclust$labels[val_hclust$order]
  )) %>%
  pivot_longer(
    cols = A_rest:C_nonrest,
    names_to = "comparison", values_to = "logres"
  ) %>%
  mutate(
    comparison = case_when(
      comparison == "A_nonrest" ~ "free kelim vs httk nonrestrictive",
      comparison == "A_rest" ~ "free kelim vs httk restrictive",
      comparison == "B_nonrest" ~ "restrictive kelim vs httk nonrestrictive",
      comparison == "B_rest" ~ "restrictive kelim vs httk restrictive",
      comparison == "C_nonrest" ~ "nonrestrictive kelim vs httk nonrestrictive",
      comparison == "C_rest" ~ "nonrestrictive kelim vs httk restrictive",
    ),
    comparison = factor(comparison,
                        levels = c(
                          "free kelim vs httk restrictive",
                          "free kelim vs httk nonrestrictive",
                          "restrictive kelim vs httk restrictive",
                          "nonrestrictive kelim vs httk nonrestrictive",
                          "nonrestrictive kelim vs httk restrictive",
                          "restrictive kelim vs httk nonrestrictive"
                        )
    )
  ) %>%
  ggplot(aes(x = comparison, y = Chemical, fill = logres)) +
  geom_tile() +
  scale_fill_gradient2(
    mid = "grey5", high = "orange", low = "green3",
    name = "Total Clearance\nlog10-scale difference",
    limits = c(-4, 4),
    oob = scales::oob_squish
  ) +
  coord_fixed(ratio = 0.3) +
  theme_bw() +
  theme(
    axis.text.x = element_text(hjust = 1, angle = 45),
    axis.ticks.y = element_blank(),
    panel.grid = element_blank(),
    axis.title.x = element_blank()
  )

ggsave(
  filename = paste0(Sys.getenv("FIG_DIR"), "Cltot_compare_heatmap_2024_0409.png"),
  height = 14,
  width = 8
)
```

# Winning models - An aside

I am a little worried that I am not selecting the best fit model for each
chemical all the time. Ultimately, the consequences for this being having parameters
for chemicals with rather poor fits in general, when there is a better fitting
model available. Here I want to do a quick analysis of the chemical
*DTXSID3022405* in mouse. The one compartment fits best in the original data
(by visual inspection) but the 2-compartment wins out in minimizing AIC.

For each model, I can find how many distinct experimental timepoints have the minimum
RMSE then investigate whether the model with the most _minimum RMSE_ timepoint
corresponds to the winning model by AIC.

*Update*: About half do not correspond. However, the lack of correspondance doesn't
mean there is bad fits declared winning model. To address this I need to understand 
_how_ different the AIC-based winning model and the RMSE-based winning models
are and whether the data leaves something to be desired.
This may be as simple as a Pearson correlation between residual values.
The output of this analysis is how related both

Another thing I am noticing in oral data: the constant _kelim_ fits sometimes are very similar
to their respective non-constant _kelim_ fits and they actually intersect/cross over.
I wonder if there is something special about the timepoints at which they intersect.
For example, a restrictive fit will reach a slightly higher Cmax, but then
dive below the fully optimized fit during elimination phase before crossing over
again near the second inflection point.

```{r oxybenzene-test}
old_AIC <- AIC(this_pk) %>%
  filter(
    Chemical %in% "DTXSID3022405",
    Species %in% "mouse"
  )

new_AIC <- new_AIC %>%
  filter(
    Chemical %in% "DTXSID3022405",
    Species %in% "mouse"
  )

plot(new_pk,
     model = c(
       "model_2comp",
       "model_1comp", "model_1comp_nonrest"
     ),
     use_scale_conc = FALSE
) %>%
  filter(
    Chemical %in% "DTXSID3022405",
    Species %in% "mouse"
  ) %>%
  pull(final_plot) %>%
  .[[1]]

old_rsquared <- rmse(this_pk) %>%
  filter(
    Chemical %in% "DTXSID3022405",
    Species %in% "mouse"
  )

# Only test residual correlations with those not the same

winmodel_conflict <- rmse(new_pk, use_scale_conc = FALSE) %>%
  group_by(Chemical, Species, Route, Media, Dose, Time) %>%
  filter(RMSE == min(RMSE)) %>%
  group_by(Chemical, Species) %>%
  count(model) %>%
  filter(n == max(n)) %>%
  inner_join(new_wm %>% rename("winmodel" = "model")) %>%
  filter(!(model == winmodel))

# need to map2 then have `residuals.pk(model = .x)` and .y
# then correlate these values for each data group
winmodel_conflict %>% names()

mod_red <- winmodel_conflict %>%
  dplyr::select(Chemical, Species, model) %>%
  left_join(new_resid,
            by = join_by(Chemical, Species, model)
  ) %>%
  rename("mod_residuals" = "Residuals") %>%
  select(
    Chemical, Species, Route, Media, Time, Detect, exclude,
    model, mod_residuals
  )

wm_red <- winmodel_conflict %>%
  dplyr::select(Chemical, Species, winmodel) %>%
  left_join(new_resid,
            by = join_by(Chemical, Species, "winmodel" == "model")
  ) %>%
  rename("wm_residuals" = "Residuals") %>%
  select(
    Chemical, Species, Route, Media, Time, Detect, exclude,
    winmodel, wm_residuals
  )
sum_red <- new_resid %>%
  group_by(Chemical, Species, model) %>%
  summarize(MSE = sum(Residuals^2) / n())

cor_mod <- mod_red %>%
  inner_join(wm_red) %>%
  select(Chemical, Species, model, mod_residuals, winmodel, wm_residuals) %>%
  group_by(Chemical, Species, model, winmodel) %>%
  summarize(cor_mods = cor(x = mod_residuals, y = wm_residuals)) %>%
  left_join(sum_red, by = join_by(Chemical, Species, model)) %>%
  rename("mod_MSE" = "MSE") %>%
  left_join(sum_red, by = join_by(Chemical, Species, "winmodel" == "model")) %>%
  rename("wm_MSE" = "MSE")

this_pl_1comp <- plot(new_pk,
                    n_interp = 11, use_scale_conc = FALSE,
                    model = c(
                      "model_1comp",
                      "model_1comp_rest",
                      "model_1comp_nonrest"
                    )
)

this_pl_2comp <- plot(new_pk,
                    n_interp = 11, use_scale_conc = FALSE,
                    model = c(
                      "model_2comp",
                      "model_2comp_rest",
                      "model_2comp_nonrest"
                    )
)

# General trends about residual correlation and MSE
cor_mod %>%
  mutate(cor_bin = ifelse(abs(cor_mods) >= 0.60,
                          "Taking win_model",
                          "Taking low RMSE model"
  )) %>%
  na.omit() %>%
  ggplot() +
  geom_point(aes(x = mod_MSE, y = wm_MSE)) +
  scale_x_log10() +
  scale_y_log10() +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted") +
  facet_grid(rows = vars(cor_bin))

cor_mod %>%
  filter(!(model %in% "model_flat")) %>%
  mutate(cor_bin = ifelse(abs(cor_mods) >= 0.60,
                          "Taking win_model",
                          "Taking low RMSE model"
  )) %>%
  na.omit() %>%
  ggplot() +
  geom_histogram(aes(x = cor_mods, fill = model),
                 binwidth = 0.05
  ) +
  facet_grid(rows = vars(winmodel)) +
  scale_fill_viridis_d() +
  theme_bw() +
  theme(strip.text.y = element_text(angle = 0)) +
  labs(
    x = "Correlation of Residuals",
    fill = "model by grouped RMSE"
  )

ggsave(
  filename = paste0(
    Sys.getenv("FIG_DIR"), "winModel_correlations_changes",
    "_2024_Apr_16.png"
  ),
  height = 6,
  width = 8
)

# Try oxybenzone fitting without dose-norm
minimal_pk <- pk(
  data = cvt %>%
    filter(analyte_dtxsid %in% "DTXSID3022405"),
  mapping = ggplot2::aes(
    Chemical = analyte_dtxsid,
    Chemical_Name = analyte_name_original,
    DTXSID = analyte_dtxsid,
    CASRN = analyte_casrn,
    Species = species,
    Reference = document_id,
    Media = conc_medium_normalized,
    Route = administration_route_normalized,
    Dose = dose_level_normalized,
    Dose.Units = "mg/kg",
    Subject_ID = subject_id,
    Series_ID = series_id,
    Study_ID = study_id,
    ConcTime_ID = conc_time_id,
    N_Subjects = n_subjects_in_series,
    Weight = weight_kg,
    Weight.Units = "kg",
    Time = time_hr,
    Time.Units = "hours",
    Value = conc,
    Value.Units = "mg/L",
    Value_SD = conc_sd_normalized,
    LOQ = loq
  )
)
new_pk <- minimal_pk +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(
    keep_data_original = TRUE,
    suppress.messages = FALSE
  ) +
  settings_optimx(method = c("L-BFGS-B")) +
  scale_conc(dose_norm = FALSE, log10_trans = FALSE) +
  stat_error_model(error_group = vars(Chemical, Species, Reference)) +
  stat_model(model = c(
    "model_flat", "model_1comp", "model_2comp",
    "model_1comp_rest", "model_1comp_nonrest",
    "model_2comp_rest", "model_2comp_nonrest"
  ))

new_pk <- do_fit(new_pk, n_cores = 3)
get_winning_model(new_pk)
plot(new_pk, best_fit = TRUE) %>% pull(final_plot)
```

# Categorizing chemicals

I first need to find the winning model for constant kelims and optimized kelims

```{r categorization}
grouped_wm <- new_AIC %>%
  mutate(model_type = case_when(
    str_detect(model, "_nonrest$") ~ "Nonrestrictive",
    str_detect(model, "_rest$") ~ "Restrictive",
    .default = "Free_kelim"
  )) %>%
  group_by(Chemical, Species, model_type) %>%
  filter(AIC == min(AIC)) %>%
  ungroup() %>%
  arrange(Chemical, Species, model_type) %>%
  select(Chemical, Species, model, method, model_type)

grouped_tkstats <- get_tkstats(new_pk) %>%
  inner_join(grouped_wm)

grouped_tkstats %>%
  select(
    Chemical, Species, Reference,
    Route, Media, model_type,
    CLtot
  ) %>%
  pivot_wider(names_from = model_type, values_from = CLtot) %>%
  mutate(
    low_invivo_CL = ifelse(
      Free_kelim < Nonrestrictive / 2 | Free_kelim < Restrictive / 2,
      TRUE, FALSE
    ),
    high_extra_hepCL = ifelse(
      Free_kelim > 1.5 * Nonrestrictive | Free_kelim > 1.5 * Restrictive,
      TRUE, FALSE
    ),
    rest_metabol = ifelse(
      Free_kelim > Restrictive / 2 & Free_kelim < 1.5 * Nonrestrictive,
      TRUE, FALSE
    ),
    combo_cat = paste0(
      low_invivo_CL + 0,
      high_extra_hepCL + 0,
      rest_metabol + 0
    )
  ) %>%
  count(combo_cat)
arrange(Chemical, combo_cat) %>%
  na.omit() %>%
  View()
```

## New reformulated models that estimate Fup!

```{r reformMod}
get_tkstats(new_pk) %>%
  distinct(Chemical, Species, model, CLtot) %>%
  pivot_wider(names_from = model, values_from = CLtot) %>%
  left_join(new_pk$prefit$par_DF %>%
              filter(param_name == "CLtot") %>%
              select(Chemical, Species, model, start) %>%
              mutate(model = paste0("start_", model)) %>%
              pivot_wider(
                names_from = model,
                values_from = start
              )) %>%
  View("Clearance")

new_cl_tkstats <- get_tkstats(new_pk)
this_tkstats <- this_pk %>%
  get_winning_model() %>%
  left_join(coef(this_pk)) %>%
  left_join(rsq(this_pk)) %>%
  distinct()
this_tkstats %>%
  select(-coefs_vector) %>%
  clipr::write_clip()
# Send the R2 too.

## Need to parameterize with gas_pbpk or new 3compartment
## To add into total clearance
# Rb2p * Qalv/Kblood2air
```

# Calculating uncertainty of parameter estimation

Below I will use the `coef_sd` method to estimate standard deviation of
the fraction unbound parameter and assess whether this may contribute
to the uncertainty of 'declaration' of clearance for chemicals that have
CvT data documented.

```{r uncertainty_estimation}
this_coefsd <- coef_sd(new_pk)

this_coefsd %>% filter(param_name == "Clint", Chemical %in% clear_chem_levels) %>%
  ggplot(aes(x = factor(Chemical, levels = clear_chem_levels), 
             y = param_value,
             color = model)) +
  geom_point(alpha = 0.5) +
  geom_linerange(aes(ymin = param_value-param_sd,
                    ymax = param_value+param_sd),
                 alpha = 0.5) +
  theme_bw() +
  coord_flip() +
  scale_y_log10()

ggsave(
  filename = paste0(
    Sys.getenv("FIG_DIR"),"RestrictiveClearance_Analysis/",
    today,
    "fup_sd_estimates.png"
  ),
  height = 8,
  width = 8,
  device = "png"
)

this_loglikelihood <- AIC(new_pk) %>%
  filter(Chemical %in% clear_chem_levels) %>%
  mutate(Chemical = factor(Chemical, levels = clear_chem_levels))

this_loglikelihood %>%
  ggplot(aes(x = model, y = Chemical, fill = log10(AIC))) +
  geom_tile() +
  scale_fill_viridis_c(option = "mako", direction = -1) +
  coord_fixed() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

ggsave(
  filename = paste0(
    Sys.getenv("FIG_DIR"),"RestrictiveClearance_Analysis/",
    today,
    "fup_AIC_heatmap.png"
  ),
  height = 10,
  width = 4,
  device = "png"
)

select_chem_info <- get_chem_info_batch(clear_chem_levels)

select_chem_info %>%
  filter(propertyId == "logkow-octanol-water") %>%
  mutate(dtxsid = factor(dtxsid, levels = clear_chem_levels)) %>%
  ggplot(aes(x = value, y = dtxsid, color = source)) +
  geom_point() +
  labs(title = "logKow") +
  geom_vline(xintercept = 5)

# Also check whether the Fup or CLint are at their respective extremes (1, 0)
# If they are, it will be difficult to tell whether a chemical is cleared in a
# particular way

```

# Slow PBPK model optimization of fup for more chemicals

There may also be an opportunity to evaluate chemicals without _in vivo_ data
by comparing them to those that do. Effectively, we can look at the uncertainty
in the call between restrictive/non-restrictive and further look at estimates
from PBPK models where $f_{up}$ is optimized and it is run through `solve_pbtk`.

```{r slow_pbpk_optimizations}

```

```{r}
sessionInfo()
```