### 2025-03-11 data-raw/ClearanceProject-1-DataRmd.Rmd
Renamed and initialized 4 analysis documents for clearance project.

---
title: "Clearance Project - 1 - Data Generation"
author: "Gilberto Padilla Mercado"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

This vignette creates and describes data for downstream analysis.

```{r library_loading}
today <- format(Sys.Date(), "%d%b%Y")
devtools::load_all()
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(ctxR)
library(ggvenn)
library(ggalluvial)
library(flextable)
library(reticulate)
library(patchwork)
source("data-raw/RC_comparison_bkgd.R") # httk & dawson_2021 loaded here
```

```{r chemical-lists}
member_list <- list(
  httk_3comp2 = unique(get_cheminfo(info = "DTXSID",
                                species = "human",
                                model = "3compartment2",
                                median.only = TRUE,
                                physchem.exclude = TRUE,
                                suppress.messages = TRUE)) %>% 
    suppressWarnings(),
  cvt = cvt %>% 
    filter(species %in% c("human", "rat"),
           conc_medium_normalized == "plasma") %>%
    pull(analyte_dtxsid)
)
# These are the chemicals we can evaluate!
target_chems <- intersect(member_list$cvt, member_list$httk_3comp2)
all_chems <- member_list$httk_3comp2

# save in an .Rdata file
save(target_chems, all_chems,
     file = "data-raw/chem_lists.RData")

```

```{r external_data}
# Here we load Kd/binding affinity data
red_assay_res <- readRDS("data-raw/2019-05-06-REDassay_HTTK.rds")

# ctxR download of all chemicals in 3comp2 model
# Chemical info and properties (averaging the latter by chemical and experimental/predicted)
chem_info <- get_chemical_details_batch(DTXSID = all_chems,
                                               rate_limit = 0L) %>%
  filter(!is.na(smiles))
# 9086 total unique chemicals with SMILES data

chem_properties <- get_chem_info_batch(
  DTXSID = chem_info$dtxsid,
  type = '', 
  rate_limit = 0L,
  verbose = TRUE
) 
# This takes a minute or so... save a reference RDS file to load
# Also get rid of any that don't have SMILES
chem_info %>% distinct(dtxsid, preferredName,
                              qsarReadySmiles, monoisotopicMass) %>%
  right_join(chem_properties) %>%
  filter(!is.na(qsarReadySmiles)) %>%
  saveRDS(file = "data-raw/ctxR_3comp2_cheminfo.rds")

functional_uses_all <- readr::read_csv(paste0(
  Sys.getenv("FIG_DIR"), "RestrictiveClearance_Analysis/",
  "ChemExpo_bulk_functional_uses_20250304.csv"
))

```

```{r inivopkfit}
# Here we fit CvTdb data that has plasma measurements
# Only use plasma for simplicity
this_pk <- pk(cvt %>% 
                dplyr::filter(conc_medium_normalized == "plasma")
              ) +
  settings_optimx(method = "bobyqa") +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(
    keep_data_original = TRUE,
    suppress.messages = FALSE) +
  scale_conc(dose_norm = TRUE, log10_trans = TRUE) +
  stat_error_model(error_group = vars(Chemical, Species))

this_pk <- do_prefit(this_pk)
this_pk <- do_fit(this_pk, n_cores = 14)
saveRDS(this_pk, "data-raw/plasma_CvTfits.rds")

# Related files for pre-calculated:
# tkstats, winning model, AIC, coef_sd, predictions, etc...
this_winmodel <- get_winning_model(this_pk)

this_tkstats <- eval_tkstats(this_pk,
                             model = c("model_1comp", "model_2comp"),
                             finite_only = TRUE,
                             dose_norm = TRUE) %>%
  inner_join(this_winmodel)
colnames(this_tkstats) <- gsub("/", "_", colnames(this_tkstats))

winning_preds <- predict(this_pk) %>%
  semi_join(this_winmodel) %>%
  distinct(Chemical, Species, Chemical_Name,
           model, method, Dose, Media, Route, Reference, pLOQ, Time,
           N_Subjects, Conc, Conc_est)
this_AIC <- AIC(this_pk)
this_coef_win <- coef(this_pk) %>% semi_join(this_winmodel)
this_coef_fit <- this_pk$fit %>% semi_join(this_winmodel)
this_coef_sd_win <- coef_sd(this_pk) %>% semi_join(this_winmodel)

save(this_winmodel, this_tkstats, winning_preds,
     this_AIC, this_coef_win, this_coef_fit, this_coef_sd_win,
     file = "data-raw/plasma_CvTfits_data.RData")
```

```{r getting_curent_params}

# rat_pars_rat <- parameterize_rat()
rat_pars_human <- parameterize_rat(human_clint_fup = TRUE)
#human_pars_human <- parameterize_human()
chosen_list <- rat_pars_human
# The following are derived from the above selected list (see environment)
this_param_rest_l <- chosen_list$named_params_restrictive
this_param_nrest_l <- chosen_list$named_params_nonrestrictive
this_param_df <- chosen_list$param_df
this_httk_clearance_df <- chosen_list$clearance_df

# HTTK predictions

# Predictions from httk(code retained but not run after 2.20.2025)
restrictive_preds <- get_httk_preds(
  parameters = this_param_rest_l,
  pk_obj = this_pk,
  species = c("rat", "human")
) %>% rename_with(
  .fn = ~paste0(.x, "_restrictive"),
  .cols = starts_with("httk_")
)

nonrestrictive_preds <- get_httk_preds(
  parameters = this_param_nrest_l,
  pk_obj = this_pk,
  species = c("rat", "human")
) %>% rename_with(
  .fn = ~paste0(.x, "_nonrestrictive"),
  .cols = starts_with("httk_")
)

names(restrictive_preds)

all_httk_preds <- inner_join(restrictive_preds, nonrestrictive_preds) %>%
  filter_targets()
rm(restrictive_preds, nonrestrictive_preds)

this_css_df <- inner_join(
  css_httk_batch(this_param_rest_l, restrictive = TRUE),
  css_httk_batch(this_param_nrest_l, restrictive = FALSE)
) %>% filter_targets()

this_clint_df <- data.frame(Chemical = names(this_param_rest_l),
                            CLint = vapply(names(this_param_rest_l),
                                           \(x) {
                                             this_param_rest_l[[x]][["Clint"]]
                                           },
                                           FUN.VALUE = numeric(1L)
                            ))

save(this_param_rest_l, this_param_nrest_l,
     this_param_df, this_httk_clearance_df,
     all_httk_preds, this_css_df, this_clint_df,
     file = "data-raw/rat_pars_human.RData")
```