### 2025-03-11 data-raw/ClearanceProject-3-TKstats.Rmd

---
title: "Clearance Project - 3 - TKstats"
author: "Gilberto Padilla Mercado"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```