### 2025-03-19 data-raw/ClearanceProject-1-DataRmd.Rmd
Updated DESCRIPTION and data.

---
title: "Clearance Project - 1 - Data Generation"
author: "Gilberto Padilla Mercado"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

This vignette creates and describes data for downstream analysis.

```{r library_loading}
today <- format(Sys.Date(), "%d%b%Y")
devtools::load_all()
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(ctxR)
library(ggvenn)
library(ggalluvial)
library(flextable)
library(reticulate)
library(patchwork)
source("data-raw/RC_comparison_bkgd.R") # httk & dawson_2021 loaded here
```

This code chunk creates and `.RData` file that contains character vectors
of DTXSIDs for:

1. All chemicals with human parameterization available for a 3-compartment model w/ inhalation.  
2. All chemicals with rat parameterization available for a 3-compartment model w/ inhalation.  
3. All chemicals with human and rat plasma concentration over time data.

```{r chemical-lists}
human_3compchems <- unique(get_cheminfo(info = "DTXSID",
                                species = "human",
                                model = "3compartment2",
                                median.only = TRUE,
                                physchem.exclude = TRUE,
                                suppress.messages = TRUE)) %>% 
    suppressWarnings()

rat_3compchems <- unique(get_cheminfo(info = "DTXSID",
                                species = "rat",
                                model = "3compartment2",
                                median.only = TRUE,
                                physchem.exclude = TRUE,
                                suppress.messages = TRUE)) %>% 
    suppressWarnings()

cvt_plasma_chems <-  cvt %>% 
    filter(species %in% c("human", "rat"),
           conc_medium_normalized == "plasma") %>%
    pull(analyte_dtxsid)

# These are the chemicals we can evaluate!
target_chems_human <- intersect(cvt_plasma_chems,
                                human_3compchems)
target_chems_rat <- intersect(cvt_plasma_chems,
                              rat_3compchems)

# save in an .Rdata file
save(target_chems_human, target_chems_rat, 
     human_3compchems, rat_3compchems,
     file = "data-raw/chem_lists.RData")

```

Below read in some external data which is publicly available already.

- `red_assay_res` is the named object that contains the data from the RED assay
including the dissociation constant $K_d$ for plasma protein binding.  
- `chem_info` is a combined data.frame from calls to `get_chemical_details_batch`
and `get_chem_info_batch`. This uses the `ctxR` API to fetch chemical information
such as physio-chemical properties and SMILES from the CompTox dashboard.  
- `cyp_data` is data from Sipes 2017 predicting the CLint fraction for each CYP enzyme.
- `functional_uses_all` is a list of functional use categories for each chemical
from the Chemical and Products Database.

```{r external_data}
# Here we load Kd/binding affinity data
red_assay_res <- readRDS("data-raw/2019-05-06-REDassay_HTTK.rds")

# ctxR download of all chemicals in 3comp2 model
# Chemical info and properties (averaging the latter by chemical and experimental/predicted)
chem_info <- get_chemical_details_batch(DTXSID = human_3compchems,
                                               rate_limit = 0L) %>%
  filter(!is.na(smiles))
# 9086 total unique chemicals with SMILES data

chem_properties <- get_chem_info_batch(
  DTXSID = chem_info$dtxsid,
  type = '', 
  rate_limit = 0L,
  verbose = TRUE
) 
# This takes a minute or so... save a reference RDS file to load
# Also get rid of any that don't have SMILES
chem_info %>% distinct(dtxsid, preferredName,
                              qsarReadySmiles, monoisotopicMass) %>%
  right_join(chem_properties) %>%
  filter(!is.na(qsarReadySmiles)) %>%
  saveRDS(file = "data-raw/RCdata/ctxR_3comp2_cheminfo.rds")

# Sipes 2017 data
cyp_data <- readr::read_csv(paste0(
  Sys.getenv("FIG_DIR"),
  "RestrictiveClearance_Analysis/",
  "CLint_CYP_Sipes2017.csv"
)) %>%
  left_join(
    chem.physical_and_invitro.data %>%
      select(CAS = CAS, Chemical = DTXSID) %>%
      distinct()
  ) %>%
  select(Chemical, CAS, starts_with("MET_")) %>%
  filter(!is.na(Chemical))
names(cyp_data) <- c("Chemical", "CAS", "CYP1A2",
                     "CYP2C9", "CYP2C19", "CYP2D6", "CYP3A4")
# Functional use categories from CPdat
functional_uses_all <- readr::read_csv(paste0(
  Sys.getenv("FIG_DIR"), "RestrictiveClearance_Analysis/",
  "ChemExpo_bulk_functional_uses_20250304.csv"
))

save(cyp_data,
     file = "data-raw/CYP_Sipes2017.Rdata")

```

Next we fit the plasma data from CvTdb, which we already have in the `invivoPKfit`
package using `cvt` object. We also get a few downstream statistics and save them
in an `.RData` file.

```{r inivopkfit}
# Here we fit CvTdb data that has plasma measurements
# Only use plasma for simplicity
this_pk <- pk(cvt %>% 
                dplyr::filter(conc_medium_normalized == "plasma")
              ) +
  settings_optimx(method = "bobyqa") +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(
    keep_data_original = TRUE,
    suppress.messages = FALSE) +
  scale_conc(dose_norm = TRUE, log10_trans = TRUE) +
  stat_error_model(error_group = vars(Chemical, Species))

this_pk <- do_prefit(this_pk)
this_pk <- do_fit(this_pk, n_cores = 14)
saveRDS(this_pk, "data-raw/plasma_CvTfits.rds")

# Related files for pre-calculated:
# tkstats, winning model, AIC, coef_sd, predictions, etc...
this_winmodel <- get_winning_model(this_pk)

this_tkstats <- eval_tkstats(this_pk,
                             model = c("model_1comp", "model_2comp"),
                             finite_only = TRUE,
                             dose_norm = TRUE) %>%
  inner_join(this_winmodel)
colnames(this_tkstats) <- gsub("/", "_", colnames(this_tkstats))

winning_preds <- predict(this_pk) %>%
  semi_join(this_winmodel) %>%
  distinct(Chemical, Species, Chemical_Name,
           model, method, Dose, Media, Route, Reference, pLOQ, Time,
           N_Subjects, Conc, Conc_est)
this_AIC <- AIC(this_pk)
this_coef_win <- coef(this_pk) %>% semi_join(this_winmodel)
this_coef_fit <- this_pk$fit %>% semi_join(this_winmodel)
this_coef_sd_win <- coef_sd(this_pk) %>% semi_join(this_winmodel)

save(this_winmodel, this_tkstats, winning_preds,
     this_AIC, this_coef_win, this_coef_fit, this_coef_sd_win,
     file = "data-raw/plasma_CvTfits_data.RData")
```

The next two chunks take a while. These will parameterize the `3compartment2`
`httk` model with restrictive and nonrestrictive clearance.
It also creates data.frames for evaluating predictions, steady-state concentration,
and other PK statistics.
In the first chunk rats are limited parameterized as rats (for _fup_ and _CLint_), and in
the second chunk human parameterization of _fup_ and _CLint_ are allowed.

```{r getting_curent_params_rpr}

rat_pars_rat <- parameterize_rat(human_clint_fup = FALSE)

rpr_param_rest_l <- rat_pars_rat$named_params_restrictive
rpr_param_nrest_l <- rat_pars_rat$named_params_nonrestrictive
rpr_param_df <- rat_pars_rat$param_df
rpr_httk_clearance_df <- rat_pars_rat$clearance_df

# HTTK predictions
# Predictions from httk(code retained but not run after 2.20.2025)
restrictive_preds <- get_httk_preds(
  parameters = rpr_param_rest_l,
  pk_obj = this_pk,
  species = c("rat", "human")
) %>% rename_with(
  .fn = ~paste0(.x, "_restrictive"),
  .cols = starts_with("httk_")
)

nonrestrictive_preds <- get_httk_preds(
  parameters = rpr_param_nrest_l,
  pk_obj = this_pk,
  species = c("rat", "human")
) %>% rename_with(
  .fn = ~paste0(.x, "_nonrestrictive"),
  .cols = starts_with("httk_")
)

names(restrictive_preds)

rpr_httk_preds <- inner_join(restrictive_preds, nonrestrictive_preds) 
rm(restrictive_preds, nonrestrictive_preds)

rpr_css_df <- inner_join(
  css_httk_batch(rpr_param_rest_l, restrictive = TRUE),
  css_httk_batch(rpr_param_nrest_l, restrictive = FALSE)
)

rpr_clint_df <- data.frame(Chemical = names(rpr_param_rest_l),
                           CLint = vapply(names(rpr_param_rest_l),
                                          \(x) {
                                            rpr_param_rest_l[[x]][["Clint"]]
                                          },
                                          FUN.VALUE = numeric(1L)
                           ))
rownames(rpr_clint_df) <- NULL
rpr_param_df <- merge(rpr_param_df, rpr_clint_df)
rpr_httk_clearance_df <- merge(rpr_httk_clearance_df, rpr_css_df)

save(rpr_param_rest_l, rpr_param_nrest_l,
     rpr_param_df, rpr_httk_clearance_df,
     rpr_httk_preds, file = "data-raw/rat_pars_rat.RData")
```

```{r getting_current_params_rph}

rat_pars_human <- parameterize_rat(human_clint_fup = TRUE)
#human_pars_human <- parameterize_human()
# The following are derived from the above selected list (see environment)
rph_param_rest_l <- rat_pars_human$named_params_restrictive
rph_param_nrest_l <- rat_pars_human$named_params_nonrestrictive
rph_param_df <- rat_pars_human$param_df
rph_httk_clearance_df <- rat_pars_human$clearance_df

# HTTK predictions
# Predictions from httk(code retained but not run after 2.20.2025)
restrictive_preds <- get_httk_preds(
  parameters = rph_param_rest_l,
  pk_obj = this_pk,
  species = c("rat", "human")
) %>% rename_with(
  .fn = ~paste0(.x, "_restrictive"),
  .cols = starts_with("httk_")
)

nonrestrictive_preds <- get_httk_preds(
  parameters = rph_param_nrest_l,
  pk_obj = this_pk,
  species = c("rat", "human")
) %>% rename_with(
  .fn = ~paste0(.x, "_nonrestrictive"),
  .cols = starts_with("httk_")
)

names(restrictive_preds)

rph_httk_preds <- inner_join(restrictive_preds, nonrestrictive_preds)
rm(restrictive_preds, nonrestrictive_preds)

rph_css_df <- inner_join(
  css_httk_batch(rph_param_rest_l, restrictive = TRUE),
  css_httk_batch(rph_param_nrest_l, restrictive = FALSE)
)

rph_clint_df <- data.frame(Chemical = names(rph_param_rest_l),
                            CLint = vapply(names(rph_param_rest_l),
                                           \(x) {
                                             rph_param_rest_l[[x]][["Clint"]]
                                           },
                                           FUN.VALUE = numeric(1L)
                            ))

rownames(rph_clint_df) <- NULL
rph_param_df <- merge(rph_param_df, rph_clint_df)
rph_httk_clearance_df <- merge(rph_httk_clearance_df, rph_css_df)

save(rph_param_rest_l, rph_param_nrest_l,
     rph_param_df, rph_httk_clearance_df,
     rph_httk_preds,
     file = "data-raw/rat_pars_human.RData")
```