### 2025-03-20 data-raw/ClearanceProject-4-PhysChemClustering.Rmd

---
title: "Clearance Project - 4 - PhysChem Clustering"
author: "Gilberto Padilla Mercado"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r data_loading}
target_chem_properties <- readRDS("data-raw/RCdata/ctxR_3comp2_cheminfo.rds")
new_pk <- readRDS("data-raw/ratCVT_wHTTK_reformulated.rds")
red_assay_res <- readRDS("data-raw/2019-05-06-REDassay_HTTK.rds")
load("data-raw/CYP_Sipes2017.RData")
```

## Lipinski's Rules

Here I will look at the whether chemicals are "drug-like" by evaluating if:

1. Molecular weight less than 500 Da.  
2. logP (octanol-water partitioning coefficient) no greater than 5.  
3. Less than 5 hydrogen bond donors and less than 10 acceptors.

```{r lipinksi_test_data}

target_chem_properties %>%
  rename(Chemical = dtxsid) %>%
  filter_targets() %>%
  distinct(Chemical, monoisotopicMass) %>%
  ggplot(aes(x = monoisotopicMass)) +
  geom_histogram(binwidth = 50, fill = "grey5",
                 boundary = 0) +
  geom_vline(xintercept = 500, linetype = "longdash") +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 100)) +
  theme_bw()

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "target_chems_molmass.png"),
       bg = "white",
       height = 3, width = 5
)

target_chem_info %>%
  count(monoisotopicMass <= 500)

target_chem_properties %>%
  ggplot(aes(x = forcats::fct_reorder(dtxsid, logkow))) +
  geom_pointrange(aes(y = logkow, ymin = logkow - logkow_sd,
                      ymax = logkow + logkow_sd), size = 0.1) +
  # facet_grid(cols = vars(type_logkow)) +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())
target_chem_pw_full %>%
  ggplot(aes(x = logkow)) +
  geom_histogram(color = "grey50", fill = "grey5",
                 binwidth = 1) +
  theme_bw()

```

```{r reticulate-prep}
smiles_list <- target_chem_properties %>% 
  rename(Chemical = dtxsid) %>%
  filter_targets() %>%
  pull(qsarReadySmiles) %>% unique()
mykit <- import("rdkit")
```

```{python get_H_bonds}
import rdkit
from rdkit import Chem
from rdkit.Chem import Descriptors
from rdkit.Chem import rdMolDescriptors

py_smiles_list = r.smiles_list
mol_from_smiles = [Chem.MolFromSmiles(x) for x in py_smiles_list]

py_h_donors = [Chem.rdMolDescriptors.CalcNumLipinskiHBD(x) for x in mol_from_smiles]
py_h_acceptors = [Chem.rdMolDescriptors.CalcNumLipinskiHBA(x) for x in mol_from_smiles]
py_tpsa = [Chem.rdMolDescriptors.CalcTPSA(x, False, True) for x in mol_from_smiles]
py_rotbonds = [Chem.rdMolDescriptors.CalcNumRotatableBonds(x) for x in mol_from_smiles]
```

```{r chemical_summary_matrix}

target_chem_summary <- target_chem_properties %>%
  filter(propertyId == "logkow-octanol-water", source == "ACD/Labs") %>%
  pivot_wider(names_from = propertyId, values_from = value) %>%
  rename(logkow = `logkow-octanol-water`) %>%
  distinct(dtxsid, monoisotopicMass, qsarReadySmiles, logkow) %>%
  filter(qsarReadySmiles %in% smiles_list) %>%
  right_join(data.frame(qsarReadySmiles = smiles_list,
                        H_donors = py_to_r(py$py_h_donors),
                        H_acceptors = py_to_r(py$py_h_acceptors),
                        TPSA = py_to_r(py$py_tpsa),
                        Rot_Bonds = py_to_r(py$py_rotbonds))
            ) %>%
  rowwise() %>%
  mutate(logD_7.4 = httk::calc_dow(dtxsid = dtxsid, pH = 7.4)) %>%
  left_join(red_assay_res %>%
              distinct(DTXSID, Affinity.Kd.Med) %>%
              na.omit(),
            by = join_by(dtxsid == DTXSID)) %>%
  distinct() %>%
  left_join(cyp_data, join_by(dtxsid == Chemical))

# Make a matrix of all properties so far
tcs_mat <- target_chem_summary %>% 
  select(dtxsid, monoisotopicMass, logkow:Affinity.Kd.Med,
         starts_with("CYP")) %>%
  distinct()
tcs_mat <- column_to_rownames(tcs_mat, var = "dtxsid") %>% as.matrix()

# z-score all properties
# Margin = 2 is columns
tcs_mat_z <- apply(tcs_mat, MARGIN = 2, FUN = scale)
rownames(tcs_mat_z) <- rownames(tcs_mat)
summary(tcs_mat_z[, 1])
summary(tcs_mat_z[(rownames(tcs_mat_z) %in% target_chems_human), 1])
tcs_mat_z[is.na(tcs_mat_z)] <- 0
pca_tcs <- prcomp(tcs_mat_z, scale = FALSE)

pca_df <- pca_tcs$x %>% as.data.frame() %>% rownames_to_column(var = "Chemical")

ggplot(data = pca_df %>% mutate(Target = (Chemical %in% target_chems_human)) %>%
         left_join(target_chem_summary, by = join_by(Chemical == dtxsid)),
       aes(x = PC1, y = PC2)) +
  geom_point(aes(color = TPSA), alpha = 0.8) +
  scale_color_viridis_c(option = "magma") +
  facet_wrap(~Target) +
  theme_bw()

target_chem_lipinski <- target_chem_summary %>%
  select(-c(qsarReadySmiles)) %>%
  group_by(dtxsid) %>%
  mutate("Mass <= 500 Da" = (monoisotopicMass <= 500),
          "H-donors < 5" = (H_donors < 5),
          "H-acceptors < 10" = (H_acceptors < 10),
          "logP < 5" = (logkow < 5),
          .keep = "none") %>%
  ungroup() %>%
  mutate(Lipinski = rowSums(dplyr::select(., where(is.logical))) >= 3)

lipinski_plot <- target_chem_lipinski %>%
  group_by(Lipinski) %>%
  count(`Mass <= 500 Da`, `H-donors < 5`, `H-acceptors < 10`, `logP < 5`,
        name = "Count") %>%
  ggplot(aes(
    axis3 = `Mass <= 500 Da`,
    axis4 = `logP < 5`,
    axis2 = `H-donors < 5`,
    axis1 = `H-acceptors < 10`,
    y = Count
  )) +
  geom_alluvium(aes(fill = Lipinski)) +
  geom_stratum() +
  scale_x_discrete(limits = c( 
    "H-acceptors < 10",
    "H-donors < 5",
    "<= 500 Da",
    "logP < 5"
  ),
  position = "top",
  expand = c(.2, .05)) +
  geom_text(stat = "stratum", aes(label = after_stat(stratum)),
            size = 1.5) +
  labs(title = "Target chemicals passing Lipinksi's Rules of 5",
       fill = "Passes Lipinksi's Rules?") +
  scale_fill_manual(values = c("grey20", "green4")) +
  theme_minimal() +
  theme(axis.text.x = element_text(face = "bold", color = "black", size = 8),
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5, size = 10),
        axis.title = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8),
        axis.title.x = element_blank(),
        legend.position = "bottom")

lipinski_plot
ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "Alluvium_Lipinski.png"),
       bg = "white",
       height = 3, width = 4
)

chem_rell_label <- this_rel_evaluated_df %>% distinct(Chemical, ll_label) %>%
  filter(ll_label %in% c("Restrictive", "Nonrestrictive"))

fup_df <- rph_param_df %>% distinct(Chemical, Funbound.plasma)
# Molecular descriptors in bins
bin_size <- 10
logP_plot <- target_chem_summary %>%
  inner_join(chem_rell_label, by = join_by(dtxsid == Chemical)) %>%
  distinct(dtxsid, logkow) %>% 
  left_join(fup_df, by = join_by(dtxsid == Chemical)) %>%
  mutate(logP_bin = factor(logkow %/% bin_size * bin_size)) %>%
  ggplot(aes(x = logP_bin, y = Funbound.plasma)) +
  labs(x = "logKow", y = "fup") +
  ggbeeswarm::geom_beeswarm() +
  theme_bw()

bin_size <- 50
TPSA_plot <- target_chem_summary %>%
  distinct(dtxsid, TPSA) %>% 
  left_join(fup_df, by = join_by(dtxsid == Chemical)) %>%
  mutate(TPSA_bin = factor(TPSA %/% bin_size * bin_size)) %>%
  ggplot(aes(x = TPSA_bin, y = Funbound.plasma)) +
  labs(x = "Topological Polar Surface Area", y = "fup") +
  ggbeeswarm::geom_beeswarm() +
  theme_bw()

bin_size <- 2
RotBond_plot <- target_chem_summary %>%
  distinct(dtxsid, Rot_Bonds) %>% 
  left_join(fup_df, by = join_by(dtxsid == Chemical)) %>%
  mutate(Rot_Bonds_bin = factor(Rot_Bonds %/% bin_size * bin_size)) %>%
  ggplot(aes(x = Rot_Bonds_bin, y = Funbound.plasma)) +
  labs(x = "Rotatable Bonds", y = "fup") +
  ggbeeswarm::geom_beeswarm() +
  theme_bw()

bin_size <- 150
MW_plot <- target_chem_summary %>%
  distinct(dtxsid, monoisotopicMass) %>% 
  left_join(fup_df, by = join_by(dtxsid == Chemical)) %>%
  mutate(MW_bin = factor(monoisotopicMass %/% bin_size * bin_size)) %>%
  ggplot(aes(x = MW_bin, y = Funbound.plasma)) +
  labs(x = "Molecular Mass", y = "fup") +
  ggbeeswarm::geom_beeswarm() +
  theme_bw()

bin_size <- 4
HBA_plot <- target_chem_summary %>%
  distinct(dtxsid, H_acceptors) %>% 
  left_join(fup_df, by = join_by(dtxsid == Chemical)) %>%
  mutate(HBA_bin = factor(H_acceptors %/% bin_size * bin_size)) %>%
  ggplot(aes(x = HBA_bin, y = Funbound.plasma)) +
  labs(x = "Hydrogen bond acceptors", y = "fup") +
  ggbeeswarm::geom_beeswarm() +
  theme_bw()

bin_size <- 2
HBD_plot <- target_chem_summary %>%
  distinct(dtxsid, H_donors) %>% 
  left_join(fup_df, by = join_by(dtxsid == Chemical)) %>%
  mutate(HBD_bin = factor(H_donors %/% bin_size * bin_size)) %>%
  ggplot(aes(x = HBD_bin, y = Funbound.plasma)) +
  labs(x = "Hydrogen bond donors", y = "fup") +
  ggbeeswarm::geom_beeswarm() +
  theme_bw()

bin_size <- 200
target_chem_summary %>%
  distinct(dtxsid, logD_7.4) %>% 
  left_join(fup_df, by = join_by(dtxsid == Chemical)) %>%
  mutate(logD_bin =  factor(logD_7.4 %/% bin_size * bin_size)) %>%
  ggplot(aes(x = logD_bin, y = Funbound.plasma)) +
  labs(x = "logD_7.4", y = "fup") +
  ggbeeswarm::geom_beeswarm() +
  theme_bw()

(MW_plot + TPSA_plot + HBA_plot) / (logP_plot + RotBond_plot + HBD_plot)

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "PhysChemProperties_rdkit.png"),
       bg = "white",
       height = 6, width = 8
)

```

Overwhelmingly, we see that chemicals (by Lipinski's rules of 5) are "drug-like".
While this may not matter effectively for clearance, it may be nice to look 
at whether chemicals deemed not "drug-like" (11 total) are more likely to be
restrictively or non-restrictively cleared.