### 2025-03-24 data-raw/ClearanceProject-3-TKstats.Rmd
Added figure investigating interspecies trends in fup and CLint.

---
title: "Clearance Project - 3 - TKstats"
author: "Gilberto Padilla Mercado"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library-loading}
today <- format(Sys.Date(), "%d%b%Y")
devtools::load_all()
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(ctxR)
library(ggdist)
library(patchwork)
library(httk)
.onUnload <- function(libpath) {
  library.dynam.unload("httk", libpath)
}
```

```{r data-loading}
# Note that you should run ...-1-Data.Rmd before this
# RCdata/ is a directory that is not tracked by git because it contains
# large files 
this_pk <- readRDS("data-raw/plasma_CvTfits.rds")
load("data-raw/chem_lists.RData")
load("data-raw/plasma_CvTfits_data.RData")
load("data-raw/rat_pars_human.RData")
load("data-raw/rat_pars_rat.RData")
all_chem_properties <- readRDS("data-raw/RCdata/ctxR_3comp2_cheminfo.rds")
full_pk <- readRDS("data-raw/all_cvt_pk.rds")
full_tkstats <- get_tkstats(full_pk)
```

## Species-specific differences in fraction unbound and intrinsic clearance

We first want to establish that there is a difference between human and rat
parameterization. Because one of our `param_df`s is only rat and the other has
substituted human parameters (scaled to a rat).

```{r rat_human_diff_fupClint}
fup_rh <- rpr_param_df %>% 
  distinct(Chemical, MW,
           Clint, Funbound.plasma) %>% 
  rename_with(\(x) {paste0(x, "_rat")},
              .cols = c(Clint, Funbound.plasma)) %>%
  left_join(
    rph_param_df %>%
      distinct(Chemical, MW,
               Clint, Funbound.plasma) %>% 
      rename_with(\(x) {paste0(x, "_human")},
                  .cols = c(Clint, Funbound.plasma))
    )

species_fup_pl <- fup_rh %>%
  ggplot(aes(
    Funbound.plasma_human,
    Funbound.plasma_rat
  )) +
  geom_abline(slope = 1) +
  geom_point() +
  geom_smooth(method = "lm", linetype = "dashed",
              color = "forestgreen") +
  labs(title = expression(f[up]),
       x = "Human", y = "Rat") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) +
  coord_equal(xlim = c(0,1), ylim = c(0,1), expand = TRUE)

fup_pl <- ggExtra::ggMarginal(
  species_fup_pl,
  type = "histogram",
  binwidth = 0.025,
  fill = "white"
)

lm(
  Funbound.plasma_rat ~ Funbound.plasma_human,
  data = fup_rh
)

cor.test(fup_rh$Funbound.plasma_human,
            fup_rh$Funbound.plasma_rat)

## Clint
species_clint_pl <- subset(
  fup_rh, Clint_human < 500, Clint_rat < 500
) %>%
  ggplot(aes(
    Clint_human,
    Clint_rat
  )) +
  geom_abline(slope = 1) +
  geom_abline(slope = 1, intercept = c(-50,50), linetype = "longdash") +
  geom_point() +
  # geom_smooth(method = "lm", linetype = "dashed", color = "forestgreen") +
  labs(title = expression(CL[int]),
       x = "Human", y = "Rat") +
  coord_equal(xlim = c(0,500), ylim = c(0,500)) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5,vjust = -1))

clint_pl <- ggExtra::ggMarginal(
  species_clint_pl,
  type = "histogram",
  fill = "white"
)

lm(
  Clint_rat ~ Clint_human,
  data = fup_rh
)

cor.test(fup_rh$Clint_human,
            fup_rh$Clint_rat)

# Both plots
species_fup_pl + species_clint_pl +
  plot_layout(axis_titles = "collect")

ggsave(filename = paste0(Sys.getenv("FIG_DIR"), "RestrictiveClearance_Analysis/",
                         "Rat_Human_Interspecies_FupCLint_compare.png"),
       height = 3, width = 6)

# Get the "close" estimates (+/-50)
close_clint <- fup_rh %>%
  filter(between(Clint_rat, Clint_human - 50, Clint_human + 50)) %>%
  pull(Chemical)

rph_rmsle <- rph_httk_preds %>%
  mutate(fupSpecies = "human",
         rest_sqerror = (log10(httk_Preds_restrictive) - log10(Conc))^2,
         nonrest_sqerror = (log10(httk_Preds_nonrestrictive) - log10(Conc))^2) %>%
  group_by(
    Chemical, Species, fupSpecies
  ) %>%
  summarize(rest_rmsle = sqrt(sum(rest_sqerror)/n()),
            nonrest_rmsle = sqrt(sum(nonrest_sqerror)/n())
  ) %>%
  ungroup()

rpr_rmsle <- rpr_httk_preds %>%
  mutate(fupSpecies = "rat",
        rest_sqerror = (log10(httk_Preds_restrictive) - log10(Conc))^2,
         nonrest_sqerror = (log10(httk_Preds_nonrestrictive) - log10(Conc))^2) %>%
  group_by(
    Chemical, Species, fupSpecies
  ) %>%
  summarize(rest_rmsle = sqrt(sum(rest_sqerror)/n()),
            nonrest_rmsle = sqrt(sum(nonrest_sqerror)/n())
  ) %>%
  ungroup()

bind_rows(rpr_rmsle, rph_rmsle) %>%
  mutate(close_clint = ifelse(Chemical %in% close_clint,
                              "|CLint| within 50", "|CLint| > 50")
  ) %>% 
  filter(Chemical %in% unique(rpr_rmsle$Chemical)) %>%
  ggplot(aes(nonrest_rmsle, rest_rmsle)) +
  geom_abline(slope = 1, color = "blue") +
  geom_point(aes(color = fupSpecies), shape = 21) +
  scale_color_manual(values = c("green4", "magenta3")) +
  facet_grid(cols = vars(close_clint)) +
  geom_line(aes(group = paste0(Chemical)),
            alpha = 0.4) +
  labs(x = "Nonrestrictive RMSLE", y = "Restrictive RMSLE") +
  theme_bw() +
  coord_equal()

interspecies_rmsle_trends <- bind_rows(rpr_rmsle, rph_rmsle) %>%
  mutate(
    close_clint = ifelse(Chemical %in% close_clint,
                         "|\U0394 CLint| within 50", "|\U0394 CLint| > 50")
  ) %>% 
  filter(Chemical %in% unique(rpr_rmsle$Chemical)) %>% 
  pivot_longer(cols = contains("_rmsle"),
               names_to = "clearance_type",
               values_to = "RMSLE") %>%
  pivot_wider(names_from = fupSpecies,
              values_from = RMSLE) %>%
  mutate(norm_diff = (human - rat)/rat)

irt_pl <- interspecies_rmsle_trends %>%
  filter(!is.na(norm_diff)) %>% 
  ggplot(aes(x = forcats::fct_reorder(Chemical, norm_diff, mean),
             y = norm_diff,
             fill = clearance_type)) +
  geom_col(
    width = 0.6,
    # fill = "grey5", 
    position = position_dodge(),
    color = NA
    ) +
  labs(
    x = "Chemical",
    y = "Fold Change RMSLE\nHuman more accurate <<     >> Rat more accurate"
  ) +
  scale_fill_manual(values = c("grey20", "seagreen"),
                    breaks = c("nonrest_rmsle", "rest_rmsle"),
                    labels = c("Nonrestrictive", "Restrictive"),
                    name = "Clearance") +
  geom_hline(yintercept = 0) +
  theme_minimal() + 
  facet_grid(cols = vars(close_clint),
             # rows = vars(clearance_type),
             space = "free_x", scales = "free_x") +
  theme(axis.text.x = element_blank(),
        axis.ticks = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(color = "grey5"),
        panel.background = element_rect(color = "grey5"),
        axis.title = element_text(size = 10),
        legend.position = "bottom")

species_clint_pl + irt_pl +
  plot_layout(widths = c(5,8))

ggsave(filename = paste0(Sys.getenv("FIG_DIR"), "RestrictiveClearance_Analysis/",
                         "Rat_Human_Interspecies_RMSLE_trends.png"),
       height = 5, width = 7,
       bg = "white")
```

From this we notice that CLint specifically can have drastic differences which
we want to explore. We also glean that there are potential drastic improvements
in model accuracy. We next want to compare how a changing human CLint towards rat
CLint by 1% steps changes RMSE.

```{r}
combo_clint <- rpr_param_df %>%
  select(Chemical, rCLint = CLint) %>% 
  left_join(rph_param_df %>%
              select(Chemical, hCLint = CLint)) %>%
  mutate(hclint_1p = (1/10)*hCLint, 
         direction = sign(hCLint - rCLint)) %>%
  mutate(Nstep = ifelse(direction == 0, 
                        0, abs(hCLint - rCLint) %/% hclint_1p))

combo_clint %>%
  left_join(interspecies_rmsle_trends) %>%
  ggplot(aes(Nstep, norm_diff)) +
  geom_point() +
  scale_x_log10() +
  theme_bw()
```

This next analysis looks at bootstrapping to look at specific variation in downstream
parameters (like Css and halflife) which depend on _kelim_ which cannot be estimated
at times. However, it is noted that even this method does not provide much more information.

```{r combination-function}
# Only need a simple bootstrapping method
combo_gen_splitter <- function(df, num = 1000L) {
  # Checks
  num = as.integer(num)
  stopifnot(
    is.data.frame(df),
    !is.na(num)
  )

  start <- 100L
  end <- 100L+num
  seed_seq <- seq.int(start, end)
  full_df_combos <- lapply(
    seed_seq,
    \(x) {
      set.seed(x)
      sample_df <- df %>%
        group_by(Chemical, Species, Reference, Route, Media, Dose, Time) %>%
        slice_sample(prop = 1, replace = TRUE) %>%
        ungroup()
      return(sample_df)
    }
  )

  message(paste("Samples: ", num))
  set.seed(NULL)
  invisible(full_df_combos)
}

```

```{r example_chemical}
# Using the example chemical
trial_coefsd <- coef_sd(full_pk)
trial_coefsd_reduced <- trial_coefsd %>% 
  filter(param_name %in% c("kelim", "Vdist", "V1")) %>%
  distinct()

trial_coefsd_reduced %>%
  filter(is.na(param_sd), model == "model_1comp") %>%
  distinct(Chemical, Species) -> selected_trial

test_chem_data <- get_data(full_pk) %>% filter(Conc > pLOQ) %>%
  count(Chemical, Species, Reference, Route, Media, Dose, Time) %>% arrange(-n) %>%
  filter(n > 4) %>%
  inner_join(selected_trial) %>%
  distinct(Chemical, Species, Reference, Route, Media, Dose) %>%
  inner_join(get_data(full_pk))

# Make 1000 sample data sets
system.time(
  test_chem_data_listsplits <- combo_gen_splitter(test_chem_data)
)

# Takes way too long, parallelize it

clust <- parallel::makeCluster(parallel::detectCores()-4L)
parallel::clusterCall(clust, function() devtools::load_all())
system.time(
  Css_list <- parallel::parLapply(
    clust,
    test_chem_data_listsplits,
    \(x) {
      tmp_pk <- pk(x,
                   mapping = ggplot2::aes(
                     Chemical = Chemical,
                     Chemical_Name = Chemical_Name,
                     DTXSID = DTXSID,
                     CASRN = CASRN,
                     Species = Species,
                     Reference = Reference,
                     Media = Media,
                     Route = Route,
                     Dose = Dose,
                     Dose.Units = "mg/kg",
                     Subject_ID = Subject_ID,
                     Series_ID = Series_ID,
                     Study_ID = Study_ID,
                     ConcTime_ID = ConcTime_ID,
                     N_Subjects = N_Subjects,
                     Weight = Weight,
                     Weight.Units = "kg",
                     Time = Time,
                     Time.Units = "hours",
                     Value = Value,
                     Value.Units = "mg/L",
                     Value_SD = Value_SD,
                     LOQ = LOQ
                   )
      ) +
        scale_conc(dose_norm = TRUE,
                   log10_trans = TRUE) +
        settings_optimx(method = c(
          "bobyqa"
        ))

      tmp_pk <- do_fit(tmp_pk)
      tmp_tkstats <- get_tkstats(tmp_pk) %>%
        distinct(Chemical, Species, Reference,
                 Route, Media,
                 model, method,
                 Css, CLtot, halflife)

      return(tmp_tkstats)
    })
)
parallel::stopCluster(clust)

Css_df_full <- bind_rows(Css_list)

```

```{r plotting-bootstraps}
num_samples <- length(Css_list)

test_chem_full_tkstats <- full_tkstats %>%
  filter(model != "model_flat") %>%
  inner_join(selected_trial) %>%
  distinct(Chemical, Species, model, Css)

Css_df_full %>%
  filter(model != "model_flat") %>%
  left_join(this_winmodel %>% 
              mutate(winmodel = TRUE) %>%
              select(-c(near_flat, preds_below_loq))) %>% 
  mutate(winmodel = ifelse(is.na(winmodel), FALSE, winmodel)) %>%
  ggplot(aes(x = halflife, y = Chemical, fill = winmodel)) +
  stat_slab() +
  # xlim(0,40) +
  scale_x_log10() +
  facet_grid(rows = vars(model),
             cols = vars(Route)) +
  scale_fill_manual(values = c("grey5", "green3")) +
  theme_bw() +
  labs(title = "halflife (hours)") +
  theme(
    axis.text.y = element_text(size = 6),
    axis.text.x = element_text(size = 10),
    plot.title = element_text(hjust = 0.5, size = 12),
    axis.title = element_blank()
  )
ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "select_",
                         "halflife_bootstrapping.png"),
       bg = "white",
       height = 4, width = 6
)

css_hist <- Css_df_full %>%
  filter(model != "model_flat") %>%
  ggplot(aes(x = Css)) +
  geom_histogram(fill = "grey5", binwidth = 0.2) +
  geom_vline(data = test_chem_full_tkstats,
             aes(xintercept = Css),
             color = "green4",
             linetype = "dashed") +
  facet_grid(rows = vars(model)) +
  labs(title = unique(test_chem_data$Chemical_Name),
       subtitle = paste("Number of samples: ", num_samples)) +
  scale_x_log10() +
  xlim(0,50) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 1),
        plot.subtitle = element_text(hjust = 1))

css_lines <- Css_df_full %>%
  filter(model != "model_flat") %>%
  group_by(model) %>%
  mutate(grouping = row_number()) %>%
  ggplot(aes(x = model, y = Css)) +
  geom_point(alpha = 0.4) +
  geom_line(aes(group=grouping), alpha = 0.02) +
  scale_y_log10() +
  labs(subtitle = "N per timepoint: 4") +
  theme_bw()

get_winning_model(full_pk) %>% filter(Chemical == test_chem) %>% glimpse()

css_hist + css_lines

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", test_chem,
                         "_Css_pseudobootstrapping.png"),
       bg = "white",
       height = 3, width = 5
)
```