### 2025-06-02 data-raw/optimizing_httk_model.Rmd
Fixing total liver flow, renal and hepatic clearance definition in 1comp_cl model functions. Added liver_mass as a parameter to facilitate hepatic clearance calculations.

---
title: "Optimizing httk's 3comp2 model"
author: "Gilberto Padilla Mercado"
date: "2025-05-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
devtools::load_all()
library(dplyr)
library(httk)
```

# Purpose

This documents how to optimize `Clint` and `Funbound.plasma` using
_invivoPKfit_ and _httk_'s `3compartment2` model. There are
a couple of issues mainly stemming from the fact that the solver often
returns an error from the underlying FORTRAN code used in `deSolve` in
multiple iterations of the optimization procedure.

```{r}
load_dawson2021()
schems <- httk::get_cheminfo("dtxsid", species = "rat",
                             model = "3compartment2") |>
  suppressWarnings()

sum(unique(cvt$analyzed_chem_dtxsid) %in% schems)

minicvt <- subset(
  cvt,
  species == "rat" &
    analyzed_chem_dtxsid %in% schems
)

table(minicvt$analyzed_chem_dtxsid)
```

First I want to try just the clearance-aware 1-compartment models

```{r}
test_pk <- pk(data = minicvt) + 
  stat_error_model(error_group = vars(Chemical, Species)) +
  stat_model(model = c("model_flat", "model_1comp_cl_rest",
                       "model_1comp_cl_nonrest", "model_1comp_fup")) +
  settings_optimx(hessian = TRUE) +
  scale_conc(dose_norm = TRUE, log10_trans = FALSE)

test_pk <- do_data_info(test_pk)
test_pk <- do_prefit(test_pk)
test_pk
gc()

test_fit <- do_fit(test_pk, n_cores = 4)
winmodel <- get_winning_model(test_fit)
winmodel %>% ungroup() %>% count(method, model)
View(winmodel)
get_fit(test_fit) %>%
  filter(convcode == 0) %>%
  ungroup() %>%
  distinct(Chemical, Species, model, method) %>% View()

pl <- plot(test_fit, use_scale_conc = FALSE,
           best_fit = FALSE, log10_C = FALSE,
           method = "L-BFGS-B")

winmodel %>%
  filter(model != "model_flat") %>%
  distinct(Chemical, model)

pl %>%
  filter(Chemical == "DTXSID0020442") %>%
  pull(final_plot) %>% .[[1]]

pl %>%
  filter(Chemical == "DTXSID2020139") %>%
  pull(final_plot) %>% .[[1]]

pl %>%
  filter(Chemical == "DTXSID4020458") %>%
  pull(final_plot) %>% .[[1]]

pl %>%
  filter(Chemical == "DTXSID4022527") %>%
  pull(final_plot) %>% .[[1]]

pl %>%
  filter(Chemical == "DTXSID4032405") %>%
  pull(final_plot) %>% .[[1]]

pl %>%
  filter(Chemical == "DTXSID5023320") %>%
  pull(final_plot) %>% .[[1]]

pl %>%
  filter(Chemical == "DTXSID6022923") %>%
  pull(final_plot) %>% .[[1]]

pl %>%
  filter(Chemical == "DTXSID7046627") %>%
  pull(final_plot) %>% .[[1]]

pl %>%
  filter(Chemical == "DTXSID8020541") %>%
  pull(final_plot) %>% .[[1]]

pl %>%
  filter(Chemical == "DTXSID8032673") %>%
  pull(final_plot) %>% .[[1]]

pl %>%
  filter(Chemical == "DTXSID9020584") %>%
  pull(final_plot) %>% .[[1]]

pl %>%
  filter(Chemical == "DTXSID9032329") %>%
  pull(final_plot) %>% .[[1]]

pl %>%
  filter(Chemical == "DTXSID1022265") %>%
  pull(final_plot) %>% .[[1]]

get_fit(test_fit) %>%
  filter(Chemical == "DTXSID0020442", model == "model_1comp_cl_nonrest",
         param_name == "Clint", method == "L-BFGS-B") %>% View()

get_fit(test_fit) %>%
  semi_join(winmodel) %>%
  filter(model != "model_flat", optimize_param) %>%
  mutate(diff_est = estimate - start) %>%
  View("diffs")

get_fit(test_fit) %>%
  semi_join(winmodel) %>%
  filter(model != "model_flat", optimize_param,
         convcode == 0) %>%
  mutate(diff_est = estimate - start) %>%
  filter(stringr::str_detect(param_name, "sigma_", negate = TRUE)) %>%
  group_by(model, method, param_name) %>%
  summarize(avg_diff = mean(diff_est),
            min_diff = min(diff_est),
            max_diff = max(diff_est))

```

We have a list of things to try:

[] Are the timepoints causing issues? Run and stop with initial parameters.
[] Should we only optimize one parameter? Modify `get_params_httk_3comp2` function.

```{r}

test_pk <- pk(data = minicvt) + 
  stat_error_model(error_group = vars(Chemical, Species)) +
  stat_model(model = c("model_flat", "model_httk_3comp2")) +
  settings_optimx(hessian = TRUE) +
  scale_conc(dose_norm = TRUE, log10_trans = FALSE)

test_pk <- do_prefit(test_pk)
test_pk
gc()

test_fit <- do_fit(test_pk)

```

```{r}
preds <- predict(test_fit)

my_fit <- get_fit(test_fit) |>
  dplyr::filter(optimize_param)

View(transform(my_fit, opt_diff = start - estimate))

```