### 2025-08-12 data-raw/pulling_oral_iv_data_cvtdb.Rmd
Moved helper functions to CvTdb query report.

---
title: "Pulling oral and IV data from CvTdb"
author: "Caroline Ring, Gilberto Padilla Mercado"
date: "2023-09-29"
output: 
  html_document:
    toc: true
  pdf_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  comment = NA,
  tidy.opts = list(width.cutoff = 75),
  tidy = TRUE
)
```

Load necessary packages: **DBI, dplyr, tidyr**

```{r}
library(DBI)
devtools::load_all() # So this loads the previous cvt object
library(dplyr)
library(tidyr)
library(readr)
library(stringr)
library(cli)
```

# Connect to database

This section is largely taken from a `Querying CvTdb` vignette from Taylor Wall 
and a previous version of this vignette authored by Caroline Ring.

For connection to CvTdb:

Connect to EPA VPN.

Set environment variables for the user name, password, host, and dbname that will be used to connect. Doing it this way allows us to not hardcode the password.

The following are some helper functions for the database connection.

```{r query_functions}
postgreSQL_connection <- function() {
  return(
    dbConnect(RPostgreSQL::PostgreSQL(),
      user = Sys.getenv("CVT_USER"),
      password = Sys.getenv("CVT_PASS"),
      host = Sys.getenv("CVT_HOST"),
      dbname = Sys.getenv("CVT_DB_NAME")
    )
  )
}

query_db <- function(query = NULL, schema) {
  if (is.null(query)) {
    return(cat("\nMust provide a query to send"))
  }
  con <- postgreSQL_connection()

  query_result <- tryCatch(
    return(dbGetQuery(con, query %>%
      gsub("FROM ", paste0("FROM ", schema, "."), .) %>%
      gsub("JOIN (?=\\w)", paste0("JOIN ", schema, "."), .,
        perl = TRUE
      ))),
    error = function(cond) {
      cat(paste0("\nError message: ", cond))
      return(NULL)
    },
    finally = {
      dbDisconnect(con)
    }
  )
  return(query_result)
}
```

```{r quick_message_funtion}
quick_message <- function(text = character(0L)) {
  stopifnot(length(text) > 0)

  this_par <- cli_par()
  cli_text(text[1])
  text <- text[-1]

  ul <- cli_ul()
  for (this_text in text) {
    cli_li(this_text)
  }
  cli_end(ul)
  cli_end(this_par)
}

cli_print <- function(x, file = "", append = FALSE) {
  cat(x, file = file, sep = "\n", append = append)
}

# Wrapper for qcount())
qcount <- function(x) length(unique(x))

quick_stats <- function(.data) {
  stopifnot(is.data.frame(.data))

  .data <- subset(.data, !is.na(analyzed_chem_dtxsid))
  cli_text(c(
    "!" = "Only data with analyzed chemical dtxsid will be considered."
  ))
  cli_text("---- Statistics on the current data ----")
  cli_text("Total number of data points: {nrow(.data)}")
  cli_text("Unique Species and data counts:")
  cli_dl(table(.data$species))

  this_par <- cli_par()
  cli_text("Number of unique studies: {qcount(.data$fk_study_id)}")
  cli_text("Number of unique series: {qcount(.data$fk_series_id)}")
  cli_text("Number of unique chemicals: {qcount(.data$analyzed_chem_dtxsid)}")
  cli_end(this_par)

  that_par <- cli_par()
  cli_text("Original Dose Units: {qcount(.data$dose_level_units_original)}")
  cli_text("Normalized Dose Units: {qcount(.data$dose_level_units_normalized)}")
  cli_text("Original Conc Units: {qcount(.data$conc_units_original)}")
  cli_text("Normalized Conc Units: {qcount(.data$conc_units_normalized)}")
  cli_end(that_par)

  cli_par()
  cli_text("Data by unique route of administration:")
  cli_dl(table(.data$administration_route_normalized))
  cli_end()
  cli_par()
  cli_text("Data by unique medium collected:")
  cli_dl(table(.data$conc_medium_normalized))
  cli_end()
  cli_par()
  cli_text("Data by unique curation set:")
  cli_dl(table(.data$curation_set_tag))
  cli_end()
}

#' Wrapper for [dplyr::case_match()] that accepts named lists as arguments
#' @param .x A vector to match against. In a data context, a column.
#' @param .lst A named list in the form `old_value = new_value`.
#' @param .default Value to be used when there are no values. Suggest setting it to `.x`
#' @return A vector of replaced values
case_match_lst <- function(.x, .lst, .default = NULL) {
  stopifnot(rlang::is_named(.lst))

  dplyr::case_match(
    .x,
    !!!lapply(paste0("\"", names(.lst), "\"~\"", .lst, "\""), stats::as.formula),
    .default = .default
  )
}

#' Wrapper for [str_replace_all()] which replaces whole words.
#' @param .x A vector to match against.
#' @param .lst A named vector in the form `old_value = new_value`.
#' @return A character vector with values substituted in.
str_replace_lst <- function(.x, .lst) {
  stopifnot(rlang::is_named(.lst))

  str_replace_all(.x, setNames(.lst, paste0("^", names(.lst), "$")))
}

chem_id_cols <- rlang::quos(
  fk_analyzed_chemical_id, fk_dosed_chemical_id,
  analyzed_chem_dtxsid, dosed_chem_dtxsid,
  analyzed_chem_name, dosed_chem_name,
  analyzed_chem_name_original, dosed_chem_name_original,
  analyzed_chem_name_secondary_original,
  dosed_chem_name_secondary_original
)

unit_values_cols <- rlang::quos(
  conc_original, conc,
  conc_units_original, conc_units_normalized,
  conc_sd_original, conc_sd,
  loq, loq_units,
  dose_level_original, dose_level_normalized,
  dose_level_units_original, dose_volume_units,
  dose_volume, weight_kg
)
```

# SQL Query

The SQL query result is built by left joining onto the `conc_time_values` table.
Additionally, because _invivoPKfit_ uses DTXSID as the standard chemical identifier,
it is important to repair some missing values for DTXSID. To do this, I 
make a separate query `SELECT distinct * FROM chemicals` and making the chemical
names to lowercase and by specifically taking the columns for name (now lowercase)
and DTXSID, I can see there are many instances of duplicated chemical entries
(which means they have separate `chemical.id` key values in this table). I then
"fill in" the DTXSID when they are missing and if any other instance of that
chemical name in the table has a DTXSID. I fill in DTXSIDs in the main
query result that I will be using using some joins with this repaired table.

```{r SQL_queries, warning=FALSE}
# Need to make this in a specified ordered
# First all IDs (so I can check them quickly)
# Note that this uses an ALIAS for each table which is later specified in
# the FROM and LEFT JOIN statements
series_chems <- paste0(
  "SELECT DISTINCT a.id, a.fk_analyzed_chemical_id, ",
  "b.analyzed_chem_dtxsid, b.analyzed_chem_name_original, ",
  "b.analyzed_chem_name_secondary_original, ",
  "b.analyzed_chem_casrn, b.analyzed_chem_name, ",
  "a.time_units_original, a.conc_units_original, ",
  "a.conc_units_normalized, a.loq, a.loq_units, ",
  "a.n_subjects_in_series, a.radiolabeled, a.fk_study_id, ",
  "a.qc_notes, a.qc_status, a.qc_flags, c.conc_medium_normalized ",
  "FROM series a ",
  "LEFT JOIN (SELECT id, dsstox_substance_id AS analyzed_chem_dtxsid, ",
  "chemical_name_original AS analyzed_chem_name_original, ",
  "chemical_name_secondary_original AS analyzed_chem_name_secondary_original, ",
  "dsstox_casrn as analyzed_chem_casrn, preferred_name AS analyzed_chem_name ",
  "FROM chemicals) AS b ON a.fk_analyzed_chemical_id = b.id ",
  "LEFT JOIN conc_medium_dict AS c ON a.fk_conc_medium_id = c.id"
)

study_docs <- paste0(
  "SELECT a.id, a.fk_extraction_document_id, a.fk_dosed_chemical_id, ",
  "b.pmid, b.doi, b.year, b.other_study_identifier, ",
  "b.url, b.extracted, b.curation_set_tag ",
  "FROM studies a ",
  "LEFT JOIN documents b ON a.fk_extraction_document_id = b.id"
)

audit_query <- query_db(
  paste0(
    "SELECT DISTINCT * FROM cvt_audit ",
    "WHERE fk_table_name IN ('documents', 'series', 'studies')"
  ),
  "cvt"
)
study_query <- query_dib(study_docs, "cvt")
series_query <- query_db(series_chems, "cvt")
dtxsid_query <- query_db("SELECT DISTINCT * FROM chemicals", "cvt") %>%
  mutate(
    across(ends_with("_id"), as.character),
    across(contains("_name"), \(x) {
      str_to_lower(x) |>
        force_ascii_quotes() |>
        # trim whitespace around hyphens
        str_replace_all("\\s*-\\s*", "-") |>
        # comma-separate forms like 3'6 into 3',6
        str_replace_all("(?<=\\d)'(?=\\d)", "',") |>
        # fix some bracket cases
        str_replace_all("\\[3\\]h", "[3h]") |>
        str_replace_all("benzo\\[a\\]", "benzo(a)") |>
        # adds a space after square bracket separator
        str_replace_all("](?!\\s|-)", "] ")
    })
  )

select_chems <- subset(
  dtxsid_query,
  select = c(
    "id",
    "dsstox_substance_id",
    "dsstox_casrn",
    "preferred_name",
    "chemical_name_original",
    "chemical_name_secondary_original"
  )
)

# Surveying the non-equal chemical id landscape:
series_query |>
  inner_join(study_query, by = join_by(fk_study_id == id)) |>
  distinct(fk_analyzed_chemical_id, fk_dosed_chemical_id, curation_set_tag) |>
  count(fk_analyzed_chemical_id == fk_dosed_chemical_id)

unequal_chem_ids <- series_query |>
  inner_join(study_query, by = join_by(fk_study_id == id)) |>
  distinct(fk_analyzed_chemical_id, fk_dosed_chemical_id, curation_set_tag) |>
  filter(fk_analyzed_chemical_id != fk_dosed_chemical_id) |>
  left_join(
    rename_with(select_chems, ~ paste0("analyzed_", .x)),
    by = join_by(fk_analyzed_chemical_id == analyzed_id)
  ) |>
  left_join(
    rename_with(select_chems, ~ paste0("dosed_", .x)),
    by = join_by(fk_dosed_chemical_id == dosed_id)
  ) |>
  glimpse()

count(
  unequal_chem_ids,
  analyzed_dsstox_substance_id == dosed_dsstox_substance_id,
  analyzed_dsstox_casrn == dosed_dsstox_casrn,
  analyzed_preferred_name == dosed_preferred_name,
  analyzed_chemical_name_original == dosed_chemical_name_original,
  analyzed_chemical_name_secondary_original == dosed_chemical_name_secondary_original
) |>
  rename(
    "Equal DTXSID" = 1, "Equal CASRN" = 2, "Equal Preffered Name" = 3,
    "Equal Recorded Name" = 4, "Equal Secondary Recorded Name" = 5, "N" = 6
  )

# Are any names equal???
count(
  unequal_chem_ids,
  analyzed_dsstox_substance_id == dosed_dsstox_substance_id |
    analyzed_dsstox_casrn == dosed_dsstox_casrn |
    analyzed_preferred_name == dosed_preferred_name |
    analyzed_chemical_name_original == dosed_chemical_name_original |
    analyzed_chemical_name_secondary_original == dosed_chemical_name_secondary_original
) |>
  rename(
    "Any" = 1, "N" = 2
  ) |>
  mutate(
    Any = ifelse(Any %in% TRUE, "Some Identical", "None Identical")
  ) |>
  group_by(Any) |>
  summarize(N = sum(N)) |>
  pivot_wider(names_from = Any, values_from = N)

# Let's parse through with those "Some Identical" identifiers first
unequal_chem_ids |>
  filter(
    analyzed_dsstox_substance_id == dosed_dsstox_substance_id |
      analyzed_dsstox_casrn == dosed_dsstox_casrn |
      analyzed_preferred_name == dosed_preferred_name |
      analyzed_chemical_name_original == dosed_chemical_name_original |
      analyzed_chemical_name_secondary_original == dosed_chemical_name_secondary_original
  ) |>
  group_by(fk_analyzed_chemical_id, fk_dosed_chemical_id) |>
  summarize(
    same_dtxsid = if_else(
      analyzed_dsstox_substance_id == dosed_dsstox_substance_id,
      analyzed_dsstox_substance_id, NA_character_, missing = NA_character_
    ),
    same_casrn = if_else(
      analyzed_dsstox_casrn == dosed_dsstox_casrn,
      analyzed_dsstox_casrn, NA_character_, NA_character_
    ),
    same_preferred_name = if_else(
      analyzed_preferred_name == dosed_preferred_name,
      analyzed_preferred_name, NA_character_, NA_character_
    ),
    same_chem_name = if_else(
      analyzed_chemical_name_original == dosed_chemical_name_original,
      analyzed_chemical_name_original, NA_character_, NA_character_
    ),
    same_chem_name_secondary = if_else(
      analyzed_chemical_name_secondary_original == dosed_chemical_name_secondary_original,
      analyzed_chemical_name_secondary_original, NA_character_, NA_character_
    )
  ) |>
  rowwise() |>
  mutate(
    likely_same = sum(!is.na(c_across(starts_with("same_"))))
  ) |>
  ungroup() |>
  arrange(
    same_dtxsid, same_casrn, same_preferred_name, same_chem_name,
    same_chem_name_secondary
  ) |>
  View("Similar IDs")
# There are two ways the identifiers may be duplicated,
# shown here between rows and columns

# There are a lot of duplicate DTXSIDs
query <- paste0(
  "SELECT distinct ",
  # Conc_Time_Values table fields
  "a.id, a.fk_series_id, a.time_original, a.time_hr, ",
  "a.conc_original, a.conc_sd_original, a.conc, a.conc_sd, ",

  # Series table fields
  "b.fk_analyzed_chemical_id, ",
  ## Chemical dictionary fields (analyzed chemical information)
  "l.analyzed_chem_dtxsid, l.analyzed_chem_name_original, ",
  "l.analyzed_chem_casrn, l.analyzed_chem_name, ",
  "l.analyzed_chem_name_secondary_original, ",
  "b.time_units_original, b.conc_units_original, ",
  "b.conc_units_normalized, b.loq, b.loq_units, ",
  "b.n_subjects_in_series, b.radiolabeled, b.fk_study_id, ",

  # Studies table fields
  "c.fk_dosed_chemical_id, ",
  ## Chemical dictionary fields (dosed chemical information)
  "k.dosed_chem_dtxsid, k.dosed_chem_name_original, ",
  "k.dosed_chem_casrn, k.dosed_chem_name, ",
  "k.dosed_chem_name_secondary_original, ",
  "c.dose_volume, c.dose_volume_units, c.dose_vehicle, ",
  "c.dose_duration, c.dose_duration_units, ",
  ## dose_frequency dictionary fields
  "j.dose_frequency_original, j.dose_frequency_normalized, ",
  "c.fasting_period, ",
  "c.dose_level_normalized, c.dose_level_original, ",
  "c.dose_level_units_original, c.dose_level_units_normalized, ",
  ## conc_medium dictionary fields
  "i.conc_medium_original, i.conc_medium_normalized, ",
  ## administration_route dictionary fields
  "h.administration_route_original, h.administration_route_normalized, ",
  ## administration_method dictionary fields
  # "g.administration_method_original, g.administration_method_normalized, ",
  ## administration_form dictionary fields
  # "f.administration_form_original, f.administration_form_normalized, ",

  # Subjects table fields
  "b.fk_subject_id, d.weight_kg, d.species, d.sex, d.age, d.age_units, d.age_category, ",
  # Documents table fields
  "c.fk_extraction_document_id, e.pmid, e.year, e.other_study_identifier, ",
  "e.url, e.doi, e.extracted, e.curation_set_tag ",

  # Main Table
  "FROM conc_time_values a ",

  # Join with series table by series ID
  "LEFT JOIN series b ON a.fk_series_id = b.id ",

  # Join to studies table by study ID
  "LEFT JOIN studies c ON b.fk_study_id = c.id ",

  # Join to subjects table by subject ID
  "LEFT JOIN subjects d ON b.fk_subject_id = d.id ",

  # Join to documents table by extraction document ID
  "LEFT JOIN documents e ON c.fk_extraction_document_id = e.id ",

  # # Join to dictionary tables
  # "LEFT JOIN administration_form_dict f ON c.fk_administration_form_id = f.id ",
  # "LEFT JOIN administration_method_dict g ON c.fk_administration_method_id = g.id ",
  "LEFT JOIN administration_route_dict h ON c.fk_administration_route_id = h.id ",
  "LEFT JOIN conc_medium_dict i ON b.fk_conc_medium_id = i.id ",
  "LEFT JOIN dose_frequency_dict j ON c.fk_dose_frequency_id = j.id ",

  # Rename chemical fields for dosed vs. analyzed chemical record foreign keys
  "LEFT JOIN (SELECT id, dsstox_substance_id AS dosed_chem_dtxsid, ",
  "chemical_name_original AS dosed_chem_name_original, ",
  "chemical_name_secondary_original AS dosed_chem_name_secondary_original, ",
  "dsstox_casrn AS dosed_chem_casrn, preferred_name AS dosed_chem_name ",
  "FROM chemicals) AS k ON c.fk_dosed_chemical_id = k.id ",
  "LEFT JOIN (SELECT id, dsstox_substance_id AS analyzed_chem_dtxsid, ",
  "chemical_name_original AS analyzed_chem_name_original, ",
  "chemical_name_secondary_original AS analyzed_chem_name_secondary_original, ",
  "dsstox_casrn AS analyzed_chem_casrn, preferred_name AS analyzed_chem_name ",
  "FROM chemicals) AS l ON b.fk_analyzed_chemical_id = l.id ",

  # Filtering steps
  "WHERE h.administration_route_normalized IN ('iv', 'oral') AND ",
  "(i.conc_medium_normalized IN ('blood', 'plasma') OR ",
  "i.conc_medium_original IN ('blood', 'plasma')) AND ",
  "(j.dose_frequency_normalized IN ('Once') OR e.curation_set_tag IN ('CVT_Showa'))"
)

sql_query <- query_db(query = query, schema = "cvt")

# Need to change the column name for this
names(sql_query)[1] <- "conc_time_id"

cvtdb_original <- sql_query
save(cvtdb_original, file = "data/cvtdb_original.rda", compress = "bzip2")
```

```{r harmonizing}
log_file <- "data-raw/creating_cvt_object.txt"
cli_print(
  "Creating the `cvt` data object for invivoPKfit, after SQL query.",
  file = log_file,
  append = FALSE # Overwrite file
)

load("data/cvtdb_original.rda")
sql_query <- cvtdb_original
rm(cvtdb_original)

columns_to_numeric <- c(
  "time_original",
  "time_hr",
  "conc_original",
  "conc_sd_original",
  "conc",
  "conc_sd",
  "dose_level_original",
  "dose_level_normalized",
  "loq"
)

# Harmonize values for easier analysis
cvt_df <- sql_query %>%
  mutate(
    # Note that sometimes fk_test_chemical_id is NA
    across(all_of(columns_to_numeric), as.numeric),
    # conc, conc_sd_original, and conc_original
    # have NA/NE/ND values that are coerced to NAs
    year = as.character(year),
    species = tolower(species),
    sex = tolower(sex),
    across(contains("chem_name"), \(x) {
      str_to_lower(x) |>
        force_ascii_quotes() |>
        # trim whitespace around hyphens
        str_replace_all("\\s*-\\s*", "-") |>
        # comma-separate forms like 3'6 into 3',6
        str_replace_all("(?<=\\d)'(?=\\d)", "',") |>
        # fix some bracket cases
        str_replace_all("\\[3\\]h", "[3h]") |>
        str_replace_all("\\[a\\]pyr", "(a)pyr") |>
        str_remove_all(" \\(aids\\)$") |>
        # adds a space after square bracket separator
        str_replace_all("](?!\\s|-)", "] ")
    }),
    dose_volume_units = ifelse(
      dose_volume_units %in% c("NR", "ND"),
      NA_character_,
      gsub(" / ", "/", gsub("ml", "mL", dose_volume_units))
    ),
    dose_volume = signif(as.numeric(dose_volume), digits = 7),
    dose_duration_units = trimws(dose_duration_units),
    dose_duration = as.numeric(str_extract(dose_duration,
      pattern = "^\\d+"
    )),
    dose_frequency_normalized = "Once", # Showa is not annotated
    radiolabeled = ifelse(is.na(radiolabeled), FALSE, as.logical(radiolabeled)),
    # for n_subject ranges, take integer value of median
    # otherwise pull number or NA
    n_subjects_normalized = case_when(
      n_subjects_in_series %in% "NR" ~ NA_real_,
      grepl("-", n_subjects_in_series) ~ floor(mean(
        as.numeric(unlist(strsplit(n_subjects_in_series, split = "-"))),
        na.rm = TRUE
      )),
      .default = as.numeric(
        str_extract(n_subjects_in_series, pattern = "\\d+")
      )
    )
  )

qm <- quick_message(c(
  "Column changes:",
  "Set {.var {columns_to_numeric}} to numeric type.",
  "Set {.var year} to character type.",
  "Set {.var {c('species', 'sex')}} to lowercase.",
  "Set {.var radiolabeled} to logical.",
  "Normalized some chemicals names. (See code for details)",
  "If `dose_volume_units` was 'NR' or 'ND' set this to {NA_character_}.",
  "Replaced ' / ', with '/' and 'ml' with 'mL' for `dose_volume_units`.",
  "Set `dose_duration` to first number and converted to numeric.",
  "Trimmed whitespace around `dose_duration_units`.",
  "Extracted integer average from `n_subjects_in_series` and set 'NR' to NA."
)) |> cli_fmt()

cli_print(qm)
cli_print(qm, file = log_file, append = TRUE)
```

## Some data malformations

There are some duplications in chemical entries such that not all unique chemical ids
represent a unique DTXSID, but rather a unique original name entry.
To fix this, when DTXSIDs for dosed and analyzed chemicals match, I will
be setting `fk_dosed_chemical_id` to `fk_analyzed_chemical_id`.

Note that we can "de-duplicate" these in two ways. Take for instance:

|analyzed_chem_id|dosed_chem_id|analyzed_dtxsid|dosed_dtxsid  |
|---------------:|:------------|--------------:|:-------------|
| 893            | 892         | DTXSID0021383 | DTXSID0021383|
| 929            | 893         | DTXSID0021383 | DTXSID0021383|

This is duplicated across rows and between rows! Ideally, we would take the
"most complete" record if we are convinced that these are the same chemical.

There is also some radiolabeling experiments that aren't described as such.
We can use `str_detect(x,"(13|14)c")` to recode `radiolabeled = TRUE`
for these. Then I filter out radiolabeled data because currently _invivoPKfit_
does not handle these.

By hand, I can add back the DTXSIDs for about 15 chemicals. Some of these will
be ChEMBLIDs actually.

```{r malformed_data}
# First quick check to see if any chemical_ids are NA
# Should be FALSE for both of these!
any(is.na(cvt_df$fk_analyzed_chemical_id))
any(is.na(cvt_df$fk_dosed_chemical_id))

cli_text("{qcount(cvt_df$analyzed_chem_dtxsid)} unique analyzed_dtxsids")

cli_text("{qcount(cvt_df$dosed_chem_dtxsid)} unique dosed dtxsids")

# Fix Radiolabeled ones!
cvt_df %>%
  distinct(
    fk_dosed_chemical_id,
    analyzed_chem_name_original,
    dosed_chem_name_original,
    radiolabeled
  ) %>%
  filter(
    !radiolabeled,
    str_detect(dosed_chem_name_original, "((13|14)c|3h)") |
      str_detect(analyzed_chem_name_original, "((13|14)c|3h)")
  )

cvt_df <- cvt_df %>%
  mutate(
    radiolabeled = ifelse(
      str_detect(dosed_chem_name_original, "((13|14)c|3h)") |
        str_detect(analyzed_chem_name_original, "((13|14)c|3h)"),
      TRUE,
      radiolabeled
    )
  )
# There is a distinct challenge when handling radiolabelled data but these will be
# filtered out later for our purposes.
# Pre-emptive filtering to cut down workload
cvt_df <- filter(cvt_df, !(radiolabeled %in% TRUE))

qm <- quick_message(c(
  "Found chemicals that were missidentified as not radiolabeled.",
  "Dosed Chemical IDs = (710, 1108)",
  "Filtered out radiolabeled data."
)) |> cli_fmt()
cli_print(qm)
cli_print(qm, file = log_file, append = TRUE)

cvt_df |>
  distinct(!!!chem_id_cols) |>
  count(is.na(analyzed_chem_dtxsid), is.na(dosed_chem_dtxsid))
# Most have both dosed and chemical dtxsid
# Are the 'NA' values for analyzed_chem_dtxsid that have dosed_chem_dtxsid the same?

name_misspellings <- c(
  "paracetmol" = "paracetamol",
  "deacetyl diltiazem" = "desacetyl diltaiazem",
  "deacetyl n-monodesmethyl diltiazem" = "desacetyl n-monodesmethyl diltiazem",
  "2,2',4,4',5,5'-hexachlorobiphenyl" = "2,4,5,2',4',5'-hexachlorobiphenyl",
  "valproate \\(vpa\\)" = "valproic acid",
  "clotiapine" = "clothiapine"
)

found_dtxsids <- c(
  "4,4'-sulfonylbisphenol-d8" = "DTXSID3022409",
  "(+-)-metoprolol tartrate" = "DTXSID9037248",
  "antipyrene" = "DTXSID6021117",
  "3'-azido-3'-deoxythymidine" = "DTXSID8020127",
  "2,4,5,2',4',5'-hexachlorobiphenyl" = "DTXSID2032180",
  "fluorotelomer acid 7+3" = "DTXSID20874028",
  "valproic acid" = "DTXSID6023733",
  "s-(-)-metoprolol" = "DTXSID60230861",
  "r-(+)-metoprolol" = "DTXSID20230862",
  "paracetamol" = "DTXSID2020006"
)

cvt_df <- cvt_df |>
  mutate(
    # fix misspellings
    analyzed_chem_name_original = str_replace_lst(
      analyzed_chem_name_original,
      name_misspellings
    ),
    dosed_chem_name_original = str_replace_lst(
      dosed_chem_name_original,
      name_misspellings
    ),
    # Missing DTXSIDs I found
    dosed_chem_dtxsid = case_match_lst(
      dosed_chem_name_original,
      found_dtxsids,
      dosed_chem_dtxsid
    ),
    analyzed_chem_dtxsid = case_match_lst(
      analyzed_chem_name_original,
      found_dtxsids,
      analyzed_chem_dtxsid
    ),
    # check name matches
    analyzed_chem_dtxsid = ifelse(
      is.na(analyzed_chem_dtxsid) &
        analyzed_chem_name_original == dosed_chem_name_original,
      dosed_chem_dtxsid,
      analyzed_chem_dtxsid
    )
  )

qm <- cli_fmt({
  cli_par()
  cli_text("Fixed the following misspellings [old: new] :")
  cli_dl(name_misspellings)
  cli_par()
  cli_text("Added missing DTXSIDs:")
  cli_dl(found_dtxsids)
  cli_end()
})
cli_print(qm)
cli_print(qm, file = log_file, append = TRUE)

quick_stats(cvt_df)
# Note unique chemical counts are

cvt_df |>
  distinct(!!!chem_id_cols) |>
  filter(is.na(analyzed_chem_dtxsid)) |>
  View("ID-less")

# Anything without an ID for analyzed_chemical is likely a metabolite/control of sorts
# Or justifiably undetermined...
# So... I filter it out.
cvt_df <- cvt_df |>
  filter(!is.na(analyzed_chem_dtxsid))

cli_print(
  cli_fmt({
    quick_message("Filtered out observations where `analyzed_chem_dtxsid` is still NA.")
  }),
  file = log_file,
  append = TRUE
)

quick_stats(cvt_df)
```

Now we need to preemptively address the chemicals that have different
IDs but actually represent the same chemical during the experiment
This includes primarily: SALTS and stereoisomers

**Small note about stereoisomers:**  
Here I give an un-elaborate treatment to racemic mixtures,
both stereoisomers/enantiomers are assumed to be 100% of the dose.
This isn't physically possible but the assumption would be that for each 
analyzed stereoisomer, the dose is representing the maximum possible dose in those
cases. For the purposes of _invivoPKfit_ the complication lies in possible
overestimation of elimination or underestimation of absorption.
Since this is rare in the data, I note this caveat but keep the data.

```{r salts-enantiomers}
cvt_df %>%
  filter(
    fk_analyzed_chemical_id != fk_dosed_chemical_id,
    analyzed_chem_dtxsid != dosed_chem_dtxsid
  ) %>%
  distinct(
    analyzed_chem_name_original, dosed_chem_name_original,
    fk_analyzed_chemical_id, fk_dosed_chemical_id,
    analyzed_chem_dtxsid, dosed_chem_dtxsid
  ) %>%
  arrange(analyzed_chem_name_original) %>%
  # head()
  View()

# The other issue with salts is sometimes they will appear in analyzed chemical names!
# Compose regular expressions looking for possible salts.
mcations <- paste(
  c(
    "(hydrogen", "lithium", "sodium", "potassium", "rubidium", "cesium",
    "silver", "beryllium", "magnesium", "strontium", "barium", "radium",
    "zinc", "calcium", "aluminum)"
  ),
  collapse = "|"
)
manions <- paste(
  c(
    "(hydride", "fluoride", "chloride", "bromide", "iodide", "oxide",
    "sulfide", "selenide", "telluride", "nitride", "phosphiide",
    "arsenide)"
  ),
  collapse = "|"
)

pa_ions <- paste(
  c(
    "(ammonium", "nitrate", "nitrite", "chromate", "dichromate", "cyanide",
    "permanganate", "hydroxide", "peroxide", " amide", "carbonate", "sulfate",
    "sulfite", "oxalate", "phosphate", "phosphite", "thiosulfate",
    "arsenate", "selenate", "silicate", "tartrate", "perchlorate", "acetate",
    "chlorate", "chlorite", "hypochlorite", "periodate", "iodate", "hypoiodite",
    "bromate", "hypobromite", "binoxolate", "borate", "hydrochloride", "thiocyanate)"
  ),
  collapse = "|"
)

cvt_df |>
  distinct(!!!chem_id_cols) |>
  filter(
    str_detect(analyzed_chem_name_original, pattern = mcations) |
      str_detect(analyzed_chem_name_original, pattern = manions) |
      str_detect(analyzed_chem_name_original, pattern = pa_ions) |
      str_detect(analyzed_chem_name, pattern = mcations) |
      str_detect(analyzed_chem_name, pattern = manions) |
      str_detect(analyzed_chem_name, pattern = pa_ions) |
      str_detect(dosed_chem_name_original, pattern = mcations) |
      str_detect(dosed_chem_name_original, pattern = manions) |
      str_detect(dosed_chem_name_original, pattern = pa_ions) |
      str_detect(dosed_chem_name, pattern = mcations) |
      str_detect(dosed_chem_name, pattern = manions) |
      str_detect(dosed_chem_name, pattern = pa_ions)
  ) |>
  View("Potential salt errors")

name_change_dtxsid <- c(
  "methadone hydrochloride" = "DTXSID7023273",
  "ammonium perfluorooctanoate" = "DTXSID40892486",
  "sodium nitrite" = "DTXSID5024219",
  "potassium perfluorobutanesulfonate" = "DTXSID5030030", # acid is more stable
  "potassium perfluorohexanesulfonate" = "DTXSID80873012",
  "bosentan" = "DTXSID7046627",
  "hexobarbital" = "DTXSID9023122",
  "quinidine" = "DTXSID4023549",
  "tamoxifen citrate" = "DTXSID1034187",
  "sodium nitrite" = "DTXSID5024219",
  "imipramine hydrochloride" = "DTXSID1043881",
  "ondansetron hydrochloride dihydrate" = "DTXSID8023393",
  "thiopental sodium" = "DTXSID1023653",
  "phenobarbital" = "DTXSID5021122",
  "phenobarbital sodium" = "DTXSID5021122"
)

original_name_change_dtxsid <- c(
  "formetanate hydrochloride" = "DTXSID9041990",
  "propamocarb hydrochloride" = "DTXSID1040295",
  "pyrithiobac sodium" = "DTXSID3037703",
  "tamoxifen citrate" = "DTXSID1034187",
  "valproic acid" = "DTXSID6023733",
  "r-(+)-metoprolol" = "DTXSID20230862",
  "s-(-)-metoprolol" = "DTXSID60230861"
)

# Method: To make DTXSID == DTXSID, salt DTXSIDs will be replaced with
# non-salt DTXSIDs.
cvt_df <- cvt_df |>
  mutate(
    analyzed_chem_dtxsid = case_match_lst(
      analyzed_chem_name,
      name_change_dtxsid,
      analyzed_chem_dtxsid
    ),
    dosed_chem_dtxsid = case_match_lst(
      analyzed_chem_name,
      name_change_dtxsid,
      dosed_chem_dtxsid
    ),
    dosed_chem_dtxsid = case_match_lst(
      dosed_chem_name,
      name_change_dtxsid,
      dosed_chem_dtxsid
    ),
    analyzed_chem_dtxsid = case_match_lst(
      analyzed_chem_name_original,
      original_name_change_dtxsid,
      analyzed_chem_dtxsid
    ),
    dosed_chem_dtxsid = case_match_lst(
      analyzed_chem_name_original,
      original_name_change_dtxsid,
      dosed_chem_dtxsid
    )
  ) # This is imperfect but will do.

qm <- cli_fmt({
  this_par <- cli_par()
  cli_text("Changed DTXSIDs based on chemical names to match unsalted chemical:")
  cli_dl(name_change_dtxsid)
  cli_end(this_par)
  that_par <- cli_par()
  cli_text("Changed DTXSID based on original chemical name to match unsalted chemical:")
  cli_dl(original_name_change_dtxsid)
  cli_end(that_par)
})
cli_print(qm)
cli_print(qm, file = log_file, append = TRUE)

cvt_df |>
  distinct(analyzed_chem_dtxsid, dosed_chem_dtxsid) |>
  count(analyzed_chem_dtxsid == dosed_chem_dtxsid)

## Now that we've added some missing DTXSIDs, redo the chemical id
cvt_df <- cvt_df %>%
  rowwise() %>%
  mutate(fk_dosed_chemical_id = ifelse(
    identical(analyzed_chem_dtxsid, dosed_chem_dtxsid),
    fk_analyzed_chemical_id,
    fk_dosed_chemical_id
  )) %>%
  ungroup()

qm <- quick_message(
  "Now, when dosed and analyzed chemical DTXSIDs are identical, replace fk_dosed_chemical_id with fk_analyzed_chemical_id."
) |> cli_fmt()
cli_print(qm, file = log_file, append = TRUE)

any(is.na(cvt_df$fk_analyzed_chemical_id))
any(is.na(cvt_df$fk_dosed_chemical_id))

quick_stats(cvt_df)
```

# Validating data

This step in the loading process allows us to correct some mis-annotations in the
data, without directly changing CvTdb (the review process there is more rigorous).
Some of these changes will be sent as a ticket item for the CvTdb team.
Other times, the annotation is correct, it's just that units need to be standardized
for `invivoPKfit` specifically.

There are some procedures needed to be taken to standardized some of this data
such that dose units are all in mg/kg of bodyweight and concentration units are in
ug/L.

## Preliminary Filtering

I begin with a couple of filtering steps:

+ Exclude data without any concentration value  
+ Exclude data with only one observation per *study_id*

```{r preliminary_filtering}
cvt_df <- cvt_df %>%
  filter(
    !is.na(analyzed_chem_dtxsid),
    (fk_analyzed_chemical_id == fk_dosed_chemical_id)
  ) %>%
  distinct()

qm <- quick_message(text = c(
  "Filtering out:",
  "Missing {.var analyzed_chem_dtxsid}",
  "Data where the analyzed chemicals and dosed chemicals are different"
)) |>
  cli_fmt()
cli_print(qm)
cli_print(qm, file = log_file, append = TRUE)

# Filtering out non-circulation mediums (these are retained in cvtdb_original)
cvt_df <- filter(cvt_df, conc_medium_normalized %in% c("blood", "plasma"))

qm <- quick_message(text = c(
  "Only keeping:",
  "Plasma or blood medium collected."
)) |>
  cli_fmt()

cli_print(qm)
cli_print(qm, file = log_file, append = TRUE)

# Filter out where all values in experiment are missing values/below LOQ
# Also filter out observations where
cvt_df <- cvt_df %>%
  mutate(
    conc = ifelse(conc < 0, NA_real_, conc),
    conc_original = ifelse(conc_original < 0, NA_real_, conc_original)
  ) %>%
  group_by(
    analyzed_chem_dtxsid, species, fk_extraction_document_id,
    administration_route_normalized, conc_medium_normalized
  ) %>%
  filter(
    !(if_all(conc, is.na) & if_all(conc_original, is.na))
  ) %>%
  ungroup() %>%
  distinct()

qm <- quick_message(text = c(
  "Filtering out:",
  "Experiments missing concentration values altogether."
)) |>
  cli_fmt()

cli_print(qm)
cli_print(qm, file = log_file, append = TRUE)

quick_stats(cvt_df)

# Need to see all the conc, conc_sd, dose, loq (original, values and units)
# Dose Units: Need mg/kg... = ug/mg = ng/ug, and kg = L
# Conc Units: Need mg/L... = ug/mL = ng/uL = pg/nL, and kg = L, g = mL
cvt_df <- cvt_df %>% mutate(
  conc_units_original = str_replace(
    conc_units_original,
    pattern = "l$",
    replacement = "L"
  ),
  conc_units_original = str_replace(
    conc_units_original,
    pattern = "\u00B5",
    replacement = "u"
  ),
  conc_units_normalized = str_replace(
    conc_units_normalized,
    pattern = "l$",
    replacement = "L"
  ),
  loq_units = str_replace(loq_units, "l$", "L"),
  invivPK_dose_level_units = "mg/kg",
  invivPK_conc_units = "ug/mL"
)

# Filtering out observations with no concentration information
cvt_df <- cvt_df |>
  filter(!(is.na(conc_original) & is.na(conc)))

qm <- quick_message(c(
  "Preparing concentration (and LOQ) units for validation:",
  "Replacing lowercase 'l' at end of unit into upper case 'L' for liters.",
  "Replacing '{'\u00B5'}' character with 'u' to standardize unit labels.",
  "Filtering out any observations with NA values in both `conc` and `conc_original`."
)) |>
  cli_fmt()
cli_print(qm)
cli_print(qm, file = log_file, append = TRUE)
```

For the majority of compounds with a missing `conc` value, it is usually a 
factor of converting `conc_units_original` to the standard `conc_units_normalized`.

While I leave most of this curation work to the CvTdb team which have been working
on this data since 2020 at least, I will try to wade through the data for
the following 5 chemicals with large amounts of data that is missing.

1. _tetralin/DTXSID1026118_: has `conc_units_original` in `ug/g` of plasma concentration.
Blood plasma has a density of approximately `1.025 g/mL` which means that the values
for `conc_original` and `loq` can be multiplied by this value to give the standard
`ug/mL`.

2. _PCB 136/DTXSID6073499_: has some `conc_units_original` in `ng/g` of whole blood
concentration. Whole blood has a density of approximately `1.05 g/mL` which means
we can multiply the concentrations by $1.05\frac{g}{mL}\times 10^{-3}\frac{\mu g}{ng}$.
For simplicity, I will discard the observations where `conc_units_original` is
given in `molar percentage`.

3. _PFOA/DTXSID40892486_: has some `conc_units_original` in `ug/g` of plasma.
I will apply the same formula as #1 above.

4. _propyzamide/DTXSID2020420_: this, and all chemicals in `pmid == 29385628`
have concentration over time values of `mg/kg`, but these are actually the
`ug/mL`. This is apparent in the Supplementary Table 1 of this paper. The `loq`
values are carefully calibrated, but should all be in `ng/mL` units.

5. _carbon disulfide/DTXSID6023947_: similar to #2 the `conc_units_original` is
in `ug/g` of whole blood, which means we can simply multiply `conc_original`
by `1.05 g/mL` to supply `conc`.

In general, if we have missing `conc` values, we are only interested in those
that have a mass value (i.e. kg, g, ug) in the denominator, so I will begin
by filtering out all other observations.

```{r}
# Densities in units: g/ml
WB_DENSITY <- 1.05
PL_DENSITY <- 1.025

# This removes around 100 observations!
cvt_df <- cvt_df |>
  filter(!is.na(conc) | str_detect(conc_units_original, "g$"))
qm <- quick_message(c(
  "Only keeping:",
  "Observations with non-NA `conc` values.",
  "Observations with `conc_units_original` ends with 'g'."
)) |>
  cli_fmt()
cli_print(qm)
cli_print(qm, file = log_file, append = TRUE)

cvt_df <- cvt_df |>
  mutate(
    conc = case_when(
      # Carry over values when pmid == 29385628
      is.na(conc) &
        pmid == 29385628L &
        conc_units_original == "mg/kg" ~ conc_original,
      is.na(conc) &
        conc_medium_normalized == "plasma" &
        conc_units_original == "ug/g" ~ conc_original * PL_DENSITY,
      is.na(conc) &
        conc_medium_normalized == "blood" &
        conc_units_original == "ug/g" ~ conc_original * WB_DENSITY,
      is.na(conc) &
        conc_medium_normalized == "blood" &
        conc_units_original == "ng/g" ~ conc_original * WB_DENSITY * 1E-3,
      is.na(conc) &
        conc_medium_normalized == "plasma" &
        conc_units_original == "ng/g" ~ conc_original * PL_DENSITY * 1E-3,
      .default = conc
    )
  )

qm <- quick_message(c(
  "Calculating missing concentrations when `conc_units_original` is per *grams:",
  "When blood is measured, multiply by {WB_DENSITY} and metric scaling factor.",
  "When plasma is measured, multiply by {PL_DENSITY} and metric scaling factor."
)) |>
  cli_fmt()
cli_print(qm)
cli_print(qm, file = log_file, append = TRUE)

quick_stats(cvt_df)

# When LOQ values are present, the units are always specified
# There is one case that is at first glance not an easy conversion, nmol/mL,
# but TAME has a mass of 102.104465 g/mol. LOQ value in ug/mL is approximately the
# same and we will treat it as such
# There is another case where loq_units is recorded as "ng". I will treat this as "ng/mL"

cvt_df <- cvt_df |>
  mutate(
    loq_units = case_match(
      other_study_identifier,
      "K02922" ~ "ng/mL",
      .default = loq_units
    )
  ) |> 
  mutate(
    loq = case_when(
      loq_units %in% c("ug/L", "ng/mL", "ng") ~ loq * 1E-3,
      loq_units == "pg/mL" ~ loq * 1E-6,
      loq_units == "ng/g" & conc_medium_normalized == "blood" ~ loq * WB_DENSITY * 1E-3,
      loq_units == "ug/g" & conc_medium_normalized == "blood" ~ loq * WB_DENSITY,
      loq_units == "ug/g" & conc_medium_normalized == "plasma" ~ loq * PL_DENSITY,
      .default = loq
    ),
    loq_units = "ng/uL"
  )

qm <- quick_message(c(
  "Calculating missing LOQs:",
  "For NTP study 'K02922', loq units are actually 'ng/mL'",
  "When units are in 'pg/mL', multiply by 10^-6",
  "When units are in 'ug/L', 'ng/mL', or 'ng', multiply by 10^-3",
  "When blood is measured and `loq_units` are per gram, multiply by {WB_DENSITY} and metric scaling factor.",
  "When plasma is measured and `loq_units` are per gram, multiply by {PL_DENSITY} and metric scaling factor."
)) |>
  cli_fmt()
cli_print(qm)
cli_print(qm, file = log_file, append = TRUE)

quick_stats(cvt_df)
```

Per the CvTdb 2.0.0 release notes, Dose is normalized to `mg/kg BW`.
However, there are some with missing values, representing 180 observations across
3 chemicals (by DTXSIDs): quinidine, clothiapine, and padsevonil. I will simply 
exclude these in this case.

```{r handling_dose}
cvt_df <- filter(cvt_df, dose_level_units_normalized == "mg/kg BW")
cli_print(
  cli_fmt({
    quick_message("Removing observations with missing dose values.")
  }),
  file = log_file,
  append = TRUE
)

quick_stats(cvt_df)
```

```{r plot_by_dtxsid}
cvt_df %>%
  ggplot(aes(
    x = time_hr, y = conc, group = analyzed_chem_dtxsid,
    color = analyzed_chem_dtxsid
  )) +
  geom_point() +
  geom_smooth(se = FALSE) +
  theme_bw() +
  scale_y_log10() +
  theme(legend.position = "none")
```

```{r final_checks}
cvt_df %>%
  count(
    is.na(dose_level_normalized),
    is.na(dose_level_original)
  ) %>%
  glimpse()

cvt_df %>%
  count(is.na(conc_sd) & n_subjects_normalized > 1)

cvt_df %>%
  count(is.na(n_subjects_normalized))

cvt_df <- cvt_df |> select(!starts_with("invivPK_"))
cvt_df <- cvt_df |> filter(time_hr >= 0)
qm <- quick_message("Filtering out all non-positive or non-zero timepoints.") |> 
  cli_fmt()
cli_print(qm)
cli_print(qm, file = log_file, append = TRUE)

# Eliminate all observations with unique timepoints below 5
low_observation_chem_names <- cvt_df |> 
  group_by(analyzed_chem_name_original) |> 
  summarize(observation_count = qcount(time_hr)) |>
  arrange(observation_count) |> 
  filter(observation_count < 5) |> 
  pull(analyzed_chem_name_original)

cvt_df <- cvt_df |> 
  filter(!analyzed_chem_name_original %in% low_observation_chem_names)

qm <- cli_fmt({
  cli_par()
  cli_text("Removing chemical entries with less than 5 unique timepoints.")
  cli_text("This eliminates the following chemicals:")
  cli_ul(low_observation_chem_names)
  cli_end()
})
cli_print(qm)
cli_print(qm, file = log_file, append = TRUE)

```

### Changes to mapping!
We now use `conc`, `conc_units_normalized`, `loq`, `loq_units`,
`dose_level_units_normalized`, and `dose_level_normalized` directly.

Save the pulled CVT data with the edits/normalizations, and also save the current date.

```{r saving_data, include=FALSE, eval=FALSE}
cvt_date <- Sys.Date()

cvt <- cvt_df
save(cvt, file = "data/cvt.rda", compress = "bzip2")
save(cvt_date, file = "data/cvt_date.rda")
cli_print(
  cli_fmt({quick_stats(cvt)}),
  file = log_file,
  append = TRUE
)

cli_print(
  paste0("Created: ", cvt_date),
  file = log_file,
  append = TRUE
)

```

# QSAR Bakeoff Run

```{r running}
my_pk <- pk(
  cvt,
  mapping = ggplot2::aes(
    Chemical = analyzed_chem_dtxsid,
    Chemical_Name = analyzed_chem_name_original,
    DTXSID = analyzed_chem_dtxsid,
    CASRN = analyzed_chem_casrn,
    Species = species,
    Reference = fk_extraction_document_id,
    Media = conc_medium_normalized,
    Route = administration_route_normalized,
    Dose = dose_level_normalized,
    Dose.Units = "mg/kg",
    Subject_ID = fk_subject_id,
    Series_ID = fk_series_id,
    Study_ID = fk_study_id,
    ConcTime_ID = conc_time_id,
    N_Subjects = n_subjects_normalized,
    Weight = weight_kg,
    Weight.Units = "kg",
    Time = time_hr,
    Time.Units = "hours",
    Value = conc,
    Value.Units = "mg/L",
    Value_SD = conc_sd,
    LOQ = loq
  )
)

my_pk <- my_pk +
  stat_error_model(Chemical, Species) +
  settings_optimx(method = "bobyqa") +
  facet_data(Chemical, Species) +
  settings_preprocess(
    keep_data_original = TRUE,
    suppress.messages = TRUE
  ) +
  scale_conc(dose_norm = FALSE, log10_trans = FALSE)

my_pk <- do_preprocess(my_pk)
apply(my_pk$data, 2, \(x) {
  any(is.na(x))
})

my_pk <- do_data_info(my_pk)
my_pk <- do_prefit(my_pk)

mirai::daemons(10L)
mirai::everywhere(devtools::load_all())
my_pk <- do_fit(my_pk, async = TRUE)
mirai::daemons(0L)

fit_pl <- plot(my_pk, use_scale_conc = FALSE)

fit_pl %>%
  filter(Chemical %in% "DTXSID1021116", Species %in% "human") %>%
  pull(final_plot) %>%
  .[[1]]
fit_pl %>%
  filter(Chemical %in% "DTXSID9025617", Species %in% "mouse") %>%
  pull(final_plot) %>%
  .[[1]]
fit_pl %>%
  filter(Chemical %in% "DTXSID9025617", Species %in% "rat") %>%
  pull(final_plot) %>%
  .[[1]]
fit_pl %>%
  filter(Chemical %in% "DTXSID2021781", Species %in% "mouse") %>%
  pull(final_plot) %>%
  .[[1]]
fit_pl %>%
  filter(Chemical %in% "DTXSID1043881") %>% # Imipramine sanity check
  pull(final_plot) %>%
  .[[1]]
fit_pl %>%
  filter(Chemical %in% "DTXSID1023653") %>% # Thiopental sanity check
  pull(final_plot) %>%
  .[[1]]
fit_pl %>%
  filter(Chemical %in% "DTXSID5021122") %>% # Phenobarbital sanity check
  pull(final_plot) %>%
  .[[1]]
fit_pl %>%
  filter(Chemical %in% "DTXSID80873012") %>% # PFHS sanity check
  pull(final_plot) %>%
  .[[1]]
fit_pl %>%
  filter(Chemical %in% "DTXSID20230862") %>% # Metoprolol sanity check
  pull(final_plot) %>%
  .[[1]]
fit_pl %>%
  filter(Chemical %in% "DTXSID60230861") %>% # Metoprolol sanity check
  pull(final_plot) %>%
  .[[1]]
fit_pl %>%
  filter(Chemical %in% "DTXSID8031865") %>% # PFOS sanity check
  pull(final_plot) %>%
  .[[1]]

final_plot <- fit_pl |> pull(final_plot)
pdf(paste0(Sys.getenv("FIG_DIR"), "20250812_plots.pdf"))
for (this_plot in seq_len(length(final_plot))) {
  print(final_plot[[this_plot]])
}
dev.off()

get_winning_model(my_pk) %>% filter(Chemical %in% "DTXSID1034187")
get_fit(my_pk) |>
  distinct(Chemical, Species, model, convergence) |>
  count(model, convergence)

eval_tkstats(my_pk) |> glimpse(80)
eval_tkstats(my_pk) |>
  pull(DATA_GROUP_ID) |>
  unique() |>
  length()

save(my_pk, file = "data-raw/temp_pk.rda")

my_tkstats <- eval_tkstats(my_pk, dose_norm = TRUE)
my_fit <- get_fit(my_pk)
my_wm <- get_winning_model(my_pk)

my_fit |>
  select(DATA_GROUP_ID:param_name, estimate) |>
  inner_join(my_wm, by = c("Chemical", "Species", "method", "model")) |>
  select(-c(near_flat, preds_below_loq)) |>
  filter(str_detect(param_name, "sigma_", negate = TRUE)) |>
  mutate(
    estimate = purrr::map_dbl(
      estimate,
      \(x) ifelse(is.null(x), NA_real_, as.numeric(x))
    )
  ) |>
  tidyr::pivot_wider(names_from = param_name, values_from = estimate) |>
  inner_join(my_tkstats,
    by = join_by(DATA_GROUP_ID, model, method, Chemical, Species)
  ) |>
  readr::write_csv(
    file = paste0(
      Sys.getenv("FIG_DIR"),
      "20250805_pk000_evalTKstats.csv"
    )
  )
```