### 2025-08-22 data-raw/ClearanceProject-1-Data.Rmd

---
title: "Clearance Project - 1 - Data Generation"
author: "Gilberto Padilla Mercado"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

This vignette creates and describes data for downstream analysis.

```{r library_loading}
today <- format(Sys.Date(), "%d%b%Y")
devtools::load_all()
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(ctxR)
library(ComplexHeatmap, quietly = TRUE)
library(ggalluvial)
library(flextable)
library(reticulate)
library(patchwork)
source("data-raw/RC_comparison_bkgd.R") # httk & dawson_2021 loaded here
```

This code chunk creates and `.RData` file that contains character vectors
of DTXSIDs for:

1. All chemicals with human parameterization available for a 3-compartment model w/ inhalation.  
2. All chemicals with rat parameterization available for a 3-compartment model w/ inhalation.  
3. All chemicals with human and rat plasma concentration over time data.

```{r chemical-lists}
human_chems <- unique(get_cheminfo(
  info = "DTXSID",
  species = "human",
  model = "gas_pbtk",
  physchem.exclude = TRUE,
  suppress.messages = TRUE
)) %>%
  suppressWarnings()

rat_chems <- unique(get_cheminfo(
  info = "DTXSID",
  species = "rat",
  model = "gas_pbtk",
  physchem.exclude = TRUE,
  suppress.messages = TRUE
)) %>%
  suppressWarnings()

cvt_plasma_chems <- cvt %>%
  filter(
    species %in% c("human", "rat"),
    conc_medium_normalized == "plasma"
  ) %>%
  pull(analyzed_chem_dtxsid) %>%
  unique()

# These are the chemicals we can evaluate!
target_chems_human <- intersect(
  cvt_plasma_chems,
  human_chems
)
target_chems_rat <- intersect(
  cvt_plasma_chems,
  rat_chems
)

# save in an .Rdata file
save(target_chems_human, target_chems_rat,
  human_chems, rat_chems,
  cvt_plasma_chems,
  file = "data-raw/chem_lists.RData"
)
```

Below read in some external data which is publicly available already.

- `red_assay_res` is the named object that contains the data from the RED assay
including the dissociation constant $K_d$ for plasma protein binding.  
- `chem_info` is a combined data.frame from calls to `get_chemical_details_batch`
and `get_chem_info_batch`. This uses the `ctxR` API to fetch chemical information
such as physio-chemical properties and SMILES from the CompTox dashboard.  
- `cyp_data` is data from Sipes 2017 predicting the CLint fraction for each CYP enzyme.
- `functional_uses_all` is a list of functional use categories for each chemical
from the Chemical and Products Database.

```{r external_data}
# Here we load Kd/binding affinity data
red_assay_res <- readRDS("data-raw/2019-05-06-REDassay_HTTK.rds")

# ctxR download of all chemicals in gas_pbtk model
# Chemical info and properties (averaging the latter by chemical and experimental/predicted)
chem_info <- get_chemical_details_batch(
  DTXSID = human_chems,
  rate_limit = 0L
) %>%
  filter(!is.na(smiles))
# 9086 total unique chemicals with SMILES data

chem_properties <- get_chem_info_batch(
  DTXSID = chem_info$dtxsid,
  type = "",
  rate_limit = 0L,
  verbose = TRUE
)
# This takes a minute or so... save a reference RDS file to load
# Also get rid of any that don't have SMILES
chem_info %>%
  distinct(
    dtxsid, preferredName,
    qsarReadySmiles, monoisotopicMass
  ) %>%
  right_join(chem_properties) %>%
  filter(!is.na(qsarReadySmiles)) %>%
  saveRDS(file = "data-raw/RCdata/ctxR_3comp2_cheminfo.rds")

# Sipes 2017 data
cyp_data <- readr::read_csv(paste0(
  Sys.getenv("FIG_DIR"),
  "RestrictiveClearance_Analysis/",
  "CLint_CYP_Sipes2017.csv"
)) %>%
  left_join(
    chem.physical_and_invitro.data %>%
      select(CAS = CAS, Chemical = DTXSID) %>%
      distinct()
  ) %>%
  select(Chemical, CAS, starts_with("MET_")) %>%
  filter(!is.na(Chemical))
names(cyp_data) <- c(
  "Chemical", "CAS", "CYP1A2",
  "CYP2C9", "CYP2C19", "CYP2D6", "CYP3A4"
)
# Functional use categories from CPdat
functional_uses_all <- readr::read_csv(paste0(
  Sys.getenv("FIG_DIR"), "RestrictiveClearance_Analysis/",
  "ChemExpo_bulk_functional_uses_20250304.csv"
))

save(cyp_data,
  file = "data-raw/CYP_Sipes2017.Rdata"
)
```

Next we fit the plasma data from CvTdb, which we already have in the `invivoPKfit`
package using `cvt` object. We also get a few downstream statistics and save them
in an `.RData` file.

```{r inivopkfit}
# Here we fit CvTdb data that has plasma measurements
# Only use plasma for simplicity
this_pk <- pk(cvt %>%
  dplyr::filter(conc_medium_normalized == "plasma")) +
  settings_optimx(method = "bobyqa") +
  facet_data(vars(Chemical, Species)) +
  settings_preprocess(
    keep_data_original = TRUE,
    suppress.messages = FALSE
  ) +
  scale_conc(dose_norm = TRUE, log10_trans = TRUE) +
  stat_error_model(error_group = vars(Chemical, Species))

this_pk <- do_prefit(this_pk)
this_pk <- do_fit(this_pk, n_cores = 14)
saveRDS(this_pk, "data-raw/plasma_CvTfits.rds")

# Related files for pre-calculated:
# tkstats, winning model, AIC, coef_sd, predictions, etc...
this_winmodel <- get_winning_model(this_pk)

this_tkstats <- eval_tkstats(this_pk,
  model = c("model_1comp", "model_2comp"),
  finite_only = TRUE,
  dose_norm = TRUE
) %>%
  inner_join(this_winmodel)
colnames(this_tkstats) <- gsub("/", "_", colnames(this_tkstats))

winning_preds <- predict(this_pk) %>%
  semi_join(this_winmodel) %>%
  distinct(
    Chemical, Species, Chemical_Name,
    model, method, Dose, Media, Route, Reference, pLOQ, Time,
    N_Subjects, Conc, Conc_est
  )
this_AIC <- AIC(this_pk)
this_coef_win <- coef(this_pk) %>% semi_join(this_winmodel)
this_coef_fit <- this_pk$fit %>% semi_join(this_winmodel)
this_coef_sd_win <- coef_sd(this_pk) %>% semi_join(this_winmodel)

save(this_winmodel, this_tkstats, winning_preds,
  this_AIC, this_coef_win, this_coef_fit, this_coef_sd_win,
  file = "data-raw/plasma_CvTfits_data.RData"
)
```

 We already have the fits from the previous vignette!